### 2013-08-21 bandwidth.tex
Original paper, must run pdflatex.

\section{Bandwidth}
I measured bandwidth reduction using a Reverse Cuthill-McKee [], implemented 
in ViennaCL []. Two kinds of grids were considered. A 2D and a 3D Cartesian mesh of sizes ranging from $32^3$ to $64^3$ and $128^2$ to $256^2$. 
Stencils of size 32 and 64 were considered and computed using a Kd-Tree. 
(we also considered an overlayed grid (NO TIMINGS YET) and various forms of space filling curves. See Bollig thesis [].)  (There we no subdomain decompostion. In Bollig notation, the figures below correspond to a single 
square matrix.

\begin{verbatim}
  N2D      reduced bandwidth  (stencil size: 32)
  128^2       1031
  256^2       2055    (bw = 8 * N)

        (stencil size 64)
  128^2      1522
  256^2      3059     (bw = 12 * N)

  N3D   reduced bandwidth (stencil size 32)
  32^3   4902
  64^3  11184    (bw = 4.8 * N^2)

  N3D   reduced bandwidth (stencil size 64) 
  32^3   6573   
  64^3   26923  (bw = 6.5 * N^2)
\end{verbatim}

Matrix representations after bandwidth reduction. (NEED FIGURE.)

\section{Register Density}
Calculate Register Density (RD) (Saule et al.). For each row of the \ttt{ col\_id} matrix, compute the number $n_c$ of cache lines touched by all the nonzero elements in the vector $x$ $(Ax)$. Divide the number of nonzeros $nnz$ by the number of elements that can be held $n_c$  cachelines, to obtained the 
$$
   RD = \frac{nnz}{n_e n_c}
$$
Each cache line is 64 bytes, which holds 16 floats or 8 doubles.

We have computed the RD for a variety of 2D and 3D matrices with 32 or 64 nonzeros per row, that correspond to 2D ($32^2$, $64^2$, $128^2$ and $256^2$ grids, and 3D grids of size $32^3$, $64^3$, and $128^3$. The results are generated before and after bandwidth reduction (assume symmetric adjacency matrix (SHOW RESULTS: slightly better results than non-symmetric adjacency matrix (EXPLAIN IN PAPER.)

\begin{verbatim}
\end{verbatim}

### 2013-08-21 notation.tex
Original paper, must run pdflatex.

\section{Notation}
In this section, we introduce some variables that will prove useful to 
describe our experiments and construct a working model to examine the pros and cons of various assumptions. While deceptively simple, the MIC architecture requires a careful consideration of the properties of cache, memory bandwidth, cores and their interaction, threads per core (1-4), parallelization and vectorization. We will also seek to perform experiments that measure the best possible performance of the SPMV with maximum and minimum cache thrashing \qes{same as cache misses?}, with and without floating point operations, and with and without memory transfer. At the onset, we expect the dominant cost to e the transfer of the vector $x$ from memory, because contiguous elements of $x$ are not accessed sequentially.

Relevant notation is contained in Table~\ref{tab:not}.

\begin{center}
\begin{tabular}{|c|l|}
%\hline
%& & \\
\hline
$b_i$ & number of bytes per int \\
$b_x$ & number of bytes per float(4)/double(8) \\
$n_z$ & number of nonzeros per row $A$ \\
$n_r$ & number of rows of $A$ \\
$n_c$ & total number of elements in $col_id$: $n_z n_r$ \\
$b_w$ & matrix bandwidth \\
$b_c$ & total L2 cache per core ($512k=2^{19}$ bytes) \\
$b_T$ & theoretical minimum number of bytes transferred  \\
$b_{wT}$ & theoretical minimum number of bytes written to memory  \\
$b_{rT}$ & theoretical minimum number of bytes read from memory  \\
$n_C$ & total number of cores used  \\
$t_C$ & total number of threads used per core \\
$\rho$ & average cache density  \\
$n_v, n_m$ & number of vectors and matrices \\
$n_F$ & total number of floating point operations  \\
\hline
\end{tabular}
\end{center}

- weights and vector elements are either all floats or all doubles.

Let us first estimate the memory (in a serial implementation) required to compute $y = Ax$ where $y$ has $n_r$ rows and $n_m$ columns, $A$ has
$n_c=n_r n_z$ nonzero elements, and $x$ is a vector of $n_r$ rows and 
$n_v$ columns. In this paper, we assume that the adjacency matrix for the
$n_m$ matrices $A$ is constant, but the values stored in the various $A$ differ. We are therefore executing $n_v n_m$ spmv operations. The objective of course, is to minimize wall-clock time on the MIC.

The total number of bytes is 
$$
   b_T = n_r (n_z b_i + n_z b_x n_m + b_x n_v)
$$
The ratio of $b_T$ to the total cache  over all cores is given by
$$
   R = \frac{b_T}{b_c}
$$
Clearly, the memory $b_T$ is composed of the memory required by the vectors $x$, $y$, the nonzeros of $A$ and the elements of the matrix \ttt{col}, which stores the nonzero columns in each row, and is an integral component of the Ellpack compressed matrix format specification. While the Ellpack format is less general than CSR, it is the idea structure for the RBBF simulations we are intersted in, wherein the number of RBF nodes per stencil is constant throughout the 2D or 3D grid and in time.

Let us consider separately the number of bytes read from and written to memory.
Both the matrix $n_m$ matrices $A$ \qes{need an index on $A$ since there are more than one?} and \ttt{col\_id} are read from memory, as is $x$, whereas, the vector$y$ are stored to memory. In a later sections, we will discuss the practical performance of these operations and estimate the dominant contributions. 
\qes{We also wish to isolate the effects of read, write and compute operations as a function of the number of cores and threads to see how they influence the results and to estimate whether our results have the potential to scale similarly on systems with a higher number of cores. We "might" also compare are results against the best implementation using OpenCL, or do so in the paper.}

The total number of bytes read into memory is
$$ 
n_{rT} = n_r n_v b_x + n_r n_z b_i + n_r n_m n_z b_x
$$
while the total number of bytes written to memory is
$$
n_{wT} = n_r n_m n_v b_x
$$

Experiments similar to those performed in Saule \etal \cite{} suggest a maximum read  memory bandwidth of $190 Gbytes/sec$ using all threads on all four cores. The maximum speed advertised by Intel is $??? Gbytes/sec$, which we have found impossible to achieve in practice. To achieve $150 Gbytes/sec$ required allocating enough arrays to fill the 32 Gbytes of memory on the MIC, enabling prefetching, adding an explicit software prefetch, storing data as nrngo (prevent a rewrite to cache from memory prior to sending data to memory), and ensuring measurements in a steady-state regime.  With prefetching turned on, the maximum performances for each number of cores can be achieved with a single thread. However, when prefetching is turned off, performance increases as the number of active threads increases. Scaling is sublinear as a function of the number of cores (Fig. 1 in Saule \etal~\cite{}).  This suggests that when one computes the spmv, software prefetching might be useful to allow efficient transfer of data from memory {\em during\/} the calculation.

Writing of data is less efficient with a maximum rate of 150 Gbytes/sec when using the No-Read hint and  no global reordering. \qes{We will return later to this issue. Prefetching with irregular access may not be practical}.
With only Vectorization and No-Read Hint, only 100 Gbytes/sec are achieved. Performance increase as the number of threads per core is increased, and the scaling is linear in the number of cores (Fig. 1b in \cite{Saule}).

\qes{Perform your memory bandwidth experiments on my machine.}
\qes{How does Erik measure store and load separately?}

\subsection{Performance}
The best compute rate measured by Saule \etal~\cite{} is on the order of 20 Gflops for matrices 13 and 18, using $-O3$ compiler optimization \qes{icc compiler?} which have on the order of 100 and 200 average nonzero elements per row and a nonzero density on the order of $10^-3$. We work with matrices with density that ranges from $10^{-3}$ to $.3\times 10^{-5}$.  Lower density implies higher level of sparsity, and usually, higher sparsity leads to decreased performance \qes{may I say this?}. Based on Table 1, matrix 14 appears to have the closest characteristics to ours, with a constant number of nonzeros per row (41 nonzeros per row and a density of $10^{-5}$). From Figure 5 (Saule \etal), we see that matrix 22 has a cache line density between 0.15 and 0.25 . After bandwidth reduction, our matrices have a cache line density of about $0.15$, in single precision (I BELIEVE THIS WOULD double in double precision (not yet checked). All results in Saule \etal are in double precision \qes{can you confirm this?}.

In the benchmarks to be described, we will demonstrate spmv at $100 Gflops$, or 5x the best performance in Saule \etal by simultaneously computing four spmvs of identical size where the four matrices have the same adjacency structure, with each spmv applied to four vectors. This would correspond in an RBFFD code to computing four derivatives of four functions, resulting in 16 vectors.

\ge{I need a figure describing our three problems. Perhaps one figure for the 1x1 problem and one for the 4x4 problems.}
Rather than rely on the Intel compiler to perform the optimization, we program all our kernels in the Intel vector  syntax, combined with OpenMP pragmas. We will compare with straightforward implementations in C++. Naturally, the low level vector implementations cannot run on the host, or on previous generations of the Intel architecture. The advantage is that one gets a better feeling for the relationship between implementation and performance. If the Intel compiler is "good enough", better performance should be achieved, since, depending on the matrix parameters, more complex coding might be implemented, such as loop unrolling.

\bold{Vector operations} \\
All matrices are aligned to 64 bytes, the size of a cache line, and the size of the vector registers. 
Each core has 32 vector registers, shared by 4 hardware threads. 
\qes{Is only a single thread responsible for loading up a vector register? So the other threads are either in a wait state or loading another register, or initiating a load/store?} \qes{I believe only two hardware threads can be involved in operations? That means 4 ops per second (using MADD) if both threads are active.}

Registers on the MIC are 512 bits wide, and can store 16 floats or 8 doubles. It is through these registers that the \#SIMD pragma is able to achieve vectorization. (In comparison to OpenCL or CUDA, one can think of these registers as a warp.) 
We use the following vector operations in our code, which we'll explain in the text.

\def\loadps{\ttt{\_mm512\_load\_ps}}
\def\loadds{\ttt{\_mm512\_load\_ds}}
\def\fmadps{\ttt{\_mm512\_fmad\_ps}}
\def\gatherps{\ttt{\_mm512\_i32gather\_ps}}
\def\swizzleps{\ttt{\_mm512\_swizzle\_ps}}
\def\storenrngops{\ttt{\_mm512\_storenrngo\_ps}}
\def\castsi{\ttt{\_mm512\_castsi512\_ps}}
\def\permute{\ttt{\_mm512\_permute4f128\_epi32}}
\def\intmask{\ttt{\_mm512\_int2mask}}
\def\loadunpack{\ttt{\_mm512\_mask\_loadunpacklo\_epi32}}
\def\castitops{\ttt{\_mm512\_castsi512\_ps}}
\def\castpstoi{\ttt{\_mm512\_castps\_si512\_ps}}
%
\begin{center}
\begin{tabular}{|l|l|}
\hline
\loadps &  load 16 consecutive floats to register\\
\fmadps &  multiply/add of 16 floats\\
\gatherps &  gather 16, possibly disconnected floats from memory\\
\swizzleps &  reorder a channel, and duplicate across channels\\
\storenrngops &  store 16 floats to memory without register rewrite or reordering\\
\castpstoi & reinterpret 16 floats as integers\\
\castitops & reinterpret 16 integers as floats\\
\permute &  permute channels; do not change individual channels\\
\intmask &  \\
\loadunpack &  \\
\hline
\end{tabular}
\end{center}
%
In the vector instructions, \ttt{ps} refers to a float (4 bytes), while \ttt{epi32} refers to a 4-byte integer. Although we do not discuss double precision in this paper, \ttt{ds} refers to an 8 byte double precision real number (i.e., \loadds). 
Each vector register is broken up into four channels of 128 bits each. It is possible to interchange these channels at a cost of a "few" \qes{exact numbers?} cycles, and it is also possible to execute swizzle operations within a channel. For example, if the 128 bytes of each channel are labelled as $ABCD$, the vector instruction 
\ttt{\_mm512\_swizzle\_ps(v, \_MM\_SWIZ\_REG\_AAAA)}
replaces each channel by $AAAA$.

Here is a function that reads in four floats (a,b,c,d) and creates the 16-float vector dddd,cccc,bbbb,aaaa: 
\begin{verbatim}
__m512 read_aaaa(float* a)                                                              
{
    int int_mask_lo = (1 << 0) + (1 << 4) + (1 << 8) + (1 << 12);
    __mmask16 mask_lo = _mm512_int2mask(int_mask_lo);
    __m512 v1_old;
    v1_old = _mm512_setzero_ps();
    v1_old = _mm512_mask_loadunpacklo_ps(v1_old, mask_lo, a);
    v1_old = _mm512_mask_loadunpackhi_ps(v1_old, mask_lo, a);
    v1_old = _mm512_swizzle_ps(v1_old, _MM_SWIZ_REG_AAAA);
    return v1_old;
}
\end{verbatim}

Here is a function that takes four floats from float register \ttt{v1}, and 
places them in each of the four lanes. Because the permulation function only 
operates on 4-byte integers, it is necessary to convert (in place) each float
to an integer (the bit structure is not changed), permute the lanes and cast
the integers back to floats (with no modification of the bit structure)
\begin{verbatim}
__m512 permute(__m512 v1, _MM_PERM_ENUM perm)                                           
{
    __m512i vi = _mm512_castps_si512(v1);
    vi = _mm512_permute4f128_epi32(vi, perm);
    v1 = _mm512_castsi512_ps(vi);
    return v1;
}
\end{verbatim}

%----------------------------------------------------------------------
\section{Base implementation}
We present an implementation written in C++ for the case of 4 matrices and 4 vectors to run on the host system and serve as base of comparison against more optimized code written with register based routines optimized for the MIC. Such an implementation is as follows:
\begin{verbatim}
\end{verbatim}
%----------------------------------------------------------------------
\section{Floating point operations}
We compute the floating point operations of the SpMV with general $n_v$ and $n_m$. Calling the total number of floating point operations $N_F$, we find
$$
  N_F = 2 n_r n_z n_v m_m
$$
Thus the number of flpos to bytes transferred is $N_F / (n_{rT} + n_{wT})$, which when written out, 
\begin{eqnarray}
N_F &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_v  + n_r n_z b_i + n_r n_m n_z b_x + n_r n_m n_v b_x} \nonb \\
N_F &=& \frac{2 n_z n_v n_m}{b_x ( n_v + n_m n_z + n_m n_v) + n_z b_i}  \nonb \\
%N_F &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_v (n_v+n_m) + n_r n_z b_i + n_r n_m n_z b_x} \nonb \\
    %&=& \frac{2 n_z n_v n_m}{b_x (n_v+n_m) + n_z (b_i + n_m b_x)} \nonb 
\end{eqnarray}
In the special case $n_v=n_m=1$, 
$$
N_F = \frac{2 n_z}{n_z b_i + (n_z +2) b_x}
$$
Neglecting $2$ compared to $n_z$, we find that $N_F = 2/(b_i+b_x)$, which leads to 
$N_F = 2/8=1/4$ and  $N_F=2/12=1/6$ for single and double precision, respectively. 
In the general case, in the condition of $n_z$ large compared to 2, 
$$
N_F =  \frac{2 n_v n_m}{n_m b_x + b_i}
$$ 
The case are interested in is $n_m=n_v=4$, so that
$$
N_F = \frac{8}{ b_x + 1}
$$
which is 8/5 and 8/9 in single and double precision, respectively. 
In single and double precision, we get $N_F=32/36=8/5$ and $N_F=32/68=8/9$, respectively.
The third column of the following table (TABLE ???) is the maximum achievable Gflop rate, assuming 
a maximum memory transfer speed of $150 Gflops$ (see earlier benchmark in this paper). 
(higher speeds can be achieved in practice,  upwards of 190 Gflops (Saule \etal\cite{}), but we use
a benchmark code similar to our current algorithm, without the indirect acessing required for $x$, but keeping the same use of vector registers used in our fastest implementation. (DO WE INCLUDE the CODE?)

\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Precision & $n_m=n_v=1$ & $n_m=n_v=4$     \\
\hline
single    &  1/4  (37.5)     &   8/9 (133) \\
double    &  1/6  (25.5)     &   8/17 (62) \\
\hline
\end{tabular}
\end{center}
(With a 190 Gflop peak memory rate, maximum performance is 47.5 and 170 Mflops for the 1/1 case, while we calculate 32 and 90 Gflops in the 4/4/ case.

Let us also calculate the flop to byte ratio for the case of 1/4 and 4/1 ($n_v$/$n_m$). One notices the symmetry with respect to $n_v$ and $n_m$. 
One gets $N_F = \frac{2*4}{5 b_x + 4}$ equal to 8/24 (=.33) and 8/44 (=.18), in single and double precision, respectively, which corresponds to 49 and 27 Gflops if a maximum of 150 Gflops is assumed.
As $n_m$ and $n_v$ keep increasing, $N_F$ tends towards $2/b_x$, equal to $0.5$ and $0.25$ respectively for single and double precision, which leads to peak speed of 75 and 37.5 Gflops.

In single precision, our best benchmarks achieve 20 Gflops for single precions for on matrix/one vector, and 100 Glops for the 4/4 matrix/vector case, or 53\% and 75\% of peak performance, respectivey. We conclude that we achieve a higher percentage of peak performance for this algorithm when calculating 16 output $y$ vectors. The results are even more impressive when measured against the peak perforance of 2Tflops of the MIC in double precision.

### 2013-08-21 problem_statement.tex
Original paper, must run pdflatex.

\section{Problem Statement}
There has been much work on the problem of efficient sparse matrix/vector (spmv) and sparse matrix  (spmm) multiplication over the years \cite{} specialized for a range of computer architectures \cite{}. In the past 10 years, researchers have addressed implementations on the GPU using mostly CUDA \cite{}, and on multicore sytems, exemplified by the chips produced by Intel~\cite{}. In general, spvm consists of multiplying the matrix $A$ of size $N\times N$ by a vector $x$ of size $N\times 1$ to produce a vector $y$ of size $N\times 1$. More succinctly, $ y = A x $ .

In the theory of Radial Basis Functions Finite-Difference (RBFFD), derivatives of a function $f(\rvec)$ at node $i$ are expressed as a linear combination of the function values at the stencil center and the nodes connected to node $i$ (Figure~\ref{fig:rbf_stencils}). Thus $y$ is a discrete derivative of the vector $x$.

When solving a one dimensional system of PDEs, one might require an $x$ derivative of multiple functions. For example, the Euler equations require the $x$ derivative of the three components of velocity and pressure. In this case, there are $n_v=4$ vectors $x^k$, $k=0,\cdots,3$.

Thus for each vector element $y_i$, we compute $y_i = \sum_j A_{ij} x_j$. If $A_i$ is row $i$ of $A$, $y_i$ is simply the dot product $A_i x$. The next level of generality is to consider $n_v$ vectors $x^k$, $k=0,\cdots,n_v-1$. Whatever the spmv implementation, one achieves improved performance if the matrix formed from the columns $x^k$ are stored in row major order. Thus, $x^0_0,x^1_0,\cdots,x^{n_v-1}_0$, are stored in consecutive memory location. The random access of the elements of $x$ is thus reduced. Maximum efficiency is achieved when $n_v=16$ floats or $8$ doubles, given that cache lines take 64 bytes. We will benchmark this case, labeled $Svn$, where $n$ refers to the number of vectors (Iv4 uses four vectors). The $S$ refers to singe precision. A double precision run is labelled $Dvn$.

Alternatively, when solving a PDE, one might require derivatives of a given scalar function with respect to coordinate directions $x$, $y$, $z$. Second order operators of often required, such as a second derivative with respect to $x$ or a Laplacian operator. In the RBFFD formulation, on can compute different derivatives using the same stencil, but with different weights. In other words, the adjacency matrix that corresponds to $A$ remains constant, but the matrix elements of $A$ change with the particular derivative. 
In this case, label with a superscript $l$ the particular matrix $A^k$. Since the adjacency matrix is assumed invariant, there is only need for a single matrix \ttt{C{ij}}. In Ellpack format, each row is of constant size (the number of nonzeros per row of $A$. $C_{ij}$ is the column number that locates the $j^{th}$ nonzero in row $i$ of $A$.

\ge{Must rewrite the above.}

### 2013-08-21 tables_and_figures.tex
Original paper, must run pdflatex.

\section{List of Tables and Figures}
\begin{enumerate}
\item
Table: Notation for formulas
\item
Table: List of relevant vector instructions
\item
Figure: Matrix-Vector multiplication
\item
\item
\item
\item
\end{enumerate}

\section{Suggested Benchmarks}
\begin{enumerate}
\item
SpMM with 16 floats (or 8 doubles) vectors (will give upper bound in speed). Implement algorithm  from Saul \etal. 
\item
SpMM with 4 floats or 4 doubles (to compare against SpMM). As a precursor to adding more matrices. 
\item
Try to figure out why our performance is so much lower than peak. 
\end{enumerate}

\newfig{figures/rbf_stencils.pdf}{0.7}{RBFFD Stencils. Each node of the mesh is connected to $n_z-1$ stencil nodes in addition to itself. In the figure, node $A$ is connected to $B$, but $B$ is {\em not\/} connected to $A$. Thus adjacency graph of $A$ is not-symmetric. This must be corrected when reducing the bandwidth of $A$ with a Cuthill\-McKee algorithm, which assumes symmetry.}
{fig:rbf_stencils}

\newfig{figures/matrix_structure.pdf}{0.7}{Matrix Structure.}{fig:mat_struct}
\newfig{figures/swizzling.pdf}{0.7}{Duplication of the first four bytes of each channel across the entire channel.}{fig:swizzling}
\newfig{figures/channel_permutation.pdf}{0.7}{Channel Permutations on the MIC}{fig:permulatio.n}
\newfig{figures/tensor_product.pdf}{0.7}{Tensor scalar product using channels and swizzling.}{fig:tensor_product}

\newfig{figures/theoretical_performance.png}{0.7}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors. We used a memory bandwidth of 190 Gbytes/sec (based on best measurements with specialized code) and 150 Gbytes/sec (based on measurements in a code similar to the SpMV kernel, with computations removed.)}{fig:peak_perf}

### 2013-08-21 toc.tex
Original paper, must run pdflatex.

\section{Table of Contents}
\begin{enumerate}
\item Introduction (1 page) 
\item Motivation, RBF derivatives, stenciles, RBFFD  (1 page)
\item D=Ax, sparse vector matrix  multiplication
  - Ellpack format only
\item Notation (1 page)
\item Multiple derivatives of a single function
\item Single derivative of multiple functions
\item Multiple derivatives of multiple functions
\item MIC architecture (1 page)
\item Basic implementation with OpenMP (1 page)
\item MIC Vector hardware instructions, gather, masks, Pack/Unpack
  (1 page)
\item Mathematical model to predict timings, taking into account latencies, 
   cache sizes, problem size, cost of memory, cost of operations. 
\item Benchmarks. Effect of (3 pages)
\begin{itemize}
 \item Problem size
 \item Single/Double precision
 \item Bandwidth reduction
 \item Breaking up the domain into smaller pieces 
 \item Cache thrashing
 \item Indirect addressing (addressed in previous papers)
\end{itemize}
\item Domain decomposition to minimize cache thrashing
\item Bandwidth reduction, effect of domain decomposition
\item Do not address
\begin{itemize}
  \item Cost of matrix assembly
  \item Cost of accessing the derivatives in an application code (future paper)
\end{itemize}
\item Bibliography (1 page)
\end{enumerate}

### 2013-08-21 chart1.py
Added chart1.py to generate bar chart.

@@ -1,0 +1,81 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+
+#----------------------------------
+def ratio(nv, nm, bx, nz):
+# per row:
+# flops divided by number bytes sent/received
+# nv : number vectors
+# nm : number matrices
+# bx = 4 or 8 (single, double precision)
+# nz = nb nonzeros per row
+# y = A*x
+# y : nv*nm*bx   writes
+# x : nv*bx      loads
+# A : nm*nz*bx   loads
+# col_id : nz*bi loads
+    bi = 4
+    r = 2.*nz*nv*nm/(bx*(nv+nm*nz+nm*nv)+nz*bi)
+    return(r)
+#----------------------------------
+nz = 32
+single = 4
+double = 8
+ulti_bandwidth = 190
+max_bandwidth = 150
+bandwidths = [150,190]
+nv = [1,4]
+nm = [1,4]
+labels = [(v+m) for v in ["v1","v4"] for m in ["m1","m4"]]
+# v loop is slowest
+# m loop is fastest
+ratios_s =  np.array([ratio(v,m,single, nz) for v in nv for m in nm])
+ratios_s_ulti = ratios_s*ulti_bandwidth
+ratios_s_max  = ratios_s*max_bandwidth
+#print("single precision")
+#print ratios_s
+#print ratios_s_ulti
+#print ratios_s_max
+
+ratios_d =  np.array([ratio(v,m,double, nz) for v in nv for m in nm])
+ratios_d_ulti = ratios_d*ulti_bandwidth
+ratios_d_max  = ratios_d*max_bandwidth
+#print("\ndouble precision")
+#print ratios_d
+#print ratios_d_ulti
+#print ratios_d_max
+
+
+
+#### BEGINNING OF CHART
+
+ngroups = len(labels)
+
+fig, ax = plt.subplots()
+index = np.arange(ngroups)
+bar_width = 0.35
+opacity = .5
+
+rects1 = plt.bar(index, ratios_s_ulti, bar_width, alpha=opacity,
+           color='b', label='single ulti')
+rects2 = plt.bar(index+bar_width, ratios_d_ulti, bar_width, alpha=opacity, color='r', label='double ulti')
+
+opacity = 1
+rects1 = plt.bar(index, ratios_s_max, bar_width, alpha=opacity,
+           color='b', label='single max')
[…]

### 2013-08-22 notation.tex
Two more figures.

@@
 $$
-Thus the number of flpos to bytes transferred is $N_F / (n_{rT} + n_{wT})$, which when written out, 
+Two cases must be distinguished in regards to $x$. In the first, in a worst case scenario, each element 
+of $x$ is transferred $n_z$ times from memory.
+In the best case, the case of infinite cache, each element is transferred only once.
+Thus, we provide best and worst case scenarios and will compare against matrices $A$
+in each of these extreme situations to evaluate the degree we approach idea performance. 
+Thus the number of flops to bytes transferred is $N_F / (n_{rT} + n_{wT})$, which when written out, 
 \begin{eqnarray}
-N_F &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_v  + n_r n_z b_i + n_r n_m n_z b_x + n_r n_m n_v b_x} \nonb \\
-N_F &=& \frac{2 n_z n_v n_m}{b_x ( n_v + n_m n_z + n_m n_v) + n_z b_i}  \nonb \\
+N_{Fw} &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_z n_v  + n_r n_z b_i + n_r n_m n_z b_x + n_r n_m n_v b_x} \nonb \\
+    &=& \frac{2 n_v n_m}{b_x ( n_v + n_m + n_m n_v n_z^{-1}) + b_i}  \nonb \\
 %N_F &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_v (n_v+n_m) + n_r n_z b_i + n_r n_m n_z b_x} \nonb \\
@@
 \end{eqnarray}
+while in the best case, 
+\begin{eqnarray}
+N_{Fb} &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_v  + n_r n_z b_i + n_r n_m n_z b_x + n_r n_m n_v b_x} \nonb \\
+    &=& \frac{2 n_v n_m}{b_x ( n_v n_z^{-1} + n_m  + n_m n_v n_z^{-1}) + b_i } 
+\end{eqnarray}
 In the special case $n_v=n_m=1$, 
 $$
-N_F = \frac{2 n_z}{n_z b_i + (n_z +2) b_x}
+N_{Fw} = \frac{2}{b_x (2+n_z^{-1}) + b_i}
+$$
+and 
+$$
+N_{Fb} = \frac{2}{b_x (1 + 2 n_z^{-1}) + b_i}
 $$

### 2013-08-22 tables_and_figures.tex
Two more figures.

@@
 
-\newfig{figures/theoretical_performance.png}{0.7}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors.
+%\newfig{figures/theoretical_performance.png}{0.7}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors.
 We used a memory bandwidth of 190 Gbytes/sec (based on best measurements with specialized code) and 150 Gbytes/sec (based on measurements in a code similar to the SpMV kernel, with computations removed.)}{fig:peak_perf}
+
+\newfig{figures/gflops_peak.png}{0.7}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors.
+We used a memory bandwidth of 150 Gbytes/sec (based on measurements in a code similar to the SpMV kernel, with computations removed.)}{fig:gflops_peak_perf}
+
+\newfig{figures/speedup_wrt_base.png}{0.7}{Speedup relative to base case using one matrix and one vector.} {fig:speedkup}

### 2013-08-22 chart1.py

@@ -16,8 +16,9 @@ def ratio(nv, nm, bx, nz):
 # A : nm*nz*bx   loads
 # col_id : nz*bi loads
     bi = 4
-    r = 2.*nz*nv*nm/(bx*(nv+nm*nz+nm*nv)+nz*bi)
-    return(r)
+    rworst = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
+    rbest = (2.*nv*nm)/(bx*(nm+(1+nm)*nv/float(nz))+bi)
+    return([rworst,rbest])
 #----------------------------------
 nz = 32
 single = 4
@@ -33,20 +34,17 @@ labels = [(v+m) for v in ["v1","v4"] for m in ["m1","m4"]]
 ratios_s =  np.array([ratio(v,m,single, nz) for v in nv for m in nm])
 ratios_s_ulti = ratios_s*ulti_bandwidth
 ratios_s_max  = ratios_s*max_bandwidth
-#print("single precision")
-#print ratios_s
-#print ratios_s_ulti
-#print ratios_s_max
 
 ratios_d =  np.array([ratio(v,m,double, nz) for v in nv for m in nm])
 ratios_d_ulti = ratios_d*ulti_bandwidth
 ratios_d_max  = ratios_d*max_bandwidth
-#print("\ndouble precision")
-#print ratios_d
-#print ratios_d_ulti
-#print ratios_d_max
 
 
+ratios_s_worst = np.transpose(ratios_s_max)[0]
+ratios_s_best = np.transpose(ratios_s_max)[1]
+
+ratios_d_worst = np.transpose(ratios_d_max)[0]
+ratios_d_best = np.transpose(ratios_d_max)[1]
 
 #### BEGINNING OF CHART
 
@@ -54,28 +52,49 @@ ngroups = len(labels)
 
 fig, ax = plt.subplots()
 index = np.arange(ngroups)
-bar_width = 0.35
-opacity = .5
+bar_width = 0.35/2
 
-rects1 = plt.bar(index, ratios_s_ulti, bar_width, alpha=opacity,
-           color='b', label='single ulti')
-rects2 = plt.bar(index+bar_width, ratios_d_ulti, bar_width, alpha=opacity, color='r', label='double ulti')
+rects1 = plt.bar(index, ratios_s_worst, bar_width, alpha=1, color='b', label='single worst case')
[…]

### 2013-08-26 tables_and_figures.tex
Updated paper.

@@
 \item
-Try to figure out why our performance is so much lower than peak. 
+Try to figure out why our performance is so much lower than peak stated in Figure \ref{fig:gflops_peak_per}.
+Expecially in the case of compact. To do this, remove permutes, and replace gather by straight forward C++
+without using vector notation, but using SIMD commands in addition to OPENPM commands. Make sure we use -O3, and contrast to -O1 and -O2. 
 \end{enumerate}
@@
 \newfig{figures/speedup_wrt_base.png}{0.7}{Speedup relative to base case using one matrix and one vector.} {fig:speedkup}
+
+\newfig{figures/test1_readwrite.png}{0.7}{Bandwidth performance under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without MIC vector instructions.}{fig:band_rw}
+
+\newfig{figures/test1_gather.png}{0.7}{Bandwidth performance under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without MIC vector      instructions.}{fig:band_gather}
+
+\newfig{figures/test3_gather.png}{0.7}{Bandwidth performance under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without MIC vector      instructions.
+The greater speed of the cpp version is obtained throught the use of \#Ivdep {\em and\/} \_\_assumed\_aligned.
+}{fig:band_gather_ivdep}

### 2013-09-29 ieee_paper.tex
Add ieee latex files. Compiles ieee_paper using pdflatex

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[10pt, conference, compsocconf]{IEEEtran}
\input{support/macros} % color is defined in macros or misc_mac
\input{support/misc_mac}
\input{support/setupicase}
% Add the compsocconf option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

%\newcommand{\red}[1]{\textbf{\textcolor{red}{#1}}}
%\newcommand{\blue}[1]{\textbf{\textcolor{blue}{#1}}}
\def\red#1{\textbf{\textcolor{red}{#1}}}
\def\blue#1{\textbf{\textcolor{blue}{#1}}}
\def\qes#1{{\blue{*** For Erik: #1 ***}}}
\def\es#1{{\blue{*** For Erik: #1 ***}}}
\def\ge#1{{\red{*** For Gordon: #1 ***}}}
\def\ttt#1{{\tt #1}}
\def\bold#1{{\bf #1}}

\def\qes#1{}
\def\es#1{}
\def\ge#1{}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% handle arbitrary number of unprocessed floats
\usepackage{morefloats}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the
   MIC Architecture}

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations

\author{\IEEEauthorblockN{Gordon Erlebacher\IEEEauthorrefmark{1},
Erik Saule\IEEEauthorrefmark{2}, Natasha Flyer\IEEEauthorrefmark{3}, 
and Evan Bollig\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Scientific Computing, 
Florida State University, Tallahassee, FL 32306-4120\\
Email: gerlebacher@fsu.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte\\
Email: erik.saule@uncc.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR \\
Email: 
flyer@ucar.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
Email: bollig@gmail.com}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
In this paper, we develop an efficient scheme for the calculation of derivatives within the context
of Radial Basis Function Finite-Difference (RBFFD). RBF methods express functions as a linear
combination of radial basis functions on an arbitrary set of nodes. The Finite-Difference component
expresses this combination over a local set of nodes neighboring the point where the derivative is sought.
The derivative at all points takes the form of a sparse matrix/vector multiplication (spmv).

In this paper, we consider the case of local stencils with the number of nodes at each point and encode the
sparse matrix in ELLPACK format. We increase the number of operations relative to memory bandwidth by
calculating four derivatives of four different functions, or 16 different derivatives. We demonstrate
a novel implementation on the MIC architecture, taking into account its advanced swizzling and channel
interchange features. We present benchmarks that show an almost order of magnitude increase in speed
compared to efficient implementations of a single derivative. We explain the results through consideration
of operation count versus memory bandwidth.
\end{abstract}

\begin{IEEEkeywords}
OpenMP; MIC; spmv; sparse matrix; Radial Basis Function;
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\input{problem_statement}
% no \IEEEPARstart

\subsection{Subsection Heading Here}
Subsection text here.

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

\input{tables_and_figures}
\input{tables_and_figures}
\input{bandwidth}
\input{notation}

%Wherever Times is specified, Times Roman or Times New Roman may be used. If neither is available on your system, please use the font closest in appearance to Times. Avoid using bit-mapped fonts if possible. True-Type 1 or Open Type fonts are preferred. Please embed symbol fonts, as well, for math, etc.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

\section{Conclusion}
The conclusion goes here. this is more of the conclusion

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank...
more thanks here

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% that's all folks
\end{document}

### 2013-09-29 notation.tex
Add ieee latex files. Compiles ieee_paper using pdflatex

@@
 
-In single precision, our best benchmarks achieve 20 Gflops for single precions for on matrix/one vector, and 100 Glops for the 4/4 matrix/vector case, or 53\% and 75\% of peak performance, respectivey.
+In single precision, our best benchmarks achieve 20 Gflops for single precisions for on matrix/one vector, and 100 Glops for the 4/4 matrix/vector case, or 53\% and 75\% of peak performance, respectivey.
 We conclude that we achieve a higher percentage of peak performance for this algorithm when calculating 16 output $y$ vectors.

### 2013-09-29 tables_and_figures.tex
Add ieee latex files. Compiles ieee_paper using pdflatex

--- 2013-08-21 tables_and_figures.tex
@@
-\section{List of Tables and Figures}
-\begin{enumerate}
-\item
-Table: Notation for formulas
-\item
-Table: List of relevant vector instructions
-\item
-Figure: Matrix-Vector multiplication
-\item
-\item
-\item
-\item
-\end{enumerate}
+%\section{List of Tables and Figures}
+%\begin{enumerate}
+%\item
+%Table: Notation for formulas
+%\item
+%Table: List of relevant vector instructions
+%\item
+%Figure: Matrix-Vector multiplication
+%\item
+%\item
+%\item
+%\item
+%\end{enumerate}
 
@@
 \item
-Try to figure out why our performance is so much lower than peak. 
+Try to figure out why our performance is so much lower than peak stated in Figure \ref{fig:gflops_peak_perf}.
+Expecially in the case of compact. To do this, remove permutes, and replace gather by straight forward C++
+without using vector notation, but using SIMD commands in addition to OPENPM commands. Make sure we use -O3, and contrast to -O1 and -O2. 
 \end{enumerate}
 
-\newfig{figures/rbf_stencils.pdf}{0.7}{RBFFD Stencils.
+\def\wide{2.5in}
+\newfig{figures/rbf_stencils.pdf}{\wide}{RBFFD Stencils.
 Each node of the mesh is connected to $n_z-1$ stencil nodes in addition to itself.
@@
 Thus adjacency graph of $A$ is not-symmetric.
-This must be corrected when reducing the bandwidth of $A$ with a Cuthill\-McKee algorithm, which assumes symmetry.}
-{fig:rbf_stencils}
+This must be corrected when reducing the bandwidth of $A$ with a Cuthill\-McKee algorithm, which assumes symmetry.}{fig:rbf_stencils}
 
-\newfig{figures/matrix_structure.pdf}{0.7}{Matrix Structure.}{fig:mat_struct}
-\newfig{figures/swizzling.pdf}{0.7}{Duplication of the first four bytes of each channel across the entire channel.}{fig:swizzling}
-\newfig{figures/channel_permutation.pdf}{0.7}{Channel Permutations on the MIC}{fig:permulatio.n}
-\newfig{figures/tensor_product.pdf}{0.7}{Tensor scalar product using channels and swizzling.}{fig:tensor_product}
+\newfig{figures/matrix_structure.pdf}{\wide}{Matrix Structure.}{fig:mat_struct}
+\newfig{figures/swizzling.pdf}{\wide}{Duplication of the first four bytes of each channel across the entire channel.}{fig:swizzling}
+\newfig{figures/channel_permutation.pdf}{\wide}{Channel Permutations on the MIC}{fig:permutation}
+\newfig{figures/tensor_product.pdf}{\wide}{Tensor scalar product using channels and swizzling.}{fig:tensor_product}
 
-\newfig{figures/theoretical_performance.png}{0.7}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors.
+%\newfig{figures/theoretical_performance.png}{\wide}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors.
 We used a memory bandwidth of 190 Gbytes/sec (based on best measurements with specialized code) and 150 Gbytes/sec (based on measurements in a code similar to the SpMV kernel, with computations removed.)}{fig:peak_perf}
+
+\newfig{figures/gflops_peak.png}{2.5in}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors.
+We used a memory bandwidth of 150 Gbytes/sec (based on measurements in a code similar to the SpMV kernel, with computations removed.)}{fig:gflops_peak_perf}
+
+\newfig{figures/speedup_wrt_base.png}{\wide}{Speedup relative to base case using one matrix and one vector.}{fig:speedup}
+
+\newfig{figures/test1_readwrite.png}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without MIC vector instructions.}{fig:band_rw}
+
+\newfig{figures/test1_gather.png}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without MIC vector instructions.}{fig:band_gather}
+
+\newfig{figures/test3_gather.png}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without MIC vector instructions.
+The greater speed of the cpp version is obtained throught the use of \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}.
+All memory is aligned on 64 bytes.}{fig:band_gather_ivdep}
+
+\newfig{figures/host_test1_readwrite_no_temporal_hint.png}{\wide}{Bandwidth performance on the host under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without AVX vector instructions.
+The speed on the CPU matches the speed with AVX instructions.
+All memory is aligned on 32 bytes.}{fig:read_write}
+
+\newfig{figures/mic_performance_nb_threads.png}{\wide}{Performance of $y=Ax$ on the MIC.
+Squares: base 1/1 case, solid circles: 4/4 case implemented in C++, solid triangles: 4/4 case implemented with MIC vector instructions.
+$-O3$ compilation options.
+Each group consists of four cases: grids of $64^3$ and $96^3$ with and without Reverse Cuthill McKee.}{fig:mic_performance}
+
+% NOTES: why is result the same at 32 processors.
+I should find out the distribution of threads.
+16 threads means one thread on each of 16 cores (YES, I BELIEVE), or 2 threads on each of 8 cores.} Must run the experiment with different nodes orderings.
+%export OMP_SCHEDULE=guided,64    (I should redo experiment with Dynamic or static and compact for KMP_AFFINITY
+%export KMP_AFFINITY=compact
+% MUST PRINT OUT OMP_SCHEDULE and KMP_AFFINITY inside the program. 
+% WHY the very high performance with very few nodes? STRANGE. 
+\newfig{figures/host_performance.png}{\wide}{Performance of $y=Ax$ on the Host.
+Squares: base 1/1 case, solid circles: 4/4 case implemented in C++, solid triangles: 4/4 case implemented with MIC vector instructions.
+$-O3$ compilation options.
+The four colors distinguish grid resolution and whether or not
+Reverse Cuthill McKee is applied.}{fig:host_performance}

### 2013-10-02 ieee_paper.tex
Forgotten file.

@@
 \section{Introduction}
+\cite{Bell08, Vuduc05, Nishtala07, Stock12-TACC, Cuthill69, cramer2012openmp, Buluc2009_SPAA, Buluc11, Im01, Mellor-Crummey04, Nishtala07, Saad94sparskit, Williams07}
 \input{problem_statement}
-% no \IEEEPARstart
-
-\subsection{Subsection Heading Here}
-Subsection text here.
-
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
-
-\input{tables_and_figures}
+\input{intel_phi}
 \input{tables_and_figures}
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
+%5\begin{thebibliography}{1}
 
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
-
-% that's all folks
+\bibliographystyle{plain} %{IEEEabrv,paper.bib}
+\bibliography{paper, bollig}
 \end{document}

### 2013-10-02 intel_phi.tex
Forgotten file.

%% COPIED from Erik Saule paper
\section{The Intel Xeon Phi Coprocessor}
(USE FRODO AT MSI)
In this work, we use a pre-release KNC card SE10P. The card has 8 memory controllers where each of them can execute 5.5 billion transactions per second and has two 32-bit channels. That is the architecture can achieve a total bandwidth of 352GB/s aggregated across all the memory controllers. There are 61 cores clocked at 1.05GHz. The cores’ memory interface are 32-bit wide with two channels and the total bandwidth is 8.4GB/s per core. Thus, the cores should be able to consume 512.4GB/s at most. However, the band- width between the cores and the memory controllers is limited by the ring network which connects them and theoretically supports at most 220GB/s.

Each core in the architecture has a 32kB L1 data cache, a 32kB L1 instruction cache, and a 512kB L2 cache. The architecture of a core is based on the Pentium architecture: though its design has been updated to 64-bit. A core can hold 4 hardware contexts at any time. And at each clock cycle, instructions from a single thread are executed. Due to the hardware constraints and to overlap latency, a core never executes two instructions from the same hardware context consecutively. In other words, if a program only uses one thread, half of the clock cycles are wasted. Since there are 4 hardware contexts available, the instructions from a single thread are executed in-order. As in the Pentium architecture, a core has two di↵erent concur- rent instruction pipelines (called U-pipe and V-pipe) which allow the execution of two instructions per cycle. However, some instructions are not available on both pipelines: only one vector or floating point instruction can be executed at each cycle, but two ALU instructions can be executed in the same cycle.

Most of the performance of the architecture comes from the vector processing unit. Each of Intel Xeon Phi’s cores has 32⇥512-bit SIMD registers which can be used for double or single precision, that is, either as a vector of 8⇥64-bit values or as a vector of 16⇥32-bit values, respectively. The vector processing unit can perform many basic instructions, such as addition or division, and mathematical operations, such as sine and sqrt, allowing to reach 8 double precision operations per cycle (16 single precision). The unit also sup- ports Fused Multiply-Add (FMA) operations which are typically counted as two operations for benchmarking purposes. Therefore, the peak performance of the SE10P card is 1.0248 Tflop/s in double precision (2.0496 Tflop/s in single precision) and half without FMA.

### 2013-10-11 problem_statement.tex
Latest version. I changed problem_statement to discuss RBF.

@@
 \section{Problem Statement}
-There has been much work on the problem of efficient sparse matrix/vector (spmv) and sparse matrix  (spmm) multiplication over the years \cite{} specialized for a range of computer architectures \cite{}.
+%Introduction: RBFs -> derivatives -> spmv -> multiple derivatives of multiple functions --> our approach
+
+Sparse Matrix-Vector Multiplication (SpMV) is an integral component of a very large
+number of computer algorithms designed to solve practical problems in most scientific disciplines.
+Unfortunately, the SpMV kernel is bandwidth limited, and can only achieve a small fraction of peak performance on current and post computer architectures.
+Deeper memory hierarchies, wider memory lanes, larger registers, are all designed to increase throughput and computational efficiency for problems where data is accessed sequentially from memory.
+However, in SpMV, not only are there insufficient number of floating point operations relative to the number of bytes transfered, but the the uneven access pattern of the source vector lead to cache misses, which are detrimental to high performance.
+As a result, a wide range
+strategies have been devised to decrease memory bandwidth through compressed matrix
+formats)\cite{}, decrease the number of cache misses\cite{} %(register and cache 
+blocking\cite{}) and multiple source vectors\cite{}.
+
+%In the theory of Radial Basis Functions Finite-Difference (RBFFD), derivatives of a function $f(\rvec)$ at node $i$ are expressed as a linear combination of the function values at the stencil center and the nodes connected to node $i$ (Figure~\ref{fig:rbf_stencils}).
+Thus $y$ is a discrete derivative of the vector $x$.
+
+In this paper, we propose yet a new idea, that is applicable to the world of 
+Finite-Difference Radial Basis functions (RBF-FD). A radial basis functions refers
+to a set of shifted radially symmetric functions $\phi_i(r) =\phi{||\xvec-\xvec_i||}$ centered at $\xvec_i$\cite{}.
+Any function $f(\xvec)$ defined in a domain $\Omega$ can be expanded in this basis according to
+\begin{equation}
+f(\xvec) =  \sum_i w_i \phi_i(\xvec) \label{eq:rbf}
+\end{equation}
+% Wrong label eq:rbf in document 
+Given function values at a set of discrete points $x_i$, one inverts linear system 
+(\ref{eq:rbf})  to obtain a set of weights $w_j$.
+Given a set of $N$ RBF nodes, the matrix to invert is dense.
+Derivatives of $f(\xvec)$ result from the differentiation of Eq.
+(\ref{eq:rbf}).
+For very large problems, this is an computationally expensive since the RBFs do not form an orthogonal basis, nor are they amenable to tensor decomposition in the three coordinate directions.
+Rather, their strength is the ability to randomly distribute nodes across complex physical domains, and have an implementation that is independent of dimensionality.
+%(???)
+
+To alleviate the cost of a global approach, researchers \cite{} expand $f(\xvec)$ in a local region of $x_i$. Thus, 
+$$
+f(\xvec) =  \sum_{i=0}^{n_s} w_i \phi_i(\xvec) \label{eq:rbf}
+$$
+where one only considers a local set of $n_s$ neighbors of of point $x_j$. 
+Finite-difference RBF expresses the derivative of $f(\xvec)$ at $x_j$ as a linear
+combination of $f(\xvec)$ at the $n_s$ neighbors. Thus, 
+$$
+f = A w
+$$
+where $f$ is the vector of function values at the $n_s$ stencil points (including the center point $x_j$. The matrix $A$ is a function of 
+$\phi_{ij}(\xvec) =\phi(||x_i-x_j||)$. This leads to the $n_s$ weight values 
+$w = A^{-1} f$. The derivative of Eq. \ref{eq:rbf} evaluated at $x_j$ gives
+$$
+   f'_j = \sum_{i=0}^{n_s} w_i \phi'_{ij}
+$$
+Substituting the vector of weights $w$ leads to 
+$$
+  f'_j = \sum_{i=0}^{n_s} \alpha_i f_i
+$$
+The derivative evaluation at all points of the domain $\Omega$, notwithstanding boundary conditions, takes the form
+$$
+  y = A x
+$$
+where $x$ is the souce vector of function values, and $y$ is the vector of derivative values. The matrix $A$ is called a derivative matrix. In the current litterature, the number of stencil points is 
+a fixed value, independent of the node under consideration\cite{}.
+The computation of a single derivative has been reduced to a SpMV, where each row has $n_s$ nonzeros.
+Because the sparsity of each row is constant, ELLPACK~\cite{} becoes the most appropriate matrix compression scheme\cite{}.
+In practice, $n_s=32$ in two-dimensional flows and 64 or 100 for three-dimensional flows.
+These numbers are similar to what is used in finite-element codes.
+
+As we will discuss in later sections, the performance of SpMV is severely constrained by both the high number of load/store operations and the irregular access patterns of $x$.
+Thus, we seek to increase the number of operations per floats transfered from memory, increase register usage, and decrease cache misses.
+To this end, we consider a combination of two appraoches: increase the number of source vectors, and increase the number of derivative matrices.
+
+Many problems in fluid dynamics and in the geosciences require the solution to
+transport equations of the form
+$$
+\pf{Q}{t} = f(Q,Q_x, Q_y, Q_z, \Laplacian{Q})
+$$
+where $Q$ is a vector of unknowns (3 components of velocity, pressure, temperature). Thus, it is often necessary to compute the $x$ derivative of 
+multiple functions, typically four for the Euler equations, or five for the Navier-Stokes equations.
+Multiple right-hand sides transform a SpMV into a SpMM (Sparse Matrix/dense Matrix multiplication), which improves register utilization and decreases cache misses by vectorizing over the multiple source vectors.
+Further improvements are possible by recognizing that different derivative matrices (x,y,z and Laplacian for example), have the same sparsity distribution; only the weight
+change.  Thus, alternative to computing a derivative of multiple functions, we
+can calculate multiple derivatives of a single functions. The increased memory bandwidth due to an increase in the number of derivaive matrices is offset by 
+better cache utilization, leading to an overall benefit.
+
+In this paper, we examine the possible improvements in efficiency of SpMV for
+the case of four source vectors and four derivative matrices. We benchmark
+performance on a Knights Corner card\cite{}, using OpenMP. The paper 
+is developed as follows. In Section 2, we describe some of the previous work, 
+relevant to this paper. This is followed by a brief description of the MIC
+architecture in Section 3, the matrices we consider in Section 4 together 
+with operation and bandwidth counts, description of our benchmarks in Section 5, 
+followed by our conclusions.
+
+%There has been much work on the problem of efficient sparse matrix/vector (spmv) and sparse matrix  (spmm) multiplication over the years \cite{} specialized for a range of computer architectures \cite{}.
 In the past 10 years, researchers have addressed implementations on the GPU using mostly CUDA \cite{}, and on multicore sytems, exemplified by the chips produced by Intel~\cite{}.
@@
 
-In the theory of Radial Basis Functions Finite-Difference (RBFFD), derivatives of a function $f(\rvec)$ at node $i$ are expressed as a linear combination of the function values at the stencil center and the nodes connected to node $i$ (Figure~\ref{fig:rbf_stencils}).
-Thus $y$ is a discrete derivative of the vector $x$.
+**** PUT ELSEWHERE OR ELIMINATE? ****
 

### 2013-10-12 chart1.py
Plots in pdf format.

@@ -68,8 +68,8 @@ plt.legend(loc=2)  # upper left
 plt.grid(True)
 
 plt.tight_layout()
-plt.savefig.format = "png"
-plt.savefig("gflops_peak.png")
+plt.savefig.format = "pdf"
+plt.savefig("gflops_peak.pdf")
 #plt.show()
 #### END OF CHART
 
@@ -96,5 +96,5 @@ plt.legend(loc=2)  # upper left
 plt.grid(True)
 
 plt.tight_layout()
-plt.savefig.format = "png"
-plt.savefig("speedup_wrt_base.png")
+plt.savefig.format = "pdf"
+plt.savefig("speedup_wrt_base.pdf")

### 2013-10-13 ieee_paper.tex
Latest version. I changed problem_statement to discuss RBF.
cleaning comments
intro

@@
-%% bare_conf.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
 \documentclass[10pt, conference, compsocconf]{IEEEtran}
@@
 \input{support/setupicase}
-% Add the compsocconf option for Computer Society conferences.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
 
-%\newcommand{\red}[1]{\textbf{\textcolor{red}{#1}}}
-%\newcommand{\blue}[1]{\textbf{\textcolor{blue}{#1}}}
 \def\red#1{\textbf{\textcolor{red}{#1}}}
@@
 \def\ge#1{}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
-% handle arbitrary number of unprocessed floats
 \usepackage{morefloats}
 
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-   \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at: 
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
+\usepackage{graphicx}
 \usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
 \usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
 \usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
 \hyphenation{op-tical net-works semi-conduc-tor}
@@
 \begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
 \title{Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the
    MIC Architecture}
-
-% author names and affiliations
-% use a multiple column layout for up to two different
-% affiliations
 
@@
 Email: bollig@gmail.com}}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-% 
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3}, 
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
 \maketitle
@@
 
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
@@
 \cite{Bell08, Vuduc05, Nishtala07, Stock12-TACC, Cuthill69, cramer2012openmp, Buluc2009_SPAA, Buluc11, Im01, Mellor-Crummey04, Nishtala07, Saad94sparskit, Williams07}
+\cite{Temam:1992:CBS:147877.148091}
+\cite{DBLP:journals/ijhpca/ShantharamCR11,
+conf/ppsc/Toledo97, Liu:2013:ESM:2464996.2465013, Molka:2009:MPC:1636712.1637764,%
+DBLP:journals/corr/abs-1101-0091, conf/ipps/BulucWOD11, conf/ipps/KreutzerHWFBB12,%
+kumar2012accelerating, journals/concurrency/VazquezFG11}
+
+SpMV is an important kernel for lots of stuff. So improving the
+performance of SpMV has caputred the interest of many researcher.
+
+The main challenge that is faced to improve good performance for SpMV
+is that the operation are conducted using memory location that are
+irregular and often unpredictible. That make that the kernels are
+mostly memory bound and there is a significant instruction overhead
+per flop.
+
+Common improvement techniques such as register blocking, bandiwdth
+reduction (matrix reordering), partitioning to fit in cache or TLB
+have impacts which are very dependent on the matrix and overall do not
+lead to dramatic improvement. Assuming register blocking does not
+apply well to the matrix at hand (which is true for most matrices),
+there is about 8 bytes of the matrix to move in per non zero (assuming
+single precision); each nonzero requires two floating point operations
+leading to a flop-to-byte ratio of at most $\frac{1}{4}$. This limits
+the obtained performance to at most a quarter of the bandwidth of the
+architecture wasting a lot of potentially useful cycles. The commonly
+used techniques are mostly designed to reach that bound rather than
+overcome it.
+
+Fortunately that fate is not inevitable. One solution would be to pair
+multiple component of an application to schedule a more instruction
+intensive kernel simultaneously with SpMV, relying on some hardware
+threading capabilities, such as HyperThreading, to reduce the cycle
+wastage. However most ot the applications that use SpMV do not
+typically have an instruction intensive kernel to run simultaneously.
+
+An other solution, and the one we pursue in this paper, is to compute
+multiple SpMV at once on matrices that have the same sparsity
+patterns. Obviously not all the applications have such a property. But
+important classes of applications such as graph
+recommendation~\cite{}, eigensolving~\cite{} and the computation of
+derivative of Radial Basis Functions(RBF)~\cite{} can use multiple
+SpMVs simultaneously. In this paper in particular, we investigate the
+case of the derivative of RBFs where four derivatives of four
+different function is expressed as the multiplication of four vectors
+by four matrices with identical sparsity patterns leading to the
+simultaneous execution of 16 SpMVs at a time.
+
+To perform our analysis, we focus our attention on the improvement
+that we can achieve on the Intel Xeon Phi processor. It follow the
+Many Integrated Core (MIC) architecture, which has a significant
+memory bandwidth and peak flop throughput thanks to its 512-bit large
+SIMD registers. The Xeon Phi processor has been shown to be promising
+for sparse linear algebra compared to more classical CPU or GPU
+architecture~\cite{}.
+
+In Section~\ref{} we present the computation of RBFs and how it can be
+expressed 16 multiplication of 4 vectors by 4 matrices. Section~\ref{}
+presents an estimation of the instruction intensity of various form of
+the computations and we show that a 5- to 6- fold improvement can be
+expected when computing the 16 multiplications simultaneously and
+reach a total of about 200 Gflop/s. This performance represents
+approximatively 10\% of the available flop/s of a Knight Corner
+coprocessor. Therefore, it is necessary to have implementations that
+perform the computation in as little amount of instructions as
+possible. We describe in Section~\ref{} the details of the MIC
+architecture and how to use specialized load, store, swizzle and
+permutation instruction to efficiently bring the data in the vector
+registers to be processed. Section~\ref{} gives some experimental
+result about the amount of bandwidth that can be achieved depending on
+how the spmv kernel is written and the actual performance of the
+various kernel on multiple classes of matrices some generated for
+analysis purpose and some extracted from and RBF
+application. Concluding remarks and perspectives are provided in
+Section~\ref{}.
+
+\section{Derivatives of Radial Basis Functions}
+
 \input{problem_statement}
@@
 
-%Wherever Times is specified, Times Roman or Times New Roman may be used.
-If neither is available on your system, please use the font closest in appearance to Times.
-Avoid using bit-mapped fonts if possible.
-True-Type 1 or Open Type fonts are preferred.
-Please embed symbol fonts, as well, for math, etc.
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation Results}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
-%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that IEEE does not put floats in the very first column - or typically
-% anywhere on the first page for that matter. Also, in-text middle ("here")
-% positioning is not used. Most IEEE journals/conferences use top floats
-% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
-% command of the stfloats package.
-
 \section{Conclusion}
@@
 
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgement
 \section*{Acknowledgment}
@@
 
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-%5\begin{thebibliography}{1}
-
 \bibliographystyle{plain} %{IEEEabrv,paper.bib}

### 2013-10-13 ieee_paper.tex
adding proper sections and labels

@@
 
-In Section~\ref{} we present the computation of RBFs and how it can be
-expressed 16 multiplication of 4 vectors by 4 matrices. Section~\ref{}
-presents an estimation of the instruction intensity of various form of
-the computations and we show that a 5- to 6- fold improvement can be
-expected when computing the 16 multiplications simultaneously and
-reach a total of about 200 Gflop/s. This performance represents
-approximatively 10\% of the available flop/s of a Knight Corner
-coprocessor. Therefore, it is necessary to have implementations that
-perform the computation in as little amount of instructions as
-possible. We describe in Section~\ref{} the details of the MIC
-architecture and how to use specialized load, store, swizzle and
-permutation instruction to efficiently bring the data in the vector
-registers to be processed. Section~\ref{} gives some experimental
-result about the amount of bandwidth that can be achieved depending on
-how the spmv kernel is written and the actual performance of the
-various kernel on multiple classes of matrices some generated for
-analysis purpose and some extracted from and RBF
-application. Concluding remarks and perspectives are provided in
-Section~\ref{}.
+In Section~\ref{sec:rbf} we present the computation of RBFs and how it
+can be expressed 16 multiplication of 4 vectors by 4
+matrices. Section~\ref{sec:model} presents an estimation of the
+instruction intensity of various form of the computations and we show
+that a 5- to 6- fold improvement can be expected when computing the 16
+multiplications simultaneously and reach a total of about 200
+Gflop/s. This performance represents approximatively 10\% of the
+available flop/s of a Knight Corner coprocessor. Therefore, it is
+necessary to have implementations that perform the computation in as
+little amount of instructions as possible. We describe in
+Section~\ref{sec:imple} the details of the MIC architecture and how to
+use specialized load, store, swizzle and permutation instruction to
+efficiently bring the data in the vector registers to be
+processed. Section~\ref{sec:expe} gives some experimental result about
+the amount of bandwidth that can be achieved depending on how the spmv
+kernel is written and the actual performance of the various kernel on
+multiple classes of matrices some generated for analysis purpose and
+some extracted from and RBF application. A performance of xxx GFlop/s
+is achieved on real scenario. Concluding remarks and perspectives are
+provided in Section~\ref{sec:ccl}.
 
 \section{Derivatives of Radial Basis Functions}
+\label{sec:rbf}
+
+\section{Modelization of the Potential Improvements}
+\label{sec:model}
+
+\section{Efficient Implementation on the Intel Xeon Phi processor}
+\label{sec:impl}
+
+\section{Experimental Validation}
+\label{sec:expe}
 
@@
 
-\section{Conclusion}
+\section{Conclusions and Perspectives}
+\label{sec:ccl}
+
 The conclusion goes here. this is more of the conclusion

### 2013-10-13 ieee_paper.tex
adding proper sections and labels

@@
 In Section~\ref{sec:rbf} we present the computation of RBFs and how it
-can be expressed 16 multiplication of 4 vectors by 4
-matrices. Section~\ref{sec:model} presents an estimation of the
-instruction intensity of various form of the computations and we show
-that a 5- to 6- fold improvement can be expected when computing the 16
-multiplications simultaneously and reach a total of about 200
-Gflop/s. This performance represents approximatively 10\% of the
-available flop/s of a Knight Corner coprocessor. Therefore, it is
-necessary to have implementations that perform the computation in as
-little amount of instructions as possible. We describe in
-Section~\ref{sec:imple} the details of the MIC architecture and how to
-use specialized load, store, swizzle and permutation instruction to
-efficiently bring the data in the vector registers to be
-processed. Section~\ref{sec:expe} gives some experimental result about
-the amount of bandwidth that can be achieved depending on how the spmv
-kernel is written and the actual performance of the various kernel on
-multiple classes of matrices some generated for analysis purpose and
-some extracted from and RBF application. A performance of xxx GFlop/s
-is achieved on real scenario. Concluding remarks and perspectives are
-provided in Section~\ref{sec:ccl}.
+can be expressed 16 multiplication of 4 vectors by 4 sparse matrices
+with a common sparsity pattern. Section~\ref{sec:model} presents an
+estimation of the instruction intensity of various form of the
+computations and we show that a 5- to 6- fold improvement can be
+expected when computing the 16 multiplications simultaneously and
+reach a total of about 200 Gflop/s. This performance represents
+approximatively 10\% of the available flop/s of a Knight Corner
+coprocessor. Therefore, it is necessary to have implementations that
+perform the computation in as little amount of instructions as
+possible. We describe in Section~\ref{sec:impl} the details of the MIC
+architecture and how to use specialized load, store, swizzle and
+permutation instruction to efficiently bring the data in the vector
+registers to be processed. Section~\ref{sec:expe} gives some
+experimental result about the amount of bandwidth that can be achieved
+depending on how the spmv kernel is written and the actual performance
+of the various kernel on multiple classes of matrices some generated
+for analysis purpose and some extracted from and RBF application. A
+performance of xxx GFlop/s is achieved on real scenario. Concluding
+remarks and perspectives are provided in Section~\ref{sec:ccl}.
 
@@
 
-\input{problem_statement}
-\input{intel_phi}
-\input{tables_and_figures}
-\input{bandwidth}
-\input{notation}
+%\input{problem_statement}
+%\input{intel_phi}
+%\input{tables_and_figures}
+%\input{bandwidth}
+%\input{notation}
 

### 2013-10-13 ieee_paper.tex
scrapping text that appears directly reusable. No longer use \input

@@
 \def\bold#1{{\bf #1}}
+
+\usepackage{soul}
+\usepackage{xspace}
+\usepackage{color} 
+\definecolor{darkgreen}{rgb}{0,0.5,0}
+\usepackage[colorlinks=true,% 
+  linkcolor=red,% 
+  citecolor=darkgreen,%
+  urlcolor=blue]{hyperref}
+
+\newcommand{\todo}[1]{\color{red}\textbf{\hl{#1}}\color{black}\xspace}
 
@@
 
+%In the theory of Radial Basis Functions Finite-Difference (RBFFD),
+%derivatives of a function $f(\rvec)$ at node $i$ are expressed as a
+%linear combination of the function values at the stencil center and
+%the nodes connected to node $i$ (Figure~\ref{fig:rbf_stencils}). Thus
+%$y$ is a discrete derivative of the vector $x$.
+
+In this paper, we propose yet a new idea, that is applicable to the
+world of Finite-Difference Radial Basis functions (RBF-FD). A radial
+basis functions refers to a set of shifted radially symmetric
+functions $\phi_i(r) =\phi{||\xvec-\xvec_i||}$ centered at
+$\xvec_i$\cite{}. Any function $f(\xvec)$ defined in a domain $\Omega$
+can be expanded in this basis according to
+\begin{equation}
+f(\xvec) =  \sum_i w_i \phi_i(\xvec) \label{eq:rbf}
+\end{equation}
+% Wrong label eq:rbf in document 
+Given function values at a set of discrete points $x_i$, one inverts
+linear system (\ref{eq:rbf}) to obtain a set of weights $w_j$. Given a
+set of $N$ RBF nodes, the matrix to invert is dense. Derivatives of
+$f(\xvec)$ result from the differentiation of Eq. (\ref{eq:rbf}). For
+very large problems, this is an computationally expensive since the
+RBFs do not form an orthogonal basis, nor are they amenable to tensor
+decomposition in the three coordinate directions. Rather, their
+strength is the ability to randomly distribute nodes across complex
+physical domains, and have an implementation that is independent of
+dimensionality.
+
+To alleviate the cost of a global approach, researchers \cite{} expand
+$f(\xvec)$ in a local region of $x_i$. Thus,
+$$
+f(\xvec) =  \sum_{i=0}^{n_s} w_i \phi_i(\xvec) \label{eq:rbf}
+$$
+where one only considers a local set of $n_s$ neighbors of of point $x_j$. 
+Finite-difference RBF expresses the derivative of $f(\xvec)$ at $x_j$ as a linear
+combination of $f(\xvec)$ at the $n_s$ neighbors. Thus, 
+$$
+f = A w
+$$ where $f$ is the vector of function values at the $n_s$ stencil
+points (including the center point $x_j$. The matrix $A$ is a function
+of $\phi_{ij}(\xvec) =\phi(||x_i-x_j||)$. This leads to the $n_s$
+weight values $w = A^{-1} f$. The derivative of Eq. \ref{eq:rbf}
+evaluated at $x_j$ gives
+$$
+   f'_j = \sum_{i=0}^{n_s} w_i \phi'_{ij}
+$$
+Substituting the vector of weights $w$ leads to 
+$$
+  f'_j = \sum_{i=0}^{n_s} \alpha_i f_i
+$$
+The derivative evaluation at all points of the domain $\Omega$, notwithstanding boundary conditions, takes the form
+$$
+  y = A x
+$$ 
+where $x$ is the souce vector of function values, and $y$ is the
+vector of derivative values. The matrix $A$ is called a derivative
+matrix. In the current litterature, the number of stencil points is a
+fixed value, independent of the node under consideration\cite{}. The
+computation of a single derivative has been reduced to a SpMV, where
+each row has $n_s$ nonzeros. Because the sparsity of each row is
+constant, ELLPACK~\cite{} becoes the most appropriate matrix
+compression scheme\cite{}. In practice, $n_s=32$ in two-dimensional
+flows and 64 or 100 for three-dimensional flows. These numbers are
+similar to what is used in finite-element codes.
+
+Many problems in fluid dynamics and in the geosciences require the solution to
+transport equations of the form
+$$
+\pf{Q}{t} = f(Q,Q_x, Q_y, Q_z, \Laplacian{Q})
+$$ 
+where $Q$ is a vector of unknowns (3 components of velocity,
+pressure, temperature). Thus, it is often necessary to compute the $x$
+derivative of multiple functions, typically four for the Euler
+equations, or five for the Navier-Stokes equations. Multiple
+right-hand sides transform a SpMV into a SpMM (Sparse Matrix/dense
+Matrix multiplication), which improves register utilization and
+decreases cache misses by vectorizing over the multiple source
+vectors. Further improvements are possible by recognizing that
+different derivative matrices (x,y,z and Laplacian for example), have
+the same sparsity distribution; only the weight change.  Thus,
+alternative to computing a derivative of multiple functions, we can
+calculate multiple derivatives of a single functions. The increased
+memory bandwidth due to an increase in the number of derivaive
+matrices is offset by better cache utilization, leading to an overall
+benefit.
+
+**** PUT ELSEWHERE OR ELIMINATE? ****
+
+When solving a one dimensional system of PDEs, one might require an
+$x$ derivative of multiple functions. For example, the Euler equations
+require the $x$ derivative of the three components of velocity and
+pressure. In this case, there are $n_v=4$ vectors $x^k$,
+$k=0,\cdots,3$.
+
+Thus for each vector element $y_i$, we compute $y_i = \sum_j A_{ij}
+x_j$. If $A_i$ is row $i$ of $A$, $y_i$ is simply the dot product $A_i
+x$. The next level of generality is to consider $n_v$ vectors $x^k$,
+$k=0,\cdots,n_v-1$. Whatever the spmv implementation, one achieves
+improved performance if the matrix formed from the columns $x^k$ are
+stored in row major order. Thus, $x^0_0,x^1_0,\cdots,x^{n_v-1}_0$, are
+stored in consecutive memory location. The random access of the
+elements of $x$ is thus reduced. Maximum efficiency is achieved when
+$n_v=16$ floats or $8$ doubles, given that cache lines take 64
+bytes. We will benchmark this case, labeled $Svn$, where $n$ refers to
+the number of vectors (Iv4 uses four vectors). The $S$ refers to singe
+precision. A double precision run is labelled $Dvn$.
+
+Alternatively, when solving a PDE, one might require derivatives of a
+given scalar function with respect to coordinate directions $x$, $y$,
+$z$. Second order operators of often required, such as a second
+derivative with respect to $x$ or a Laplacian operator. In the RBFFD
+formulation, on can compute different derivatives using the same
+stencil, but with different weights. In other words, the adjacency
+matrix that corresponds to $A$ remains constant, but the matrix
+elements of $A$ change with the particular derivative.  In this case,
+label with a superscript $l$ the particular matrix $A^k$. Since the
+adjacency matrix is assumed invariant, there is only need for a single
+matrix \ttt{C{ij}}. In Ellpack format, each row is of constant size
+(the number of nonzeros per row of $A$. $C_{ij}$ is the column number
+that locates the $j^{th}$ nonzero in row $i$ of $A$.
+
+\todo{Must rewrite the above to make it more focused on the
+  application. Remove the architectural/implementation details such as
+  using ELLPACK. Let's focus on on why this computation is important
+  and how does the 16 multiplication appears in the equations.}
+
+\def\wide{2.5in}
+\newfig{figures/rbf_stencils.pdf}{\wide}{RBFFD Stencils.
+Each node of the mesh is connected to $n_z-1$ stencil nodes in addition to itself.
+In the figure, node $A$ is connected to $B$, but $B$ is {\em not\/} connected to $A$.
+Thus adjacency graph of $A$ is not-symmetric.
+This must be corrected when reducing the bandwidth of $A$ with a Cuthill\-McKee algorithm, which assumes symmetry.}{fig:rbf_stencils}
+
+\newfig{figures/matrix_structure.pdf}{\wide}{Matrix Structure.}{fig:mat_struct}
+
 \section{Modelization of the Potential Improvements}
 \label{sec:model}
+
+In this section, we introduce some variables that will prove useful to 
+describe our experiments and construct a working model to examine the pros and cons of various assumptions.
+While deceptively simple, the MIC architecture requires a careful consideration of the properties of cache, memory bandwidth, cores and their interaction, threads per core (1-4), parallelization and vectorization.
+We will also seek to perform experiments that measure the best possible performance of the SPMV with maximum and minimum cache thrashing \qes{same as cache misses?}, with and without floating point operations, and with and without memory transfer.
+At the onset, we expect the dominant cost to e the transfer of the vector $x$ from memory, because contiguous elements of $x$ are not accessed sequentially.
+
+Relevant notation is contained in Table~\ref{tab:not}.
+
+\begin{center}
+\begin{tabular}{|c|l|}
+%\hline
+%& & \\
+\hline
+$b_i$ & number of bytes per int \\
+$b_x$ & number of bytes per float(4)/double(8) \\
+$n_z$ & number of nonzeros per row $A$ \\
+$n_r$ & number of rows of $A$ \\
+$n_c$ & total number of elements in $col_id$: $n_z n_r$ \\
+$b_w$ & matrix bandwidth \\
+$b_c$ & total L2 cache per core ($512k=2^{19}$ bytes) \\
+$b_T$ & theoretical minimum number of bytes transferred  \\
+$b_{wT}$ & theoretical minimum number of bytes written to memory  \\
+$b_{rT}$ & theoretical minimum number of bytes read from memory  \\
+$n_C$ & total number of cores used  \\
+$t_C$ & total number of threads used per core \\
+$\rho$ & average cache density  \\
+$n_v, n_m$ & number of vectors and matrices \\
+$n_F$ & total number of floating point operations  \\
+\hline
+\end{tabular}
+\end{center}
+
+- weights and vector elements are either all floats or all doubles.
+
+Let us first estimate the memory (in a serial implementation) required to compute $y = Ax$ where $y$ has $n_r$ rows and $n_m$ columns, $A$ has
+$n_c=n_r n_z$ nonzero elements, and $x$ is a vector of $n_r$ rows and 
+$n_v$ columns. In this paper, we assume that the adjacency matrix for the
+$n_m$ matrices $A$ is constant, but the values stored in the various $A$ differ.
+We are therefore executing $n_v n_m$ spmv operations.
+The objective of course, is to minimize wall-clock time on the MIC.
+
+The total number of bytes is 
+$$
+   b_T = n_r (n_z b_i + n_z b_x n_m + b_x n_v)
+$$
+The ratio of $b_T$ to the total cache  over all cores is given by
+$$
+   R = \frac{b_T}{b_c}
+$$
+Clearly, the memory $b_T$ is composed of the memory required by the vectors $x$, $y$, the nonzeros of $A$ and the elements of the matrix \ttt{col}, which stores the nonzero columns in each row, and is an integral component of the Ellpack compressed matrix format specification.
+While the Ellpack format is less general than CSR, it is the idea structure for the RBBF simulations we are intersted in, wherein the number of RBF nodes per stencil is constant throughout the 2D or 3D grid and in time.
+
+Let us consider separately the number of bytes read from and written to memory.
+Both the matrix $n_m$ matrices $A$ \qes{need an index on $A$ since there are more than one?} and \ttt{col\_id} are read from memory, as is $x$, whereas, the vector$y$ are stored to memory.
+In a later sections, we will discuss the practical performance of these operations and estimate the dominant contributions.
+\qes{We also wish to isolate the effects of read, write and compute operations as a function of the number of cores and threads to see how they influence the results and to estimate whether our results have the potential to scale similarly on systems with a higher number of cores.
+We "might" also compare are results against the best implementation using OpenCL, or do so in the paper.}
+
+The total number of bytes read into memory is
+$$ 
+n_{rT} = n_r n_v b_x + n_r n_z b_i + n_r n_m n_z b_x
+$$
+while the total number of bytes written to memory is
+$$
+n_{wT} = n_r n_m n_v b_x
+$$
+
+We compute the floating point operations of the SpMV with general $n_v$ and $n_m$. Calling the total number of floating point operations $N_F$, we find
+$$
+  N_F = 2 n_r n_z n_v m_m
+$$
+Two cases must be distinguished in regards to $x$. In the first, in a worst case scenario, each element 
+of $x$ is transferred $n_z$ times from memory.
+In the best case, the case of infinite cache, each element is transferred only once.
+Thus, we provide best and worst case scenarios and will compare against matrices $A$
+in each of these extreme situations to evaluate the degree we approach idea performance. 
+Thus the number of flops to bytes transferred is $N_F / (n_{rT} + n_{wT})$, which when written out, 
+\begin{eqnarray}
+N_{Fw} &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_z n_v  + n_r n_z b_i + n_r n_m n_z b_x + n_r n_m n_v b_x} \nonb \\
+    &=& \frac{2 n_v n_m}{b_x ( n_v + n_m + n_m n_v n_z^{-1}) + b_i}  \nonb \\
+%N_F &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_v (n_v+n_m) + n_r n_z b_i + n_r n_m n_z b_x} \nonb \\
+    %&=& \frac{2 n_z n_v n_m}{b_x (n_v+n_m) + n_z (b_i + n_m b_x)} \nonb 
+\end{eqnarray}
+while in the best case, 
+\begin{eqnarray}
+N_{Fb} &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_v  + n_r n_z b_i + n_r n_m n_z b_x + n_r n_m n_v b_x} \nonb \\
+    &=& \frac{2 n_v n_m}{b_x ( n_v n_z^{-1} + n_m  + n_m n_v n_z^{-1}) + b_i } 
+\end{eqnarray}
+In the special case $n_v=n_m=1$, 
+$$
+N_{Fw} = \frac{2}{b_x (2+n_z^{-1}) + b_i}
+$$
+and 
+$$
+N_{Fb} = \frac{2}{b_x (1 + 2 n_z^{-1}) + b_i}
+$$
+Neglecting $2$ compared to $n_z$, we find that $N_F = 2/(b_i+b_x)$, which leads to 
+$N_F = 2/8=1/4$ and  $N_F=2/12=1/6$ for single and double precision, respectively. 
+In the general case, in the condition of $n_z$ large compared to 2, 
+$$
+N_F =  \frac{2 n_v n_m}{n_m b_x + b_i}
+$$ 
+The case are interested in is $n_m=n_v=4$, so that
+$$
+N_F = \frac{8}{ b_x + 1}
+$$
+which is 8/5 and 8/9 in single and double precision, respectively. 
+In single and double precision, we get $N_F=32/36=8/5$ and $N_F=32/68=8/9$, respectively.
+The third column of the following table (TABLE ???) is the maximum achievable Gflop rate, assuming 
+a maximum memory transfer speed of $150 Gflops$ (see earlier benchmark in this paper). 
+(higher speeds can be achieved in practice,  upwards of 190 Gflops (Saule \etal\cite{}), but we use
+a benchmark code similar to our current algorithm, without the indirect acessing required for $x$, but keeping the same use of vector registers used in our fastest implementation.
+(DO WE INCLUDE the CODE?)
+
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+Precision & $n_m=n_v=1$ & $n_m=n_v=4$     \\
+\hline
+single    &  1/4  (37.5)     &   8/9 (133) \\
+double    &  1/6  (25.5)     &   8/17 (62) \\
+\hline
+\end{tabular}
+\end{center}
+(With a 190 Gflop peak memory rate, maximum performance is 47.5 and 170 Mflops for the 1/1 case, while we calculate 32 and 90 Gflops in the 4/4/ case.
+
+Let us also calculate the flop to byte ratio for the case of 1/4 and 4/1 ($n_v$/$n_m$). One notices the symmetry with respect to $n_v$ and $n_m$. 
+One gets $N_F = \frac{2*4}{5 b_x + 4}$ equal to 8/24 (=.33) and 8/44 (=.18), in single and double precision, respectively, which corresponds to 49 and 27 Gflops if a maximum of 150 Gflops is assumed.
+As $n_m$ and $n_v$ keep increasing, $N_F$ tends towards $2/b_x$, equal to $0.5$ and $0.25$ respectively for single and double precision, which leads to peak speed of 75 and 37.5 Gflops.
+
+In single precision, our best benchmarks achieve 20 Gflops for single precisions for on matrix/one vector, and 100 Glops for the 4/4 matrix/vector case, or 53\% and 75\% of peak performance, respectivey.
+We conclude that we achieve a higher percentage of peak performance for this algorithm when calculating 16 output $y$ vectors.
+The results are even more impressive when measured against the peak perforance of 2Tflops of the MIC in double precision.
+
+\newfig{figures/gflops_peak.png}{2.5in}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors.
+We used a memory bandwidth of 150 Gbytes/sec (based on measurements in a code similar to the SpMV kernel, with computations removed.)}{fig:gflops_peak_perf}
+
+\newfig{figures/speedup_wrt_base.png}{\wide}{Speedup relative to base case using one matrix and one vector.}{fig:speedup}
 
@@
 
+\subsection{Knights Corner}
+
+In this work, we use a pre-release KNC card SE10P. The card has 8
+memory controllers where each of them can execute 5.5 billion
+transactions per second and has two 32-bit channels. That is the
+architecture can achieve a total bandwidth of 352GB/s aggregated
+across all the memory controllers. There are 61 cores clocked at
+1.05GHz. The cores’ memory interface are 32-bit wide with two channels
+and the total bandwidth is 8.4GB/s per core. Thus, the cores should be
+able to consume 512.4GB/s at most. However, the bandwidth between
+the cores and the memory controllers is limited by the ring network
+which connects them and theoretically supports at most 220GB/s.
+
+Each core in the architecture has a 32kB L1 data cache, a 32kB L1
+instruction cache, and a 512kB L2 cache. The architecture of a core is
+based on the Pentium architecture: though its design has been updated
+to 64-bit. A core can hold 4 hardware contexts at any time. And at
+each clock cycle, instructions from a single thread are executed. Due
+to the hardware constraints and to overlap latency, a core never
+executes two instructions from the same hardware context
+consecutively. In other words, if a program only uses one thread, half
+of the clock cycles are wasted. Since there are 4 hardware contexts
+available, the instructions from a single thread are executed
+in-order. As in the Pentium architecture, a core has two di↵erent
+concur- rent instruction pipelines (called U-pipe and V-pipe) which
+allow the execution of two instructions per cycle. However, some
+instructions are not available on both pipelines: only one vector or
+floating point instruction can be executed at each cycle, but two ALU
+instructions can be executed in the same cycle.
+
+Most of the performance of the architecture comes from the vector
+processing unit. Each of Intel Xeon Phi’s cores has 32 512-bit SIMD
+registers which can be used for double or single precision, that is,
+either as a vector of 8 64-bit values or as a vector of 16 32-bit
+values, respectively. The vector processing unit can perform many
+basic instructions, such as addition or division, and mathematical
+operations, such as sine and sqrt, allowing to reach 8 double
+precision operations per cycle (16 single precision). The unit also
+sup- ports Fused Multiply-Add (FMA) operations which are typically
+counted as two operations for benchmarking purposes. Therefore, the
+peak performance of the SE10P card is 1.0248 Tflop/s in double
+precision (2.0496 Tflop/s in single precision) and half without FMA.
+
+\todo{rewrite that to make sure it fits the description of the card
+  used in the experiments. This text with copy/pasted from the PPAM
+  paper}
+
+\subsection{Bringing the data in vector register}
+
+Registers on the MIC are 512 bits wide, and can store 16 floats or 8
+doubles. It is through these registers that the \#SIMD pragma is able
+to achieve vectorization. (In comparison to OpenCL or CUDA, one can
+think of these registers as a warp.)  We use the following vector
+operations in our code, which we'll explain in the text.
+
+\def\loadps{\ttt{\_mm512\_load\_ps}}
+\def\loadds{\ttt{\_mm512\_load\_ds}}
+\def\fmadps{\ttt{\_mm512\_fmad\_ps}}
+\def\gatherps{\ttt{\_mm512\_i32gather\_ps}}
+\def\swizzleps{\ttt{\_mm512\_swizzle\_ps}}
+\def\storenrngops{\ttt{\_mm512\_storenrngo\_ps}}
+\def\castsi{\ttt{\_mm512\_castsi512\_ps}}
+\def\permute{\ttt{\_mm512\_permute4f128\_epi32}}
+\def\intmask{\ttt{\_mm512\_int2mask}}
+\def\loadunpack{\ttt{\_mm512\_mask\_loadunpacklo\_epi32}}
+\def\castitops{\ttt{\_mm512\_castsi512\_ps}}
+\def\castpstoi{\ttt{\_mm512\_castps\_si512\_ps}}
+%
+\begin{center}
+\begin{tabular}{|l|l|}
+\hline
+\loadps &  load 16 consecutive floats to register\\
+\fmadps &  multiply/add of 16 floats\\
+\gatherps &  gather 16, possibly disconnected floats from memory\\
+\swizzleps &  reorder a channel, and duplicate across channels\\
+\storenrngops &  store 16 floats to memory without register rewrite or reordering\\
+\castpstoi & reinterpret 16 floats as integers\\
+\castitops & reinterpret 16 integers as floats\\
+\permute &  permute channels; do not change individual channels\\
+\intmask &  \\
+\loadunpack &  \\
+\hline
+\end{tabular}
+\end{center}
+%
+In the vector instructions, \ttt{ps} refers to a float (4 bytes), while \ttt{epi32} refers to a 4-byte integer.
+Although we do not discuss double precision in this paper, \ttt{ds} refers to an 8 byte double precision real number (i.e., \loadds).
+Each vector register is broken up into four channels of 128 bits each.
+It is possible to interchange these channels at a cost of a "few" \qes{exact numbers?} cycles, and it is also possible to execute swizzle operations within a channel.
+For example, if the 128 bytes of each channel are labelled as $ABCD$, the vector instruction
+\ttt{\_mm512\_swizzle\_ps(v, \_MM\_SWIZ\_REG\_AAAA)}
+replaces each channel by $AAAA$.
+
+Here is a function that reads in four floats (a,b,c,d) and creates the 16-float vector dddd,cccc,bbbb,aaaa: 
+\begin{verbatim}
+__m512 read_aaaa(float* a)                                                              
+{
+    int int_mask_lo = (1 << 0) + (1 << 4) + (1 << 8) + (1 << 12);
+    __mmask16 mask_lo = _mm512_int2mask(int_mask_lo);
+    __m512 v1_old;
+    v1_old = _mm512_setzero_ps();
+    v1_old = _mm512_mask_loadunpacklo_ps(v1_old, mask_lo, a);
+    v1_old = _mm512_mask_loadunpackhi_ps(v1_old, mask_lo, a);
+    v1_old = _mm512_swizzle_ps(v1_old, _MM_SWIZ_REG_AAAA);
+    return v1_old;
+}
+\end{verbatim}
+
+Here is a function that takes four floats from float register \ttt{v1}, and 
+places them in each of the four lanes. Because the permulation function only 
+operates on 4-byte integers, it is necessary to convert (in place) each float
+to an integer (the bit structure is not changed), permute the lanes and cast
+the integers back to floats (with no modification of the bit structure)
+\begin{verbatim}
+__m512 permute(__m512 v1, _MM_PERM_ENUM perm)                                           
+{
+    __m512i vi = _mm512_castps_si512(v1);
+    vi = _mm512_permute4f128_epi32(vi, perm);
+    v1 = _mm512_castsi512_ps(vi);
+    return v1;
+}
+\end{verbatim}
+
+\newfig{figures/swizzling.pdf}{\wide}{Duplication of the first four bytes of each channel across the entire channel.}{fig:swizzling}
+\newfig{figures/channel_permutation.pdf}{\wide}{Channel Permutations on the MIC}{fig:permutation}
+\newfig{figures/tensor_product.pdf}{\wide}{Tensor scalar product using channels and swizzling.}{fig:tensor_product}
+
 \section{Experimental Validation}
@@
 
-%\input{problem_statement}
-%\input{intel_phi}
-%\input{tables_and_figures}
-%\input{bandwidth}
-%\input{notation}
+\subsection{Instances}
+
+\todo{How were instances generated. Detail the different type of
+  instances and the rational for trying them. Detail how the realistic
+  instances were generated. Explain matrix ordering (Kutill McKee).}
+
+\subsection{Bandwidth}
+
+\newfig{figures/test1_readwrite.png}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without MIC vector instructions.}{fig:band_rw}
+
+\newfig{figures/test1_gather.png}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without MIC vector instructions.}{fig:band_gather}
+
+\newfig{figures/test3_gather.png}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without MIC vector instructions.
+The greater speed of the cpp version is obtained throught the use of \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}.
+All memory is aligned on 64 bytes.}{fig:band_gather_ivdep}
+
+\subsection{Computations}
+
+\newfig{figures/host_test1_readwrite_no_temporal_hint.png}{\wide}{Bandwidth performance on the host under idealized conditions as a function of matrix row size.
+Entries with "cpp" denote cases where coding was performed without AVX vector instructions.
+The speed on the CPU matches the speed with AVX instructions.
+All memory is aligned on 32 bytes.}{fig:read_write}
+
+\newfig{figures/mic_performance_nb_threads.png}{\wide}{Performance of $y=Ax$ on the MIC.
+Squares: base 1/1 case, solid circles: 4/4 case implemented in C++, solid triangles: 4/4 case implemented with MIC vector instructions.
+$-O3$ compilation options.
+Each group consists of four cases: grids of $64^3$ and $96^3$ with and without Reverse Cuthill McKee.}{fig:mic_performance}
+
+% NOTES: why is result the same at 32 processors.
+I should find out the distribution of threads.
+16 threads means one thread on each of 16 cores (YES, I BELIEVE), or 2 threads on each of 8 cores.} Must run the experiment with different nodes orderings.
+%export OMP_SCHEDULE=guided,64    (I should redo experiment with Dynamic or static and compact for KMP_AFFINITY
+%export KMP_AFFINITY=compact
+% MUST PRINT OUT OMP_SCHEDULE and KMP_AFFINITY inside the program. 
+% WHY the very high performance with very few nodes? STRANGE. 
+\newfig{figures/host_performance.png}{\wide}{Performance of $y=Ax$ on the Host.
+Squares: base 1/1 case, solid circles: 4/4 case implemented in C++, solid triangles: 4/4 case implemented with MIC vector instructions.
+$-O3$ compilation options.
+The four colors distinguish grid resolution and whether or not
+Reverse Cuthill McKee is applied.}{fig:host_performance}
 

### 2013-10-13 ieee_paper.tex
minor + using pdf figures

@@
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
+\documentclass[10pt,conference,compsocconf]{IEEEtran}
 \input{support/macros} % color is defined in macros or misc_mac
@@
 \def\ge#1{}
-\usepackage{morefloats}
+%\usepackage{morefloats}
 
-\usepackage{graphicx}
-\usepackage[tight,footnotesize]{subfigure}
-\usepackage{fixltx2e}
+\usepackage[pdftex]{graphicx}
+\usepackage{subfigure}
+%\usepackage{fixltx2e}
 \usepackage{url}
@@
 \begin{document}
-\title{Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the
-   MIC Architecture}
+\title{Sparse Matrix Vector Multiplication with Multiple vectors and
+  Multiple Matrices on the MIC Architecture}
 
@@
 \IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte\\
-Email: erik.saule@uncc.edu}
+Email: esaule@uncc.edu}
 \IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR \\
-Email: 
-flyer@ucar.edu}
+Email: flyer@ucar.edu}
 \IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
@@
 \begin{abstract}
-In this paper, we develop an efficient scheme for the calculation of derivatives within the context
-of Radial Basis Function Finite-Difference (RBFFD). RBF methods express functions as a linear
-combination of radial basis functions on an arbitrary set of nodes. The Finite-Difference component
-expresses this combination over a local set of nodes neighboring the point where the derivative is sought.
-The derivative at all points takes the form of a sparse matrix/vector multiplication (spmv).
+In this paper, we develop an efficient scheme for the calculation of
+derivatives within the context of Radial Basis Function
+Finite-Difference (RBFFD). RBF methods express functions as a linear
+combination of radial basis functions on an arbitrary set of
+nodes. The Finite-Difference component expresses this combination over
+a local set of nodes neighboring the point where the derivative is
+sought.  The derivative at all points takes the form of a sparse
+matrix/vector multiplication (spmv).
 
-In this paper, we consider the case of local stencils with the number of nodes at each point and encode the
-sparse matrix in ELLPACK format. We increase the number of operations relative to memory bandwidth by
-calculating four derivatives of four different functions, or 16 different derivatives. We demonstrate
-a novel implementation on the MIC architecture, taking into account its advanced swizzling and channel
-interchange features. We present benchmarks that show an almost order of magnitude increase in speed
-compared to efficient implementations of a single derivative. We explain the results through consideration
-of operation count versus memory bandwidth.
+In this paper, we consider the case of local stencils with the number
+of nodes at each point and encode the sparse matrix in ELLPACK
+format. We increase the number of operations relative to memory
+bandwidth by calculating four derivatives of four different functions,
+or 16 different derivatives. We demonstrate a novel implementation on
+the MIC architecture, taking into account its advanced swizzling and
+channel interchange features. We present benchmarks that show an
+almost order of magnitude increase in speed compared to efficient
+implementations of a single derivative. We explain the results through
+consideration of operation count versus memory bandwidth.
 \end{abstract}
@@
 N_F = \frac{8}{ b_x + 1}
-$$
-which is 8/5 and 8/9 in single and double precision, respectively. 
-In single and double precision, we get $N_F=32/36=8/5$ and $N_F=32/68=8/9$, respectively.
-The third column of the following table (TABLE ???) is the maximum achievable Gflop rate, assuming 
-a maximum memory transfer speed of $150 Gflops$ (see earlier benchmark in this paper). 
-(higher speeds can be achieved in practice,  upwards of 190 Gflops (Saule \etal\cite{}), but we use
-a benchmark code similar to our current algorithm, without the indirect acessing required for $x$, but keeping the same use of vector registers used in our fastest implementation.
-(DO WE INCLUDE the CODE?)
+$$ which is 8/5 and 8/9 in single and double precision, respectively.
+In single and double precision, we get $N_F=32/36=8/5$ and
+$N_F=32/68=8/9$, respectively.  The third column of the following
+table (TABLE ???) is the maximum achievable Gflop rate, assuming a
+maximum memory transfer speed of $150 Gflops$ (see earlier benchmark
+in this paper).  (higher speeds can be achieved in practice, upwards
+of 190 Gflops (Saule \etal\cite{}), but we use a benchmark code
+similar to our current algorithm, without the indirect acessing
+required for $x$, but keeping the same use of vector registers used in
+our fastest implementation. (DO WE INCLUDE the CODE?)
 
@@
 \end{center}
-(With a 190 Gflop peak memory rate, maximum performance is 47.5 and 170 Mflops for the 1/1 case, while we calculate 32 and 90 Gflops in the 4/4/ case.
+(With a 190 Gflop peak memory rate, maximum performance is 47.5 and
+170 Mflops for the 1/1 case, while we calculate 32 and 90 Gflops in
+the 4/4/ case.
 
-Let us also calculate the flop to byte ratio for the case of 1/4 and 4/1 ($n_v$/$n_m$). One notices the symmetry with respect to $n_v$ and $n_m$. 
-One gets $N_F = \frac{2*4}{5 b_x + 4}$ equal to 8/24 (=.33) and 8/44 (=.18), in single and double precision, respectively, which corresponds to 49 and 27 Gflops if a maximum of 150 Gflops is assumed.
-As $n_m$ and $n_v$ keep increasing, $N_F$ tends towards $2/b_x$, equal to $0.5$ and $0.25$ respectively for single and double precision, which leads to peak speed of 75 and 37.5 Gflops.
+Let us also calculate the flop to byte ratio for the case of 1/4 and
+4/1 ($n_v$/$n_m$). One notices the symmetry with respect to $n_v$ and
+$n_m$.  One gets $N_F = \frac{2*4}{5 b_x + 4}$ equal to 8/24 (=.33)
+and 8/44 (=.18), in single and double precision, respectively, which
+corresponds to 49 and 27 Gflops if a maximum of 150 Gflops is assumed.
+As $n_m$ and $n_v$ keep increasing, $N_F$ tends towards $2/b_x$, equal
+to $0.5$ and $0.25$ respectively for single and double precision,
+which leads to peak speed of 75 and 37.5 Gflops.
 
-In single precision, our best benchmarks achieve 20 Gflops for single precisions for on matrix/one vector, and 100 Glops for the 4/4 matrix/vector case, or 53\% and 75\% of peak performance, respectivey.
-We conclude that we achieve a higher percentage of peak performance for this algorithm when calculating 16 output $y$ vectors.
-The results are even more impressive when measured against the peak perforance of 2Tflops of the MIC in double precision.
+In single precision, our best benchmarks achieve 20 Gflops for single
+precisions for on matrix/one vector, and 100 Glops for the 4/4
+matrix/vector case, or 53\% and 75\% of peak performance,
+respectivey. We conclude that we achieve a higher percentage of peak
+performance for this algorithm when calculating 16 output $y$
+vectors. The results are even more impressive when measured against
+the peak perforance of 2Tflops of the MIC in double precision.
 
-\newfig{figures/gflops_peak.png}{2.5in}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors.
-We used a memory bandwidth of 150 Gbytes/sec (based on measurements in a code similar to the SpMV kernel, with computations removed.)}{fig:gflops_peak_perf}
+\newfig{figures/gflops_peak.pdf}{2.5in}{Peak performance of SpMV
+  algorithm assuming either 1 or 4 matrices and/or vectors. We used a
+  memory bandwidth of 150 Gbytes/sec (based on measurements in a code
+  similar to the SpMV kernel, with computations
+  removed.)}{fig:gflops_peak_perf}
 
-\newfig{figures/speedup_wrt_base.png}{\wide}{Speedup relative to base case using one matrix and one vector.}{fig:speedup}
+\newfig{figures/speedup_wrt_base.pdf}{\wide}{Speedup relative to base
+  case using one matrix and one vector.}{fig:speedup}
 
@@
 %
-In the vector instructions, \ttt{ps} refers to a float (4 bytes), while \ttt{epi32} refers to a 4-byte integer.
-Although we do not discuss double precision in this paper, \ttt{ds} refers to an 8 byte double precision real number (i.e., \loadds).
-Each vector register is broken up into four channels of 128 bits each.
-It is possible to interchange these channels at a cost of a "few" \qes{exact numbers?} cycles, and it is also possible to execute swizzle operations within a channel.
-For example, if the 128 bytes of each channel are labelled as $ABCD$, the vector instruction
-\ttt{\_mm512\_swizzle\_ps(v, \_MM\_SWIZ\_REG\_AAAA)}
-replaces each channel by $AAAA$.
+In the vector instructions, \ttt{ps} refers to a float (4 bytes),
+while \ttt{epi32} refers to a 4-byte integer. Although we do not
+discuss double precision in this paper, \ttt{ds} refers to an 8 byte
+double precision real number (i.e., \loadds).  Each vector register is
+broken up into four channels of 128 bits each. It is possible to
+interchange these channels at a cost of a "few" \qes{exact numbers?}
+cycles, and it is also possible to execute swizzle operations within a
+channel. For example, if the 128 bytes of each channel are labelled as
+$ABCD$, the vector instruction \ttt{\_mm512\_swizzle\_ps(v,
+  \_MM\_SWIZ\_REG\_AAAA)} replaces each channel by $AAAA$.
 
@@
 
-\newfig{figures/swizzling.pdf}{\wide}{Duplication of the first four bytes of each channel across the entire channel.}{fig:swizzling}
-\newfig{figures/channel_permutation.pdf}{\wide}{Channel Permutations on the MIC}{fig:permutation}
-\newfig{figures/tensor_product.pdf}{\wide}{Tensor scalar product using channels and swizzling.}{fig:tensor_product}
+\newfig{figures/swizzling.pdf}{\wide}{Duplication of the first four
+  bytes of each channel across the entire channel.}{fig:swizzling}
+
+\newfig{figures/channel_permutation.pdf}{\wide}{Channel Permutations
+  on the MIC}{fig:permutation}
+
+\newfig{figures/tensor_product.pdf}{\wide}{Tensor scalar product using
+  channels and swizzling.}{fig:tensor_product}
 
@@
 
-\newfig{figures/test1_readwrite.png}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
-Entries with "cpp" denote cases where coding was performed without MIC vector instructions.}{fig:band_rw}
+\newfig{figures/test1_readwrite.png}{\wide}{Bandwidth performance
+  under idealized conditions as a function of matrix row size. Entries
+  with "cpp" denote cases where coding was performed without MIC
+  vector instructions.}{fig:band_rw}
 
-\newfig{figures/test1_gather.png}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
-Entries with "cpp" denote cases where coding was performed without MIC vector instructions.}{fig:band_gather}
+\newfig{figures/test1_gather.png}{\wide}{Bandwidth performance under
+  idealized conditions as a function of matrix row size. Entries with
+  "cpp" denote cases where coding was performed without MIC vector
+  instructions.}{fig:band_gather}
 
-\newfig{figures/test3_gather.png}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
-Entries with "cpp" denote cases where coding was performed without MIC vector instructions.
-The greater speed of the cpp version is obtained throught the use of \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}.
-All memory is aligned on 64 bytes.}{fig:band_gather_ivdep}
+\newfig{figures/test3_gather.png}{\wide}{Bandwidth performance under
+  idealized conditions as a function of matrix row size. Entries with
+  "cpp" denote cases where coding was performed without MIC vector
+  instructions. The greater speed of the cpp version is obtained
+  throught the use of \ttt{\#Ivdep} {\em and\/}
+  \ttt{\_\_assumed\_aligned}. All memory is aligned on 64
+  bytes.}{fig:band_gather_ivdep}
 
@@
 
-\newfig{figures/host_test1_readwrite_no_temporal_hint.png}{\wide}{Bandwidth performance on the host under idealized conditions as a function of matrix row size.
-Entries with "cpp" denote cases where coding was performed without AVX vector instructions.
-The speed on the CPU matches the speed with AVX instructions.
-All memory is aligned on 32 bytes.}{fig:read_write}
+\newfig{figures/host_test1_readwrite_no_temporal_hint.png}{\wide}{Bandwidth
+  performance on the host under idealized conditions as a function of
+  matrix row size. Entries with "cpp" denote cases where coding was
+  performed without AVX vector instructions. The speed on the CPU
+  matches the speed with AVX instructions. All memory is aligned on 32
+  bytes.}{fig:read_write}
 
-\newfig{figures/mic_performance_nb_threads.png}{\wide}{Performance of $y=Ax$ on the MIC.
-Squares: base 1/1 case, solid circles: 4/4 case implemented in C++, solid triangles: 4/4 case implemented with MIC vector instructions.
-$-O3$ compilation options.
-Each group consists of four cases: grids of $64^3$ and $96^3$ with and without Reverse Cuthill McKee.}{fig:mic_performance}
+\newfig{figures/mic_performance_nb_threads.png}{\wide}{Performance of
+  $y=Ax$ on the MIC. Squares: base 1/1 case, solid circles: 4/4 case
+  implemented in C++, solid triangles: 4/4 case implemented with MIC
+  vector instructions. $-O3$ compilation options. Each group consists
+  of four cases: grids of $64^3$ and $96^3$ with and without Reverse
+  Cuthill McKee.}{fig:mic_performance}
 
@@
 % WHY the very high performance with very few nodes? STRANGE. 
-\newfig{figures/host_performance.png}{\wide}{Performance of $y=Ax$ on the Host.
-Squares: base 1/1 case, solid circles: 4/4 case implemented in C++, solid triangles: 4/4 case implemented with MIC vector instructions.
-$-O3$ compilation options.
-The four colors distinguish grid resolution and whether or not
-Reverse Cuthill McKee is applied.}{fig:host_performance}
+\newfig{figures/host_performance.png}{\wide}{Performance of $y=Ax$ on
+  the Host. Squares: base 1/1 case, solid circles: 4/4 case
+  implemented in C++, solid triangles: 4/4 case implemented with MIC
+  vector instructions. $-O3$ compilation options. The four colors
+  distinguish grid resolution and whether or not Reverse Cuthill McKee
+  is applied.}{fig:host_performance}
+
+\begin{figure*}[t]
+  \centering
+  \subfigure[a]{\includegraphics[width=.49\linewidth]{figures/supercompact_max_perf.pdf}}
+  \subfigure[b]{\includegraphics[width=.49\linewidth]{figures/compact_max_perf.pdf}}
+
+  \subfigure[c]{\includegraphics[width=.49\linewidth]{figures/random_max_perf.pdf}}
+  \subfigure[d]{\includegraphics[width=.49\linewidth]{figures/rbf_max_perf_run1.pdf}}
+  \caption{perf accross instance types}
+\end{figure*}
 

### 2013-10-13 ieee_paper.tex
organizing figures

@@
 
-\newcommand{\todo}[1]{\color{red}\textbf{\hl{#1}}\color{black}\xspace}
+\newcommand{\todo}[1]{{\color{red}\textbf{\hl{#1}}\xspace}}
 
@@
 
-\newfig{figures/gflops_peak.pdf}{2.5in}{Peak performance of SpMV
-  algorithm assuming either 1 or 4 matrices and/or vectors. We used a
-  memory bandwidth of 150 Gbytes/sec (based on measurements in a code
-  similar to the SpMV kernel, with computations
-  removed.)}{fig:gflops_peak_perf}
+\begin{figure*}
 
-\newfig{figures/speedup_wrt_base.pdf}{\wide}{Speedup relative to base
-  case using one matrix and one vector.}{fig:speedup}
+  \centering 
+  \subfigure[Peak performance of SpMV algorithm assuming either 1 or 4
+    matrices and/or vectors. We used a memory bandwidth of 150
+    Gbytes/sec (based on measurements in a code similar to the SpMV
+    kernel, with computations
+    removed).]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops_peak_perf}}
+%
+  \subfigure[Speedup relative to base case using one matrix and one
+    vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}
+
+  \caption{\todo{to write}}
+
+\end{figure*}
 
@@
 
-\newfig{figures/test1_readwrite.png}{\wide}{Bandwidth performance
-  under idealized conditions as a function of matrix row size. Entries
-  with "cpp" denote cases where coding was performed without MIC
-  vector instructions.}{fig:band_rw}
+\begin{figure*}
 
-\newfig{figures/test1_gather.png}{\wide}{Bandwidth performance under
+  \subfigure[Bandwidth performance under idealized conditions as a
+    function of matrix row size. Entries with "cpp" denote cases where
+    coding was performed without MIC vector
+    instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_readwrite.png}\label{fig:band_rw}}
+%
+  \subfigure[Bandwidth performance under
   idealized conditions as a function of matrix row size. Entries with
   "cpp" denote cases where coding was performed without MIC vector
-  instructions.}{fig:band_gather}
+  instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_gather.png}\label{fig:band_gather}}
 
-\newfig{figures/test3_gather.png}{\wide}{Bandwidth performance under
-  idealized conditions as a function of matrix row size. Entries with
-  "cpp" denote cases where coding was performed without MIC vector
-  instructions. The greater speed of the cpp version is obtained
-  throught the use of \ttt{\#Ivdep} {\em and\/}
-  \ttt{\_\_assumed\_aligned}. All memory is aligned on 64
-  bytes.}{fig:band_gather_ivdep}
+  \subfigure[Bandwidth performance under idealized conditions as a
+    function of matrix row size. Entries with "cpp" denote cases where
+    coding was performed without MIC vector instructions. The greater
+    speed of the cpp version is obtained throught the use of
+    \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}. All memory
+    is aligned on 64
+    bytes.]{\includegraphics[width=.49\linewidth]{figures/test3_gather.png}\label{fig:band_gather_ivdep}}
+%
+  \subfigure[Bandwidth performance on the host under idealized
+    conditions as a function of matrix row size. Entries with "cpp"
+    denote cases where coding was performed without AVX vector
+    instructions. The speed on the CPU matches the speed with AVX
+    instructions. All memory is aligned on 32
+    bytes.]{\includegraphics[width=.49\linewidth]{figures/host_test1_readwrite_no_temporal_hint.png}\label{fig:read_write}}
+
+\caption{Bandwidth Analysis.\todo{not 100\% sure what this plots are.}
+  \todo{can we harmozie y-axis?} \todo{ylabel should read "Bandwidth (GB/s)".}
+  \todo{remove title.}}
+
+\end{figure*}
 
@@
 
-\newfig{figures/host_test1_readwrite_no_temporal_hint.png}{\wide}{Bandwidth
-  performance on the host under idealized conditions as a function of
-  matrix row size. Entries with "cpp" denote cases where coding was
-  performed without AVX vector instructions. The speed on the CPU
-  matches the speed with AVX instructions. All memory is aligned on 32
-  bytes.}{fig:read_write}
+\begin{figure*}
+  \centering
 
-\newfig{figures/mic_performance_nb_threads.png}{\wide}{Performance of
-  $y=Ax$ on the MIC. Squares: base 1/1 case, solid circles: 4/4 case
-  implemented in C++, solid triangles: 4/4 case implemented with MIC
-  vector instructions. $-O3$ compilation options. Each group consists
-  of four cases: grids of $64^3$ and $96^3$ with and without Reverse
-  Cuthill McKee.}{fig:mic_performance}
+  \subfigure[Performance of $y=Ax$ on the MIC. Squares: base 1/1 case,
+    solid circles: 4/4 case implemented in C++, solid triangles: 4/4
+    case implemented with MIC vector instructions. $-O3$ compilation
+    options. Each group consists of four cases: grids of $64^3$ and
+    $96^3$ with and without Reverse Cuthill
+    McKee.\todo{I don't think this is supercompact but RBF 3D.}]{\includegraphics[width=.49\linewidth]{figures/mic_performance_nb_threads.png}\label{}}
+%
+  \subfigure[performance on the host on a 64x64x64 matrix. Manual
+    vectoization brings significant improvement there as
+    well.]{\includegraphics[width=.49\linewidth]{figures/64x64x64.pdf}\label{}}
 
-% NOTES: why is result the same at 32 processors.
-I should find out the distribution of threads.
-16 threads means one thread on each of 16 cores (YES, I BELIEVE), or 2 threads on each of 8 cores.} Must run the experiment with different nodes orderings.
-%export OMP_SCHEDULE=guided,64    (I should redo experiment with Dynamic or static and compact for KMP_AFFINITY
-%export KMP_AFFINITY=compact
-% MUST PRINT OUT OMP_SCHEDULE and KMP_AFFINITY inside the program. 
-% WHY the very high performance with very few nodes? STRANGE. 
-\newfig{figures/host_performance.png}{\wide}{Performance of $y=Ax$ on
-  the Host. Squares: base 1/1 case, solid circles: 4/4 case
-  implemented in C++, solid triangles: 4/4 case implemented with MIC
-  vector instructions. $-O3$ compilation options. The four colors
-  distinguish grid resolution and whether or not Reverse Cuthill McKee
-  is applied.}{fig:host_performance}
+  \caption{Vectorization matters in practice. Differences in BW
+    translate in different performance.\todo{this two plots should ``correlate'' }}
+\end{figure*}
 
@@
   \centering
-  \subfigure[a]{\includegraphics[width=.49\linewidth]{figures/supercompact_max_perf.pdf}}
-  \subfigure[b]{\includegraphics[width=.49\linewidth]{figures/compact_max_perf.pdf}}
+  \subfigure[Supercompact\todo{is the title? this looks like super compact.}]{\includegraphics[width=.49\linewidth]{figures/supercompact_max_perf.pdf}}
+  \subfigure[Compact]{\includegraphics[width=.49\linewidth]{figures/compact_max_perf.pdf}}
 
-  \subfigure[c]{\includegraphics[width=.49\linewidth]{figures/random_max_perf.pdf}}
-  \subfigure[d]{\includegraphics[width=.49\linewidth]{figures/rbf_max_perf_run1.pdf}}
-  \caption{perf accross instance types}
+  \subfigure[Random]{\includegraphics[width=.49\linewidth]{figures/random_max_perf.pdf}}
+  \subfigure[RBF 3D]{\includegraphics[width=.49\linewidth]{figures/rbf_max_perf_run1.pdf}}
+  \caption{Performance of "best code" accross instance types. \todo{Why stop at 1.6M
+      rows. Is that because of memory limitation? I guess that's what
+      it is}\todo{can we remove the title?}\todo{can we harmonize y-axis?}}
 \end{figure*}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
+
+  \caption{Performance on a MIC of derivative computation for RBFFD
+    with 884,736 nodes distributed quasi-randomly in a cube. Single
+    derivative of a single function (base case, triangle) and four
+    derivatives of four functions (multi case, circles). No bandwidth
+    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
+\end{figure}
+
+Derivatives in an RBFFD formulation are expressed as a sparse
+matrix/vector multiplication (SpMV). Using the full 61 nodes of a MIC
+processor, and all four threads of each node, we achieve lowly 18
+Gflops when calculating a single vector of a single
+function. Bandwidth reduction has minimal effect since the cache is
+large enough to store both the derivative matrix and the solution
+vector. We increase the number of computations per byte transferred by
+calculating four different derivatives of four different
+functions. With bandwidth reduction, we speed the calculation by a
+factor of 8, and achieve upwards of 140 Gflops.
 
@@
 The conclusion goes here. this is more of the conclusion
+
+future works: gpu. more applications. more nnz/row.
 

### 2013-10-13 ieee_paper.tex
section 3 (partial)

@@
 %\usepackage{morefloats}
+
+\newcommand{\ceil}[1]{\left\lceil#1\right\rceil}
 
@@
 
-In this section, we introduce some variables that will prove useful to 
-describe our experiments and construct a working model to examine the pros and cons of various assumptions.
-While deceptively simple, the MIC architecture requires a careful consideration of the properties of cache, memory bandwidth, cores and their interaction, threads per core (1-4), parallelization and vectorization.
-We will also seek to perform experiments that measure the best possible performance of the SPMV with maximum and minimum cache thrashing \qes{same as cache misses?}, with and without floating point operations, and with and without memory transfer.
-At the onset, we expect the dominant cost to e the transfer of the vector $x$ from memory, because contiguous elements of $x$ are not accessed sequentially.
+We saw in the previous section that one can express the RBF problem as
+a multiplication of four matrices by four vectors. We present here an
+estimation of the variation on the flop intensity of the computation
+and its impact on the expected performance of the
+application. Relevant notations are given in Table~\ref{tab:not}.
 
-Relevant notation is contained in Table~\ref{tab:not}.
-
+\begin{figure}
   \begin{center}
+    \scalebox{.9}{
       \begin{tabular}{|c|l|}
@@
         \hline
-$b_i$ & number of bytes per int \\
-$b_x$ & number of bytes per float(4)/double(8) \\
+        $b_i$ & number of bytes per index \\
+        $b_x$ & number of bytes per value \\
         $n_z$ & number of nonzeros per row $A$ \\
-$n_r$ & number of rows of $A$ \\
-$n_c$ & total number of elements in $col_id$: $n_z n_r$ \\
-$b_w$ & matrix bandwidth \\
-$b_c$ & total L2 cache per core ($512k=2^{19}$ bytes) \\
-$b_T$ & theoretical minimum number of bytes transferred  \\
-$b_{wT}$ & theoretical minimum number of bytes written to memory  \\
-$b_{rT}$ & theoretical minimum number of bytes read from memory  \\
-$n_C$ & total number of cores used  \\
-$t_C$ & total number of threads used per core \\
-$\rho$ & average cache density  \\
-$n_v, n_m$ & number of vectors and matrices \\
-$n_F$ & total number of floating point operations  \\
+        $n_r$ & number of column/rows of $A$ \\
+        $n_c$ & total number of non-zero\\
+        $n_v$ & number of {\tt x} vectors \\
+        $n_m$ & number of matrices \\
+        $s_M$ & size of the matrice(s)\\
+        $s_x$ & size of the {\tt x} vector(s)\\
+        $s_y$ & size of the {\tt y} vector(s)\\
+        \hline
+        $cl$    & size of a cache line \\
+        $b_{wT}$ & number of bytes written to memory  \\
+        $b_{rT}$ & minimum number of bytes read from memory  \\
+        $b_T$   & minimum number of bytes transferred  \\
+        $B_{rT}$ & maximum number of bytes read from memory \\
+        $B_T$   & maximum number of bytes transferred  \\
+        $O$     & number of floating point operations \\
+        $I_b$   & maximal computational intensity\\
+        $I_w$   & minimum computational intensity\\        
         \hline
       \end{tabular}
+    }
   \end{center}
+  \caption{Notations}
+  \label{tab:not}
+\end{figure}
 
-- weights and vector elements are either all floats or all doubles.
+Each vector in the problem is of dimension $n_r$ and each dimension
+tasks $b_x$ bytes. There are $n_v$ {\tt x} vectors and $n_v n_m$ {\tt
+  y} vectors, which lead to the size of the {\tt x} and {\tt y} vectors:
+$$s_x = n_v b_x n_r$$ $$s_y = n_v n_m b_x n_r$$
 
-Let us first estimate the memory (in a serial implementation) required to compute $y = Ax$ where $y$ has $n_r$ rows and $n_m$ columns, $A$ has
-$n_c=n_r n_z$ nonzero elements, and $x$ is a vector of $n_r$ rows and 
-$n_v$ columns. In this paper, we assume that the adjacency matrix for the
-$n_m$ matrices $A$ is constant, but the values stored in the various $A$ differ.
-We are therefore executing $n_v n_m$ spmv operations.
-The objective of course, is to minimize wall-clock time on the MIC.
+The matrix is composed of $n_r$ rows and columns with $n_z$ non-zero
+per row leading to a total of $$n_c = n_r n_z$$ non-zero entries in
+the matrix. Each of these non-zero entry have one index of size $b_i$
+and $n_m$ values of size $b_x$. The matrices have a total size
+of $$s_M = n_c (b_i + b_x n_m) = n_r n_z (b_i + b_x n_m)$$
 
-The total number of bytes is 
-$$
-   b_T = n_r (n_z b_i + n_z b_x n_m + b_x n_v)
-$$
-The ratio of $b_T$ to the total cache  over all cores is given by
-$$
-   R = \frac{b_T}{b_c}
-$$
-Clearly, the memory $b_T$ is composed of the memory required by the vectors $x$, $y$, the nonzeros of $A$ and the elements of the matrix \ttt{col}, which stores the nonzero columns in each row, and is an integral component of the Ellpack compressed matrix format specification.
-While the Ellpack format is less general than CSR, it is the idea structure for the RBBF simulations we are intersted in, wherein the number of RBF nodes per stencil is constant throughout the 2D or 3D grid and in time.
+If we assume an algorithm where the rows are processed the one after
+the one, the amount of memory that need to be written is exactly the
+size of the {\tt y} vector. (This assumption removes the possibility of
+cache partitioning techniques.) $$b_{wT} = s_y = n_v n_m b_x n_r$$
 
-Let us consider separately the number of bytes read from and written to memory.
-Both the matrix $n_m$ matrices $A$ \qes{need an index on $A$ since there are more than one?} and \ttt{col\_id} are read from memory, as is $x$, whereas, the vector$y$ are stored to memory.
-In a later sections, we will discuss the practical performance of these operations and estimate the dominant contributions.
-\qes{We also wish to isolate the effects of read, write and compute operations as a function of the number of cores and threads to see how they influence the results and to estimate whether our results have the potential to scale similarly on systems with a higher number of cores.
-We "might" also compare are results against the best implementation using OpenCL, or do so in the paper.}
+The amount of data read from memory depends highly on how the
+algorithm execute, how the matrix is structured. But in the best case
+the whole matrix is read a single time and the whole {\tt x} vector is
+read a single time from the main memory. (Assuming the whole vector
+{\tt x} is used is reasonnable since only oddly shaped matrix will
+left a dimension of the vector unused.) $$b_{rT} = s_M + s_x = n_r n_z
+(b_i + b_x n_m) + n_v b_x n_r$$
+ $$b_T = b_{rT} + b_{wT} =  n_r n_z (b_i + b_x n_m) + n_v b_x n_r (1 + n_m)$$
 
-The total number of bytes read into memory is
-$$ 
-n_{rT} = n_r n_v b_x + n_r n_z b_i + n_r n_m n_z b_x
-$$
-while the total number of bytes written to memory is
-$$
-n_{wT} = n_r n_m n_v b_x
-$$
+Notice that there is no reason for a piece of the matrix to be read
+multiple times. But assuming that each element of the {\tt x} vector
+is read a single time is a strong assumption. If using a single core,
+it assumes that either the cache of the architecture can store all the
+{\tt x} vector or that the matrix is sufficiently well structured for
+causing no cache trashing. If using multiple core, this assumes that
+no element of the {\tt x} vectors will be used by multiple
+cores. \cite{Saule12} showed that there is very little cache trashing
+in practice, but it showed that having elements of the vectors used by
+multiple cores can have a significant impact on the performance
+(growing with $n_v$).
 
-We compute the floating point operations of the SpMV with general $n_v$ and $n_m$. Calling the total number of floating point operations $N_F$, we find
-$$
-  N_F = 2 n_r n_z n_v m_m
-$$
-Two cases must be distinguished in regards to $x$. In the first, in a worst case scenario, each element 
-of $x$ is transferred $n_z$ times from memory.
-In the best case, the case of infinite cache, each element is transferred only once.
-Thus, we provide best and worst case scenarios and will compare against matrices $A$
-in each of these extreme situations to evaluate the degree we approach idea performance. 
-Thus the number of flops to bytes transferred is $N_F / (n_{rT} + n_{wT})$, which when written out, 
-\begin{eqnarray}
-N_{Fw} &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_z n_v  + n_r n_z b_i + n_r n_m n_z b_x + n_r n_m n_v b_x} \nonb \\
-    &=& \frac{2 n_v n_m}{b_x ( n_v + n_m + n_m n_v n_z^{-1}) + b_i}  \nonb \\
-%N_F &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_v (n_v+n_m) + n_r n_z b_i + n_r n_m n_z b_x} \nonb \\
-    %&=& \frac{2 n_z n_v n_m}{b_x (n_v+n_m) + n_z (b_i + n_m b_x)} \nonb 
-\end{eqnarray}
-while in the best case, 
-\begin{eqnarray}
-N_{Fb} &=& \frac{2 n_r n_z n_v n_m}{b_x n_r n_v  + n_r n_z b_i + n_r n_m n_z b_x + n_r n_m n_v b_x} \nonb \\
-    &=& \frac{2 n_v n_m}{b_x ( n_v n_z^{-1} + n_m  + n_m n_v n_z^{-1}) + b_i } 
-\end{eqnarray}
-In the special case $n_v=n_m=1$, 
-$$
-N_{Fw} = \frac{2}{b_x (2+n_z^{-1}) + b_i}
-$$
-and 
-$$
-N_{Fb} = \frac{2}{b_x (1 + 2 n_z^{-1}) + b_i}
-$$
-Neglecting $2$ compared to $n_z$, we find that $N_F = 2/(b_i+b_x)$, which leads to 
-$N_F = 2/8=1/4$ and  $N_F=2/12=1/6$ for single and double precision, respectively. 
-In the general case, in the condition of $n_z$ large compared to 2, 
-$$
-N_F =  \frac{2 n_v n_m}{n_m b_x + b_i}
-$$ 
-The case are interested in is $n_m=n_v=4$, so that
-$$
-N_F = \frac{8}{ b_x + 1}
-$$ which is 8/5 and 8/9 in single and double precision, respectively.
-In single and double precision, we get $N_F=32/36=8/5$ and
-$N_F=32/68=8/9$, respectively.  The third column of the following
-table (TABLE ???) is the maximum achievable Gflop rate, assuming a
-maximum memory transfer speed of $150 Gflops$ (see earlier benchmark
-in this paper).  (higher speeds can be achieved in practice, upwards
-of 190 Gflops (Saule \etal\cite{}), but we use a benchmark code
-similar to our current algorithm, without the indirect acessing
-required for $x$, but keeping the same use of vector registers used in
-our fastest implementation. (DO WE INCLUDE the CODE?)
+On the other hand, in the worst case, every time the {\tt x} vector is
+accessed, the value needs to be transfered from memory again. So in
+total, there are as many transfers as the number of non zero in the
+matrix. Notice that most architectures do not allow to read the memory
+one byte at a time but always in a set of multiple bytes. On most
+cached architecture, that minimum size corresponds to the size $cl$ of
+a cacheline (typically 64 bytes). Notice that when there are multiple
+vectors, each non-zero element uses $n_v$ consecutive entries. The
+worst case number of bytes read and transfered are 
+$$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}} $$ 
+$$B_T = n_v n_m b_x n_r + n_r n_z \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)$$
 
-\begin{center}
-\begin{tabular}{|c|c|c|}
+In SpMV, each non-zero of the matrix requires two floating point
+operations: one for performing the multiplication and one for
+accumulating the result row-wise. Here we are dealing with $n_v n_m$
+simultaneous SpMVs and the number of floating point operations is
+$$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$$
+
+The computation intensity is the amount of computations performed per
+byte transfered. In the worst case and in the best case, we have
+$$I_b = \frac{O}{b_T} = \frac{2 n_v n_m n_z}{n_z (b_i + b_x n_m) + n_v b_x (1 + n_m)}$$
+$$I_w = \frac{O}{B_T} = \frac{2 n_v n_m n_z}{n_v n_m b_x + n_z \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)}$$
+
+In the special case $n_v=n_m=1$ in single precision
+$$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
+$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
+
+In the special case $n_v=n_m=4$ in single precision
+$$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
+$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
+
+\begin{table}
+  \centering
+  \begin{tabular}{|l|r|r|r|r|}
     \hline
-Precision & $n_m=n_v=1$ & $n_m=n_v=4$     \\
+                           & v1m1  & v1m4  & v4m1   & v4m4  \\
     \hline
-single    &  1/4  (37.5)     &   8/9 (133) \\
-double    &  1/6  (25.5)     &   8/17 (62) \\
+    Best Single Precision  & 36.36 & 58.18 & 133.33 & 213.33\\
+    Worst Single Precision &  4.15 & 14.20 &  16.55 &  55.81\\
+    Best Double Precision  & 24.00 & 32.21 &  85.71 & 117.07\\
+    Worst Double Precision &  3.93 & 11.88 &  15.58 &  46.15\\
     \hline
   \end{tabular}
-\end{center}
-(With a 190 Gflop peak memory rate, maximum performance is 47.5 and
-170 Mflops for the 1/1 case, while we calculate 32 and 90 Gflops in
-the 4/4/ case.
 
-Let us also calculate the flop to byte ratio for the case of 1/4 and
-4/1 ($n_v$/$n_m$). One notices the symmetry with respect to $n_v$ and
-$n_m$.  One gets $N_F = \frac{2*4}{5 b_x + 4}$ equal to 8/24 (=.33)
-and 8/44 (=.18), in single and double precision, respectively, which
-corresponds to 49 and 27 Gflops if a maximum of 150 Gflops is assumed.
-As $n_m$ and $n_v$ keep increasing, $N_F$ tends towards $2/b_x$, equal
-to $0.5$ and $0.25$ respectively for single and double precision,
-which leads to peak speed of 75 and 37.5 Gflops.
-
-In single precision, our best benchmarks achieve 20 Gflops for single
-precisions for on matrix/one vector, and 100 Glops for the 4/4
-matrix/vector case, or 53\% and 75\% of peak performance,
-respectivey. We conclude that we achieve a higher percentage of peak
-performance for this algorithm when calculating 16 output $y$
-vectors. The results are even more impressive when measured against
-the peak perforance of 2Tflops of the MIC in double precision.
+  \caption{Estimation of performance in GFlop/s at 150GB/s for a matrix of $n_z = 32$}
+\end{table}
 
@@
 \def\castpstoi{\ttt{\_mm512\_castps\_si512\_ps}}
+
+\begin{figure*}
   %
@@
   %
+  \caption{List of MIC vectorial instructions used}
+\end{figure*}
 In the vector instructions, \ttt{ps} refers to a float (4 bytes),

### 2013-10-14 ieee_paper.tex
comments

@@
   \end{tabular}
-
-  \caption{Estimation of performance in GFlop/s at 150GB/s for a matrix of $n_z = 32$}
+  \caption{Estimation of performance in GFlop/s at 150GB/s for a
+    matrix of $n_z = 32$}
 \end{table}
@@
 \begin{figure*}
-
   \centering 
-  \subfigure[Peak performance of SpMV algorithm assuming either 1 or 4
-    matrices and/or vectors. We used a memory bandwidth of 150
-    Gbytes/sec (based on measurements in a code similar to the SpMV
-    kernel, with computations
-    removed).]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops_peak_perf}}
+  \subfigure[Absolute performance ]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops_peak_perf}}
 %
-  \subfigure[Speedup relative to base case using one matrix and one
+  \subfigure[Relative performance to base case using one matrix and one
     vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}
 
-  \caption{\todo{to write}}
+  \caption{Estimation of the reachable performance one can
+      achieve using a device with a memory to computational units of
+      150GB/s when varying the number of vectors and matrices}
+  \label{fig:perf_predict}
+\end{figure*}
 
-\end{figure*}
+\todo{conclude on the evaluation using the figure/table. Note that we
+  base prediction on 150GB/s which is not the peak of saule12. Note
+  that if you are in best in v1m1 you might go to bad v4m4, but if you
+  are in bad v1m1, you will essentially stay there.}
 

### 2013-10-14 ieee_paper.tex
Added that the size was in bytes in the tables. ` I checked Eric's formulas. They are correct. Changed some english.

@@
         $n_m$ & number of matrices \\
-        $s_M$ & size of the matrice(s)\\
-        $s_x$ & size of the {\tt x} vector(s)\\
-        $s_y$ & size of the {\tt y} vector(s)\\
+        $s_M$ & size of the matrice(s) in bytes\\
+        $s_x$ & size of the {\tt x} vector(s) in bytes\\
+        $s_y$ & size of the {\tt y} vector(s) in bytes\\
         \hline
-        $cl$    & size of a cache line \\
+        $cl$    & size of a cache line in bytes\\
         $b_{wT}$ & number of bytes written to memory  \\
@@
 
-The matrix is composed of $n_r$ rows and columns with $n_z$ non-zero
+The matrix is composed of $n_r$ rows and columns with $n_z$ non-zeros
 per row leading to a total of $$n_c = n_r n_z$$ non-zero entries in
-the matrix. Each of these non-zero entry have one index of size $b_i$
+the matrix. Each of these non-zero entries has one index of size $b_i$
 and $n_m$ values of size $b_x$. The matrices have a total size
@@
 
-If we assume an algorithm where the rows are processed the one after
-the one, the amount of memory that need to be written is exactly the
+If we assume an algorithm where the rows are processed one after
+the other, the amount of memory written is precisely the
 size of the {\tt y} vector. (This assumption removes the possibility of
@@
 
-The amount of data read from memory depends highly on how the
-algorithm execute, how the matrix is structured. But in the best case
-the whole matrix is read a single time and the whole {\tt x} vector is
-read a single time from the main memory. (Assuming the whole vector
-{\tt x} is used is reasonnable since only oddly shaped matrix will
-left a dimension of the vector unused.) $$b_{rT} = s_M + s_x = n_r n_z
+The amount of data read from memory depends highly on both the algorithm's
+execution path, and on %algorithm execute, 
+how the matrix is structured. But in the best case
+both the matrix $A$ %is read a single time 
+and the source vector {\tt x} are read once
+%read a single time 
+from the main memory. (We assume that all the elements of $\tt x$ are involved
+in the SpMV.)  %is used is reasonnable since only oddly shaped matrix will
+%left a dimension of the vector unused.) 
+Thus, $$b_{rT} = s_M + s_x = n_r n_z
 (b_i + b_x n_m) + n_v b_x n_r$$
@@
 is read a single time is a strong assumption. If using a single core,
-it assumes that either the cache of the architecture can store all the
-{\tt x} vector or that the matrix is sufficiently well structured for
+it assumes that either the cache of the architecture can store the full
+{\tt x} vector or that the matrix is sufficiently well structured 
 causing no cache trashing. If using multiple core, this assumes that
@@
 accessed, the value needs to be transfered from memory again. So in
-total, there are as many transfers as the number of non zero in the
-matrix. Notice that most architectures do not allow to read the memory
-one byte at a time but always in a set of multiple bytes. On most
-cached architecture, that minimum size corresponds to the size $cl$ of
-a cacheline (typically 64 bytes). Notice that when there are multiple
+total, there are as many transfers as the number of non-zeros in the
+matrix. Note however that most architectures cannot read memory a single 
+byte at a time. Instead, a minimum number of bytes,equal to the size o a cacheline, 
+are transfered at once. 
+%multiple bytes, with each chunk equal to the data do not allow to read the memory
+%one byte at a time but always in a set of multiple bytes. On most
+%cached architecture, that 
+%the minimum size corresponds to the size $cl$ of
+%a cacheline (typically 64 bytes). 
+When there are multiple
 vectors, each non-zero element uses $n_v$ consecutive entries. The
@@
 $$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
-$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
+%$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
+$$I_w = \frac{n_z }{2  + (4+cl/2)  n_z} \approx \frac{1}{4+cl/2}$$
 
@@
 $$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
-$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
+%$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
+$$I_w = \frac{32 n_z }{64 + (20+cl) n_z} \approx \frac{32}{84} = \frac{8}{21} $$
+The above formulas have been applied to MIC, which has a cache size $cl=64$ bytes.
 

### 2013-10-14 chart1.py

@@ -16,7 +16,19 @@ def ratio(nv, nm, bx, nz):
 # A : nm*nz*bx   loads
 # col_id : nz*bi loads
     bi = 4
-    rworst = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
+    #rworst = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
+    # in worst case, every vector element brings in an entire cache line (256 bits = 32 floats),
+    # so nv=4 vectors (1/8th cache line), requires a transfer of 32 bytes = 8 floats
+    # 16 vectors = 2 cache lines (, so the formula becomes
+    if nv == 1:
+        rworst = (2.*nv*nm)/(bx*(16.+nm+nm*nv/float(nz))+bi)
+    elif nv == 4:
+        rworst = (2.*nv*nm)/(bx*(16.*nv+nm+nm*nv/float(nz))+bi)
+    elif nv == 16:
+        rworst= (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
+    else:
+        print "nv case not considered"
+
     rbest = (2.*nv*nm)/(bx*(nm+(1+nm)*nv/float(nz))+bi)
     return([rworst,rbest])
 #----------------------------------
@@ -29,16 +41,19 @@ bandwidths = [150,190]
 nv = [1,4]
 nm = [1,4]
 labels = [(v+m) for v in ["v1","v4"] for m in ["m1","m4"]]
+labels.extend(["v16m1"])
+print labels
 # v loop is slowest
 # m loop is fastest
-ratios_s =  np.array([ratio(v,m,single, nz) for v in nv for m in nm])
+ratios_s = np.array([ratio(v,m,single, nz) for v in nv for m in nm])
+ratios_s = np.reshape(np.append(ratios_s, ratio(16,1,single, nz)),[5,2])
 ratios_s_ulti = ratios_s*ulti_bandwidth
 ratios_s_max  = ratios_s*max_bandwidth
 
 ratios_d =  np.array([ratio(v,m,double, nz) for v in nv for m in nm])
+ratios_d = np.reshape(np.append(ratios_d, ratio(16,1,double, nz)),[5,2])
 ratios_d_ulti = ratios_d*ulti_bandwidth
 ratios_d_max  = ratios_d*max_bandwidth
-
 
 ratios_s_worst = np.transpose(ratios_s_max)[0]
 ratios_s_best = np.transpose(ratios_s_max)[1]
[…]

### 2013-10-14 gflops_speedup.py
Corrected figures and python script to generate them.

@@ -23,7 +23,8 @@ def ratio(nv, nm, bx, nz):
     if nv == 1:
         rworst = (2.*nv*nm)/(bx*(16.+nm+nm*nv/float(nz))+bi)
     elif nv == 4:
-        rworst = (2.*nv*nm)/(bx*(16.*nv+nm+nm*nv/float(nz))+bi)
+        rworst = (2.*nv*nm)/(bx*(16.+nm+nm*nv/float(nz))+bi)
+        #rworst = (2.*nv*nm)/(bx*(16.*nv+nm+nm*nv/float(nz))+bi) # error
     elif nv == 16:
         rworst= (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
     else:

### 2013-10-14 ieee_paper.tex
updated plots.
Updated papers to refer to pdf plots and modifed pdf legend positions.

@@
 \def\es#1{}
-\def\ge#1{}
+%\def\ge#1{}
 %\usepackage{morefloats}
@@
     coding was performed without MIC vector
-    instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_readwrite.png}\label{fig:band_rw}}
+    instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_readwrite.pdf}\label{fig:band_rw}}
 %
@@
   "cpp" denote cases where coding was performed without MIC vector
-  instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_gather.png}\label{fig:band_gather}}
+  instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_gather.pdf}\label{fig:band_gather}}
 
@@
     is aligned on 64
-    bytes.]{\includegraphics[width=.49\linewidth]{figures/test3_gather.png}\label{fig:band_gather_ivdep}}
+    bytes.]{\includegraphics[width=.49\linewidth]{figures/test3_gather.pdf}\label{fig:band_gather_ivdep}}
 %
@@
     instructions. All memory is aligned on 32
-    bytes.]{\includegraphics[width=.49\linewidth]{figures/host_test1_readwrite_no_temporal_hint.png}\label{fig:read_write}}
+    bytes.]{\includegraphics[width=.49\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:read_write}}
 
@@
     $96^3$ with and without Reverse Cuthill
-    McKee.\todo{I don't think this is supercompact but RBF 3D.}]{\includegraphics[width=.49\linewidth]{figures/mic_performance_nb_threads.png}\label{}}
+    McKee.\todo{I don't think this is supercompact but RBF 3D.}]{\includegraphics[width=.49\linewidth]{figures/mic_performance_nb_threads.pdf}\label{}}
 %

### 2013-10-14 ieee_paper.tex
Added conclusion.

@@
 
-\section{Conclusions and Perspectives}
+\section{Conclusion}
 \label{sec:ccl}
+In the previous sections, we have explored in the practical implementation on a MIC architecture of multiple derivative operators acting on multiple vectors within the context of RBF-RD.
+Each derivative has an associated sparse matrix with a fixed number
+of non-zeros on each row.
+While computing a single derivative of multiple functions is rather common (vectorization occurs over the vectors), we accelerate the algorithm further by considering multiple derivatives with corresponding
+matrices with identical adjacency graphs. We specialize to four matrices and four derivatives, with 16 outputs, 
+computed as a sum of outer products.
 
-The conclusion goes here. this is more of the conclusion
+Using a matrix with $64^3$ (or $96^3$ rows, we achieve a speedup of xxx relative to the base case of one derivative and one function (1/1/ case), releative to a potential speedup of 16.
+We have achieved a speed 100 Gflops on 60 nodes using 240 threads, or a speed of 61\% of maximum possible performance.
+The maximum possible speed for the 1/1 case is around 40 Gflops, of which we achieve 20 Gflops.
+On matrices that exceed cache, a reverse Cuthill-McKee is applied
+to reduce the bandwidth, with an associated speedup of 30\% up to 135 Gflops. Our optimal implementation makes 
+use of the AVX instruction set, which makes use of swizzling and channel swapping operations for an extremely 
+efficient tensor product implementation. A straightforward implementation without AVX run at around 35 Gflops, 
+or 16\% of peak performance. RCM has no effect because the number of floating point operations is excessive (the 
+algorithm does not fill the vector registers to capacity).
 
-future works: gpu. more applications. more nnz/row.
+In a future work, we will examine the effect of larger stencil sizes, double precision, and the actual cost
+of computing the derivatives within a fluid simulation using RBF-FD.
 
 \section*{Acknowledgment}
-
-The authors would like to thank...
-more thanks here
+The first and last authors acknowledge NSF funding under NSF grant DMS-\#0934331 (FSU).
 

### 2013-10-14 ieee_paper.tex
4.a

--- 2013-10-14 ieee_paper.tex
@@
 
-In this work, we use a pre-release KNC card SE10P. The card has 8
-memory controllers where each of them can execute 5.5 billion
-transactions per second and has two 32-bit channels. That is the
-architecture can achieve a total bandwidth of 352GB/s aggregated
-across all the memory controllers. There are 61 cores clocked at
-1.05GHz. The cores’ memory interface are 32-bit wide with two channels
-and the total bandwidth is 8.4GB/s per core. Thus, the cores should be
-able to consume 512.4GB/s at most. However, the bandwidth between
-the cores and the memory controllers is limited by the ring network
-which connects them and theoretically supports at most 220GB/s.
+In this work, we use an Intel Xeon Phi 5110P coprocessor. This card
+has 8 memory controllers where each of them can execute 5 billion
+transactions per second and has two 32-bit channels, achieving a total
+bandwidth of 320GB/s aggregated across all the memory
+controllers. There are 60 cores clocked at 1.05GHz. Their memory
+interfaces are 32-bit wide with two channels and the total bandwidth
+is 8.4GB/s per core. Thus, the cores should be able to consume 504GB/s
+at most. However, the bandwidth between the cores and the memory
+controllers is limited by the ring network that connect the cores and
+the memory controller. Its precise bandwidth is unknown for this
+precise card, but it would be coehrent that it is between 200GB/s and
+250GB/s.
 
@@
 instruction cache, and a 512kB L2 cache. The architecture of a core is
-based on the Pentium architecture: though its design has been updated
-to 64-bit. A core can hold 4 hardware contexts at any time. And at
-each clock cycle, instructions from a single thread are executed. Due
-to the hardware constraints and to overlap latency, a core never
-executes two instructions from the same hardware context
-consecutively. In other words, if a program only uses one thread, half
-of the clock cycles are wasted. Since there are 4 hardware contexts
-available, the instructions from a single thread are executed
-in-order. As in the Pentium architecture, a core has two di↵erent
-concur- rent instruction pipelines (called U-pipe and V-pipe) which
-allow the execution of two instructions per cycle. However, some
-instructions are not available on both pipelines: only one vector or
-floating point instruction can be executed at each cycle, but two ALU
-instructions can be executed in the same cycle.
+based the traditional Pentium architecture which have been extended to
+64-bit. A core can hold 4 hardware contexts at any time. And at each
+clock cycle, instructions from a single thread are executed. Due to
+some hardware constraints two hardware context must be used to reach
+the peak instruction throughput in that architecture. As in the
+Pentium architecture, a core has two different concurrent instruction
+pipelines which allow the execution of two instructions per
+cycle. However, only one vector or floating point instruction can be
+executed at each cycle.
 
 Most of the performance of the architecture comes from the vector
-processing unit. Each of Intel Xeon Phi’s cores has 32 512-bit SIMD
-registers which can be used for double or single precision, that is,
-either as a vector of 8 64-bit values or as a vector of 16 32-bit
-values, respectively. The vector processing unit can perform many
-basic instructions, such as addition or division, and mathematical
-operations, such as sine and sqrt, allowing to reach 8 double
-precision operations per cycle (16 single precision). The unit also
-sup- ports Fused Multiply-Add (FMA) operations which are typically
-counted as two operations for benchmarking purposes. Therefore, the
-peak performance of the SE10P card is 1.0248 Tflop/s in double
-precision (2.0496 Tflop/s in single precision) and half without FMA.
+processing unit. Each core has 32 512-bit SIMD registers which can be
+used for double or single precision, that is, either as a vector of 8
+64-bit values or as a vector of 16 32-bit values, respectively. The
+vector processing unit can perform many basic instructions, such as
+addition or division, and mathematical operations, such as sine and
+sqrt, allowing to reach 8 double precision operations per cycle (16
+single precision). The unit also supports Fused Multiply-Add (FMA)
+operations which are typically counted as two operations for
+benchmarking purposes. Therefore, the peak performance of the 5110P
+card is 1 Tflop/s in double precision and 2 Tflop/s in single
+precision. If FMA can not be used and half of these number can be
+achieved.
 

### 2013-10-14 ieee_paper.tex
format

@@
 clock cycle, instructions from a single thread are executed. Due to
-some hardware constraints two hardware context must be used to reach
+some hardware constraints, two hardware contexts must be used to reach
 the peak instruction throughput in that architecture. As in the
@@
 achieved.
-
-\todo{rewrite that to make sure it fits the description of the card
-  used in the experiments. This text with copy/pasted from the PPAM
-  paper}
 
@@
 \label{sec:ccl}
-In the previous sections, we have explored in the practical implementation on a MIC architecture of multiple derivative operators acting on multiple vectors within the context of RBF-RD.
-Each derivative has an associated sparse matrix with a fixed number
-of non-zeros on each row.
-While computing a single derivative of multiple functions is rather common (vectorization occurs over the vectors), we accelerate the algorithm further by considering multiple derivatives with corresponding
-matrices with identical adjacency graphs. We specialize to four matrices and four derivatives, with 16 outputs, 
+
+In the previous sections, we have explored in the practical
+implementation on a MIC architecture of multiple derivative operators
+acting on multiple vectors within the context of RBF-RD. Each
+derivative has an associated sparse matrix with a fixed number of
+non-zeros on each row. While computing a single derivative of multiple
+functions is rather common (vectorization occurs over the vectors), we
+accelerate the algorithm further by considering multiple derivatives
+with corresponding matrices with identical adjacency graphs. We
+specialize to four matrices and four derivatives, with 16 outputs,
 computed as a sum of outer products.
 
-Using a matrix with $64^3$ (or $96^3$ rows, we achieve a speedup of xxx relative to the base case of one derivative and one function (1/1/ case), releative to a potential speedup of 16.
-We have achieved a speed 100 Gflops on 60 nodes using 240 threads, or a speed of 61\% of maximum possible performance.
-The maximum possible speed for the 1/1 case is around 40 Gflops, of which we achieve 20 Gflops.
-On matrices that exceed cache, a reverse Cuthill-McKee is applied
-to reduce the bandwidth, with an associated speedup of 30\% up to 135 Gflops. Our optimal implementation makes 
-use of the AVX instruction set, which makes use of swizzling and channel swapping operations for an extremely 
-efficient tensor product implementation. A straightforward implementation without AVX run at around 35 Gflops, 
-or 16\% of peak performance. RCM has no effect because the number of floating point operations is excessive (the 
-algorithm does not fill the vector registers to capacity).
+Using a matrix with $64^3$ (or $96^3$ rows, we achieve a speedup of
+xxx relative to the base case of one derivative and one function (1/1/
+case), releative to a potential speedup of 16. We have achieved a
+speed 100 Gflops on 60 nodes using 240 threads, or a speed of 61\% of
+maximum possible performance. The maximum possible speed for the 1/1
+case is around 40 Gflops, of which we achieve 20 Gflops. On matrices
+that exceed cache, a reverse Cuthill-McKee is applied to reduce the
+bandwidth, with an associated speedup of 30\% up to 135 Gflops. Our
+optimal implementation makes use of the AVX instruction set, which
+makes use of swizzling and channel swapping operations for an
+extremely efficient tensor product implementation. A straightforward
+implementation without AVX run at around 35 Gflops, or 16\% of peak
+performance. RCM has no effect because the number of floating point
+operations is excessive (the algorithm does not fill the vector
+registers to capacity).
 
-In a future work, we will examine the effect of larger stencil sizes, double precision, and the actual cost
-of computing the derivatives within a fluid simulation using RBF-FD.
+In a future work, we will examine the effect of larger stencil sizes,
+double precision, and the actual cost of computing the derivatives
+within a fluid simulation using RBF-FD.
 

### 2013-10-14 ieee_paper.tex
pass on ccl

@@
 Using a matrix with $64^3$ (or $96^3$ rows, we achieve a speedup of
-xxx relative to the base case of one derivative and one function (1/1/
-case), releative to a potential speedup of 16. We have achieved a
-speed 100 Gflops on 60 nodes using 240 threads, or a speed of 61\% of
+xxx relative to the base case of one derivative and one function (1/1
+case), relative to a potential speedup of 16. We have achieved a
+speed of 100 Gflop/s on 60 cores using 240 threads, or a speed of 61\% of
 maximum possible performance. The maximum possible speed for the 1/1
-case is around 40 Gflops, of which we achieve 20 Gflops. On matrices
-that exceed cache, a reverse Cuthill-McKee is applied to reduce the
-bandwidth, with an associated speedup of 30\% up to 135 Gflops. Our
-optimal implementation makes use of the AVX instruction set, which
+case is around 40 Gflop/s, of which we achieve 20 Gflop/s. On matrices
+that exceed the cache, a reverse Cuthill-McKee is applied to reduce the
+bandwidth, with an associated speedup of 30\% up to 135 Gflop/s. Our
+optimal implementation makes use of the AVX instruction set \todo{that's not AVX on MIC, not sure what the name is}, which
 makes use of swizzling and channel swapping operations for an
 extremely efficient tensor product implementation. A straightforward
-implementation without AVX run at around 35 Gflops, or 16\% of peak
+implementation without vectorization runs at around 35 Gflops, or 16\% of peak
 performance. RCM has no effect because the number of floating point

### 2013-10-14 ieee_paper.tex
updated instruction set in conclusion.

@@
 bandwidth, with an associated speedup of 30\% up to 135 Gflop/s. Our
-optimal implementation makes use of the AVX instruction set \todo{that's not AVX on MIC, not sure what the name is}, which
-makes use of swizzling and channel swapping operations for an
+optimal implementation makes use of the (as yet unreleased) IMCI MIC instruction set 
+%\todo{that's not AVX on MIC, not sure what the name is}, 
+which includes a number of swizzling and channel swapping operations, for an
 extremely efficient tensor product implementation. A straightforward

### 2013-10-14 ieee_paper.tex
temp

--- 2013-10-14 ieee_paper.tex
@@
 
+As explained in Section~XXX, we expect to achieve a performance of
+about 210Gflop/s in single precision in the best case. We will focus
+on the single precision case with 4 vectors and 4 matrices. But
+similar techniques apply for other combinations. This target
+performance represents 10\% of the peak performance of the
+architecture which is impossible without an efficient vectorization of
+the kernel. Indeed, that means a Fused Multiply-Add instruction on  
+fully loaded registers must be executed at most every 10 cycles.
+
 Registers on the MIC are 512 bits wide, and can store 16 floats or 8
@@
 to achieve vectorization. (In comparison to OpenCL or CUDA, one can
-think of these registers as a warp.)  We use the following vector
-operations in our code, which we'll explain in the text.
+think of these registers as a warp.)  We use the vector operations
+presented in Figure~\ref{fig:MIC-vect-instr} in our code, which we'll
+explain in the text.
 
@@
   %
-  \caption{List of MIC vectorial instructions used}
+  \caption{List of MIC vectorial instructions used.}
+\label{fig:MIC-vect-instr}
 \end{figure*}

### 2013-10-15 ieee_paper.tex
basic description of the swizzling

@@
 architecture which is impossible without an efficient vectorization of
-the kernel. Indeed, that means a Fused Multiply-Add instruction on  
-fully loaded registers must be executed at most every 10 cycles.
+the kernel. Indeed, to reach such a performance, a Fused Multiply-Add
+instruction on fully loaded registers must be executed at most every
+10 cycles.
 
-Registers on the MIC are 512 bits wide, and can store 16 floats or 8
-doubles. It is through these registers that the \#SIMD pragma is able
-to achieve vectorization. (In comparison to OpenCL or CUDA, one can
-think of these registers as a warp.)  We use the vector operations
-presented in Figure~\ref{fig:MIC-vect-instr} in our code, which we'll
-explain in the text.
+We discuss how the implementation of the v4m4 kernel works using
+vectorial instruction. (We will show that an implementation relying on
+compiler vectorization does not lead to desirable performance.) The
+registers in the MIC architecture are 512 bits wide and can store 16
+floats or integers. (In comparison to OpenCL or CUDA, one can think of
+these registers as a warp.) The end goal is to be able to load and
+format the non zero values $a,b,c,d$ and the vector entries $p,q,r,s$
+in two vectorial register to be able to apply Fused Multiply-Add on
+them, which is depicted in Figure~\ref{fig:tensor_product}. Bringing
+the data in the vector register as depicted in the figure is the
+difficult part that we now explain.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/tensor_product.pdf}
+  \caption{Structure of the vector registers to perform an efficient
+    tensor scalar product. See Figure~\ref{fig:mat_struct} for
+    reference.}
+  \label{fig:tensor_product}
+\end{figure}
+
+The core of the technique is the use of swizzling and permutation
+features of the vectors in the Intel MIC architecture. A vector of 512
+bits is composed of 4 lanes of 128 bits which are made of 4 elements
+of 32 bits. A swizzling operation allows to reorder the elements
+within each lane. While a permutation allow to change the orders of
+the lanes. Figure~\ref{fig:swizzling_permutation} illustrates these
+two possibilities. Both permutation and swizzling support common
+reordering and broadcast.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/swizzling.pdf}
+  \includegraphics[width=\linewidth]{figures/channel_permutation.pdf}
+
+  \caption{Swizzling enables to reorder or broadcast elements within
+    each 128 bit group. Permutation enables to reorder or broadcast
+    while 128 bit groups.}
+  \label{fig:swizzling_permutation}
+\end{figure}
+
+We use the vector operations presented in
+Figure~\ref{fig:MIC-vect-instr} in our code, which we'll explain in
+the text.
 
@@
 
-Here is a function that reads in four floats (a,b,c,d) and creates the 16-float vector dddd,cccc,bbbb,aaaa: 
+Figure~\ref{code:read_aaaa} is a function that reads in four floats
+(a,b,c,d) and creates the 16-float vector dddd,cccc,bbbb,aaaa:
+
+\begin{figure*}
 \begin{verbatim}
@@
 {
-    int int_mask_lo = (1 << 0) + (1 << 4) + (1 << 8) + (1 << 12);
-    __mmask16 mask_lo = _mm512_int2mask(int_mask_lo);
-    __m512 v1_old;
-    v1_old = _mm512_setzero_ps();
-    v1_old = _mm512_mask_loadunpacklo_ps(v1_old, mask_lo, a);
-    v1_old = _mm512_mask_loadunpackhi_ps(v1_old, mask_lo, a);
-    v1_old = _mm512_swizzle_ps(v1_old, _MM_SWIZ_REG_AAAA);
+    int int_mask = 0x1111;
+    __mmask16 mask = _mm512_int2mask(int_mask);
+    __m512 v1;
+    v1 = _mm512_setzero_ps();
+    v1 = _mm512_mask_loadunpacklo_ps(v1, mask, a);
+    v1 = _mm512_mask_loadunpackhi_ps(v1, mask, a);
+    v1 = _mm512_swizzle_ps(v1, _MM_SWIZ_REG_AAAA);
     return v1_old;
@@
 \end{verbatim}
+\caption{read\_aaaa code}
+\label{code:read_aaaa}
+\end{figure*}
 
-Here is a function that takes four floats from float register \ttt{v1}, and 
-places them in each of the four lanes. Because the permulation function only 
-operates on 4-byte integers, it is necessary to convert (in place) each float
-to an integer (the bit structure is not changed), permute the lanes and cast
-the integers back to floats (with no modification of the bit structure)
+Figure~\ref{code:permute} is a function that takes four floats from
+float register \ttt{v1}, and places them in each of the four
+lanes. Because the permulation function only operates on 4-byte
+integers, it is necessary to convert (in place) each float to an
+integer (the bit structure is not changed), permute the lanes and cast
+the integers back to floats (with no modification of the bit
+structure)
+
+\begin{figure*}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\newfig{figures/swizzling.pdf}{\wide}{Duplication of the first four
-  bytes of each channel across the entire channel.}{fig:swizzling}
-
-\newfig{figures/channel_permutation.pdf}{\wide}{Channel Permutations
-  on the MIC}{fig:permutation}
-
-\newfig{figures/tensor_product.pdf}{\wide}{Tensor scalar product using
-  channels and swizzling.}{fig:tensor_product}
+\caption{permute code}
+\label{code:permute}
+\end{figure*}
 

### 2013-10-15 ieee_paper.tex
typo
section 4 minor

@@
 the data in the vector register as depicted in the figure is the
-difficult part that we now explain.
+difficult part that we now explain. A thread will perform the
+multiplications one row at a time, and parallelism is achieved by
+giving blocks of rows to each core usign an OpenMP construct.
 
@@
 features of the vectors in the Intel MIC architecture. A vector of 512
-bits is composed of 4 lanes of 128 bits which are made of 4 elements
-of 32 bits. A swizzling operation allows to reorder the elements
-within each lane. While a permutation allow to change the orders of
-the lanes. Figure~\ref{fig:swizzling_permutation} illustrates these
-two possibilities. Both permutation and swizzling support common
-reordering and broadcast.
+bits is composed of 4 lanes (sometimes called channels) of 128 bits
+which are made of 4 elements of 32 bits. A swizzling operation allows
+to reorder the elements within each lane. While a permutation allow to
+change the orders of the lanes. Figure~\ref{fig:swizzling_permutation}
+illustrates these two possibilities. Both permutation and swizzling
+support common reordering and broadcast.
 
@@
     each 128 bit group. Permutation enables to reorder or broadcast
-    while 128 bit groups.}
+    whole 128 bit groups.}
   \label{fig:swizzling_permutation}
 \end{figure}
+
+All the operations that loads data from the memory into a register
+allows to load 512 bits, so 16 floats at a time. A single non zero in
+the matrix will cause 16 operations to be performed, but it only uses
+4 floats from the matrix and 4 floats from the {\tt x} vector. So we
+need to load at once the data for 4 non zeroes in the matrix, so that
+we load 16 flaots from the matrix and 16 floats from the vectors at
+once. (The number of non zeroes on a row in our application is always
+a multiple of 4, if it was, we would need to add some explicit zeroes
+in the data structure to pad it.)
 

### 2013-10-15 ieee_paper.tex
loading the matrix nnz

@@
 
+The values from the non zero element in the matrix are the simplest to
+load. One can load the 16 floating point values for the 4 non zero of
+the matrix at once. The 4 floats coming from the four matrices for
+each non zero are naturally grouped within each lane of the SIMD
+vector. One can use lane permutation to broad a lane to take the whole
+vector. So one 1 permutations and a quarter load are necessary to put
+the data of the matrix in the required format.
+
 We use the vector operations presented in

### 2013-10-15 ieee_paper.tex
reformating code

@@
 each non zero are naturally grouped within each lane of the SIMD
-vector. One can use lane permutation to broad a lane to take the whole
-vector. So one 1 permutations and a quarter load are necessary to put
-the data of the matrix in the required format.
+vector. One can use a swizzle operation to broadcast the 4 elements of
+a single matrix to take the whole vector.
+
+Loading the entries from the {\tt x} vector is a little bit more
+complicated.
 
@@
 \begin{verbatim}
-__m512 read_aaaa(float* a)                                                              
-{
+const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
+const __m512i four = _mm512_set4_epi32(4,4,4,4); 
 int int_mask = 0x1111;
 __mmask16 mask = _mm512_int2mask(int_mask);
-    __m512 v1;
-    v1 = _mm512_setzero_ps();
-    v1 = _mm512_mask_loadunpacklo_ps(v1, mask, a);
-    v1 = _mm512_mask_loadunpackhi_ps(v1, mask, a);
-    v1 = _mm512_swizzle_ps(v1, _MM_SWIZ_REG_AAAA);
-    return v1_old;
+
+#pragma omp parallel for
+for (int r=0; r < nb_rows; r++) {
+  __m512 accu = _mm512_setzero_ps();
+
+  for (int n=0; n < nz; n+=4) {
+    float* a = &dom.col_id_t[0]+n+nz*r;
+    __m512i vect_i = _mm512_setzero_ps();
+    vect_i = _mm512_mask_loadunpacklo_ps(vect_i, mask, a);
+    vect_i = _mm512_mask_loadunpackhi_ps(vect_i, mask, a);
+    vect_i = _mm512_swizzle_ps(vect_i, _MM_SWIZ_REG_AAAA);
+    vect_i = _mm512_fmadd_epi32(vect_i, four, offsets);
+
+    all_vect = _mm512_i32gather_ps(vect_i, dom.vec_vt, scale);
+
+    mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));
+
+    //perform first tensor multiplication
+    correct_vect = permute(all_vect, _MM_PERM_AAAA);
+    correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
+    accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);
+
+    //three more times with BBBB, CCCC, and DDDD
+  }
+  _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
 } 
 \end{verbatim}
-\caption{read\_aaaa code}
-\label{code:read_aaaa}
+\caption{mat mul code}
+\label{code:mat_mul}
 \end{figure*}
 
-Figure~\ref{code:permute} is a function that takes four floats from
-float register \ttt{v1}, and places them in each of the four
-lanes. Because the permulation function only operates on 4-byte
-integers, it is necessary to convert (in place) each float to an
-integer (the bit structure is not changed), permute the lanes and cast
-the integers back to floats (with no modification of the bit
-structure)
-
-\begin{figure*}
-\begin{verbatim}
-__m512 permute(__m512 v1, _MM_PERM_ENUM perm)                                           
-{
-    __m512i vi = _mm512_castps_si512(v1);
-    vi = _mm512_permute4f128_epi32(vi, perm);
-    v1 = _mm512_castsi512_ps(vi);
-    return v1;
-}
-\end{verbatim}
-\caption{permute code}
-\label{code:permute}
-\end{figure*}
+Note that the ``permute'' function is using the {\tt
+  \_mm512\_permute4f128\_epi32} instruction and therefore typecast are
+necessary to convert to a vector of int type.
 

### 2013-10-15 additional_figures.tex
Updated figures in text. See end of text.

\clearpage

\def\ww{.14\textwidth}
\begin{figure}
\begin{center}
    \subfigure[Supercompact.]{%
    \label{fig:supercompact}
	\includegraphics[width=\ww]{figures/supercompact_matrix-crop.pdf}}%
    \subfigure[Compact.]{%
    \label{fig:compact}
	\includegraphics[width=\ww]{figures/compact_matrix-crop.pdf}}%
    \subfigure[Random.]{%
    \label{fig:random}
	\includegraphics[width=\ww]{figures/random_matrix-crop.pdf}}%
    \end{center}
	\caption{Different sparsity distributions. In all cases, there are 32 nonzeros per line and 300 rows.}
\end{figure}

\def\ww{.20\textwidth}
\begin{figure}
\begin{center}
    \subfigure[2D, no RCM.]{%
    \label{fig:supercompact}
	\includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.pdf}}%
    \subfigure[2D, RCM.]{%
    \label{fig:compact}
	\includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.pdf}} \linebreak
    \subfigure[3D, RCM.]{%
    \label{fig:random} 
	\includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.pdf}}%
    \subfigure[3D, RCM.]{%
    \label{fig:random}
	\includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.pdf}}%
    \end{center}
	\caption{Matrix A that corresponds to derivative stencils in 2D (a-b) and 3D (c-d) 
      RBF-FD calculations.}
\end{figure}

### 2013-10-15 ieee_paper.tex
Updated figures in text. See end of text.
formatting
text of section 4

@@
 We discuss how the implementation of the v4m4 kernel works using
-vectorial instruction. (We will show that an implementation relying on
-compiler vectorization does not lead to desirable performance.) The
-registers in the MIC architecture are 512 bits wide and can store 16
-floats or integers. (In comparison to OpenCL or CUDA, one can think of
-these registers as a warp.) The end goal is to be able to load and
-format the non zero values $a,b,c,d$ and the vector entries $p,q,r,s$
-in two vectorial register to be able to apply Fused Multiply-Add on
-them, which is depicted in Figure~\ref{fig:tensor_product}. Bringing
-the data in the vector register as depicted in the figure is the
-difficult part that we now explain. A thread will perform the
-multiplications one row at a time, and parallelism is achieved by
-giving blocks of rows to each core usign an OpenMP construct.
+vectorial instruction. The code is given for reference in
+Figure~\ref{code:mat_mul}. (We will show that an implementation
+relying on compiler vectorization does not lead to desirable
+performance.) The registers in the MIC architecture are 512 bits wide
+and can store 16 floats or integers. (In comparison to OpenCL or CUDA,
+one can think of these registers as a warp.) The end goal is to be
+able to load and format the non zero values $a,b,c,d$ and the vector
+entries $p,q,r,s$ in two vectorial register to be able to apply Fused
+Multiply-Add on them, which is depicted in
+Figure~\ref{fig:tensor_product}. Bringing the data in the vector
+register as depicted in the figure is the difficult part that we now
+explain. A thread will perform the multiplications one row at a time,
+and parallelism is achieved by giving blocks of rows to each core
+usign an OpenMP construct.
 
@@
 illustrates these two possibilities. Both permutation and swizzling
-support common reordering and broadcast.
+support common reordering and broadcast. In the code, the ``permute''
+function needs to use the {\tt \_mm512\_permute4f128\_epi32}
+instruction which is only defined on integers and therefore typecast
+are necessary to convert to a vector to/from the integer type.
 
@@
 Loading the entries from the {\tt x} vector is a little bit more
-complicated.
-
-We use the vector operations presented in
-Figure~\ref{fig:MIC-vect-instr} in our code, which we'll explain in
-the text.
-
-\def\loadps{\ttt{\_mm512\_load\_ps}}
-\def\loadds{\ttt{\_mm512\_load\_ds}}
-\def\fmadps{\ttt{\_mm512\_fmad\_ps}}
-\def\gatherps{\ttt{\_mm512\_i32gather\_ps}}
-\def\swizzleps{\ttt{\_mm512\_swizzle\_ps}}
-\def\storenrngops{\ttt{\_mm512\_storenrngo\_ps}}
-\def\castsi{\ttt{\_mm512\_castsi512\_ps}}
-\def\permute{\ttt{\_mm512\_permute4f128\_epi32}}
-\def\intmask{\ttt{\_mm512\_int2mask}}
-\def\loadunpack{\ttt{\_mm512\_mask\_loadunpacklo\_epi32}}
-\def\castitops{\ttt{\_mm512\_castsi512\_ps}}
-\def\castpstoi{\ttt{\_mm512\_castps\_si512\_ps}}
-
-\begin{figure*}
-  %
-  \begin{center}
-    \begin{tabular}{|l|l|}
-      \hline
-      \loadps &  load 16 consecutive floats to register\\
-      \fmadps &  multiply/add of 16 floats\\
-      \gatherps &  gather 16, possibly disconnected floats from memory\\
-      \swizzleps &  reorder a channel, and duplicate across channels\\
-      \storenrngops &  store 16 floats to memory without register rewrite or reordering\\
-      \castpstoi & reinterpret 16 floats as integers\\
-      \castitops & reinterpret 16 integers as floats\\
-      \permute &  permute channels; do not change individual channels\\
-      \intmask &  \\
-      \loadunpack &  \\
-      \hline
-    \end{tabular}
-  \end{center}
-  %
-  \caption{List of MIC vectorial instructions used.}
-\label{fig:MIC-vect-instr}
-\end{figure*}
-In the vector instructions, \ttt{ps} refers to a float (4 bytes),
-while \ttt{epi32} refers to a 4-byte integer. Although we do not
-discuss double precision in this paper, \ttt{ds} refers to an 8 byte
-double precision real number (i.e., \loadds).  Each vector register is
-broken up into four channels of 128 bits each. It is possible to
-interchange these channels at a cost of a "few" \qes{exact numbers?}
-cycles, and it is also possible to execute swizzle operations within a
-channel. For example, if the 128 bytes of each channel are labelled as
-$ABCD$, the vector instruction \ttt{\_mm512\_swizzle\_ps(v,
-  \_MM\_SWIZ\_REG\_AAAA)} replaces each channel by $AAAA$.
-
-Figure~\ref{code:read_aaaa} is a function that reads in four floats
-(a,b,c,d) and creates the 16-float vector dddd,cccc,bbbb,aaaa:
+complicated. First the columns pointers of 4 different non zeroes are
+loaded in a vector register and the values are distributed one per
+lane of the vector register using an {\tt unpack}
+operation\footnote{Notice that there are two calls to unpack for the
+  LSB and MSB. Both are mandatory even according the documentation of
+  the hardware}. They replicated so that each fill a vector lane using
+a swizzle operation. Then each value is multiplied by 4 (because there
+are 4 vectors) and the offsets $(3,2,1,0)$ is added each lane to
+obtain the correct index of the elements of the {\tt x} vector within
+each lane. Then a {\tt gather} operation is performed bring the 16
+entries of the {\tt x} vector into the SIMD register. At this point we
+have a SIMD register where each lane is the four vector entries for a
+non zero element. Using a broadcast permutation, one can replicate an
+entry of the four vectors accross each lane.
 
@@
 \end{verbatim}
-\caption{mat mul code}
+\caption{mat mul code.}
 \label{code:mat_mul}
 \end{figure*}
-
-Note that the ``permute'' function is using the {\tt
-  \_mm512\_permute4f128\_epi32} instruction and therefore typecast are
-necessary to convert to a vector of int type.
 
@@
 
+\input{additional_figures}
+
 \section*{Acknowledgment}

### 2013-10-15 gflops_speedup.py
modified figures.

@@ -28,17 +28,21 @@ def ratio(nv, nm, bx, nz):
 # A : nm*nz*bx   loads
 # col_id : nz*bi loads
     bi = 4
+    cl = 64.
     #rworst = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
     # in worst case, every vector element brings in an entire cache line (256 bits = 32 floats),
     # so nv=4 vectors (1/8th cache line), requires a transfer of 32 bytes = 8 floats
     # 16 vectors = 2 cache lines (, so the formula becomes
     if nv == 1:
-        rworst = (2.*nv*nm)/(bx*(16.+nm+nm*nv/float(nz))+bi)
+        #rworst = (2.*nv*nm)/(bx*(16.+nm+nm*nv/float(nz))+bi)
+        rworst = (2.*nv*nm)/(cl*np.ceil((bx*nv)/cl)+bx*(nm+nm*nv/float(nz))+bi)
     elif nv == 4:
-        rworst = (2.*nv*nm)/(bx*(16.+nm+nm*nv/float(nz))+bi)
+        #rworst = (2.*nv*nm)/(bx*(16.+nm+nm*nv/float(nz))+bi)
+        rworst = (2.*nv*nm)/(cl*np.ceil((bx*nv)/cl)+bx*(nm+nm*nv/float(nz))+bi)
         #rworst = (2.*nv*nm)/(bx*(16.*nv+nm+nm*nv/float(nz))+bi) # error
     elif nv == 16:
-        rworst= (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
+        #rworst= (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
+        rworst = (2.*nv*nm)/(cl*np.ceil((bx*nv)/cl)+bx*(nm+nm*nv/float(nz))+bi)
     else:
         print "nv case not considered"
 

### 2013-10-15 tables_and_figures.tex
Added combined swizzling_lane figure.

@@
 
-\newfig{figures/gflops_peak.png}{2.5in}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors.
+\newfig{figures/gflops_peak.pdf}{2.5in}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors.
 We used a memory bandwidth of 150 Gbytes/sec (based on measurements in a code similar to the SpMV kernel, with computations removed.)}{fig:gflops_peak_perf}
 
-\newfig{figures/speedup_wrt_base.png}{\wide}{Speedup relative to base case using one matrix and one vector.}{fig:speedup}
+\newfig{figures/speedup_wrt_base.pdf}{\wide}{Speedup relative to base case using one matrix and one vector.}{fig:speedup}
 
-\newfig{figures/test1_readwrite.png}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
+\newfig{figures/test1_readwrite.pdf}{\wide}{Bandwidth performance under idealized conditions as a function of matrix row size.
 Entries with "cpp" denote cases where coding was performed without MIC vector instructions.}{fig:band_rw}

### 2013-10-15 ieee_paper.tex
Added combined swizzling_lane figure.
sply plots

@@
   \centering
-  \includegraphics[width=\linewidth]{figures/swizzling.pdf}
-  \includegraphics[width=\linewidth]{figures/channel_permutation.pdf}
-
+  \includegraphics[width=\linewidth]{figures/swizzling_lanes.pdf}
   \caption{Swizzling enables to reorder or broadcast elements within
@@
 
-\input{additional_figures}
+\def\ww{.13\textwidth}
+\begin{figure*}
+  \begin{center}
+    \subfigure[Supercompact.]{%
+      \label{fig:supercompact}
+      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.pdf}}%
+    \subfigure[Compact.]{%
+      \label{fig:compact}
+      \includegraphics[width=\ww]{figures/compact_matrix-crop.pdf}}%
+    \subfigure[Random.]{%
+      \label{fig:random}
+      \includegraphics[width=\ww]{figures/random_matrix-crop.pdf}}%
+    \subfigure[2D, no RCM.]{%
+      \label{fig:supercompact}
+      \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.pdf}}%
+    \subfigure[2D, RCM.]{%
+      \label{fig:compact}
+      \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.pdf}} 
+    \subfigure[3D, RCM.]{%
+      \label{fig:random} 
+      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.pdf}}%
+    \subfigure[3D, RCM.]{%
+      \label{fig:random}
+      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.pdf}}%
+  \end{center}
+
+  \caption{Different sparsity distributions. In all cases, there are
+    32 nonzeros per line and 300 rows. The last four matrices
+    corresponds to derivative stencils in 2D (a-b) and 3D (c-d) RBF-FD
+    calculations.}
+  \label{fig:spy_plots}
+\end{figure*}
 

### 2013-10-16 ieee_paper.tex
instance description

@@
 
-\todo{How were instances generated. Detail the different type of
-  instances and the rational for trying them. Detail how the realistic
-  instances were generated. Explain matrix ordering (Kutill McKee).}
+\def\ww{.13\textwidth}
+\begin{figure*}
+  \begin{center}
+    \subfigure[Supercompact.]{%
+      \label{fig:supercompact}
+      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.pdf}}%
+    \subfigure[Compact.]{%
+      \label{fig:compact}
+      \includegraphics[width=\ww]{figures/compact_matrix-crop.pdf}}%
+    \subfigure[Random.]{%
+      \label{fig:random}
+      \includegraphics[width=\ww]{figures/random_matrix-crop.pdf}}%
+    \subfigure[2D, no RCM.]{%
+      \label{fig:rbf2dnorcm}
+      \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.pdf}}%
+    \subfigure[2D, RCM.]{%
+      \label{fig:rbf2drcm}
+      \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.pdf}} 
+    \subfigure[3D, RCM.]{%
+      \label{fig:rbf3dnorcm} 
+      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.pdf}}%
+    \subfigure[3D, RCM.]{%
+      \label{fig:rbf3drcm}
+      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.pdf}}%
+  \end{center}
+
+  \caption{Different sparsity distributions. In all cases, there are
+    32 nonzeros per line and 300 rows. The last four matrices
+    corresponds to derivative stencils in 2D and 3D RBF-FD
+    calculations.}
+  \label{fig:spy_plots}
+\end{figure*}
+
+To perform our experiments and analysis, we use multiple kinds of
+matrices which are presented in Figure~\ref{fig:spy_plots}. The
+Supercompact matrices (Figure~\ref{fig:supercompact}) have been
+generated to only have non zero elements in the first 32 columns of
+the matrix. That way we ensure the simplest form of multiplication,
+the amount of elements from the {\tt x} vector data transfered from
+the memory will be minimum. The Compact matrices
+(Figure~\ref{fig:compact}) are generated to have 32 non zeroes per row
+centered around the diagonal and it represents the ideal case for most
+application relying on sparse matrix vector multiplication. The Random
+matrices (Figure~\ref{fig:compact}) see their non zero element
+randomly (uniformly) distributed in the matrix, they represent the
+worst case scenario for cache based architecture where the cache
+reutilisation is the lowest from one row to the next.
+
+The other two types of matrices are used in derivative stencils in 2D
+and 3D RBF-FD calculations (Figure~\ref{fig:rbf2dnorcm} pictures a
+$32$-points stencil of a $23^2$ 2D grid and
+Figure~\ref{fig:rbf2dnorcm} shows a $32$-points stencil of a $8^3$
+grid). We also apply Cuthill-Mckee reordering to all the matrices This
+ordering technique is designed to reduce the distance between the non
+zeroes and the diagonal, hopefully allowing a better cache
+reutilisation. The reordered version of these two matrices can be seen
+in Figure~\ref{fig:rbf2drcm} and in Figure~\ref{fig:rbf3drcm}.
 
@@
 
-\def\ww{.13\textwidth}
-\begin{figure*}
-  \begin{center}
-    \subfigure[Supercompact.]{%
-      \label{fig:supercompact}
-      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.pdf}}%
-    \subfigure[Compact.]{%
-      \label{fig:compact}
-      \includegraphics[width=\ww]{figures/compact_matrix-crop.pdf}}%
-    \subfigure[Random.]{%
-      \label{fig:random}
-      \includegraphics[width=\ww]{figures/random_matrix-crop.pdf}}%
-    \subfigure[2D, no RCM.]{%
-      \label{fig:supercompact}
-      \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.pdf}}%
-    \subfigure[2D, RCM.]{%
-      \label{fig:compact}
-      \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.pdf}} 
-    \subfigure[3D, RCM.]{%
-      \label{fig:random} 
-      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.pdf}}%
-    \subfigure[3D, RCM.]{%
-      \label{fig:random}
-      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.pdf}}%
-  \end{center}
-
-  \caption{Different sparsity distributions. In all cases, there are
-    32 nonzeros per line and 300 rows. The last four matrices
-    corresponds to derivative stencils in 2D (a-b) and 3D (c-d) RBF-FD
-    calculations.}
-  \label{fig:spy_plots}
-\end{figure*}
-
 \section*{Acknowledgment}

### 2013-10-16 ieee_paper.tex
Section 2 done.  Natasha's part.

--- 2013-10-15 ieee_paper.tex
@@
 
+\usepackage{listings}
+
 \newcommand{\todo}[1]{{\color{red}\textbf{\hl{#1}}\xspace}}
@@
 \newcommand{\ceil}[1]{\left\lceil#1\right\rceil}
+
+\usepackage{amsmath}
+\usepackage{amssymb}
 
@@
 
-In this paper, we propose yet a new idea, that is applicable to the
-world of Finite-Difference Radial Basis functions (RBF-FD). A radial
-basis functions refers to a set of shifted radially symmetric
-functions $\phi_i(r) =\phi{||\xvec-\xvec_i||}$ centered at
-$\xvec_i$\cite{}. Any function $f(\xvec)$ defined in a domain $\Omega$
-can be expanded in this basis according to
+In this paper, we propose yet a new idea that is applicable to radial basis functions (RBFs).
+Their strength is the ability to randomly distribute points across complex physical domains, and have an implementation that is independent of dimensionality.
+RBFs approximate a function $f(\xvec)\subset \mathbb{R}^d$ sampled at a set of $N$ distinct point locations, $x_j$, by linearly combining translates of a single radially symmetric function $\phi(r)$, where  $r = \|\xvec-\xvec_{j}\|$ denotes the Euclidean distance (e.g., in 2-D $\sqrt{(x-x_j)^2+(y-y_j)^2}$) between where the function is evaluated $\xvec$ and where the RBF is centered $\xvec_{j}$.
+That is, the interpolant is $s(\xvec) =  \sum_{j=1}^{N} w_j \phi_i(\|\xvec-\xvec_{j}\|)$.
+The weights $w_j$ are obtained by inverting the system
 \begin{equation}
-f(\xvec) =  \sum_i w_i \phi_i(\xvec) \label{eq:rbf}
+\parray{lccr}{
+\phi_{11} & \phi_{12} & \cdots & \phi_{1N} \\
+\vdots & \ddots & \vdots & \vdots \\
+\phi_{n1} & \phi_{n2} & \cdots & \phi_{NN} 
+}
+\parray{c}{ w_{1} \\ \vdots \\ w_{N} }
+=
+\parray{c}{ f(\xvec_1) \\ \vdots\\ f(\xvec_N) }. 
+\label{eq:rbf}
 \end{equation}
-% Wrong label eq:rbf in document 
-Given function values at a set of discrete points $x_i$, one inverts
-linear system (\ref{eq:rbf}) to obtain a set of weights $w_j$. Given a
-set of $N$ RBF nodes, the matrix to invert is dense. Derivatives of
-$f(\xvec)$ result from the differentiation of Eq. (\ref{eq:rbf}). For
-very large problems, this is an computationally expensive since the
-RBFs do not form an orthogonal basis, nor are they amenable to tensor
-decomposition in the three coordinate directions. Rather, their
-strength is the ability to randomly distribute nodes across complex
-physical domains, and have an implementation that is independent of
-dimensionality.
+where $\phi_{ij} = \phi(\|\xvec_i-\xvec_j\|)$. 
+%\end{equation}
+%\begin{equation}
+%\begin{bmatrix}
+%\phi(\|\xvec_{1}-\xvec_{1}\|) & \phi(\|\xvec_{1}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{1}-\xvec_{N}\|) \\
+%\vdots & \ddots & \vdots  \\
+%\phi (\|\xvec_{n}-\xvec_{1}\|) & \phi(\|\xvec_{n}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{N}-\xvec_{N}\|)
+%\end{bmatrix}
 
-To alleviate the cost of a global approach, researchers \cite{} expand
-$f(\xvec)$ in a local region of $x_i$. Thus,
+The RBF differentiation matrix, $D_N$, is derived by applying the desired analytic derivative operator $L$ to the RBF interpolant $s(\xvec)$ above and evaluating it at the point locations.
+For very large problems, this is a computationally expensive since the matrix in (\ref{eq:rbf}) is full and inversion requires O$(N^3)$ operations.
+To alleviate the cost of this global approach (i.e.
+using every node in the domain to calculate the derivative at a given node $\xvec_i$), RBF-generated finite differences (RBF-FD) have been derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}.
+RBF-FD use only a local set of the $n_z-1$ nearest neighbors to the point $\xvec_i$ to approximate the derivative.
+In other words, $Lf(\xvec_i)=\sum_{j=1}^{n_z}a_jf(\xvec_j)$.
+The differentiation weights, $a_j$, are calculated by enforcing that this linear combination should be exact for RBFs, $\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^n_z$, centered at each of the node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences (FD) would enforce that it be exact for polynomials instead).
+Similar to FD, as the stencil size $n_z$ increases so does the order of the method.
+
+For a total of $N$ points, there will be $N$ linear systems to solve, each of size $n_z \times n_z$.
+Each linear solve produces a row of the RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \ times N$ matrix with $n_zN$ nonzero entries.
+To evaluate the derivative at all points in the domain, takes the form
 $$
-f(\xvec) =  \sum_{i=0}^{n_s} w_i \phi_i(\xvec) \label{eq:rbf}
+  \mathbf{g} = D_{n_z} \mathbf{f}
 $$
-where one only considers a local set of $n_s$ neighbors of of point $x_j$. 
-Finite-difference RBF expresses the derivative of $f(\xvec)$ at $x_j$ as a linear
-combination of $f(\xvec)$ at the $n_s$ neighbors. Thus, 
-$$
-f = A w
-$$ where $f$ is the vector of function values at the $n_s$ stencil
-points (including the center point $x_j$. The matrix $A$ is a function
-of $\phi_{ij}(\xvec) =\phi(||x_i-x_j||)$. This leads to the $n_s$
-weight values $w = A^{-1} f$. The derivative of Eq. \ref{eq:rbf}
-evaluated at $x_j$ gives
-$$
-   f'_j = \sum_{i=0}^{n_s} w_i \phi'_{ij}
-$$
-Substituting the vector of weights $w$ leads to 
-$$
-  f'_j = \sum_{i=0}^{n_s} \alpha_i f_i
-$$
-The derivative evaluation at all points of the domain $\Omega$, notwithstanding boundary conditions, takes the form
-$$
-  y = A x
-$$ 
-where $x$ is the souce vector of function values, and $y$ is the
-vector of derivative values. The matrix $A$ is called a derivative
-matrix. In the current litterature, the number of stencil points is a
-fixed value, independent of the node under consideration\cite{}. The
-computation of a single derivative has been reduced to a SpMV, where
-each row has $n_s$ nonzeros. Because the sparsity of each row is
-constant, ELLPACK~\cite{} becoes the most appropriate matrix
-compression scheme\cite{}. In practice, $n_s=32$ in two-dimensional
-flows and 64 or 100 for three-dimensional flows. These numbers are
-similar to what is used in finite-element codes.
+where $f$ is the source vector of function values and $g$ is the resulting vector of derivative values.
+The computation of a single derivative has been reduced to a SpMV, where each row has $n_z$ nonzeros.
+Since the sparsity of each row is constant, ELLPACK~\cite{} becomes the most appropriate matrix compression scheme\cite{}.
+In practice, $n_z=32$ in two-dimensional flows and 64 or 100 for three-dimensional flows.
+These numbers are similar to what is used in finite-element codes.
 
-Many problems in fluid dynamics and in the geosciences require the solution to
-transport equations of the form
+Many problems in fluid dynamics and in the geosciences require the solution to transport equations of the form
 $$
@@
 where $Q$ is a vector of unknowns (3 components of velocity,
-pressure, temperature). Thus, it is often necessary to compute the $x$
+pressure, temperature). For example, in solving a system of equations, it is often necessary to compute the $x$
 derivative of multiple functions, typically four for the Euler
@@
 calculate multiple derivatives of a single functions. The increased
-memory bandwidth due to an increase in the number of derivaive
+memory bandwidth due to an increase in the number of derivative
 matrices is offset by better cache utilization, leading to an overall
 benefit.
+
+% GORDON I WOULD ELIMINATE BELOW AND END HERE.
 
@@
 \newfig{figures/rbf_stencils.pdf}{\wide}{RBFFD Stencils.
-Each node of the mesh is connected to $n_z-1$ stencil nodes in addition to itself.
+Each node of the stencil is connected to $n_z-1$ stencil nodes in addition to itself.
 In the figure, node $A$ is connected to $B$, but $B$ is {\em not\/} connected to $A$.
@@
 
-\begin{figure*}
-\begin{verbatim}
+\begin{figure}[h]
+\begin{center}
+\lstset{language=c++,basicstyle=\scriptsize,numbers=left,numbersep=5pt}
+\begin{lstlisting}{float=h, Caption}
 const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
@@
   _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
-} 
-\end{verbatim}
-\caption{mat mul code.}
-\label{code:mat_mul}
-\end{figure*}
+\end{lstlisting}
+\end{center}
+\end{figure}
+
+%\begin{figure*}
+%\begin{verbatim}
+%const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
+%const __m512i four = _mm512_set4_epi32(4,4,4,4); 
+%int int_mask = 0x1111;
+%__mmask16 mask = _mm512_int2mask(int_mask);
+%
+%#pragma omp parallel for
+%for (int r=0; r < nb_rows; r++) {
+%  __m512 accu = _mm512_setzero_ps();
+%
+%  for (int n=0; n < nz; n+=4) {
+%    float* a = &dom.col_id_t[0]+n+nz*r;
+%    __m512i vect_i = _mm512_setzero_ps();
+%    vect_i = _mm512_mask_loadunpacklo_ps(vect_i, mask, a);
+%    vect_i = _mm512_mask_loadunpackhi_ps(vect_i, mask, a);
+%    vect_i = _mm512_swizzle_ps(vect_i, _MM_SWIZ_REG_AAAA);
+%    vect_i = _mm512_fmadd_epi32(vect_i, four, offsets);
+%
+%    all_vect = _mm512_i32gather_ps(vect_i, dom.vec_vt, scale);
+%
+%    mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));
+%
+%    //perform first tensor multiplication
+%    correct_vect = permute(all_vect, _MM_PERM_AAAA);
+%    correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
+%    accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);
+%
+%    //three more times with BBBB, CCCC, and DDDD
+%  }
+%  _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
+%} 
+%\end{verbatim}
+%\caption{mat mul code.}
+%\label{code:mat_mul}
+%\end{figure*}
 
@@
     case implemented with MIC vector instructions. $-O3$ compilation
-    options. Each group consists of four cases: grids of $64^3$ and
-    $96^3$ with and without Reverse Cuthill
-    McKee.\todo{I don't think this is supercompact but RBF 3D.}]{\includegraphics[width=.49\linewidth]{figures/mic_performance_nb_threads.pdf}\label{}}
+    options. Each symbol has two cases: with (red) and without blue) 
+    with and without Reverse Cuthill-McKee. The grid is $96^3$.}
+    {\includegraphics[width=.49\linewidth]{figures/mic_performance_nb_threads.pdf}\label{}}
 %
-  \subfigure[performance on the host on a 64x64x64 matrix. Manual
+  \subfigure[performance on the host on a sparse matrix with $64^3$ rows. Manual
     vectoization brings significant improvement there as
     well.]{\includegraphics[width=.49\linewidth]{figures/64x64x64.pdf}\label{}}
-
   \caption{Vectorization matters in practice. Differences in BW
@@
 The first and last authors acknowledge NSF funding under NSF grant DMS-\#0934331 (FSU). 
+NCAR is sponsored by the National Science Foundation. N. Flyer      
+acknowledges support of NSF grant DMS-0934317.
 

### 2013-10-16 ieee_paper.tex
Modified source code with omnigraffle. Why doesn't figure span both columns?

@@
 In other words, $Lf(\xvec_i)=\sum_{j=1}^{n_z}a_jf(\xvec_j)$.
-The differentiation weights, $a_j$, are calculated by enforcing that this linear combination should be exact for RBFs, $\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^n_z$, centered at each of the node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences (FD) would enforce that it be exact for polynomials instead).
+The differentiation weights, $a_j$, are calculated by enforcing that this linear combination should be exact for RBFs, $\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences (FD) would enforce that it be exact for polynomials instead).
 Similar to FD, as the stencil size $n_z$ increases so does the order of the method.
@@
 
-**** PUT ELSEWHERE OR ELIMINATE? ****
-
-When solving a one dimensional system of PDEs, one might require an
-$x$ derivative of multiple functions. For example, the Euler equations
-require the $x$ derivative of the three components of velocity and
-pressure. In this case, there are $n_v=4$ vectors $x^k$,
-$k=0,\cdots,3$.
-
-Thus for each vector element $y_i$, we compute $y_i = \sum_j A_{ij}
-x_j$. If $A_i$ is row $i$ of $A$, $y_i$ is simply the dot product $A_i
-x$. The next level of generality is to consider $n_v$ vectors $x^k$,
-$k=0,\cdots,n_v-1$. Whatever the spmv implementation, one achieves
-improved performance if the matrix formed from the columns $x^k$ are
-stored in row major order. Thus, $x^0_0,x^1_0,\cdots,x^{n_v-1}_0$, are
-stored in consecutive memory location. The random access of the
-elements of $x$ is thus reduced. Maximum efficiency is achieved when
-$n_v=16$ floats or $8$ doubles, given that cache lines take 64
-bytes. We will benchmark this case, labeled $Svn$, where $n$ refers to
-the number of vectors (Iv4 uses four vectors). The $S$ refers to singe
-precision. A double precision run is labelled $Dvn$.
-
-Alternatively, when solving a PDE, one might require derivatives of a
-given scalar function with respect to coordinate directions $x$, $y$,
-$z$. Second order operators of often required, such as a second
-derivative with respect to $x$ or a Laplacian operator. In the RBFFD
-formulation, on can compute different derivatives using the same
-stencil, but with different weights. In other words, the adjacency
-matrix that corresponds to $A$ remains constant, but the matrix
-elements of $A$ change with the particular derivative.  In this case,
-label with a superscript $l$ the particular matrix $A^k$. Since the
-adjacency matrix is assumed invariant, there is only need for a single
-matrix \ttt{C{ij}}. In Ellpack format, each row is of constant size
-(the number of nonzeros per row of $A$. $C_{ij}$ is the column number
-that locates the $j^{th}$ nonzero in row $i$ of $A$.
-
-\todo{Must rewrite the above to make it more focused on the
-  application. Remove the architectural/implementation details such as
-  using ELLPACK. Let's focus on on why this computation is important
-  and how does the 16 multiplication appears in the equations.}
+%**** PUT ELSEWHERE OR ELIMINATE? ****
+%
+%When solving a one dimensional system of PDEs, one might require an
+%$x$ derivative of multiple functions. For example, the Euler equations
+%require the $x$ derivative of the three components of velocity and
+%pressure. In this case, there are $n_v=4$ vectors $x^k$,
+%$k=0,\cdots,3$.
+%
+%Thus for each vector element $y_i$, we compute $y_i = \sum_j A_{ij}
+%x_j$. If $A_i$ is row $i$ of $A$, $y_i$ is simply the dot product $A_i
+%x$. The next level of generality is to consider $n_v$ vectors $x^k$,
+%$k=0,\cdots,n_v-1$. Whatever the spmv implementation, one achieves
+%improved performance if the matrix formed from the columns $x^k$ are
+%stored in row major order. Thus, $x^0_0,x^1_0,\cdots,x^{n_v-1}_0$, are
+%stored in consecutive memory location. The random access of the
+%elements of $x$ is thus reduced. Maximum efficiency is achieved when
+%$n_v=16$ floats or $8$ doubles, given that cache lines take 64
+%bytes. We will benchmark this case, labeled $Svn$, where $n$ refers to
+%the number of vectors (Iv4 uses four vectors). The $S$ refers to singe
+%precision. A double precision run is labelled $Dvn$.
+%
+%Alternatively, when solving a PDE, one might require derivatives of a
+%given scalar function with respect to coordinate directions $x$, $y$,
+%$z$. Second order operators of often required, such as a second
+%derivative with respect to $x$ or a Laplacian operator. In the RBFFD
+%formulation, on can compute different derivatives using the same
+%stencil, but with different weights. In other words, the adjacency
+%matrix that corresponds to $A$ remains constant, but the matrix
+%elements of $A$ change with the particular derivative.  In this case,
+%label with a superscript $l$ the particular matrix $A^k$. Since the
+%adjacency matrix is assumed invariant, there is only need for a single
+%matrix \ttt{C{ij}}. In Ellpack format, each row is of constant size
+%(the number of nonzeros per row of $A$. $C_{ij}$ is the column number
+%that locates the $j^{th}$ nonzero in row $i$ of $A$.
+%
+%
+%\todo{Must rewrite the above to make it more focused on the
+% application. Remove the architectural/implementation details such as
+%  using ELLPACK. Let's focus on on why this computation is important
+%  and how does the 16 multiplication appears in the equations.}
 
@@
 
-\begin{figure}[h]
-\begin{center}
-\lstset{language=c++,basicstyle=\scriptsize,numbers=left,numbersep=5pt}
-\begin{lstlisting}{float=h, Caption}
-const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
-const __m512i four = _mm512_set4_epi32(4,4,4,4); 
-int int_mask = 0x1111;
-__mmask16 mask = _mm512_int2mask(int_mask);
+\newfig{figures/source_code.pdf}{6in}{Source code}{fig:source}
 
-#pragma omp parallel for
-for (int r=0; r < nb_rows; r++) {
-  __m512 accu = _mm512_setzero_ps();
-
-  for (int n=0; n < nz; n+=4) {
-    float* a = &dom.col_id_t[0]+n+nz*r;
-    __m512i vect_i = _mm512_setzero_ps();
-    vect_i = _mm512_mask_loadunpacklo_ps(vect_i, mask, a);
-    vect_i = _mm512_mask_loadunpackhi_ps(vect_i, mask, a);
-    vect_i = _mm512_swizzle_ps(vect_i, _MM_SWIZ_REG_AAAA);
-    vect_i = _mm512_fmadd_epi32(vect_i, four, offsets);
-
-    all_vect = _mm512_i32gather_ps(vect_i, dom.vec_vt, scale);
-
-    mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));
-
-    //perform first tensor multiplication
-    correct_vect = permute(all_vect, _MM_PERM_AAAA);
-    correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
-    accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);
-
-    //three more times with BBBB, CCCC, and DDDD
-  }
-  _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
-\end{lstlisting}
-\end{center}
-\end{figure}
+%\begin{figure}[h]
+%\begin{center}
+%\lstset{language=c++,basicstyle=\scriptsize,numbers=left,numbersep=5pt}
+%\begin{lstlisting}{float=h, Caption}
+%const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
+%const __m512i four = _mm512_set4_epi32(4,4,4,4); 
+%int int_mask = 0x1111;
+%__mmask16 mask = _mm512_int2mask(int_mask);
+%
+%#pragma omp parallel for
+%for (int r=0; r < nb_rows; r++) {
+%  __m512 accu = _mm512_setzero_ps();
+%
+%  for (int n=0; n < nz; n+=4) {
+%    float* a = &dom.col_id_t[0]+n+nz*r;
+%    __m512i vect_i = _mm512_setzero_ps();
+%    vect_i = _mm512_mask_loadunpacklo_ps(vect_i, mask, a);
+%    vect_i = _mm512_mask_loadunpackhi_ps(vect_i, mask, a);
+%    vect_i = _mm512_swizzle_ps(vect_i, _MM_SWIZ_REG_AAAA);
+%    vect_i = _mm512_fmadd_epi32(vect_i, four, offsets);
+%
+%    all_vect = _mm512_i32gather_ps(vect_i, dom.vec_vt, scale);
+%
+%    mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));
+%
+%    //perform first tensor multiplication
+%    correct_vect = permute(all_vect, _MM_PERM_AAAA);
+%    correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
+%    accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);
+%
+%    //three more times with BBBB, CCCC, and DDDD
+%  }
+%  _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
+%\end{lstlisting}
+%\end{center}
+%\end{figure}
+%
 
@@
   \centering
-
-  \subfigure[Performance of $y=Ax$ on the MIC. Squares: base 1/1 case,
+  \subfigure[Performance of $y=Ax$ on the MIC. Squares: base 1/1 case],
     solid circles: 4/4 case implemented in C++, solid triangles: 4/4
@@
     with and without Reverse Cuthill-McKee. The grid is $96^3$.}
-    {\includegraphics[width=.49\linewidth]{figures/mic_performance_nb_threads.pdf}\label{}}
+    {\includegraphics[width=.49\linewidth]{figures/mic_performance_nb_threads.pdf}\label{fig:perf_mic}}
 %
   \subfigure[performance on the host on a sparse matrix with $64^3$ rows. Manual
-    vectoization brings significant improvement there as
-    well.]{\includegraphics[width=.49\linewidth]{figures/64x64x64.pdf}\label{}}
+    vectorization brings significant improvement there as
+    well.]{\includegraphics[width=.49\linewidth]{figures/64x64x64.pdf}\label{fig:perf_host}}
   \caption{Vectorization matters in practice. Differences in BW

### 2013-10-16 ieee_paper.tex
Updated paper with all figures. Figure 12 is probably not required.

@@
 
-\newfig{figures/source_code.pdf}{6in}{Source code}{fig:source}
+\begin{figure*}
+  \centering
+  %\subfigure[Source code}
+  \includegraphics[width=0.8\textwidth]{figures/source_code.pdf}%
+  \label{fig:source}
+  \caption{Source Code}
+\end{figure*}
+%\newfig{figures/source_code.pdf}{6in}{Source code}{fig:source}
 
@@
 \begin{figure*}
+\centering
+  \subfigure[
+    Entries with "cpp" denote cases where
+    coding was performed without MIC vector
+    instructions.]{\includegraphics[width=.40\linewidth]{figures/test1_readwrite.pdf}\label{fig:band_rw}}
+%
+  \subfigure[ xxx ] 
+  {\includegraphics[width=.40\linewidth]{figures/test1_gather.pdf}\label{fig:band_gather}} \\
 
-  \subfigure[Bandwidth performance under idealized conditions as a
-    function of matrix row size. Entries with "cpp" denote cases where
-    coding was performed without MIC vector
-    instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_readwrite.pdf}\label{fig:band_rw}}
-%
-  \subfigure[Bandwidth performance under
-  idealized conditions as a function of matrix row size. Entries with
-  "cpp" denote cases where coding was performed without MIC vector
-  instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_gather.pdf}\label{fig:band_gather}}
-
-  \subfigure[Bandwidth performance under idealized conditions as a
-    function of matrix row size. Entries with "cpp" denote cases where
-    coding was performed without MIC vector instructions. The greater
+  \subfigure[The greater
     speed of the cpp version is obtained throught the use of
@@
     is aligned on 64
-    bytes.]{\includegraphics[width=.49\linewidth]{figures/test3_gather.pdf}\label{fig:band_gather_ivdep}}
+    bytes.]{\includegraphics[width=.40\linewidth]{figures/test3_gather.pdf}\label{fig:band_gather_ivdep}}
 %
-  \subfigure[Bandwidth performance on the host under idealized
-    conditions as a function of matrix row size. Entries with "cpp"
-    denote cases where coding was performed without AVX vector
-    instructions. The speed on the CPU matches the speed with AVX
-    instructions. All memory is aligned on 32
-    bytes.]{\includegraphics[width=.49\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:read_write}}
+  \subfigure[
+    The speed on the CPU matches the speed with vector instructions.
+    All memory is aligned on 32
+    bytes.]{\includegraphics[width=.40\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:read_write}}
 
-\caption{Bandwidth Analysis.\todo{not 100\% sure what this plots are.}
-  \todo{can we harmozie y-axis?} \todo{ylabel should read "Bandwidth (GB/s)".}
-  \todo{remove title.}}
-
+\caption{Bandwidth performance on the host under idealized conditions as a function of matrix row size.
+  Entries with "cpp" denote cases where coding was performed without MIC vector intructions. 
+\todo{not 100\% sure what this plots are.}}
 \end{figure*}
@@
   \centering
-  \subfigure[Performance of $y=Ax$ on the MIC. Squares: base 1/1 case],
+  \subfigure[Performance of $y=Ax$ on the MIC. Squares: base 1/1 case,
     solid circles: 4/4 case implemented in C++, solid triangles: 4/4
@@
     options. Each symbol has two cases: with (red) and without blue) 
-    with and without Reverse Cuthill-McKee. The grid is $96^3$.}
-    {\includegraphics[width=.49\linewidth]{figures/mic_performance_nb_threads.pdf}\label{fig:perf_mic}}
+    with and without Reverse Cuthill-McKee. The grid is $96^3$.]
+    {\includegraphics[width=.40\linewidth]{figures/mic_performance_nb_threads.pdf}\label{fig:perf_mic}}
 %
-  \subfigure[performance on the host on a sparse matrix with $64^3$ rows. Manual
+  \subfigure[Performance on the host on a sparse matrix with $64^3$ rows. Manual
     vectorization brings significant improvement there as
-    well.]{\includegraphics[width=.49\linewidth]{figures/64x64x64.pdf}\label{fig:perf_host}}
+    well.]{\includegraphics[width=.40\linewidth]{figures/64x64x64.pdf}\label{fig:perf_host}}
   \caption{Vectorization matters in practice. Differences in BW
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
   \centering
-  \subfigure[Supercompact\todo{is the title? this looks like super compact.}]{\includegraphics[width=.49\linewidth]{figures/supercompact_max_perf.pdf}}
-  \subfigure[Compact]{\includegraphics[width=.49\linewidth]{figures/compact_max_perf.pdf}}
-
-  \subfigure[Random]{\includegraphics[width=.49\linewidth]{figures/random_max_perf.pdf}}
-  \subfigure[RBF 3D]{\includegraphics[width=.49\linewidth]{figures/rbf_max_perf_run1.pdf}}
+  \subfigure[Supercompact]{\includegraphics[width=.48\linewidth]{figures/supercompact_max_perf.pdf}}
+  \subfigure[Compact]{\includegraphics[width=.48\linewidth]{figures/compact_max_perf.pdf}} \\
+  \subfigure[Random]{\includegraphics[width=.48\linewidth]{figures/random_max_perf.pdf}}
+  \subfigure[RBF 3D]{\includegraphics[width=.48\linewidth]{figures/rbf_max_perf_run1.pdf}}
   \caption{Performance of "best code" accross instance types. \todo{Why stop at 1.6M
       rows. Is that because of memory limitation? I guess that's what
-      it is}\todo{can we remove the title?}\todo{can we harmonize y-axis?}}
-\end{figure*}
+      it is}}
+\end{figure}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
+  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
 
@@
 \end{figure}
+
+\begin{figure*}
+\centering
+\subfigure[No cache effects.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
+\subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}} %\label{fig:ratio_best}}
+\subfigure[worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
+\caption{Ratio of flops to bytes; (a) simple count of flops and bytes that require transfer from memory;(b) best and (c) worst
+case taking into account that bringing one element of $x$ into memory brings along the remainder of the cache line.}
+\label{fig:ratio_bytes_flops}
+\end{figure*}
 

### 2013-10-16 ieee_paper.tex
using raster version of spyplots

@@
       \label{fig:supercompact}
-      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.pdf}}%
+      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.png}}%
     \subfigure[Compact.]{%
       \label{fig:compact}
-      \includegraphics[width=\ww]{figures/compact_matrix-crop.pdf}}%
+      \includegraphics[width=\ww]{figures/compact_matrix-crop.png}}%
     \subfigure[Random.]{%
       \label{fig:random}
-      \includegraphics[width=\ww]{figures/random_matrix-crop.pdf}}%
+      \includegraphics[width=\ww]{figures/random_matrix-crop.png}}%
     \subfigure[2D, no RCM.]{%
       \label{fig:rbf2dnorcm}
-      \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.pdf}}%
+      \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.png}}%
     \subfigure[2D, RCM.]{%
       \label{fig:rbf2drcm}
-      \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.pdf}} 
+      \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
     \subfigure[3D, RCM.]{%
       \label{fig:rbf3dnorcm} 
-      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.pdf}}%
+      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
     \subfigure[3D, RCM.]{%
       \label{fig:rbf3drcm}
-      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.pdf}}%
+      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.png}}%
   \end{center}

### 2013-10-16 ieee_paper.tex
reformat text (no actual changes)

@@
 
-In this paper, we propose yet a new idea that is applicable to radial basis functions (RBFs).
-Their strength is the ability to randomly distribute points across complex physical domains, and have an implementation that is independent of dimensionality.
-RBFs approximate a function $f(\xvec)\subset \mathbb{R}^d$ sampled at a set of $N$ distinct point locations, $x_j$, by linearly combining translates of a single radially symmetric function $\phi(r)$, where  $r = \|\xvec-\xvec_{j}\|$ denotes the Euclidean distance (e.g., in 2-D $\sqrt{(x-x_j)^2+(y-y_j)^2}$) between where the function is evaluated $\xvec$ and where the RBF is centered $\xvec_{j}$.
-That is, the interpolant is $s(\xvec) =  \sum_{j=1}^{N} w_j \phi_i(\|\xvec-\xvec_{j}\|)$.
-The weights $w_j$ are obtained by inverting the system
+In this paper, we propose yet a new idea that is applicable to radial
+basis functions (RBFs). Their strength is the ability to randomly
+distribute points across complex physical domains, and have an
+implementation that is independent of dimensionality. RBFs approximate
+a function $f(\xvec)\subset \mathbb{R}^d$ sampled at a set of $N$
+distinct point locations, $x_j$, by linearly combining translates of a
+single radially symmetric function $\phi(r)$, where $r =
+\|\xvec-\xvec_{j}\|$ denotes the Euclidean distance (e.g., in 2-D
+$\sqrt{(x-x_j)^2+(y-y_j)^2}$) between where the function is evaluated
+$\xvec$ and where the RBF is centered $\xvec_{j}$. That is, the
+interpolant is $s(\xvec) = \sum_{j=1}^{N} w_j
+\phi_i(\|\xvec-\xvec_{j}\|)$. The weights $w_j$ are obtained by
+inverting the system
 \begin{equation}
@@
 
-The RBF differentiation matrix, $D_N$, is derived by applying the desired analytic derivative operator $L$ to the RBF interpolant $s(\xvec)$ above and evaluating it at the point locations.
-For very large problems, this is a computationally expensive since the matrix in (\ref{eq:rbf}) is full and inversion requires O$(N^3)$ operations.
-To alleviate the cost of this global approach (i.e.
-using every node in the domain to calculate the derivative at a given node $\xvec_i$), RBF-generated finite differences (RBF-FD) have been derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}.
-RBF-FD use only a local set of the $n_z-1$ nearest neighbors to the point $\xvec_i$ to approximate the derivative.
-In other words, $Lf(\xvec_i)=\sum_{j=1}^{n_z}a_jf(\xvec_j)$.
-The differentiation weights, $a_j$, are calculated by enforcing that this linear combination should be exact for RBFs, $\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences (FD) would enforce that it be exact for polynomials instead).
-Similar to FD, as the stencil size $n_z$ increases so does the order of the method.
+The RBF differentiation matrix, $D_N$, is derived by applying the
+desired analytic derivative operator $L$ to the RBF interpolant
+$s(\xvec)$ above and evaluating it at the point locations. For very
+large problems, this is a computationally expensive since the matrix
+in (\ref{eq:rbf}) is full and inversion requires O$(N^3)$
+operations. To alleviate the cost of this global approach (i.e. using
+every node in the domain to calculate the derivative at a given node
+$\xvec_i$), RBF-generated finite differences (RBF-FD) have been
+derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBF-FD use only
+a local set of the $n_z-1$ nearest neighbors to the point $\xvec_i$ to
+approximate the derivative. In other words,
+$Lf(\xvec_i)=\sum_{j=1}^{n_z}a_jf(\xvec_j)$. The differentiation
+weights, $a_j$, are calculated by enforcing that this linear
+combination should be exact for RBFs,
+$\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the
+node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences
+(FD) would enforce that it be exact for polynomials instead). Similar
+to FD, as the stencil size $n_z$ increases so does the order of the
+method.
 
-For a total of $N$ points, there will be $N$ linear systems to solve, each of size $n_z \times n_z$.
-Each linear solve produces a row of the RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \ times N$ matrix with $n_zN$ nonzero entries.
-To evaluate the derivative at all points in the domain, takes the form
+For a total of $N$ points, there will be $N$ linear systems to solve,
+each of size $n_z \times n_z$. Each linear solve produces a row of the
+RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \ times N$
+matrix with $n_zN$ nonzero entries. To evaluate the derivative at all
+points in the domain, takes the form
 $$
@@
 $$
-where $f$ is the source vector of function values and $g$ is the resulting vector of derivative values.
-The computation of a single derivative has been reduced to a SpMV, where each row has $n_z$ nonzeros.
-Since the sparsity of each row is constant, ELLPACK~\cite{} becomes the most appropriate matrix compression scheme\cite{}.
-In practice, $n_z=32$ in two-dimensional flows and 64 or 100 for three-dimensional flows.
-These numbers are similar to what is used in finite-element codes.
+where $f$ is the source vector of function values and $g$ is the
+resulting vector of derivative values. The computation of a single
+derivative has been reduced to a SpMV, where each row has $n_z$
+nonzeros. Since the sparsity of each row is constant, ELLPACK~\cite{}
+becomes the most appropriate matrix compression scheme\cite{}. In
+practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
+three-dimensional flows. These numbers are similar to what is used in
+finite-element codes.
 
-Many problems in fluid dynamics and in the geosciences require the solution to transport equations of the form
+Many problems in fluid dynamics and in the geosciences require the
+solution to transport equations of the form
 $$
@@
 $$
-where $Q$ is a vector of unknowns (3 components of velocity,
-pressure, temperature). For example, in solving a system of equations, it is often necessary to compute the $x$
-derivative of multiple functions, typically four for the Euler
-equations, or five for the Navier-Stokes equations. Multiple
-right-hand sides transform a SpMV into a SpMM (Sparse Matrix/dense
-Matrix multiplication), which improves register utilization and
-decreases cache misses by vectorizing over the multiple source
-vectors. Further improvements are possible by recognizing that
-different derivative matrices (x,y,z and Laplacian for example), have
-the same sparsity distribution; only the weight change.  Thus,
-alternative to computing a derivative of multiple functions, we can
-calculate multiple derivatives of a single functions. The increased
-memory bandwidth due to an increase in the number of derivative
-matrices is offset by better cache utilization, leading to an overall
-benefit.
+where $Q$ is a vector of unknowns (3 components of velocity, pressure,
+temperature). For example, in solving a system of equations, it is
+often necessary to compute the $x$ derivative of multiple functions,
+typically four for the Euler equations, or five for the Navier-Stokes
+equations. Multiple right-hand sides transform a SpMV into a SpMM
+(Sparse Matrix/dense Matrix multiplication), which improves register
+utilization and decreases cache misses by vectorizing over the
+multiple source vectors. Further improvements are possible by
+recognizing that different derivative matrices (x,y,z and Laplacian
+for example), have the same sparsity distribution; only the weight
+change.  Thus, alternative to computing a derivative of multiple
+functions, we can calculate multiple derivatives of a single
+functions. The increased memory bandwidth due to an increase in the
+number of derivative matrices is offset by better cache utilization,
+leading to an overall benefit.
 

### 2013-10-16 ieee_paper.tex
minor

@@
 
-In this paper, we propose yet a new idea that is applicable to radial
+In this paper, we propose a new idea that is applicable to radial
 basis functions (RBFs). Their strength is the ability to randomly
@@
 $\xvec_i$), RBF-generated finite differences (RBF-FD) have been
-derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBF-FD use only
+derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBF-FD uses only
 a local set of the $n_z-1$ nearest neighbors to the point $\xvec_i$ to
@@
 each of size $n_z \times n_z$. Each linear solve produces a row of the
-RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \ times N$
+RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \times N$
 matrix with $n_zN$ nonzero entries. To evaluate the derivative at all

### 2013-10-16 ieee_paper.tex
citations

@@
 \section{Introduction}
-\cite{Bell08, Vuduc05, Nishtala07, Stock12-TACC, Cuthill69, cramer2012openmp, Buluc2009_SPAA, Buluc11, Im01, Mellor-Crummey04, Nishtala07, Saad94sparskit, Williams07}
-\cite{Temam:1992:CBS:147877.148091}
-\cite{DBLP:journals/ijhpca/ShantharamCR11,
-conf/ppsc/Toledo97, Liu:2013:ESM:2464996.2465013, Molka:2009:MPC:1636712.1637764,%
-DBLP:journals/corr/abs-1101-0091, conf/ipps/BulucWOD11, conf/ipps/KreutzerHWFBB12,%
-kumar2012accelerating, journals/concurrency/VazquezFG11}
+\cite{Stock12-TACC} is all dense. \cite{
+  Molka:2009:MPC:1636712.1637764} is about memory modeling not
+directly spmv.
 
-SpMV is an important kernel for lots of stuff. So improving the
-performance of SpMV has caputred the interest of many researcher.
+\cite{Im01, Saad94sparskit} \cite{DBLP:journals/ijhpca/ShantharamCR11,
+  DBLP:journals/corr/abs-1101-0091 }
 
-The main challenge that is faced to improve good performance for SpMV
-is that the operation are conducted using memory location that are
-irregular and often unpredictible. That make that the kernels are
-mostly memory bound and there is a significant instruction overhead
-per flop.
+The multiplication of sparse matrix by a dense vector (SpMV) is an
+important kernel in many fields from scientific computing~\cite{} to
+graph analysis (recommender systems~\cite{Brin98}, graph
+drawing~\cite{Koren05}). So improving the performance of SpMV has
+captured the interest of many researcher; including the development of
+various implemetation for multi-core
+CPUs~\cite{Buluc2009SPAA,Williams07} and
+GPUs~\cite{Bell08, conf/ipps/KreutzerHWFBB12,
+  journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
+challenge that is faced to obtain good performance for SpMV is that
+the operation are conducted using memory location that are irregular
+and often unpredictible. That make that the kernels are mostly memory
+bound and there is a significant instruction overhead per flop.
 
-Common improvement techniques such as register blocking, bandiwdth
-reduction (matrix reordering), partitioning to fit in cache or TLB
-have impacts which are very dependent on the matrix and overall do not
-lead to dramatic improvement. Assuming register blocking does not
-apply well to the matrix at hand (which is true for most matrices),
-there is about 8 bytes of the matrix to move in per non zero (assuming
-single precision); each nonzero requires two floating point operations
-leading to a flop-to-byte ratio of at most $\frac{1}{4}$. This limits
-the obtained performance to at most a quarter of the bandwidth of the
-architecture wasting a lot of potentially useful cycles. The commonly
-used techniques are mostly designed to reach that bound rather than
-overcome it.
+Common improvement techniques such as register blocking, bandwidth
+reduction (matrix reordering~\cite{Cuthill69}), partitioning to fit in
+cache or TLB~\cite{Nishtala07,Temam:1992:CBS:147877.148091,conf/ppsc/Toledo97},
+unrolling~\cite{Mellor-Crummey04} have impacts which are very
+dependent on the matrix and overall do not lead to dramatic
+improvement. (The state-of-the-art techniques in OSKI~\cite{Vuduc05}
+provide some useful yet limited improvements). Register blocking~\cite{conf/ppsc/Toledo97} does
+not apply well to most the matrix at hand (despite it can be applied
+with virtually no overhead in all cases thanks to compressed
+representation~\cite{Buluc11}). there is about 8 bytes of the matrix to move
+in per non zero (assuming single precision); each nonzero requires two
+floating point operations leading to a flop-to-byte ratio of at most
+$\frac{1}{4}$. This limits the obtained performance to at most a
+quarter of the bandwidth of the architecture wasting a lot of
+potentially useful cycles. The commonly used techniques are mostly
+designed to reach that bound rather than overcome it.
 
@@
 for sparse linear algebra compared to more classical CPU or GPU
-architecture~\cite{}.
+architecture~\cite{Saule13-ARXIV, Liu:2013:ESM:2464996.2465013, cramer2012openmp}.
 
@@
 $$
+
 where $f$ is the source vector of function values and $g$ is the
@@
 three-dimensional flows. These numbers are similar to what is used in
-finite-element codes.
+finite-element codes. Notice that the non zero elements of the matrix
+correspond to a relation of nearest neighbor. This relationship is not
+necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} which
+means that the matrix sparsity pattern might not be symmetric either.
+
+\begin{figure}[tbh]
+  \centering
+  \includegraphics[width=\linewidth]{figures/rbf_stencils.pdf}
+  \caption{RBFFD Stencils. Each node of the stencil is connected to
+    $n_z-1$ stencil nodes in addition to itself. In the figure, node
+    $A$ is connected to $B$, but $B$ is {\em not\/} connected to
+    $A$. Thus adjacency graph of $A$ is not-symmetric.}
+  \label{fig:rbf_stencils}
+\end{figure}
 
@@
 
-\def\wide{2.5in}
-\newfig{figures/rbf_stencils.pdf}{\wide}{RBFFD Stencils.
-Each node of the stencil is connected to $n_z-1$ stencil nodes in addition to itself.
-In the figure, node $A$ is connected to $B$, but $B$ is {\em not\/} connected to $A$.
-Thus adjacency graph of $A$ is not-symmetric.
-This must be corrected when reducing the bandwidth of $A$ with a Cuthill\-McKee algorithm, which assumes symmetry.}{fig:rbf_stencils}
-
-\newfig{figures/matrix_structure.pdf}{\wide}{Matrix Structure.}{fig:mat_struct}
+\newfig{figures/matrix_structure.pdf}{\linewidth}{Matrix Structure.}{fig:mat_struct}
 

### 2013-10-16 ieee_paper.tex
matrix plots with annotation.

--- 2013-10-16 ieee_paper.tex
@@
 
-In this paper, we propose yet a new idea that is applicable to radial
-basis functions (RBFs). Their strength is the ability to randomly
-distribute points across complex physical domains, and have an
-implementation that is independent of dimensionality. RBFs approximate
-a function $f(\xvec)\subset \mathbb{R}^d$ sampled at a set of $N$
-distinct point locations, $x_j$, by linearly combining translates of a
-single radially symmetric function $\phi(r)$, where $r =
-\|\xvec-\xvec_{j}\|$ denotes the Euclidean distance (e.g., in 2-D
-$\sqrt{(x-x_j)^2+(y-y_j)^2}$) between where the function is evaluated
-$\xvec$ and where the RBF is centered $\xvec_{j}$. That is, the
-interpolant is $s(\xvec) = \sum_{j=1}^{N} w_j
-\phi_i(\|\xvec-\xvec_{j}\|)$. The weights $w_j$ are obtained by
-inverting the system
+In this paper, we propose yet a new idea that is applicable to radial basis functions (RBFs).
+Their strength is the ability to randomly distribute points across complex physical domains, and have an implementation that is independent of dimensionality.
+RBFs approximate a function $f(\xvec)\subset \mathbb{R}^d$ sampled at a set of $N$ distinct point locations, $x_j$, by linearly combining translates of a single radially symmetric function $\phi(r)$, where  $r = \|\xvec-\xvec_{j}\|$ denotes the Euclidean distance (e.g., in 2-D $\sqrt{(x-x_j)^2+(y-y_j)^2}$) between where the function is evaluated $\xvec$ and where the RBF is centered $\xvec_{j}$.
+That is, the interpolant is $s(\xvec) =  \sum_{j=1}^{N} w_j \phi_i(\|\xvec-\xvec_{j}\|)$.
+The weights $w_j$ are obtained by inverting the system
 \begin{equation}
@@
 
-The RBF differentiation matrix, $D_N$, is derived by applying the
-desired analytic derivative operator $L$ to the RBF interpolant
-$s(\xvec)$ above and evaluating it at the point locations. For very
-large problems, this is a computationally expensive since the matrix
-in (\ref{eq:rbf}) is full and inversion requires O$(N^3)$
-operations. To alleviate the cost of this global approach (i.e. using
-every node in the domain to calculate the derivative at a given node
-$\xvec_i$), RBF-generated finite differences (RBF-FD) have been
-derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBF-FD use only
-a local set of the $n_z-1$ nearest neighbors to the point $\xvec_i$ to
-approximate the derivative. In other words,
-$Lf(\xvec_i)=\sum_{j=1}^{n_z}a_jf(\xvec_j)$. The differentiation
-weights, $a_j$, are calculated by enforcing that this linear
-combination should be exact for RBFs,
-$\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the
-node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences
-(FD) would enforce that it be exact for polynomials instead). Similar
-to FD, as the stencil size $n_z$ increases so does the order of the
-method.
+The RBF differentiation matrix, $D_N$, is derived by applying the desired analytic derivative operator $L$ to the RBF interpolant $s(\xvec)$ above and evaluating it at the point locations.
+For very large problems, this is a computationally expensive since the matrix in (\ref{eq:rbf}) is full and inversion requires O$(N^3)$ operations.
+To alleviate the cost of this global approach (i.e.
+using every node in the domain to calculate the derivative at a given node $\xvec_i$), RBF-generated finite differences (RBF-FD) have been derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}.
+RBF-FD use only a local set of the $n_z-1$ nearest neighbors to the point $\xvec_i$ to approximate the derivative.
+In other words, $Lf(\xvec_i)=\sum_{j=1}^{n_z}a_jf(\xvec_j)$.
+The differentiation weights, $a_j$, are calculated by enforcing that this linear combination should be exact for RBFs, $\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences (FD) would enforce that it be exact for polynomials instead).
+Similar to FD, as the stencil size $n_z$ increases so does the order of the method.
 
-For a total of $N$ points, there will be $N$ linear systems to solve,
-each of size $n_z \times n_z$. Each linear solve produces a row of the
-RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \ times N$
-matrix with $n_zN$ nonzero entries. To evaluate the derivative at all
-points in the domain, takes the form
+For a total of $N$ points, there will be $N$ linear systems to solve, each of size $n_z \times n_z$.
+Each linear solve produces a row of the RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \ times N$ matrix with $n_zN$ nonzero entries.
+To evaluate the derivative at all points in the domain, takes the form
 $$
@@
 $$
-where $f$ is the source vector of function values and $g$ is the
-resulting vector of derivative values. The computation of a single
-derivative has been reduced to a SpMV, where each row has $n_z$
-nonzeros. Since the sparsity of each row is constant, ELLPACK~\cite{}
-becomes the most appropriate matrix compression scheme\cite{}. In
-practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
-three-dimensional flows. These numbers are similar to what is used in
-finite-element codes.
+where $f$ is the source vector of function values and $g$ is the resulting vector of derivative values.
+The computation of a single derivative has been reduced to a SpMV, where each row has $n_z$ nonzeros.
+Since the sparsity of each row is constant, ELLPACK~\cite{} becomes the most appropriate matrix compression scheme\cite{}.
+In practice, $n_z=32$ in two-dimensional flows and 64 or 100 for three-dimensional flows.
+These numbers are similar to what is used in finite-element codes.
 
-Many problems in fluid dynamics and in the geosciences require the
-solution to transport equations of the form
+Many problems in fluid dynamics and in the geosciences require the solution to transport equations of the form
 $$
@@
 $$
-where $Q$ is a vector of unknowns (3 components of velocity, pressure,
-temperature). For example, in solving a system of equations, it is
-often necessary to compute the $x$ derivative of multiple functions,
-typically four for the Euler equations, or five for the Navier-Stokes
-equations. Multiple right-hand sides transform a SpMV into a SpMM
-(Sparse Matrix/dense Matrix multiplication), which improves register
-utilization and decreases cache misses by vectorizing over the
-multiple source vectors. Further improvements are possible by
-recognizing that different derivative matrices (x,y,z and Laplacian
-for example), have the same sparsity distribution; only the weight
-change.  Thus, alternative to computing a derivative of multiple
-functions, we can calculate multiple derivatives of a single
-functions. The increased memory bandwidth due to an increase in the
-number of derivative matrices is offset by better cache utilization,
-leading to an overall benefit.
+where $Q$ is a vector of unknowns (3 components of velocity,
+pressure, temperature). For example, in solving a system of equations, it is often necessary to compute the $x$
+derivative of multiple functions, typically four for the Euler
+equations, or five for the Navier-Stokes equations. Multiple
+right-hand sides transform a SpMV into a SpMM (Sparse Matrix/dense
+Matrix multiplication), which improves register utilization and
+decreases cache misses by vectorizing over the multiple source
+vectors. Further improvements are possible by recognizing that
+different derivative matrices (x,y,z and Laplacian for example), have
+the same sparsity distribution; only the weight change.  Thus,
+alternative to computing a derivative of multiple functions, we can
+calculate multiple derivatives of a single functions. The increased
+memory bandwidth due to an increase in the number of derivative
+matrices is offset by better cache utilization, leading to an overall
+benefit.
 
@@
 byte transfered. In the worst case and in the best case, we have
-$$I_b = \frac{O}{b_T} = \frac{2 n_v n_m n_z}{n_z (b_i + b_x n_m) + n_v b_x (1 + n_m)}$$
-$$I_w = \frac{O}{B_T} = \frac{2 n_v n_m n_z}{n_v n_m b_x + n_z \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)}$$
+$$I_b = \frac{O}{b_T} = \frac{2 n_v n_m}{ (b_i + b_x n_m) + n_v n_m b_x n_z^{-1} + n_v b_x n_z^{-1} }$$
+%$$I_w = \frac{O}{B_T} = \frac{2 n_v n_m}{n_v n_m b_x n_z^{-1} + \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)}$$
+$$I_w = \frac{O}{B_T} = \frac{2 n_v n_m}{(b_i+b_x n_m) + n_v n_m b_x n_z^{-1} + cl \ceil{\frac{n_vb_x}{cl}} }$$
 
@@
       \label{fig:supercompact}
-      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.png}}%
+      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.pdf}}%
     \subfigure[Compact.]{%
       \label{fig:compact}
-      \includegraphics[width=\ww]{figures/compact_matrix-crop.png}}%
+      \includegraphics[width=\ww]{figures/compact_matrix-crop.pdf}}%
     \subfigure[Random.]{%
       \label{fig:random}
-      \includegraphics[width=\ww]{figures/random_matrix-crop.png}}%
+      \includegraphics[width=\ww]{figures/random_matrix-crop.pdf}}%
     \subfigure[2D, no RCM.]{%
       \label{fig:rbf2dnorcm}
-      \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.png}}%
+      \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.pdf}}%
     \subfigure[2D, RCM.]{%
       \label{fig:rbf2drcm}
-      \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
+      \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.pdf}} 
     \subfigure[3D, RCM.]{%
       \label{fig:rbf3dnorcm} 
-      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
+      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.pdf}}%
     \subfigure[3D, RCM.]{%
       \label{fig:rbf3drcm}
-      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.png}}%
+      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.pdf}}%
   \end{center}

### 2013-10-16 ieee_paper.tex
more citations

--- 2013-10-16 ieee_paper.tex
@@
 
-\cite{Im01, Saad94sparskit} \cite{DBLP:journals/ijhpca/ShantharamCR11,
+\cite{DBLP:journals/ijhpca/ShantharamCR11,
   DBLP:journals/corr/abs-1101-0091 }
@@
 captured the interest of many researcher; including the development of
-various implemetation for multi-core
+various implementations for multi-core
 CPUs~\cite{Buluc2009SPAA,Williams07} and
@@
 by four matrices with identical sparsity patterns leading to the
-simultaneous execution of 16 SpMVs at a time.
+simultaneous execution of 16 SpMVs at a time. Using multiple vectors
+at a time has been investigated before to densify the
+computations~\cite{Im01} but this is the first time that the
+densification comes from adding both vectors and matrices.
 

### 2013-10-16 ieee_paper.tex
reorderng plots

@@
 \centering
-\subfigure[No cache effects.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
 \subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}} %\label{fig:ratio_best}}
-\subfigure[worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
+\subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
+\subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
 \caption{Ratio of flops to bytes; (a) simple count of flops and bytes that require transfer from memory;(b) best and (c) worst

### 2013-10-16 ieee_paper.tex
more citation

@@
 \section{Introduction}
-\cite{Stock12-TACC} is all dense. \cite{
-  Molka:2009:MPC:1636712.1637764} is about memory modeling not
-directly spmv.
-
-\cite{DBLP:journals/ijhpca/ShantharamCR11,
-  DBLP:journals/corr/abs-1101-0091 }
-
 The multiplication of sparse matrix by a dense vector (SpMV) is an
-important kernel in many fields from scientific computing~\cite{} to
-graph analysis (recommender systems~\cite{Brin98}, graph
+important kernel in many fields from scientific computing~\cite{} \todo{Do you have a good citation for that?}
+to graph analysis (recommender systems~\cite{Brin98}, graph
 drawing~\cite{Koren05}). So improving the performance of SpMV has
@@
 various implementations for multi-core
-CPUs~\cite{Buluc2009SPAA,Williams07} and
-GPUs~\cite{Bell08, conf/ipps/KreutzerHWFBB12,
+CPUs~\cite{Buluc2009SPAA,Williams07} and GPUs~\cite{Bell08,
+  conf/ipps/KreutzerHWFBB12,
   journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
@@
 important classes of applications such as graph
-recommendation~\cite{}, eigensolving~\cite{} and the computation of
-derivative of Radial Basis Functions(RBF)~\cite{} can use multiple
+recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of
+derivative of Radial Basis Functions(RBF)~\cite{} \todo{a good citation?} can use multiple
 SpMVs simultaneously. In this paper in particular, we investigate the

### 2013-10-16 ieee_paper.tex
intro

@@
 the operation are conducted using memory location that are irregular
-and often unpredictible. That make that the kernels are mostly memory
+and often unpredictible. That makes the kernels mostly memory
 bound and there is a significant instruction overhead per flop.
@@
 with virtually no overhead in all cases thanks to compressed
-representation~\cite{Buluc11}). there is about 8 bytes of the matrix to move
+representation~\cite{Buluc11}). There is about 8 bytes of the matrix to move
 in per non zero (assuming single precision); each nonzero requires two
@@
 
-Fortunately that fate is not inevitable. One solution would be to pair
-multiple component of an application to schedule a more instruction
-intensive kernel simultaneously with SpMV, relying on some hardware
-threading capabilities, such as HyperThreading, to reduce the cycle
-wastage. However most ot the applications that use SpMV do not
-typically have an instruction intensive kernel to run simultaneously.
+Fortunately that fate is not inevitable. One solution would be to
+schedule a more instruction intensive kernel simultaneously with the
+execution of SpMV, relying on some hardware threading capabilities,
+such as HyperThreading, to reduce the cycle wastage. However, most ot
+the applications that use SpMV do not typically have an instruction
+intensive kernel to run simultaneously.
 
@@
 case of the derivative of RBFs where four derivatives of four
-different function is expressed as the multiplication of four vectors
+different functions is expressed as the multiplication of four vectors
 by four matrices with identical sparsity patterns leading to the
@@
 estimation of the instruction intensity of various form of the
-computations and we show that a 5- to 6- fold improvement can be
+computations and we show that a 6- fold improvement can be
 expected when computing the 16 multiplications simultaneously and
@@
 experimental result about the amount of bandwidth that can be achieved
-depending on how the spmv kernel is written and the actual performance
+depending on how the SpMV kernel is written and the actual performance
 of the various kernel on multiple classes of matrices some generated
 for analysis purpose and some extracted from and RBF application. A
-performance of xxx GFlop/s is achieved on real scenario. Concluding
+performance of 135 Gflop/s is achieved on real scenario. Concluding
 remarks and perspectives are provided in Section~\ref{sec:ccl}.

### 2013-10-16 ieee_paper.tex
section 2
sec 3 (in progress)

@@
 derivative has been reduced to a SpMV, where each row has $n_z$
-nonzeros. Since the sparsity of each row is constant, ELLPACK~\cite{}
-becomes the most appropriate matrix compression scheme\cite{}. In
+nonzeros. In
 practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
@@
 
+\todo{describe the data structure} Since the sparsity of each row is
+constant, ELLPACK becomes the most appropriate matrix
+compression scheme.
+
 \newfig{figures/matrix_structure.pdf}{\linewidth}{Matrix Structure.}{fig:mat_struct}
@@
 
-\begin{figure}
+\begin{figure}[bth]
   \begin{center}
@@
 
-Each vector in the problem is of dimension $n_r$ and each dimension
-tasks $b_x$ bytes. There are $n_v$ {\tt x} vectors and $n_v n_m$ {\tt
+Each vector in the problem is of dimension $n_r$ and each entry
+takes $b_x$ bytes. There are $n_v$ {\tt x} vectors and $n_v n_m$ {\tt
   y} vectors, which lead to the size of the {\tt x} and {\tt y} vectors:
@@
 
-If we assume an algorithm where the rows are processed one after
+If we assume an algorithm where the rows are processed the one after
 the other, the amount of memory written is precisely the
 size of the {\tt y} vector. (This assumption removes the possibility of
-cache partitioning techniques.) $$b_{wT} = s_y = n_v n_m b_x n_r$$
+blocking or cache partitioning techniques.) $$b_{wT} = s_y = n_v n_m b_x n_r$$
 
-The amount of data read from memory depends highly on both the algorithm's
-execution path, and on %algorithm execute, 
-how the matrix is structured. But in the best case
-both the matrix $A$ %is read a single time 
-and the source vector {\tt x} are read once
-%read a single time 
-from the main memory. (We assume that all the elements of $\tt x$ are involved
-in the SpMV.)  %is used is reasonnable since only oddly shaped matrix will
-%left a dimension of the vector unused.) 
-Thus, $$b_{rT} = s_M + s_x = n_r n_z
-(b_i + b_x n_m) + n_v b_x n_r$$
+The amount of data read from memory depends highly on both the
+algorithm's execution path, and on how the matrix is structured. But
+in the best case both the matrix $A$ and the source vector {\tt x} are
+read once from the main memory. (We assume that all the elements of
+$\tt x$ are involved in the SpMV.)  Thus, $$b_{rT} = s_M + s_x = n_r
+n_z (b_i + b_x n_m) + n_v b_x n_r$$
  $$b_T = b_{rT} + b_{wT} =  n_r n_z (b_i + b_x n_m) + n_v b_x n_r (1 + n_m)$$
@@
 {\tt x} vector or that the matrix is sufficiently well structured 
-causing no cache trashing. If using multiple core, this assumes that
+to cause no cache trashing. If using multiple cores, this assumes that
 no element of the {\tt x} vectors will be used by multiple
-cores. \cite{Saule12} showed that there is very little cache trashing
+cores. \cite{Saule13-ARXIV} showed that there is very little cache trashing
 in practice, but it showed that having elements of the vectors used by
@@
 total, there are as many transfers as the number of non-zeros in the
-matrix. Note however that most architectures cannot read memory a single 
-byte at a time. Instead, a minimum number of bytes,equal to the size o a cacheline, 
-are transfered at once. 
-%multiple bytes, with each chunk equal to the data do not allow to read the memory
-%one byte at a time but always in a set of multiple bytes. On most
-%cached architecture, that 
-%the minimum size corresponds to the size $cl$ of
-%a cacheline (typically 64 bytes). 
-When there are multiple
-vectors, each non-zero element uses $n_v$ consecutive entries. The
-worst case number of bytes read and transfered are 
+matrix. Note however that most architectures cannot read memory a
+single byte at a time. Instead, a minimum number of bytes, equal to
+the size of a cacheline $cl$, are transfered at once.  When there are
+multiple vectors, each non-zero element uses $n_v$ consecutive
+entries. The worst case number of bytes read and transfered are
 $$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}} $$ 
@@
 
-\begin{table}
-  \centering
-  \begin{tabular}{|l|r|r|r|r|}
-    \hline
-                           & v1m1  & v1m4  & v4m1   & v4m4  \\
-    \hline
-    Best Single Precision  & 36.36 & 58.18 & 133.33 & 213.33\\
-    Worst Single Precision &  4.15 & 14.20 &  16.55 &  55.81\\
-    Best Double Precision  & 24.00 & 32.21 &  85.71 & 117.07\\
-    Worst Double Precision &  3.93 & 11.88 &  15.58 &  46.15\\
-    \hline
-  \end{tabular}
-  \caption{Estimation of performance in GFlop/s at 150GB/s for a
-    matrix of $n_z = 32$}
-\end{table}
+%% \begin{table}
+%%   \centering
+%%   \begin{tabular}{|l|r|r|r|r|}
+%%     \hline
+%%                            & v1m1  & v1m4  & v4m1   & v4m4  \\
+%%     \hline
+%%     Best Single Precision  & 36.36 & 58.18 & 133.33 & 213.33\\
+%%     Worst Single Precision &  4.15 & 14.20 &  16.55 &  55.81\\
+%%     Best Double Precision  & 24.00 & 32.21 &  85.71 & 117.07\\
+%%     Worst Double Precision &  3.93 & 11.88 &  15.58 &  46.15\\
+%%     \hline
+%%   \end{tabular}
+%%   \caption{Estimation of performance in GFlop/s at 150GB/s for a
+%%     matrix of $n_z = 32$}
+%% \end{table}
 

### 2013-10-16 ieee_paper.tex
moving heat map in section 3
fixing caption of heat maps

@@
 
-In the special case $n_v=n_m=1$ in single precision
-$$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
-%$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
-$$I_w = \frac{n_z }{2  + (4+cl/2)  n_z} \approx \frac{1}{4+cl/2}$$
+%% In the special case $n_v=n_m=1$ in single precision
+%% $$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
+%% %$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
+%% $$I_w = \frac{n_z }{2  + (4+cl/2)  n_z} \approx \frac{1}{4+cl/2}$$
 
-In the special case $n_v=n_m=4$ in single precision
-$$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
-%$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
-$$I_w = \frac{32 n_z }{64 + (20+cl) n_z} \approx \frac{32}{84} = \frac{8}{21} $$
-The above formulas have been applied to MIC, which has a cache size $cl=64$ bytes.
+%% In the special case $n_v=n_m=4$ in single precision
+%% $$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
+%% %$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
+%% $$I_w = \frac{32 n_z }{64 + (20+cl) n_z} \approx \frac{32}{84} = \frac{8}{21} $$
+%% The above formulas have been applied to MIC, which has a cache size $cl=64$ bytes.
 
@@
 %% \end{table}
+
+Figure~\ref{fig:ratio_bytes_flops} presents the flop to byte ratios in
+the best and the worst case.
+
+\begin{figure*}
+  \centering
+  \subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
+  \subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
+  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}
+
+  \caption{Ratio of flops to bytes: in the best case in
+    \ref{fig:ratio_best} best; in the worst case when $cl=64$ in
+    \ref{fig:ratio_worst}; and in the worst case neglecting that the
+    memory transfers are by batch of size $cl$ in
+    \ref{fig:ratio_worst_nocache}}
+  \label{fig:ratio_bytes_flops}
+\end{figure*}
 
@@
 
-\begin{figure*}
-\centering
-\subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}} %\label{fig:ratio_best}}
-\subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
-\subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
-\caption{Ratio of flops to bytes; (a) simple count of flops and bytes that require transfer from memory;(b) best and (c) worst
-case taking into account that bringing one element of $x$ into memory brings along the remainder of the cache line.}
-\label{fig:ratio_bytes_flops}
-\end{figure*}
-
 Derivatives in an RBFFD formulation are expressed as a sparse

### 2013-10-16 ieee_paper.tex
section 3 complete?

@@
 
-Figure~\ref{fig:ratio_bytes_flops} presents the flop to byte ratios in
-the best and the worst case.
+Figure~\ref{fig:ratio_bytes_flops} presents the flop to byte ratios
+for single precision computation in the best and the worst case on a
+classical cache based architecture ($cl = 64$) and assuming $cl=1$. We
+can easily see the potential improvement in the computational
+intensity when the number of matrices of vectors increase. There is a
+significant difference between the best and the worst case: there is a
+$8$ fold difference in the 1 matrix 1 vector case but that gap closes
+with the increase in the number of vectors and matrices to a 4 fold
+difference in the 4 matrices and 4 vectors case and 1.6 fold in the 16
+matrices and 16 vectors. The ratios computed with $cl = 1$ are mostly
+similar to the the one with $cl=64$ when the number of vectors is
+large. But the differences are important when the number of vector is
+low: this highlight that accessing the memory per batch is the main
+problem faced when performing a standard SpMV computation.
 
@@
 
-  \caption{Ratio of flops to bytes: in the best case in
+  \caption{Ratio of flops to bytes in single precision: in the best case in
     \ref{fig:ratio_best} best; in the worst case when $cl=64$ in
     \ref{fig:ratio_worst}; and in the worst case neglecting that the
-    memory transfers are by batch of size $cl$ in
-    \ref{fig:ratio_worst_nocache}}
+    memory transfers are by batch (equivalent to $cl = 1$) in
+    \ref{fig:ratio_worst_nocache}.}
   \label{fig:ratio_bytes_flops}
 \end{figure*}
+
+Figure~\ref{fig:ratio_bytes_flops} shows how these ratios translate
+into actual performance. This figure provides projected best and worst
+Gflop/s achievable assuming the computation is memory bound and the
+architecture reaches 150Gflop/s. Notice that \cite{Saule13-ARXIV}
+showed a higher peak bandwidth, but we will show in
+Section~\ref{sec:expe} why 150Gflop/s is a better estimate of what one
+might achieve in these kernels. In single precision, the worse that
+one could achieve by using 4 vectors and 4 matrices is much higher
+than the best achievable using a classical SpMV computation. The best
+performance reachable is 210Gflop/s: almost 6 times higher than the
+peak of the classical SpMV case.
 
@@
 %
-  \subfigure[Relative performance to base case using one matrix and one
+  \subfigure[Relative performance to using one matrix and one
     vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}
 
-  \caption{Estimation of the reachable performance one can
-      achieve using a device with a memory to computational units of
-      150GB/s when varying the number of vectors and matrices}
+  \caption{Estimation of the reachable performance one can achieve
+    using a device with a memory to computational units of 150GB/s
+    when varying the number of vectors and matrices. \todo{we could
+      remove the speedup plot if we need space.}  \todo{can the legend
+      read "Single Precision (best)"} \todo{there is a missing
+      parenthesis in the title; alternatively remove title.}}
   \label{fig:perf_predict}
 \end{figure*}
-
-\todo{conclude on the evaluation using the figure/table. Note that we
-  base prediction on 150GB/s which is not the peak of saule12. Note
-  that if you are in best in v1m1 you might go to bad v4m4, but if you
-  are in bad v1m1, you will essentially stay there.}
 

### 2013-10-16 ieee_paper.tex
pass on section 1 complete

@@
 expected when computing the 16 multiplications simultaneously and
-reach a total of about 200 Gflop/s. This performance represents
+reach a total of about 210 Gflop/s. This performance represents
 approximatively 10\% of the available flop/s of a Knight Corner

### 2013-10-16 ieee_paper.tex
section 4 partial

@@
 controllers is limited by the ring network that connect the cores and
-the memory controller. Its precise bandwidth is unknown for this
-precise card, but it would be coehrent that it is between 200GB/s and
-250GB/s.
+the memory controller. Its precise bandwidth is unknown but it is
+believed to be between 200GB/s and 250GB/s.
 
@@
 vectorial instruction. The code is given for reference in
-Figure~\ref{code:mat_mul}. (We will show that an implementation
+Figure~\ref{code:mat_mul}. (We will show in Section~\ref{sec:expe} that an implementation
 relying on compiler vectorization does not lead to desirable
-performance.) The registers in the MIC architecture are 512 bits wide
+performance, which confirms the results of~\cite{Saule13-ARXIV}.) The registers in the MIC architecture are 512 bits wide
 and can store 16 floats or integers. (In comparison to OpenCL or CUDA,
@@
 able to load and format the non zero values $a,b,c,d$ and the vector
-entries $p,q,r,s$ in two vectorial register to be able to apply Fused
+entries $p,q,r,s$ in two vector registers to be able to apply Fused
 Multiply-Add on them, which is depicted in
@@
 and parallelism is achieved by giving blocks of rows to each core
-usign an OpenMP construct.
+using an OpenMP construct.
 
@@
 which are made of 4 elements of 32 bits. A swizzling operation allows
-to reorder the elements within each lane. While a permutation allow to
+to reorder the elements within each lane. While a permutation allows to
 change the orders of the lanes. Figure~\ref{fig:swizzling_permutation}
@@
 All the operations that loads data from the memory into a register
-allows to load 512 bits, so 16 floats at a time. A single non zero in
-the matrix will cause 16 operations to be performed, but it only uses
-4 floats from the matrix and 4 floats from the {\tt x} vector. So we
-need to load at once the data for 4 non zeroes in the matrix, so that
-we load 16 flaots from the matrix and 16 floats from the vectors at
-once. (The number of non zeroes on a row in our application is always
-a multiple of 4, if it was, we would need to add some explicit zeroes
-in the data structure to pad it.)
+allows to transfer 512 bits, so 16 floats at a time. A single non zero
+in the matrix will cause 16 operations to be performed, but it only
+uses 4 floats from the matrix and 4 floats from the {\tt x} vector. So
+we need to load at once the data for 4 non zeroes in the matrix,
+representing 16 floats from the matrix and 16 floats from the vectors
+at once. (The number of non zeroes on a row in our application is
+always a multiple of 4, if it was, we would need to add some explicit
+zeroes in the data structure to pad it.)
 

### 2013-10-17 ieee_paper.tex
I am confused. I possibly screwd things up.

--- 2013-10-16 ieee_paper.tex
@@
 \section{Introduction}
+\cite{Stock12-TACC} is all dense. \cite{
+  Molka:2009:MPC:1636712.1637764} is about memory modeling not
+directly spmv.
+
+\cite{DBLP:journals/ijhpca/ShantharamCR11,
+  DBLP:journals/corr/abs-1101-0091 }
+
 The multiplication of sparse matrix by a dense vector (SpMV) is an
-important kernel in many fields from scientific computing~\cite{} \todo{Do you have a good citation for that?}
-to graph analysis (recommender systems~\cite{Brin98}, graph
+important kernel in many fields from scientific computing~\cite{} to
+graph analysis (recommender systems~\cite{Brin98}, graph
 drawing~\cite{Koren05}). So improving the performance of SpMV has
@@
 various implementations for multi-core
-CPUs~\cite{Buluc2009SPAA,Williams07} and GPUs~\cite{Bell08,
-  conf/ipps/KreutzerHWFBB12,
+CPUs~\cite{Buluc2009SPAA,Williams07} and
+GPUs~\cite{Bell08, conf/ipps/KreutzerHWFBB12,
   journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
@@
 the operation are conducted using memory location that are irregular
-and often unpredictible. That makes the kernels mostly memory
+and often unpredictible. That make that the kernels are mostly memory
 bound and there is a significant instruction overhead per flop.
@@
 with virtually no overhead in all cases thanks to compressed
-representation~\cite{Buluc11}). There is about 8 bytes of the matrix to move
+representation~\cite{Buluc11}). there is about 8 bytes of the matrix to move
 in per non zero (assuming single precision); each nonzero requires two
@@
 
-Fortunately that fate is not inevitable. One solution would be to
-schedule a more instruction intensive kernel simultaneously with the
-execution of SpMV, relying on some hardware threading capabilities,
-such as HyperThreading, to reduce the cycle wastage. However, most ot
-the applications that use SpMV do not typically have an instruction
-intensive kernel to run simultaneously.
+Fortunately that fate is not inevitable. One solution would be to pair
+multiple component of an application to schedule a more instruction
+intensive kernel simultaneously with SpMV, relying on some hardware
+threading capabilities, such as HyperThreading, to reduce the cycle
+wastage. However most ot the applications that use SpMV do not
+typically have an instruction intensive kernel to run simultaneously.
 
@@
 important classes of applications such as graph
-recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of
-derivative of Radial Basis Functions(RBF)~\cite{} \todo{a good citation?} can use multiple
+recommendation~\cite{}, eigensolving~\cite{} and the computation of
+derivative of Radial Basis Functions(RBF)~\cite{} can use multiple
 SpMVs simultaneously. In this paper in particular, we investigate the
 case of the derivative of RBFs where four derivatives of four
-different functions is expressed as the multiplication of four vectors
+different function is expressed as the multiplication of four vectors
 by four matrices with identical sparsity patterns leading to the
@@
 estimation of the instruction intensity of various form of the
-computations and we show that a 6- fold improvement can be
+computations and we show that a 5- to 6- fold improvement can be
 expected when computing the 16 multiplications simultaneously and
-reach a total of about 210 Gflop/s. This performance represents
+reach a total of about 200 Gflop/s. This performance represents
 approximatively 10\% of the available flop/s of a Knight Corner
@@
 experimental result about the amount of bandwidth that can be achieved
-depending on how the SpMV kernel is written and the actual performance
+depending on how the spmv kernel is written and the actual performance
 of the various kernel on multiple classes of matrices some generated
 for analysis purpose and some extracted from and RBF application. A
-performance of 135 Gflop/s is achieved on real scenario. Concluding
+performance of xxx GFlop/s is achieved on real scenario. Concluding
 remarks and perspectives are provided in Section~\ref{sec:ccl}.
@@
 derivative has been reduced to a SpMV, where each row has $n_z$
-nonzeros. In
+nonzeros. Since the sparsity of each row is constant, ELLPACK~\cite{}
+becomes the most appropriate matrix compression scheme\cite{}. In
 practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
@@
 
-\todo{describe the data structure} Since the sparsity of each row is
-constant, ELLPACK becomes the most appropriate matrix
-compression scheme.
-
 \newfig{figures/matrix_structure.pdf}{\linewidth}{Matrix Structure.}{fig:mat_struct}
@@
 
-\begin{figure}[bth]
+\begin{figure}
   \begin{center}
@@
 
-Each vector in the problem is of dimension $n_r$ and each entry
-takes $b_x$ bytes. There are $n_v$ {\tt x} vectors and $n_v n_m$ {\tt
+Each vector in the problem is of dimension $n_r$ and each dimension
+tasks $b_x$ bytes. There are $n_v$ {\tt x} vectors and $n_v n_m$ {\tt
   y} vectors, which lead to the size of the {\tt x} and {\tt y} vectors:
@@
 
-If we assume an algorithm where the rows are processed the one after
+If we assume an algorithm where the rows are processed one after
 the other, the amount of memory written is precisely the
 size of the {\tt y} vector. (This assumption removes the possibility of
-blocking or cache partitioning techniques.) $$b_{wT} = s_y = n_v n_m b_x n_r$$
+cache partitioning techniques.) $$b_{wT} = s_y = n_v n_m b_x n_r$$
 
-The amount of data read from memory depends highly on both the
-algorithm's execution path, and on how the matrix is structured. But
-in the best case both the matrix $A$ and the source vector {\tt x} are
-read once from the main memory. (We assume that all the elements of
-$\tt x$ are involved in the SpMV.)  Thus, $$b_{rT} = s_M + s_x = n_r
-n_z (b_i + b_x n_m) + n_v b_x n_r$$
+The amount of data read from memory depends highly on both the algorithm's
+execution path, and on %algorithm execute, 
+how the matrix is structured. But in the best case
+both the matrix $A$ %is read a single time 
+and the source vector {\tt x} are read once
+%read a single time 
+from the main memory. (We assume that all the elements of $\tt x$ are involved
+in the SpMV.)  %is used is reasonnable since only oddly shaped matrix will
+%left a dimension of the vector unused.) 
+Thus, $$b_{rT} = s_M + s_x = n_r n_z
+(b_i + b_x n_m) + n_v b_x n_r$$
  $$b_T = b_{rT} + b_{wT} =  n_r n_z (b_i + b_x n_m) + n_v b_x n_r (1 + n_m)$$
@@
 {\tt x} vector or that the matrix is sufficiently well structured 
-to cause no cache trashing. If using multiple cores, this assumes that
+causing no cache trashing. If using multiple core, this assumes that
 no element of the {\tt x} vectors will be used by multiple
-cores. \cite{Saule13-ARXIV} showed that there is very little cache trashing
+cores. \cite{Saule12} showed that there is very little cache trashing
 in practice, but it showed that having elements of the vectors used by
@@
 total, there are as many transfers as the number of non-zeros in the
-matrix. Note however that most architectures cannot read memory a
-single byte at a time. Instead, a minimum number of bytes, equal to
-the size of a cacheline $cl$, are transfered at once.  When there are
-multiple vectors, each non-zero element uses $n_v$ consecutive
-entries. The worst case number of bytes read and transfered are
+matrix. Note however that most architectures cannot read memory a single 
+byte at a time. Instead, a minimum number of bytes,equal to the size o a cacheline, 
+are transfered at once. 
+%multiple bytes, with each chunk equal to the data do not allow to read the memory
+%one byte at a time but always in a set of multiple bytes. On most
+%cached architecture, that 
+%the minimum size corresponds to the size $cl$ of
+%a cacheline (typically 64 bytes). 
+When there are multiple
+vectors, each non-zero element uses $n_v$ consecutive entries. The
+worst case number of bytes read and transfered are 
 $$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}} $$ 
@@
 
-%% In the special case $n_v=n_m=1$ in single precision
-%% $$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
-%% %$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
-%% $$I_w = \frac{n_z }{2  + (4+cl/2)  n_z} \approx \frac{1}{4+cl/2}$$
+In the special case $n_v=n_m=1$ in single precision
+$$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
+%$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
+$$I_w = \frac{n_z }{2  + (4+cl/2)  n_z} \approx \frac{1}{4+cl/2}$$
 
-%% In the special case $n_v=n_m=4$ in single precision
-%% $$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
-%% %$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
-%% $$I_w = \frac{32 n_z }{64 + (20+cl) n_z} \approx \frac{32}{84} = \frac{8}{21} $$
-%% The above formulas have been applied to MIC, which has a cache size $cl=64$ bytes.
+In the special case $n_v=n_m=4$ in single precision
+$$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
+%$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
+$$I_w = \frac{32 n_z }{64 + (20+cl) n_z} \approx \frac{32}{84} = \frac{8}{21} $$
+The above formulas have been applied to MIC, which has a cache size $cl=64$ bytes.
 
-%% \begin{table}
-%%   \centering
-%%   \begin{tabular}{|l|r|r|r|r|}
-%%     \hline
-%%                            & v1m1  & v1m4  & v4m1   & v4m4  \\
-%%     \hline
-%%     Best Single Precision  & 36.36 & 58.18 & 133.33 & 213.33\\
-%%     Worst Single Precision &  4.15 & 14.20 &  16.55 &  55.81\\
-%%     Best Double Precision  & 24.00 & 32.21 &  85.71 & 117.07\\
-%%     Worst Double Precision &  3.93 & 11.88 &  15.58 &  46.15\\
-%%     \hline
-%%   \end{tabular}
-%%   \caption{Estimation of performance in GFlop/s at 150GB/s for a
-%%     matrix of $n_z = 32$}
-%% \end{table}
-
-Figure~\ref{fig:ratio_bytes_flops} presents the flop to byte ratios
-for single precision computation in the best and the worst case on a
-classical cache based architecture ($cl = 64$) and assuming $cl=1$. We
-can easily see the potential improvement in the computational
-intensity when the number of matrices of vectors increase. There is a
-significant difference between the best and the worst case: there is a
-$8$ fold difference in the 1 matrix 1 vector case but that gap closes
-with the increase in the number of vectors and matrices to a 4 fold
-difference in the 4 matrices and 4 vectors case and 1.6 fold in the 16
-matrices and 16 vectors. The ratios computed with $cl = 1$ are mostly
-similar to the the one with $cl=64$ when the number of vectors is
-large. But the differences are important when the number of vector is
-low: this highlight that accessing the memory per batch is the main
-problem faced when performing a standard SpMV computation.
-
-\begin{figure*}
+\begin{table}
   \centering
-  \subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
-  \subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
-  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}
-
-  \caption{Ratio of flops to bytes in single precision: in the best case in
-    \ref{fig:ratio_best} best; in the worst case when $cl=64$ in
-    \ref{fig:ratio_worst}; and in the worst case neglecting that the
-    memory transfers are by batch (equivalent to $cl = 1$) in
-    \ref{fig:ratio_worst_nocache}.}
-  \label{fig:ratio_bytes_flops}
-\end{figure*}
-
-Figure~\ref{fig:ratio_bytes_flops} shows how these ratios translate
-into actual performance. This figure provides projected best and worst
-Gflop/s achievable assuming the computation is memory bound and the
-architecture reaches 150Gflop/s. Notice that \cite{Saule13-ARXIV}
-showed a higher peak bandwidth, but we will show in
-Section~\ref{sec:expe} why 150Gflop/s is a better estimate of what one
-might achieve in these kernels. In single precision, the worse that
-one could achieve by using 4 vectors and 4 matrices is much higher
-than the best achievable using a classical SpMV computation. The best
-performance reachable is 210Gflop/s: almost 6 times higher than the
-peak of the classical SpMV case.
+  \begin{tabular}{|l|r|r|r|r|}
+    \hline
+                           & v1m1  & v1m4  & v4m1   & v4m4  \\
+    \hline
+    Best Single Precision  & 36.36 & 58.18 & 133.33 & 213.33\\
+    Worst Single Precision &  4.15 & 14.20 &  16.55 &  55.81\\
+    Best Double Precision  & 24.00 & 32.21 &  85.71 & 117.07\\
+    Worst Double Precision &  3.93 & 11.88 &  15.58 &  46.15\\
+    \hline
+  \end{tabular}
+  \caption{Estimation of performance in GFlop/s at 150GB/s for a
+    matrix of $n_z = 32$}
+\end{table}
 
@@
 %
-  \subfigure[Relative performance to using one matrix and one
+  \subfigure[Relative performance to base case using one matrix and one
     vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}
 
-  \caption{Estimation of the reachable performance one can achieve
-    using a device with a memory to computational units of 150GB/s
-    when varying the number of vectors and matrices. \todo{we could
-      remove the speedup plot if we need space.}  \todo{can the legend
-      read "Single Precision (best)"} \todo{there is a missing
-      parenthesis in the title; alternatively remove title.}}
+  \caption{Estimation of the reachable performance one can
+      achieve using a device with a memory to computational units of
+      150GB/s when varying the number of vectors and matrices}
   \label{fig:perf_predict}
 \end{figure*}
+
+\todo{conclude on the evaluation using the figure/table. Note that we
+  base prediction on 150GB/s which is not the peak of saule12. Note
+  that if you are in best in v1m1 you might go to bad v4m4, but if you
+  are in bad v1m1, you will essentially stay there.}
 
@@
 %
-  \subfigure[ xxx ] 
-  {\includegraphics[width=.40\linewidth]{figures/test1_gather.pdf}\label{fig:band_gather}} \\
-
-  \subfigure[The greater
-    speed of the cpp version is obtained throught the use of
-    \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}. All memory
-    is aligned on 64
-    bytes.]{\includegraphics[width=.40\linewidth]{figures/test3_gather.pdf}\label{fig:band_gather_ivdep}}
+  \hspace{.3in}
 %
@@
 
+\begin{figure*}
+\centering
+  \subfigure[ xxx ] 
+  {\includegraphics[width=.20\linewidth]{figures/test1_gather.pdf}\label{fig:band_gather}} \\
+
+  \subfigure[The greater
+    speed of the cpp version is obtained throught the use of
+    \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}. All memory
+    is aligned on 64
+    bytes.]{\includegraphics[width=.20\linewidth]{figures/test3_gather.pdf}\label{fig:band_gather_ivdep}}
+\end{figure*}
+%
+
 \subsection{Computations}
@@
     {\includegraphics[width=.40\linewidth]{figures/mic_performance_nb_threads.pdf}\label{fig:perf_mic}}
+%
+\hspace{.4in}
 %
@@
 
-\begin{figure}
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
+%
+%  \caption{Performance on a MIC of derivative computation for RBFFD
+%    with 884,736 nodes distributed quasi-randomly in a cube. Single
+%    derivative of a single function (base case, triangle) and four
+%    derivatives of four functions (multi case, circles). No bandwidth
+%    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
+%\end{figure}
+%
+\begin{figure*}
 \centering
-  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
-
-  \caption{Performance on a MIC of derivative computation for RBFFD
-    with 884,736 nodes distributed quasi-randomly in a cube. Single
-    derivative of a single function (base case, triangle) and four
-    derivatives of four functions (multi case, circles). No bandwidth
-    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
-\end{figure}
+\subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}} %\label{fig:ratio_best}}
+\subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
+\subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
+\caption{Ratio of flops to bytes; (a) simple count of flops and bytes that require transfer from memory;(b) best and (c) worst
+case taking into account that bringing one element of $x$ into memory brings along the remainder of the cache line.}
+\label{fig:ratio_bytes_flops}
+\end{figure*}
 

### 2013-10-17 ieee_paper.tex

--- 2013-10-16 ieee_paper.tex
@@
 
-\begin{figure}
+%<<<<<<< HEAD
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
+%
+%  \caption{Performance on a MIC of derivative computation for RBFFD
+%    with 884,736 nodes distributed quasi-randomly in a cube. Single
+%    derivative of a single function (base case, triangle) and four
+%    derivatives of four functions (multi case, circles). No bandwidth
+%    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
+%\end{figure}
+%
+\begin{figure*}
 \centering
-  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
-
-  \caption{Performance on a MIC of derivative computation for RBFFD
-    with 884,736 nodes distributed quasi-randomly in a cube. Single
-    derivative of a single function (base case, triangle) and four
-    derivatives of four functions (multi case, circles). No bandwidth
-    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
-\end{figure}
+\subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}} %\label{fig:ratio_best}}
+\subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
+\subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
+\caption{Ratio of flops to bytes; (a) simple count of flops and bytes that require transfer from memory;(b) best and (c) worst
+case taking into account that bringing one element of $x$ into memory brings along the remainder of the cache line.}
+\label{fig:ratio_bytes_flops}
+\end{figure*}
+%=======
+%\begin{figure}
+%  \centering
+%  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
+%
+%  \caption{Performance on a MIC of derivative computation for RBFFD
+%    with 884,736 nodes distributed quasi-randomly in a cube. Single
+%    derivative of a single function (base case, triangle) and four
+%    derivatives of four functions (multi case, circles). No bandwidth
+%    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
+%\end{figure}
+%
+%>>>>>>> 619115cea6da9466f09a3b01456eacf7991606e2
 

### 2013-10-17 ieee_paper.tex
Updated figures.

@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
   \centering
-  \subfigure[Supercompact]{\includegraphics[width=.48\linewidth]{figures/supercompact_max_perf.pdf}}
-  \subfigure[Compact]{\includegraphics[width=.48\linewidth]{figures/compact_max_perf.pdf}} \\
-  \subfigure[Random]{\includegraphics[width=.48\linewidth]{figures/random_max_perf.pdf}}
-  \subfigure[RBF 3D]{\includegraphics[width=.48\linewidth]{figures/rbf_max_perf_run1.pdf}}
+  \subfigure[Supercompact and random.]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}}
+  \subfigure[Compact and RBF-FD]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}}
+
+  %\subfigure[Supercompact]{\includegraphics[width=.48\linewidth]{figures/supercompact_max_perf.pdf}}
+  %\subfigure[Compact]{\includegraphics[width=.48\linewidth]{figures/compact_max_perf.pdf}} \\
+  %\subfigure[Random]{\includegraphics[width=.48\linewidth]{figures/random_max_perf.pdf}}
+  %\subfigure[RBF 3D]{\includegraphics[width=.48\linewidth]{figures/rbf_max_perf_run1.pdf}}
   \caption{Performance of "best code" accross instance types. \todo{Why stop at 1.6M
@@
       it is}}
-\end{figure}
+\end{figure*}
 

### 2013-10-17 gflops_speedup.py
Updated figures

@@ -115,19 +115,21 @@ bar_width = 0.35/2
 print np.shape(index), np.shape(ratios_d_worst), np.shape(ratios_d_best)
 print np.shape(index), np.shape(ratios_s_worst), np.shape(ratios_s_best)
 
-rects1 = plt.bar(index, ratios_s_worst, bar_width, alpha=1, color='b', label='single worst case')
-rects1 = plt.bar(index+bar_width, ratios_s_best, bar_width, alpha=.5, color='b', label='single best case')
+rects1 = plt.bar(index, ratios_s_worst, bar_width, alpha=1, color='b', label='single precision worst case')
+rects1 = plt.bar(index+bar_width, ratios_s_best, bar_width, alpha=.5, color='b', label='single precision best case')
 print ratios_s_best
 
-rects2 = plt.bar(index+2*bar_width, ratios_d_worst, bar_width, alpha=1, color='r', label='double worst case')
-rects2 = plt.bar(index+3*bar_width, ratios_d_best, bar_width, alpha=.5, color='r', label='double best case')
+rects2 = plt.bar(index+2*bar_width, ratios_d_worst, bar_width, alpha=1, color='r', label='double precision worst case')
+rects2 = plt.bar(index+3*bar_width, ratios_d_best, bar_width, alpha=.5, color='r', label='double precision best case')
 
-plt.xlabel('Vector/Matrix sizes')
-plt.ylabel('Peak Gflops')
-plt.title('Best possible performance (max application bandwidth: 150 Gbytes/sec')
+plt.xlabel('Vector/Matrix sizes', fontsize=14)
+plt.ylabel('Peak Gflops', fontsize=14)
+#plt.title('Best possible performance (max application bandwidth: 150 Gbytes/sec')
 plt.xticks(index + bar_width, labels)
 plt.legend(loc=2)  # upper left
 plt.grid(True)
+plt.rc('xtick', labelsize=16)
+plt.rc('ytick', labelsize=16)
 
 plt.tight_layout()
 plt.savefig.format = "pdf"
@@ -150,10 +152,12 @@ rects1 = plt.bar(index+bar_width, speedup_s_best, bar_width, alpha=.5, color='b'
 rects2 = plt.bar(index+2*bar_width, speedup_d_worst, bar_width, alpha=1, color='r', label='double worst double case')
[…]

### 2013-10-17 ieee_paper.tex
sec 2 notations

@@
 
-\todo{describe the data structure} Since the sparsity of each row is
-constant, ELLPACK becomes the most appropriate matrix
-compression scheme.
+In term of memory layout, we choose a format similar to the ELLPACK
+format to store the matrix since the number of non zero per row is
+constant throughout the matrix. The matrix is given in two arrays. The
+first one is the classical {\tt col\_id} array that describes the
+sparsity structure of the matrix. For each row it consists of $n_z$
+consecutive entries which give in which column is the non zero. The
+other array is the {\tt data} array which gives the values of the non
+zero in the four matrices. The values of the matrices are interleaved
+so that the value $A^l_{col\_id[k]}$ of the $k$th non zero of the
+$l$th matrix is at index $(k-1)*4+l-1$. The vectors {\tt x} are also
+similarly interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
+consecutive in memory. The vector {\tt y} is similarly
+interleaved. This is schematically shown in Figure~\ref{fig:mat_struct}.
 
-\newfig{figures/matrix_structure.pdf}{\linewidth}{Matrix Structure.}{fig:mat_struct}
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/matrix_not.pdf}
+  \caption{Matrix Structure.}
+  \label{fig:mat_struct}
+\end{figure}
 

### 2013-10-17 ieee_paper.tex
Added 2 citations and mods to figures.

@@
 The multiplication of sparse matrix by a dense vector (SpMV) is an
-important kernel in many fields from scientific computing~\cite{} \todo{Do you have a good citation for that?}
+important kernel in many applied fields such as fluid dynamics 
+\cite{journals/tog/BolzFGS03} 
 to graph analysis (recommender systems~\cite{Brin98}, graph
@@
 recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of
-derivative of Radial Basis Functions(RBF)~\cite{} \todo{a good citation?} can use multiple
+derivatives of Radial Basis Functions(RBF)~\cite{FLBWSC12} can use multiple
 SpMVs simultaneously. In this paper in particular, we investigate the
@@
     when varying the number of vectors and matrices. \todo{we could
-      remove the speedup plot if we need space.}  \todo{can the legend
-      read "Single Precision (best)"} \todo{there is a missing
-      parenthesis in the title; alternatively remove title.}}
+      remove the speedup plot if we need space.}}
   \label{fig:perf_predict}
@@
 %
-\begin{figure*}
-\centering
-\subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}} %\label{fig:ratio_best}}
-\subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
-\subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
-\caption{Ratio of flops to bytes; (a) simple count of flops and bytes that require transfer from memory;(b) best and (c) worst
-case taking into account that bringing one element of $x$ into memory brings along the remainder of the cache line.}
-\label{fig:ratio_bytes_flops}
-\end{figure*}
+%\begin{figure*}
+%\centering
+%\subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}} %\label{fig:ratio_best}}
+%\subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
+%\subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
+%\caption{Ratio of flops to bytes; (a) simple count of flops and bytes that require transfer from memory;(b) best and (c) worst
+%case taking into account that bringing one element of $x$ into memory brings along the remainder of the cache line.}
+%\label{fig:ratio_bytes_flops}
+%\end{figure*}
 %=======

### 2013-10-17 code.tex
working on code presentation

\documentclass{article}

\usepackage{soul}
\usepackage{xspace}
\usepackage{color} 
\definecolor{darkgreen}{rgb}{0,0.5,0}
\usepackage[colorlinks=true,% 
  linkcolor=red,% 
  citecolor=darkgreen,%
  urlcolor=blue]{hyperref}

\usepackage{listings}

\begin{document}

\thispagestyle{empty}

\begin{figure}[h]
\begin{center}
\begin{verbatim}
const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
const __m512i four = _mm512_set4_epi32(4,4,4,4); 
int int_mask = 0x1111;
__mmask16 mask = _mm512_int2mask(int_mask);

#pragma omp parallel for
for (int r=0; r < nb_rows; r++) {
 __m512 accu = _mm512_setzero_ps();

 for (int n=0; n < nz; n+=4) {
   float* a = &dom.col_id_t[0]+n+nz*r;
   __m512i vect_i = _mm512_setzero_ps();
   vect_i = _mm512_mask_loadunpacklo_ps(vect_i, mask, a);
   vect_i = _mm512_mask_loadunpackhi_ps(vect_i, mask, a);
   vect_i = _mm512_swizzle_ps(vect_i, _MM_SWIZ_REG_AAAA);
   vect_i = _mm512_fmadd_epi32(vect_i, four, offsets);

   all_vect = _mm512_i32gather_ps(vect_i, dom.vec_vt, scale);

   mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));

   //perform first tensor multiplication
   correct_vect = permute(all_vect, _MM_PERM_AAAA);
   correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
   accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);

   //three more times with BBBB, CCCC, and DDDD
 }
 _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
\end{verbatim}
\end{center}
\end{figure}

\end{document}

### 2013-10-17 code.tex
code almost ready

@@
    //perform first tensor multiplication
-   correct_vect = permute(all_vect, _MM_PERM_AAAA);
-   correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
-   accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);
+   vect = permute(all_vect, _MM_PERM_AAAA);
+   mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
+   accu = _mm512_fmadd_ps(vect, mat, accu);
 

### 2013-10-18 ieee_paper.tex
working on code presentation
temp
code almost ready
final figure 8?
captioning in small touch up

@@
   %\subfigure[Source code}
-  \includegraphics[width=0.8\textwidth]{figures/source_code.pdf}%
-  \label{fig:source}
-  \caption{Source Code}
+%  \includegraphics[width=0.8\textwidth]{figures/source_code.pdf}%
+
+\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{code.pdf}}%
+\subfigure[Content of the vector registers]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
+
+\caption{Code snippet of the multiplication of 4 vectors with 4
+  matrices. When a line changes the content of a vector register its
+  content is shown on the right. The effects of swizzling and
+  permutations are highlighted using a color code.}
+\label{fig:code}
+
 \end{figure*}

### 2013-10-18 ieee_paper.tex
small touch up

@@
 
+\begin{figure*}
+  \centering
+\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{code.pdf}}%
+\subfigure[Content of the vector registers]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
+
+\caption{Code snippet of the multiplication of 4 vectors with 4
+  matrices. When a line changes the content of a vector register its
+  new content is shown on the right. The effects of swizzling and
+  permutations are highlighted using a color code.}
+\label{code:mat_mul}
+
+\end{figure*}
+
 \begin{figure}
@@
 
-\begin{figure*}
-  \centering
-  %\subfigure[Source code}
-%  \includegraphics[width=0.8\textwidth]{figures/source_code.pdf}%
-
-\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{code.pdf}}%
-\subfigure[Content of the vector registers]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
-
-\caption{Code snippet of the multiplication of 4 vectors with 4
-  matrices. When a line changes the content of a vector register its
-  content is shown on the right. The effects of swizzling and
-  permutations are highlighted using a color code.}
-\label{fig:code}
-
-\end{figure*}
 %\newfig{figures/source_code.pdf}{6in}{Source code}{fig:source}
@@
 \def\ww{.13\textwidth}
-\begin{figure*}
+\begin{figure*}[tbh]
   \begin{center}
@@
   %\subfigure[RBF 3D]{\includegraphics[width=.48\linewidth]{figures/rbf_max_perf_run1.pdf}}
-  \caption{Performance of "best code" accross instance types. \todo{Why stop at 1.6M
-      rows. Is that because of memory limitation? I guess that's what
-      it is}}
+  \caption{Performance of "best code" accross instance types.}
 \end{figure*}

### 2013-10-18 ieee_paper.tex
fixing typo in figure

@@
 
-As explained in Section~XXX, we expect to achieve a performance of
+As explained in Section~\ref{sec:model}, we expect to achieve a performance of
 about 210Gflop/s in single precision in the best case. We will focus
@@
 vectorial instruction. The code is given for reference in
-Figure~\ref{code:mat_mul}. (We will show in Section~\ref{sec:expe} that an implementation
-relying on compiler vectorization does not lead to desirable
-performance, which confirms the results of~\cite{Saule13-ARXIV}.) The registers in the MIC architecture are 512 bits wide
-and can store 16 floats or integers. (In comparison to OpenCL or CUDA,
-one can think of these registers as a warp.) The end goal is to be
-able to load and format the non zero values $a,b,c,d$ and the vector
-entries $p,q,r,s$ in two vector registers to be able to apply Fused
-Multiply-Add on them, which is depicted in
+Figure~\ref{code:mat_mul}. (We will show in Section~\ref{sec:expe}
+that an implementation relying on compiler vectorization does not lead
+to desirable performance, which confirms the results
+of~\cite{Saule13-ARXIV}.) The registers in the MIC architecture are
+512 bits wide and can store 16 floats or integers. (In comparison to
+OpenCL or CUDA, one can think of these registers as a warp.) The end
+goal is to be able to load and format the non zero values $a,b,c,d$
+and the vector entries $p,q,r,s$ in two vector registers to be able to
+apply Fused Multiply-Add on them, which is depicted in
 Figure~\ref{fig:tensor_product}. Bringing the data in the vector
@@
 The values from the non zero element in the matrix are the simplest to
-load. One can load the 16 floating point values for the 4 non zero of
-the matrix at once. The 4 floats coming from the four matrices for
-each non zero are naturally grouped within each lane of the SIMD
-vector. One can use a swizzle operation to broadcast the 4 elements of
-a single matrix to take the whole vector.
+load. One can load the 16 floating point values for 4 non zeroes of
+the matrix at once. The 4 floats coming from each matrix four 4
+consecutive non zeroes are naturally grouped within each lane of the
+SIMD vector. One can use a swizzle operation to broadcast the 4
+elements of a single matrix to take the whole vector (values
+$A_h^1,A_h^2,A_h^3,A_h^4$ are broadcasted in Figure~\ref{code:mat_mul}).
 
@@
 operation\footnote{Notice that there are two calls to unpack for the
-  LSB and MSB. Both are mandatory even according the documentation of
-  the hardware}. They replicated so that each fill a vector lane using
-a swizzle operation. Then each value is multiplied by 4 (because there
-are 4 vectors) and the offsets $(3,2,1,0)$ is added each lane to
-obtain the correct index of the elements of the {\tt x} vector within
-each lane. Then a {\tt gather} operation is performed bring the 16
-entries of the {\tt x} vector into the SIMD register. At this point we
-have a SIMD register where each lane is the four vector entries for a
-non zero element. Using a broadcast permutation, one can replicate an
-entry of the four vectors accross each lane.
-
-%\newfig{figures/source_code.pdf}{6in}{Source code}{fig:source}
-
-%\begin{figure}[h]
-%\begin{center}
-%\lstset{language=c++,basicstyle=\scriptsize,numbers=left,numbersep=5pt}
-%\begin{lstlisting}{float=h, Caption}
-%const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
-%const __m512i four = _mm512_set4_epi32(4,4,4,4); 
-%int int_mask = 0x1111;
-%__mmask16 mask = _mm512_int2mask(int_mask);
-%
-%#pragma omp parallel for
-%for (int r=0; r < nb_rows; r++) {
-%  __m512 accu = _mm512_setzero_ps();
-%
-%  for (int n=0; n < nz; n+=4) {
-%    float* a = &dom.col_id_t[0]+n+nz*r;
-%    __m512i vect_i = _mm512_setzero_ps();
-%    vect_i = _mm512_mask_loadunpacklo_ps(vect_i, mask, a);
-%    vect_i = _mm512_mask_loadunpackhi_ps(vect_i, mask, a);
-%    vect_i = _mm512_swizzle_ps(vect_i, _MM_SWIZ_REG_AAAA);
-%    vect_i = _mm512_fmadd_epi32(vect_i, four, offsets);
-%
-%    all_vect = _mm512_i32gather_ps(vect_i, dom.vec_vt, scale);
-%
-%    mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));
-%
-%    //perform first tensor multiplication
-%    correct_vect = permute(all_vect, _MM_PERM_AAAA);
-%    correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
-%    accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);
-%
-%    //three more times with BBBB, CCCC, and DDDD
-%  }
-%  _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
-%\end{lstlisting}
-%\end{center}
-%\end{figure}
-%
-
-%\begin{figure*}
-%\begin{verbatim}
-%const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
-%const __m512i four = _mm512_set4_epi32(4,4,4,4); 
-%int int_mask = 0x1111;
-%__mmask16 mask = _mm512_int2mask(int_mask);
-%
-%#pragma omp parallel for
-%for (int r=0; r < nb_rows; r++) {
-%  __m512 accu = _mm512_setzero_ps();
-%
-%  for (int n=0; n < nz; n+=4) {
-%    float* a = &dom.col_id_t[0]+n+nz*r;
-%    __m512i vect_i = _mm512_setzero_ps();
-%    vect_i = _mm512_mask_loadunpacklo_ps(vect_i, mask, a);
-%    vect_i = _mm512_mask_loadunpackhi_ps(vect_i, mask, a);
-%    vect_i = _mm512_swizzle_ps(vect_i, _MM_SWIZ_REG_AAAA);
-%    vect_i = _mm512_fmadd_epi32(vect_i, four, offsets);
-%
-%    all_vect = _mm512_i32gather_ps(vect_i, dom.vec_vt, scale);
-%
-%    mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));
-%
-%    //perform first tensor multiplication
-%    correct_vect = permute(all_vect, _MM_PERM_AAAA);
-%    correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
-%    accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);
-%
-%    //three more times with BBBB, CCCC, and DDDD
-%  }
-%  _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
-%} 
-%\end{verbatim}
-%\caption{mat mul code.}
-%\label{code:mat_mul}
-%\end{figure*}
+  LSB and MSB. Both are mandatory even if one is known not to be
+  useful according the documentation of the hardware}. (The column pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They replicated
+so that each fill a vector lane using a swizzle operation. Then each
+value is multiplied by 4 (because there are 4 vectors) and the offsets
+$(3,2,1,0)$ is added each lane to obtain the correct index of the
+elements of the {\tt x} vector within each lane. Then a {\tt gather}
+operation is performed bring the 16 entries of the {\tt x} vector into
+the SIMD register. At this point we have a SIMD register where each
+lane is the four vector entries for a non zero element. Using a
+broadcast permutation, one can replicate an entry of the four vectors
+accross each lane.
 

### 2013-10-18 ieee_paper.tex
alignment

@@
 SIMD vector. One can use a swizzle operation to broadcast the 4
-elements of a single matrix to take the whole vector (values
+elements of a single matrix to take the whole vector. (The values
 $A_h^1,A_h^2,A_h^3,A_h^4$ are broadcasted in Figure~\ref{code:mat_mul}).
@@
   LSB and MSB. Both are mandatory even if one is known not to be
-  useful according the documentation of the hardware}. (The column pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They replicated
-so that each fill a vector lane using a swizzle operation. Then each
-value is multiplied by 4 (because there are 4 vectors) and the offsets
-$(3,2,1,0)$ is added each lane to obtain the correct index of the
-elements of the {\tt x} vector within each lane. Then a {\tt gather}
-operation is performed bring the 16 entries of the {\tt x} vector into
-the SIMD register. At this point we have a SIMD register where each
-lane is the four vector entries for a non zero element. Using a
-broadcast permutation, one can replicate an entry of the four vectors
-accross each lane.
+  useful according the documentation of the hardware}. (The column
+pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They
+replicated so that each fill a vector lane using a swizzle
+operation. Then each value is multiplied by 4 (because there are 4
+vectors) and the offsets $(3,2,1,0)$ is added each lane to obtain the
+correct index of the elements of the {\tt x} vector within each lane
+using a Fused Multiply-Add operation. Then a {\tt gather} operation is
+performed bring the 16 entries of the {\tt x} vector into the SIMD
+register. At this point we have a SIMD register where each lane is the
+four vector entries for a non zero element. Using a broadcast
+permutation, one can replicate an entry of the four vectors accross
+each lane.
 

### 2013-10-18 ieee_paper.tex
more code comment

@@
 SIMD vector. One can use a swizzle operation to broadcast the 4
-elements of a single matrix to take the whole vector. (The values
-$A_h^1,A_h^2,A_h^3,A_h^4$ are broadcasted in Figure~\ref{code:mat_mul}).
+elements of a single matrix to take the whole vector. (The broadcast of 
+$A_h^1,A_h^2,A_h^3,A_h^4$ is shown in Figure~\ref{code:mat_mul}.)
 
@@
 permutation, one can replicate an entry of the four vectors accross
-each lane.
+each lane. (The broadcast of $x_m^1,x_m^2,x_m^3,x_m^4$ is shown in
+Figure~\ref{code:mat_mul}.)
+
+Now that $A_h^1,A_h^2,A_h^3,A_h^4$ and $x_m^1,x_m^2,x_m^3,x_m^4$ are
+properly laid out in the vectors, a partial value for $y_i$ can be
+computed. Once all the non zeroes of the row have been processed, the
+exact $y_i$ value can be sent to memory using a the No Read No Global
+Order hint to optimize traffic on the memory bus.
 

### 2013-10-18 code.tex
code touch up

@@
 
-   //perform first tensor multiplication
+   //perform first tensor multiplication with AAAA pattern
    vect = permute(all_vect, _MM_PERM_AAAA);
@@
 
-   //three more times with BBBB, CCCC, and DDDD
+   //three more times with BBBB, CCCC, and DDDD patterns
  }

### 2013-10-18 ieee_paper.tex
Small corrections.

@@
 representation~\cite{Buluc11}). There is about 8 bytes of the matrix to move
-in per non zero (assuming single precision); each nonzero requires two
+in per nonzero (assuming single precision); each nonzero requires two
 floating point operations leading to a flop-to-byte ratio of at most
@@
 three-dimensional flows. These numbers are similar to what is used in
-finite-element codes. Notice that the non zero elements of the matrix
+finite-element codes. Notice that the nonzero elements of the matrix
 correspond to a relation of nearest neighbor. This relationship is not
@@
 In term of memory layout, we choose a format similar to the ELLPACK
-format to store the matrix since the number of non zero per row is
+format to store the matrix since the number of nonzero per row is
 constant throughout the matrix. The matrix is given in two arrays. The
@@
 sparsity structure of the matrix. For each row it consists of $n_z$
-consecutive entries which give in which column is the non zero. The
+consecutive entries which give in which column is the nonzero. The
 other array is the {\tt data} array which gives the values of the non
 zero in the four matrices. The values of the matrices are interleaved
-so that the value $A^l_{col\_id[k]}$ of the $k$th non zero of the
+so that the value $A^l_{col\_id[k]}$ of the $k$th nonzero of the
 $l$th matrix is at index $(k-1)*4+l-1$. The vectors {\tt x} are also
@@
         $n_r$ & number of column/rows of $A$ \\
-        $n_c$ & total number of non-zero\\
+        $n_c$ & total number of nonzeros\\
         $n_v$ & number of {\tt x} vectors \\
@@
 
-The matrix is composed of $n_r$ rows and columns with $n_z$ non-zeros
-per row leading to a total of $$n_c = n_r n_z$$ non-zero entries in
-the matrix. Each of these non-zero entries has one index of size $b_i$
+The matrix is composed of $n_r$ rows and columns with $n_z$ nonzeros
+per row leading to a total of $$n_c = n_r n_z$$ nonzero entries in
+the matrix. Each of these nonzero entries has one index of size $b_i$
 and $n_m$ values of size $b_x$. The matrices have a total size
@@
 accessed, the value needs to be transfered from memory again. So in
-total, there are as many transfers as the number of non-zeros in the
+total, there are as many transfers as the number of nonzeros in the
 matrix. Note however that most architectures cannot read memory a
@@
 the size of a cacheline $cl$, are transfered at once.  When there are
-multiple vectors, each non-zero element uses $n_v$ consecutive
+multiple vectors, each nonzero element uses $n_v$ consecutive
 entries. The worst case number of bytes read and transfered are
@@
 
-In SpMV, each non-zero of the matrix requires two floating point
+In SpMV, each nonzero of the matrix requires two floating point
 operations: one for performing the multiplication and one for
@@
 OpenCL or CUDA, one can think of these registers as a warp.) The end
-goal is to be able to load and format the non zero values $a,b,c,d$
+goal is to be able to load and format the nonzero values $a,b,c,d$
 and the vector entries $p,q,r,s$ in two vector registers to be able to
@@
 All the operations that loads data from the memory into a register
-allows to transfer 512 bits, so 16 floats at a time. A single non zero
+allows to transfer 512 bits, so 16 floats at a time. A single nonzero
 in the matrix will cause 16 operations to be performed, but it only
 uses 4 floats from the matrix and 4 floats from the {\tt x} vector. So
-we need to load at once the data for 4 non zeroes in the matrix,
+we need to load at once the data for four nonzeros in the matrix,
 representing 16 floats from the matrix and 16 floats from the vectors
-at once. (The number of non zeroes on a row in our application is
+at once. (The number of nonzeros on a row in our application is
 always a multiple of 4, if it was, we would need to add some explicit
@@
 
-The values from the non zero element in the matrix are the simplest to
-load. One can load the 16 floating point values for 4 non zeroes of
+The values from the nonzero element in the matrix are the simplest to
+load. One can load the 16 floating point values for four nonzeroes of
 the matrix at once. The 4 floats coming from each matrix four 4
-consecutive non zeroes are naturally grouped within each lane of the
+consecutive nonzeros are naturally grouped within each lane of the
 SIMD vector. One can use a swizzle operation to broadcast the 4
@@
 Loading the entries from the {\tt x} vector is a little bit more
-complicated. First the columns pointers of 4 different non zeroes are
+complicated. First the columns pointers of 4 different nonzeros are
 loaded in a vector register and the values are distributed one per
@@
 register. At this point we have a SIMD register where each lane is the
-four vector entries for a non zero element. Using a broadcast
+four vector entries for a nonzero element. Using a broadcast
 permutation, one can replicate an entry of the four vectors accross
@@
 properly laid out in the vectors, a partial value for $y_i$ can be
-computed. Once all the non zeroes of the row have been processed, the
+computed. Once all the nonzeros of the row have been processed, the
 exact $y_i$ value can be sent to memory using a the No Read No Global
@@
 Supercompact matrices (Figure~\ref{fig:supercompact}) have been
-generated to only have non zero elements in the first 32 columns of
+generated to only have nonzero elements in the first 32 columns of
 the matrix. That way we ensure the simplest form of multiplication,
@@
 the memory will be minimum. The Compact matrices
-(Figure~\ref{fig:compact}) are generated to have 32 non zeroes per row
+(Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
 centered around the diagonal and it represents the ideal case for most
 application relying on sparse matrix vector multiplication. The Random
-matrices (Figure~\ref{fig:compact}) see their non zero element
+matrices (Figure~\ref{fig:compact}) see their nonzero element
 randomly (uniformly) distributed in the matrix, they represent the
@@
 derivative has an associated sparse matrix with a fixed number of
-non-zeros on each row. While computing a single derivative of multiple
+nonzeros on each row. While computing a single derivative of multiple
 functions is rather common (vectorization occurs over the vectors), we

### 2013-10-18 ieee_paper.tex
Updated paper, with grammar fixes. Still working.

@@
 matrix with $n_zN$ nonzero entries. To evaluate the derivative at all
-points in the domain, takes the form
+points in the domain, one multiplies the source vector $f$ by the 
+derivative matrix $D_{n_z}$ to obtain the vector $g$ of derivative values, 
 $$
-  \mathbf{g} = D_{n_z} \mathbf{f}
+  g = D_{n_z} f
 $$
-
-where $f$ is the source vector of function values and $g$ is the
-resulting vector of derivative values. The computation of a single
-derivative has been reduced to a SpMV, where each row has $n_z$
-nonzeros. In
+computation of a single derivative has been reduced to a SpMV, 
+where each row has $n_z$ nonzeros. In
 practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
-three-dimensional flows. These numbers are similar to what is used in
+three-dimensional flows. These numbers are on a par with what is used in
 finite-element codes. Notice that the nonzero elements of the matrix
@@
 necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} which
-means that the matrix sparsity pattern might not be symmetric either.
+implies a non-symmetric sparsity pattern in the derivative matrix.
 
@@
 $$
-where $Q$ is a vector of unknowns (3 components of velocity, pressure,
-temperature). For example, in solving a system of equations, it is
+where $Q$ is a vector of unknowns %(e.g., 3D Navier-Stokes has
+%three components of velocity, pressure and temperature). 
+For example, in solving a system of equations, it is
 often necessary to compute the $x$ derivative of multiple functions,
@@
 multiple source vectors. Further improvements are possible by
-recognizing that different derivative matrices (x,y,z and Laplacian
-for example), have the same sparsity distribution; only the weight
-change.  Thus, alternative to computing a derivative of multiple
-functions, we can calculate multiple derivatives of a single
-functions. The increased memory bandwidth due to an increase in the
+recognizing that different derivative matrices (e.g., $x$, $y$, $z$ 
+and Laplacian operators), 
+have the same sparsity distribution as long as the derivative stencils
+remain unchanged; only the values in the sparse matrix change. 
+change.  Thus, rather than computing a derivative of multiple
+functions, one can calculate multiple derivatives of a single
+function. The increased memory bandwidth due to an increase in the
 number of derivative matrices is offset by better cache utilization,
-leading to an overall benefit.
+leading to an overall performance benefit. In practice however, the number of 
+vectors whose derivatives are required is limited, limiting the 
+performance of the matrix/vector multiplication. Similarly, the number
+of different derivatives in a particular computation is limited. It is 
+for this reason that we seek to combine the benefits of multiple
+vectors and multiple matrices simultaneously. We limit this initial 
+study to four vectors and four matrices.
 
@@
 
-In term of memory layout, we choose a format similar to the ELLPACK
-format to store the matrix since the number of nonzero per row is
+In term of memory layout, we adopt a format similar to the ELLPACK
+format to store the matrix since the number of nonzeros per row is
 constant throughout the matrix. The matrix is given in two arrays. The
 first one is the classical {\tt col\_id} array that describes the
-sparsity structure of the matrix. For each row it consists of $n_z$
-consecutive entries which give in which column is the nonzero. The
-other array is the {\tt data} array which gives the values of the non
-zero in the four matrices. The values of the matrices are interleaved
+sparsity structure of the matrix. Each row of {\tt col\_id} consists of $n_z$
+entries that label the column index of each nonzero in$A$. The
+other array is the {\tt data} array that gives the values of the 
+nonzeros in the four matrices. The values of the matrices are interleaved
 so that the value $A^l_{col\_id[k]}$ of the $k$th nonzero of the
-$l$th matrix is at index $(k-1)*4+l-1$. The vectors {\tt x} are also
-similarly interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
+$l$th matrix is at index $(k-1)*4+l-1$. The vectors {\tt x} are 
+interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
 consecutive in memory. The vector {\tt y} is similarly
-interleaved. This is schematically shown in Figure~\ref{fig:mat_struct}.
+interleaved. This is schematically shown in Figure~\ref{fig:mat_struct}, 
+where the elements of $y$ are grouped in sets of 16, which correspond
+to four derivatives applied to four functions, at a single point.
 
@@
         $n_m$ & number of matrices \\
-        $s_M$ & size of the matrice(s) in bytes\\
-        $s_x$ & size of the {\tt x} vector(s) in bytes\\
-        $s_y$ & size of the {\tt y} vector(s) in bytes\\
+        $s_M$ & size of the $n_m$ matrices in bytes\\
+        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
+        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
         \hline
@@
         $O$     & number of floating point operations \\
-        $I_b$   & maximal computational intensity\\
+        $I_b$   & maximum computational intensity\\
         $I_w$   & minimum computational intensity\\        
@@
   \end{center}
-  \caption{Notations}
+  \caption{Notation relative to the matrices (upper section) and to the benchmarks (lower section.)}
   \label{tab:not}
@@
     coding was performed without MIC vector
-    instructions.]{\includegraphics[width=.40\linewidth]{figures/test1_readwrite.pdf}\label{fig:band_rw}}
+    instructions.]{\includegraphics[width=.40\linewidth]{figures/bandwidth_read_write.pdf}\label{fig:band_rw}}
 %
@@
 
-\begin{figure*}
+\begin{figure}
 \centering
-  \subfigure[ xxx ] 
-  {\includegraphics[width=.20\linewidth]{figures/test1_gather.pdf}\label{fig:band_gather}} \\
-
-  \subfigure[The greater
-    speed of the cpp version is obtained throught the use of
-    \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}. All memory
-    is aligned on 64
-    bytes.]{\includegraphics[width=.20\linewidth]{figures/test3_gather.pdf}\label{fig:band_gather_ivdep}}
-\end{figure*}
+  \subfigure[Performance of gather and unpack operations on the MIC. ] 
+  {\includegraphics[width=.8\linewidth]{figures/bandwidth_gather_unpack.pdf}\label{fig:band_gather}} \\
+\end{figure}
 %

### 2013-10-18 ieee_paper.tex
small mods.

@@
 
-If we assume an algorithm where the rows are processed the one after
+If we assume an algorithm where the rows are processed one after
 the other, the amount of memory written is precisely the
@@
 multiple vectors, each nonzero element uses $n_v$ consecutive
-entries. The worst case number of bytes read and transfered are
+entries. In the worst case, the number of bytes read and transfered is
 $$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}} $$ 
@@
 significant difference between the best and the worst case: there is a
-$8$ fold difference in the 1 matrix 1 vector case but that gap closes
-with the increase in the number of vectors and matrices to a 4 fold
-difference in the 4 matrices and 4 vectors case and 1.6 fold in the 16
-matrices and 16 vectors. The ratios computed with $cl = 1$ are mostly
-similar to the the one with $cl=64$ when the number of vectors is
-large. But the differences are important when the number of vector is
-low: this highlight that accessing the memory per batch is the main
+eightfold difference in the one matrix -- one vector (1/1) case but that gap closes
+with the increase in the number of vectors and matrices to a fourfold
+difference in the four4matrices and four vectors (4/4) case and 1.6 fold in the 
+16/16 case. 
+The ratios computed with $cl = 1$ are mostly
+similar to the one with $cl=64$ when the number of vectors is
+large. But the differences are important when the number of vectors is
+low: this highlights that accessing the memory per batch \todo{what do you mean by 
+accessing the memory per batch} is the main
 problem faced when performing a standard SpMV computation.
@@
 
-  \caption{Ratio of flops to bytes in single precision: in the best case in
-    \ref{fig:ratio_best} best; in the worst case when $cl=64$ in
-    \ref{fig:ratio_worst}; and in the worst case neglecting that the
-    memory transfers are by batch (equivalent to $cl = 1$) in
-    \ref{fig:ratio_worst_nocache}.}
+  \caption{Ratio of flops to bytes in single precision:  
+    \ref{fig:ratio_best} best case; \ref{fig:ratio_worst} worst case when $cl=64$ 
+    ; and \ref{fig:ratio_worst_nocache} worst case neglecting that the
+    memory transfers are by batch (equivalent to $cl = 1$)}
   \label{fig:ratio_bytes_flops}
@@
 into actual performance. This figure provides projected best and worst
-Gflop/s achievable assuming the computation is memory bound and the
-architecture reaches 150Gflop/s. Notice that \cite{Saule13-ARXIV}
+Gflop/s achievable assuming the computation is memory-bound and the
+architecture reaches 150 Gflop/s. Notice that \cite{Saule13-ARXIV}
 showed a higher peak bandwidth, but we will show in
@@
 than the best achievable using a classical SpMV computation. The best
-performance reachable is 210Gflop/s: almost 6 times higher than the
+performance reachable is 210Gflop/s: almost six times higher than the
 peak of the classical SpMV case.

### 2013-10-18 ieee_paper.tex
More mods.

@@
 
-  \caption{Estimation of the reachable performance one can achieve
+  \caption{Estimation of the maximum achievable performance 
     using a device with a memory to computational units of 150GB/s
@@
 In this work, we use an Intel Xeon Phi 5110P coprocessor. This card
-has 8 memory controllers where each of them can execute 5 billion
-transactions per second and has two 32-bit channels, achieving a total
+has 8 memory controllers that can each execute 5 billion
+transactions per second. Each has two 32-bit channels, achieving a total
 bandwidth of 320GB/s aggregated across all the memory
@@
 instruction cache, and a 512kB L2 cache. The architecture of a core is
-based the traditional Pentium architecture which have been extended to
-64-bit. A core can hold 4 hardware contexts at any time. And at each
+based the traditional Pentium architecture, which has been extended to
+64-bit. A core can hold four hardware contexts at any given time. At each
 clock cycle, instructions from a single thread are executed. Due to
 some hardware constraints, two hardware contexts must be used to reach
-the peak instruction throughput in that architecture. As in the
+the peak instruction throughput in that architecture. Similar to the
 Pentium architecture, a core has two different concurrent instruction
-pipelines which allow the execution of two instructions per
+pipelines that allow the execution of two instructions per
 cycle. However, only one vector or floating point instruction can be
@@
 Most of the performance of the architecture comes from the vector
-processing unit. Each core has 32 512-bit SIMD registers which can be
+processing unit. Each core has 32 512-bit SIMD registers that can be
 used for double or single precision, that is, either as a vector of 8
@@
 addition or division, and mathematical operations, such as sine and
-sqrt, allowing to reach 8 double precision operations per cycle (16
-single precision). The unit also supports Fused Multiply-Add (FMA)
-operations which are typically counted as two operations for
-benchmarking purposes. Therefore, the peak performance of the 5110P
+sqrt, making it possible to reach 8 double precision (or 16 singe precision)
+operations per cycle.  
+The unit also supports Fused Multiply-Add (FMA)
+operations, which are typically counted as two operations for
+when benchmarking. Therefore, the peak performance of the 5110P
 card is 1 Tflop/s in double precision and 2 Tflop/s in single
-precision. If FMA can not be used and half of these number can be
+precision. If FMA cannot be used, only half of these rates can be
 achieved.
 
-\subsection{Bringing the data in vector register}
+\subsection{Bringing the data into the vector register}
 
@@
 about 210Gflop/s in single precision in the best case. We will focus
-on the single precision case with 4 vectors and 4 matrices. But
-similar techniques apply for other combinations. This target
+on the single precision case with four vectors and four matrices. 
+Similar techniques apply for other matrix/vector combinations. This target
 performance represents 10\% of the peak performance of the
@@
 We discuss how the implementation of the v4m4 kernel works using
-vectorial instruction. The code is given for reference in
+vectorial instructions. The code is given for reference in
 Figure~\ref{code:mat_mul}. (We will show in Section~\ref{sec:expe}
-that an implementation relying on compiler vectorization does not lead
+that an implementation that relies on compiler vectorization does not lead
 to desirable performance, which confirms the results
@@
 OpenCL or CUDA, one can think of these registers as a warp.) The end
-goal is to be able to load and format the nonzero values $a,b,c,d$
-and the vector entries $p,q,r,s$ in two vector registers to be able to
+goal is to load and format the nonzero values $a,b,c,d$
+and the vector entries $p,q,r,s$ in two vector registers and 
 apply Fused Multiply-Add on them, which is depicted in
-Figure~\ref{fig:tensor_product}. Bringing the data in the vector
-register as depicted in the figure is the difficult part that we now
-explain. A thread will perform the multiplications one row at a time,
+Figure~\ref{fig:tensor_product}. Bringing the data into the vector
+register as depicted in the figure is the difficult component of our 
+proposed algorith, which we proceed to explain further. 
+A thread will perform the multiplications one row at a time,
 and parallelism is achieved by giving blocks of rows to each core

### 2013-10-18 ieee_paper.tex
small changes.

@@
 features of the vectors in the Intel MIC architecture. A vector of 512
-bits is composed of 4 lanes (sometimes called channels) of 128 bits
-which are made of 4 elements of 32 bits. A swizzling operation allows
-to reorder the elements within each lane. While a permutation allows to
-change the orders of the lanes. Figure~\ref{fig:swizzling_permutation}
+bits is composed of four lanes (sometimes called channels) of 128 bits, 
+which are made of four segments of 32 bits. A swizzling operation allows
+the reorder of the elements within each lane, while a permutation allows 
+changes the order of entire lanes. 
+Figure~\ref{fig:swizzling_permutation}
 illustrates these two possibilities. Both permutation and swizzling
@@
 function needs to use the {\tt \_mm512\_permute4f128\_epi32}
-instruction which is only defined on integers and therefore typecast
+instruction which is only defined on integers and therefore in place typecasts
 are necessary to convert to a vector to/from the integer type.
+\todo{I added a comment stating we preferred swizzling.}
+(Note that we prefer swizzling to permutation if possible because we believe 
+it is cheaper. Many instructions have swizzling capability, while permuation
+is its own instruction.} \todo{reference?}
 
@@
   \includegraphics[width=\linewidth]{figures/swizzling_lanes.pdf}
-  \caption{Swizzling enables to reorder or broadcast elements within
+  \caption{Swizzling enables the reorder or broadcast of elements within
     each 128 bit group. Permutation enables to reorder or broadcast
-    whole 128 bit groups.}
+    whole 128 bit groups. All lanes must be sizzeled identically.}
   \label{fig:swizzling_permutation}
 \end{figure}
+\todo{I added to the caption.}
 
-All the operations that loads data from the memory into a register
-allows to transfer 512 bits, so 16 floats at a time. A single nonzero
+Every operation that loads data from memory into a register
+transfers 512 bits, or 16 floats at a time. A single nonzero
 in the matrix will cause 16 operations to be performed, but it only
 uses 4 floats from the matrix and 4 floats from the {\tt x} vector. So
-we need to load at once the data for four nonzeros in the matrix,
-representing 16 floats from the matrix and 16 floats from the vectors
+we would like to execute a single instruction to load the data for four 
+nonzeros in the matrix, representing 16 floats from the matrix and 16 floats from the vectors
 at once. (The number of nonzeros on a row in our application is
-always a multiple of 4, if it was, we would need to add some explicit
-zeroes in the data structure to pad it.)
+always a multiple of 4; if it weren't, we would add some explicit
+zeroes in the data structure for padding.)
+\todo{ reread paragraph above. single instruction to load 16 floats. Confused.}
 
-The values from the nonzero element in the matrix are the simplest to
+The values from the nonzero elements in the matrix are the simplest to
 load. One can load the 16 floating point values for four nonzeroes of
-the matrix at once. The 4 floats coming from each matrix four 4
+the matrix at once. The four floats coming from each matrix four 4
+\todo{something wrong with sentence I am in. Did you discuss matrix ordering?}
 consecutive nonzeros are naturally grouped within each lane of the
@@
 Loading the entries from the {\tt x} vector is a little bit more
-complicated. First the columns pointers of 4 different nonzeros are
-loaded in a vector register and the values are distributed one per
+involved. First, the columns pointers of four different nonzeros are
+loaded into a vector register and the values are distributed one per
 lane of the vector register using an {\tt unpack}
@@
 pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They
-replicated so that each fill a vector lane using a swizzle
-operation. Then each value is multiplied by 4 (because there are 4
-vectors) and the offsets $(3,2,1,0)$ is added each lane to obtain the
+are replicated so that each value fills an single vector lane 
+using a swizzle operation. Then each value is multiplied by 4 
+(because there are 4 vectors) 
+and the offset vector $(3,2,1,0)$ is added to each lane to obtain the
 correct index of the elements of the {\tt x} vector within each lane
 using a Fused Multiply-Add operation. Then a {\tt gather} operation is
-performed bring the 16 entries of the {\tt x} vector into the SIMD
+performed to bring the 16 entries of the {\tt x} vector into the SIMD
 register. At this point we have a SIMD register where each lane is the
 four vector entries for a nonzero element. Using a broadcast
-permutation, one can replicate an entry of the four vectors accross
+permutation, one can replicate an entry of the four vectors across
 each lane. (The broadcast of $x_m^1,x_m^2,x_m^3,x_m^4$ is shown in
@@
 exact $y_i$ value can be sent to memory using a the No Read No Global
-Order hint to optimize traffic on the memory bus.
+Order hint to optimize write traffic on the memory bus.
+\todo{ added "write" to traffic }
 

### 2013-10-18 ieee_paper.tex
Updated paper. Still stuff to do.  I fixed graphs, and grammar.

@@
 study to four vectors and four matrices.
+
+\todo{Add few sentence with discretization}
 
@@
 it is cheaper. Many instructions have swizzling capability, while permuation
-is its own instruction.} \todo{reference?}
+is its own instruction. \todo{reference?}
 
@@
 
-To perform our experiments and analysis, we use multiple kinds of
-matrices which are presented in Figure~\ref{fig:spy_plots}. The
+The different matrices used to perform our experiments and analysis
+are presented in Figure~\ref{fig:spy_plots}. The
 Supercompact matrices (Figure~\ref{fig:supercompact}) have been
@@
 the matrix. That way we ensure the simplest form of multiplication,
-the amount of elements from the {\tt x} vector data transfered from
-the memory will be minimum. The Compact matrices
+the number of elements from the {\tt x} vector data transfered from
+the memory is be minimum, they are transferred efficiently and remain 
+in cache. The Compact matrices
 (Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
-centered around the diagonal and it represents the ideal case for most
-application relying on sparse matrix vector multiplication. The Random
-matrices (Figure~\ref{fig:compact}) see their nonzero element
-randomly (uniformly) distributed in the matrix, they represent the
-worst case scenario for cache based architecture where the cache
+centered around the diagonal and it represents the ideal case for many
+applications that rely on sparse matrix vector multiplication. The nonzero
+elements of the Random
+matrices (Figure~\ref{fig:compact}) see their nonzero elements
+randomly (uniformly) distributed in the matrix; they represent the
+worst case scenario for a cache based architecture where the cache
 reutilisation is the lowest from one row to the next.
@@
 Figure~\ref{fig:rbf2dnorcm} shows a $32$-points stencil of a $8^3$
-grid). We also apply Cuthill-Mckee reordering to all the matrices This
-ordering technique is designed to reduce the distance between the non
-zeroes and the diagonal, hopefully allowing a better cache
+grid). We also apply a Cuthill-Mckee reordering to all the matrices. This
+ordering technique is designed to reduce the distance between the 
+nonzeros and the diagonal, hopefully allowing a better cache
 reutilisation. The reordered version of these two matrices can be seen

### 2013-10-18 ieee_paper.tex
changes.

@@
 each of size $n_z \times n_z$. Each linear solve produces a row of the
-RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \times N$
+RBF-FD differentiation matrix $A_{n_z}$, resulting in a $N \times N$
 matrix with $n_zN$ nonzero entries. To evaluate the derivative at all
-points in the domain, one multiplies the source vector $f$ by the 
-derivative matrix $D_{n_z}$ to obtain the vector $g$ of derivative values, 
+points in the domain, one multiplies the source vector $x$ by the 
+derivative matrix $A_{n_z}$ to obtain the vector $y$ of derivative values, 
 $$
-  g = D_{n_z} f
+  y = A_{n_z} x
 $$
 computation of a single derivative has been reduced to a SpMV, 
-where each row has $n_z$ nonzeros. In
-practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
+where each row has $n_z$ nonzeros\footnote{We use non-conventional 
+symbology for differentiation matrix, source and destination vectors
+anticipating what follows.}. 
+In practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
 three-dimensional flows. These numbers are on a par with what is used in
@@
 solution to transport equations of the form
-$$
-\pf{Q}{t} = f(Q,Q_x, Q_y, Q_z, \Laplacian{Q})
-$$
+\begin{equation}
+\pf{Q}{t} = f(Q,\grad{Q},\Laplacian{Q})  \label{eq:Q}
+\end{equation}
 where $Q$ is a vector of unknowns %(e.g., 3D Navier-Stokes has
@@
 For example, in solving a system of equations, it is
-often necessary to compute the $x$ derivative of multiple functions,
+often necessary to compute the derivative of multiple functions along
+the same coordinate direction,
 typically four for the Euler equations, or five for the Navier-Stokes
@@
 multiple source vectors. Further improvements are possible by
-recognizing that different derivative matrices (e.g., $x$, $y$, $z$ 
-and Laplacian operators), 
+recognizing that different derivative matrices (e.g., gradient and 
+Laplacian operators), 
 have the same sparsity distribution as long as the derivative stencils
@@
 study to four vectors and four matrices. 
+\todo{I have remove reference to $x$, $y$, ... to avoid confusion with what follows.}
 
-\todo{Add few sentence with discretization}
+\todo{Add few sentence with discretization. Check Erik}
+In discrete form, Equation (\ref{eq:Q}) takes the form
+$$
+  y^{i,j} = A^i x^j 
+$$
+where $j$ ranges over the number of source vectors, $i$ ranges of the number of derivative
+matrices and there are $i*j$ output vectors $y$.
 
@@
 sparsity structure of the matrix. Each row of {\tt col\_id} consists of $n_z$
-entries that label the column index of each nonzero in$A$. The
+entries that label the column index of each nonzero in $A$. The
 other array is the {\tt data} array that gives the values of the 
@@
         $b_x$ & number of bytes per value \\
-        $n_z$ & number of nonzeros per row $A$ \\
+        $n_z$ & number of nonzeros per row of $A$ \\
         $n_r$ & number of column/rows of $A$ \\
@@
   \centering 
-  \subfigure[Absolute performance ]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops_peak_perf}}
+  \subfigure[Absolute performance ]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}}
 %
@@
 which are made of four segments of 32 bits. A swizzling operation allows
-the reorder of the elements within each lane, while a permutation allows 
-changes the order of entire lanes. 
+the reorder of the elements within each lane, while a permutation reorders
+entire lanes, without changing their contents. 
 Figure~\ref{fig:swizzling_permutation}
@@
 it is cheaper. Many instructions have swizzling capability, while permuation
-is its own instruction. \todo{reference?}
+is its own instruction. \todo{do we need a reference to my note?})
 
@@
     each 128 bit group. Permutation enables to reorder or broadcast
-    whole 128 bit groups. All lanes must be sizzeled identically.}
+    whole 128 bit groups. All lanes must be sizzeled identically.
+    \todo{I added to the caption.}}
   \label{fig:swizzling_permutation}
 \end{figure}
-\todo{I added to the caption.}
 
@@
 zeroes in the data structure for padding.)
-\todo{ reread paragraph above. single instruction to load 16 floats. Confused.}
+\todo{Erik, reread paragraph above. single instruction to load 16 floats. Confused.}
 
@@
 Order hint to optimize write traffic on the memory bus.
-\todo{ added "write" to traffic }
+\todo{I added "write" to traffic }
 
@@
 \label{sec:expe}
+In this section, we describe a series of numerical experiments to confirm 
+the performance predictions of Section~\ref{sec:model}. To better
+understand our results, we first perform some bandwidth measurements, followed
+by the actual SPmV.  \todo{We might be criticized for not comparing with other work.}
 
@@
 generated to only have nonzero elements in the first 32 columns of
-the matrix. That way we ensure the simplest form of multiplication,
-the number of elements from the {\tt x} vector data transfered from
-the memory is be minimum, they are transferred efficiently and remain 
-in cache. The Compact matrices
+the matrix. As a result, only 32 values of {\tt x} are read
+from memory, and the expense associated with cache misses and gather
+operation is removed from consideration. The cost associated with 
+{\tt A} remains.   \todo{reread the above.}
+The Compact matrices
 (Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
@@
 Figure~\ref{fig:rbf2dnorcm} shows a $32$-points stencil of a $8^3$
-grid). We also apply a Cuthill-Mckee reordering to all the matrices. This
-ordering technique is designed to reduce the distance between the 
+grid). We also apply a Cuthill-Mckee reordering to all the matrices. 
+The 2D and 3D grid sizes are chosen to have comparable matrix sizes. One notes
+that the bandwidth reduction is higher in the 2D configuration, which leads
+to better performance.  \todo{we do not show performance results in 2D. Do we
+need to change anything? Do we have a reference for RCM?} 
+This ordering technique is designed to reduce the distance between the 
 nonzeros and the diagonal, hopefully allowing a better cache
@@
 \centering
-  \subfigure[
-    Entries with "cpp" denote cases where
-    coding was performed without MIC vector
-    instructions.]{\includegraphics[width=.40\linewidth]{figures/bandwidth_read_write.pdf}\label{fig:band_rw}}
+  \subfigure[Run on the MIC with 240 threads, 64-byte memory alignment.]{\includegraphics[width=.40\linewidth]{figures/bandwidth_read_write.pdf}\label{fig:band_rw}}
 %
@@
   \subfigure[
-    The speed on the CPU matches the speed with vector instructions.
-    All memory is aligned on 32
-    bytes.]{\includegraphics[width=.40\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:read_write}}
+    Run on the host with 32 threads, 32-byte memory alignment.]{\includegraphics[width=.40\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:read_write}}
 
@@
     case implemented with MIC vector instructions. $-O3$ compilation
-    options. Each symbol has two cases: with (red) and without blue) 
-    with and without Reverse Cuthill-McKee. The grid is $96^3$.]
+    options. Each symbol has two cases: with (red) and without (blue) 
+    Reverse Cuthill-McKee bandwidth reduction. The grid is $96^3$.]
     {\includegraphics[width=.40\linewidth]{figures/mic_performance_nb_threads.pdf}\label{fig:perf_mic}}
@@
   \centering
-  \subfigure[Supercompact and random.]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}}
+  \subfigure[Supercompact and Random]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}}
   \subfigure[Compact and RBF-FD]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}}
@@
   %\subfigure[RBF 3D]{\includegraphics[width=.48\linewidth]{figures/rbf_max_perf_run1.pdf}}
-  \caption{Performance of "best code" accross instance types.}
+  \caption{Performance of "best code" accross matrix types. The two horizontal lines depict worst and 
+  best case performance (Figure fig:gflops-peak-perf). RCM (blue, $\blacktriangle$) and no RCM (red,$\blacktriangledown$).}
 \end{figure*}

### 2013-10-18 ieee_paper.tex
forgotten file.

@@
 \subsection{Computations}
+%----------------------
+Scaling results as a function of the number of processors, on the MIC and on the host. The host calculations
+are performed for reference only, to bring out the dominance of the MIC due to its advanced swizzling and channel 
+operations, not available on the host. The performance on the host is 3x slower than on the MIC. The large cache
+of the host alleviates somewhat the lower memory bandwidth, but this only reduces the cost of transfer {\tt x}. 
+The cost memory cost for {\tt A} and {\tt col\_id} is unaffected by the larger cache size.
+
+The performance on the MIC scales linearly with the number of threads.
+We compare three cases, each with and without bandwidth reduction using Reverse Cuthil-McKee (RCM).
+There are $96^3$ RBF nodes, leading to a matrix with 884,736 rows.
+The four source vectors occupy more than 14M bytes, which exceeds the aggregate L2 cache of 31 Mbytes (7.8 Mfloats.)
+
+\todo{Is the host architecture described? Problem: I had to change hosts a few weeks ago, and it has different characteristics}. 
+%----------------------
 

### 2013-10-18 ieee_paper.tex
basic bandwidth

@@
 understand our results, we first perform some bandwidth measurements, followed
-by the actual SPmV.  \todo{We might be criticized for not comparing with other work.}
+by the actual SpMV.
 
@@
 
-\def\ww{.13\textwidth}
+\def\ww{.19\textwidth}
 \begin{figure*}[tbh]
@@
       \includegraphics[width=\ww]{figures/random_matrix-crop.png}}%
-    \subfigure[2D, no RCM.]{%
-      \label{fig:rbf2dnorcm}
-      \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.png}}%
-    \subfigure[2D, RCM.]{%
-      \label{fig:rbf2drcm}
-      \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
+    %% \subfigure[2D, no RCM.]{%
+    %%   \label{fig:rbf2dnorcm}
+    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.png}}%
+    %% \subfigure[2D, RCM.]{%
+    %%   \label{fig:rbf2drcm}
+    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
     \subfigure[3D, RCM.]{%
@@
 
-The other two types of matrices are used in derivative stencils in 2D
-and 3D RBF-FD calculations (Figure~\ref{fig:rbf2dnorcm} pictures a
-$32$-points stencil of a $23^2$ 2D grid and
-Figure~\ref{fig:rbf2dnorcm} shows a $32$-points stencil of a $8^3$
-grid). We also apply a Cuthill-Mckee reordering to all the matrices. 
-The 2D and 3D grid sizes are chosen to have comparable matrix sizes. One notes
-that the bandwidth reduction is higher in the 2D configuration, which leads
-to better performance.  \todo{we do not show performance results in 2D. Do we
-need to change anything? Do we have a reference for RCM?} 
-This ordering technique is designed to reduce the distance between the 
-nonzeros and the diagonal, hopefully allowing a better cache
-reutilisation. The reordered version of these two matrices can be seen
-in Figure~\ref{fig:rbf2drcm} and in Figure~\ref{fig:rbf3drcm}.
+The other two types of matrices are used in derivative stencils in 3D
+RBF-FD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-points
+stencil of a 3D $8^3$ grid). We also apply a Cuthill-Mckee reordering to
+all the matrices.  This ordering technique is designed to reduce the
+distance between the nonzeros and the diagonal, hopefully allowing a
+better cache reutilisation. The reordered version of the matrix can be
+seen in Figure~\ref{fig:rbf3drcm}.
 
@@
 
+In this experimental section, as a point of reference we provide some
+results on a typical CPU architecture. This system is ...\todo{include
+  a short description of the host.}
+All the codes are compiled with icc xxx and -O3 optimization.
+
 \begin{figure*}
 \centering
-  \subfigure[Run on the MIC with 240 threads, 64-byte memory alignment.]{\includegraphics[width=.40\linewidth]{figures/bandwidth_read_write.pdf}\label{fig:band_rw}}
+%
+ \subfigure[On MIC with 240 threads, 64-byte memory
+   alignment.]{\includegraphics[width=.40\linewidth]{figures/bandwidth_read_write.pdf}\label{fig:band_rw_MIC}}
 %
@@
 %
-  \subfigure[
-    Run on the host with 32 threads, 32-byte memory alignment.]{\includegraphics[width=.40\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:read_write}}
+  \subfigure[On CPU with 32 threads, 32-byte memory
+    alignment.]{\includegraphics[width=.40\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:band_rw_CPU}}
 
-\caption{Bandwidth performance on the host under idealized conditions as a function of matrix row size.
-  Entries with "cpp" denote cases where coding was performed without MIC vector intructions. 
-\todo{not 100\% sure what this plots are.}}
+  \caption{Bandwidth performance on the host under idealized
+    conditions as a function of matrix row size.  Entries with "cpp"
+    denote cases where coding was performed without MIC vector
+    intructions.  \todo{Can we label the x-axis "number of floats" or
+      "size in MB"?}\todo{Can we use the same symbols in both charts?}
+    \todo{Can we change the color scheme? in (a), I do not see the
+      difference between "read\_cpp" and "read/write cpp"} }
+  \label{fig:bandwidth}
 \end{figure*}
+
+Figure~\ref{fig:bandwidth} gives some benchmark that compute the
+bandwidth achieved when reading and/or writing large arrays on both
+the MIC card and the CPU architecture. The read benchmark performs a
+simple sum of the array. The write benchmark sets whole the elemtns of
+the array to 0. While the read/write benchmarks copies the array into
+another one. We also investigate the difference between a straight C++
+implementation (compiled with -O3 with proper tagging to inform the
+compiler of memory alignments) and a simple implementation using
+vector register explicitly.
+
+When using the MIC card (Figure~\ref{fig:band_rw_MIC}) the {\tt
+  read\_cpp} implementation reaches a performance of 160GB/s
+\todo{gordon check exact number} which is better than the {\tt read}
+implementation which only achieves 145GB/s \todo{and that one}. This
+is explained by the compiler adding loop unrolling and software
+prefetching in the assembly code in the {\tt read\_cpp} case which is
+not added in the {\tt read} case. On the other hand, the {\tt write}
+case reaches a performance of 155GB/s\todo{and that one} while the
+{\tt write\_cpp} implementation never passes 55GB/s. This difference
+is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction
+which bypasses the Read For Ownership protocol and allows the writes
+to be performed in any order. The compiler can not use that
+instruction on its own because using would break the memory convention
+useful for some parallel algorithms. The {\tt read/write} benchmark
+performs at 140GB/s while the {\tt read/write cpp} benchmark performs
+at about 90GB/s.
 
@@
   \centering
-  \subfigure[Performance of gather and unpack operations on the MIC. ] 
-  {\includegraphics[width=.8\linewidth]{figures/bandwidth_gather_unpack.pdf}\label{fig:band_gather}} \\
+  \includegraphics[width=\linewidth]{figures/bandwidth_gather_unpack.pdf}
+  \caption{Performance of gather and unpack operations on the MIC.}
+  \label{fig:band_gather} 
 \end{figure}

### 2013-10-18 ieee_paper.tex
bandwidth

@@
 results on a typical CPU architecture. This system is ...\todo{include
-  a short description of the host.}
-All the codes are compiled with icc xxx and -O3 optimization.
+  a short description of the host.}  All the codes are compiled with
+icc xxx and -O3 optimization. The MIC codes use the IMCI instructions
+while the CPU codes use AVX instructions.
 
@@
 which bypasses the Read For Ownership protocol and allows the writes
-to be performed in any order. The compiler can not use that
+to be performed in any order. The compiler can not use this
 instruction on its own because using would break the memory convention
@@
 at about 90GB/s.
+
+Looking at the results on the host in Figure~\ref{fig:band_rw_CPU}
+show no significant difference between manual vectorization and the
+compiler generated one (there is no store nrngo on the host). It also
+shows quite stable performance with about 70GB/s in the {\tt read}
+benchmark, about 40GB/s in the {\tt write} benchmark and about 49 in
+the {\tt read write} benchmark. The bandwidth achieved on MIC is about
+2 to 3 times faster than the bandwidth achieved on the CPU. The
+kernels we consider being memory bound, we expect MIC to be 2 to 3
+times faster.
+
+Note that SpMV kernels will likely use only simple read and simple
+writes on the MIC architecture. We also benchmark the performance we
+can achieve using {\tt gather} and {\tt unpack} operations and we
+present the results in Figure~\ref{fig:band_gather}. The {\tt unpack}
+instruction can reach a performance of 140GB/s. (There is no easy way
+to implement the equivalent of the {\tt unpack} operation without
+using explicit registers.) The performance of the {\tt gather}
+operations depends in which order the array is read. If the indices
+are ordered ({\tt compact}) then a performance of 140GB/s is
+reached. If the indices are completely randomized, the performance
+drops to 2GB/s\todo{correct?}. We can see that implementing the
+indirection mechanism without vector instructions makes the
+performance drop significantly. (In the {\tt compact} case, it never
+passes 100GB/s and is close to 50GB/s most of the time.)
 
@@
 %
+
+To summarize, using vector instruction seems mandatory to reach the
+highest bandwidth on the MIC architecture. For the instructions used
+in a typical SpMV computation, one can expect a bandwidth between
+140GB/s to 160GB/s. That is why we set in our estimation the best
+achievable bandwidth to 150GB/s despite carefully crafted code to
+maximize bandwidth can reach higher
+values~\cite{Saule13-ARXIV}. Comparing to a CPU architecture, we
+expect MIC to be 2 to 3 times faster.
 
@@
 \end{figure*}
-
-%<<<<<<< HEAD
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
-%
-%  \caption{Performance on a MIC of derivative computation for RBFFD
-%    with 884,736 nodes distributed quasi-randomly in a cube. Single
-%    derivative of a single function (base case, triangle) and four
-%    derivatives of four functions (multi case, circles). No bandwidth
-%    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
-%\end{figure}
-%
-%\begin{figure*}
-%\centering
-%\subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}} %\label{fig:ratio_best}}
-%\subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
-%\subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
-%\caption{Ratio of flops to bytes; (a) simple count of flops and bytes that require transfer from memory;(b) best and (c) worst
-%case taking into account that bringing one element of $x$ into memory brings along the remainder of the cache line.}
-%\label{fig:ratio_bytes_flops}
-%\end{figure*}
-%=======
-%\begin{figure}
-%  \centering
-%  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
-%
-%  \caption{Performance on a MIC of derivative computation for RBFFD
-%    with 884,736 nodes distributed quasi-randomly in a cube. Single
-%    derivative of a single function (base case, triangle) and four
-%    derivatives of four functions (multi case, circles). No bandwidth
-%    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
-%\end{figure}
-%
-%>>>>>>> 619115cea6da9466f09a3b01456eacf7991606e2
 

### 2013-10-18 ieee_paper.tex
bandwidth before instances

@@
 \label{sec:expe}
-In this section, we describe a series of numerical experiments to confirm 
-the performance predictions of Section~\ref{sec:model}. To better
-understand our results, we first perform some bandwidth measurements, followed
-by the actual SpMV.
-
-\subsection{Instances}
-
-\def\ww{.19\textwidth}
-\begin{figure*}[tbh]
-  \begin{center}
-    \subfigure[Supercompact.]{%
-      \label{fig:supercompact}
-      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.png}}%
-    \subfigure[Compact.]{%
-      \label{fig:compact}
-      \includegraphics[width=\ww]{figures/compact_matrix-crop.png}}%
-    \subfigure[Random.]{%
-      \label{fig:random}
-      \includegraphics[width=\ww]{figures/random_matrix-crop.png}}%
-    %% \subfigure[2D, no RCM.]{%
-    %%   \label{fig:rbf2dnorcm}
-    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.png}}%
-    %% \subfigure[2D, RCM.]{%
-    %%   \label{fig:rbf2drcm}
-    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
-    \subfigure[3D, RCM.]{%
-      \label{fig:rbf3dnorcm} 
-      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
-    \subfigure[3D, RCM.]{%
-      \label{fig:rbf3drcm}
-      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.png}}%
-  \end{center}
-
-  \caption{Different sparsity distributions. In all cases, there are
-    32 nonzeros per line and 300 rows. The last four matrices
-    corresponds to derivative stencils in 2D and 3D RBF-FD
-    calculations.}
-  \label{fig:spy_plots}
-\end{figure*}
-
-The different matrices used to perform our experiments and analysis
-are presented in Figure~\ref{fig:spy_plots}. The
-Supercompact matrices (Figure~\ref{fig:supercompact}) have been
-generated to only have nonzero elements in the first 32 columns of
-the matrix. As a result, only 32 values of {\tt x} are read
-from memory, and the expense associated with cache misses and gather
-operation is removed from consideration. The cost associated with 
-{\tt A} remains.   \todo{reread the above.}
-The Compact matrices
-(Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
-centered around the diagonal and it represents the ideal case for many
-applications that rely on sparse matrix vector multiplication. The nonzero
-elements of the Random
-matrices (Figure~\ref{fig:compact}) see their nonzero elements
-randomly (uniformly) distributed in the matrix; they represent the
-worst case scenario for a cache based architecture where the cache
-reutilisation is the lowest from one row to the next.
-
-The other two types of matrices are used in derivative stencils in 3D
-RBF-FD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-points
-stencil of a 3D $8^3$ grid). We also apply a Cuthill-Mckee reordering to
-all the matrices.  This ordering technique is designed to reduce the
-distance between the nonzeros and the diagonal, hopefully allowing a
-better cache reutilisation. The reordered version of the matrix can be
-seen in Figure~\ref{fig:rbf3drcm}.
+In this section, we describe a series of numerical experiments to
+confirm the performance predictions of Section~\ref{sec:model}. To
+better understand our results, we first perform some bandwidth
+measurements, followed by the actual SpMV computations.
 
@@
 
-\begin{figure*}
+\begin{figure*}[bth]
   \centering
-%
   \subfigure[On MIC with 240 threads, 64-byte memory
     alignment.]{\includegraphics[width=.40\linewidth]{figures/bandwidth_read_write.pdf}\label{fig:band_rw_MIC}}
-%
-  \hspace{.3in}
   %
@@
     denote cases where coding was performed without MIC vector
-    intructions.  \todo{Can we label the x-axis "number of floats" or
+    instructions.  \todo{Can we label the x-axis "number of floats" or
       "size in MB"?}\todo{Can we use the same symbols in both charts?}
@@
 expect MIC to be 2 to 3 times faster.
+
+\subsection{Instances}
+
+\def\ww{.19\textwidth}
+\begin{figure*}[tbh]
+  \begin{center}
+    \subfigure[Supercompact.]{%
+      \label{fig:supercompact}
+      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.png}}%
+    \subfigure[Compact.]{%
+      \label{fig:compact}
+      \includegraphics[width=\ww]{figures/compact_matrix-crop.png}}%
+    \subfigure[Random.]{%
+      \label{fig:random}
+      \includegraphics[width=\ww]{figures/random_matrix-crop.png}}%
+    %% \subfigure[2D, no RCM.]{%
+    %%   \label{fig:rbf2dnorcm}
+    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.png}}%
+    %% \subfigure[2D, RCM.]{%
+    %%   \label{fig:rbf2drcm}
+    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
+    \subfigure[3D, RCM.]{%
+      \label{fig:rbf3dnorcm} 
+      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
+    \subfigure[3D, RCM.]{%
+      \label{fig:rbf3drcm}
+      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.png}}%
+  \end{center}
+
+  \caption{Different sparsity distributions. In all cases, there are
+    32 nonzeros per line and 300 rows. The last four matrices
+    corresponds to derivative stencils in 2D and 3D RBF-FD
+    calculations.}
+  \label{fig:spy_plots}
+\end{figure*}
+
+The different matrices used to perform our experiments and analysis
+are presented in Figure~\ref{fig:spy_plots}. The
+Supercompact matrices (Figure~\ref{fig:supercompact}) have been
+generated to only have nonzero elements in the first 32 columns of
+the matrix. As a result, only 32 values of {\tt x} are read
+from memory, and the expense associated with cache misses and gather
+operation is removed from consideration. The cost associated with 
+{\tt A} remains.   \todo{reread the above.}
+The Compact matrices
+(Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
+centered around the diagonal and it represents the ideal case for many
+applications that rely on sparse matrix vector multiplication. The nonzero
+elements of the Random
+matrices (Figure~\ref{fig:compact}) see their nonzero elements
+randomly (uniformly) distributed in the matrix; they represent the
+worst case scenario for a cache based architecture where the cache
+reutilisation is the lowest from one row to the next.
+
+The other two types of matrices are used in derivative stencils in 3D
+RBF-FD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-points
+stencil of a 3D $8^3$ grid). We also apply a Cuthill-Mckee reordering to
+all the matrices.  This ordering technique is designed to reduce the
+distance between the nonzeros and the diagonal, hopefully allowing a
+better cache reutilisation. The reordered version of the matrix can be
+seen in Figure~\ref{fig:rbf3drcm}.
 

### 2013-10-18 ieee_paper.tex
description of the host can go out of bandwidth
performance section

@@
 
+In this experimental section, as a point of reference we also provide
+some results on a typical CPU architecture. This system is
+...\todo{include a short description of the host.}  All the codes are
+compiled with icc xxx and -O3 optimization. The MIC codes use the IMCI
+instructions while the CPU codes use AVX instructions.
+
 \subsection{Bandwidth}
-
-In this experimental section, as a point of reference we provide some
-results on a typical CPU architecture. This system is ...\todo{include
-  a short description of the host.}  All the codes are compiled with
-icc xxx and -O3 optimization. The MIC codes use the IMCI instructions
-while the CPU codes use AVX instructions.
 
@@
 \subsection{Computations}
-%----------------------
-Scaling results as a function of the number of processors, on the MIC and on the host. The host calculations
-are performed for reference only, to bring out the dominance of the MIC due to its advanced swizzling and channel 
-operations, not available on the host. The performance on the host is 3x slower than on the MIC. The large cache
-of the host alleviates somewhat the lower memory bandwidth, but this only reduces the cost of transfer {\tt x}. 
-The cost memory cost for {\tt A} and {\tt col\_id} is unaffected by the larger cache size.
-
-The performance on the MIC scales linearly with the number of threads.
-We compare three cases, each with and without bandwidth reduction using Reverse Cuthil-McKee (RCM).
-There are $96^3$ RBF nodes, leading to a matrix with 884,736 rows.
-The four source vectors occupy more than 14M bytes, which exceeds the aggregate L2 cache of 31 Mbytes (7.8 Mfloats.)
-
-\todo{Is the host architecture described? Problem: I had to change hosts a few weeks ago, and it has different characteristics}. 
-%----------------------
+
+\begin{figure*}[t]
+  \centering
+  \subfigure[Supercompact and Random]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
+  \subfigure[Compact and RBF-FD]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
+
+  \caption{Performance of "best code" accross matrix types. The two
+    horizontal lines depict worst and best case performance
+    (Figure~\ref{fig:gflops-peak-perf}). RCM (blue, $\blacktriangle$)
+    and no RCM (red,$\blacktriangledown$).\todo{can we reorder the
+      legend to give supercompact first in (a) and compact first in
+      (b)?}}
+  \label{fig:expe_types}
+\end{figure*}
+
+We now investigate the actual performance that we can obtain when
+multiplying 4 vectors by 4 matrices in single
+precision. Figure~\ref{fig:expe_types} present results for different
+types of matrices and also shows the minimum and maximum performance
+predicted in Section~\ref{sec:model}. Figure~\ref{fig:perf_ideal}
+gives the results for the {\tt supercompact} and {\tt random} matrices
+which represent both the best and the worst case for such a
+computation. The {\tt supercompact} case peaks at
+205Gflop/s\todo{check that} being very close to the predicted peak
+performance of 213Gflop/s. Conversely the performance of the {\tt
+  random} case decreases to 59Gflop/s\todo{check} which is close to
+the lowest predicted performance of 55Gflop/s. One can see that RCM
+ordering helps the {\tt random} case but the impact decreases when the
+size of the matrix increase.
+
+Figure~\ref{fig:perf_realistic} gives the results on the {\tt compact}
+case which represents the best realistic matrix one could find and in
+an {\tt rbf} matrix extracted from a real application. In the {\tt
+  compact} case, the performance can be as high as 195Gflop/s being
+within 15\% of the peak predicted performance. The performance of the
+{\tt rbf} case varies between 100Gflop/s and 140Gflop/s. Most of the
+time, the RCM ordering provide an improvement which can be as high as
+30Gflop/s.
 
@@
     translate in different performance.\todo{this two plots should ``correlate'' }}
+  \label{fig:comparison_spmv}
 \end{figure*}
 
-\begin{figure*}[t]
-  \centering
-  \subfigure[Supercompact and Random]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}}
-  \subfigure[Compact and RBF-FD]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}}
-
-  %\subfigure[Supercompact]{\includegraphics[width=.48\linewidth]{figures/supercompact_max_perf.pdf}}
-  %\subfigure[Compact]{\includegraphics[width=.48\linewidth]{figures/compact_max_perf.pdf}} \\
-  %\subfigure[Random]{\includegraphics[width=.48\linewidth]{figures/random_max_perf.pdf}}
-  %\subfigure[RBF 3D]{\includegraphics[width=.48\linewidth]{figures/rbf_max_perf_run1.pdf}}
-  \caption{Performance of "best code" accross matrix types. The two horizontal lines depict worst and 
-  best case performance (Figure fig:gflops-peak-perf). RCM (blue, $\blacktriangle$) and no RCM (red,$\blacktriangledown$).}
-\end{figure*}
-
-Derivatives in an RBFFD formulation are expressed as a sparse
-matrix/vector multiplication (SpMV). Using the full 61 nodes of a MIC
-processor, and all four threads of each node, we achieve lowly 18
-Gflops when calculating a single vector of a single
-function. Bandwidth reduction has minimal effect since the cache is
-large enough to store both the derivative matrix and the solution
-vector. We increase the number of computations per byte transferred by
-calculating four different derivatives of four different
-functions. With bandwidth reduction, we speed the calculation by a
-factor of 8, and achieve upwards of 140 Gflops.
+We finally investigate the central questions of this paper. Do we gain
+actual performance by transforming classical SpMV computation into the
+multiplication of 4 vectors by 4 matrices for the computation of the
+derivative of RBFs? Does using manual vectorization improve actual
+application performance? Figure~\ref{fig:comparison_spmv} compares the
+performance achieved by a classical SpMV and by the multiplication of
+4 vectors by 4 matrices using either standard C++ code and our
+optimized implementation. On MIC (Figure~\ref{fig:perf_mic}) the
+classical SpMV computation reaches 15Gflop/s\todo{check}. The standard
+C++ implementation reaches a performance of 40Gflop/s while our
+optimized implementation reaches 137Gflop/s\todo{check}. Notice that
+RCM ordering has no impact in the standard C++ implementation while it
+provides an significant improvement in our manually vectorized
+version. This highlights the C++ implementation is instruction bound
+while our manually vectorized implementation is most likely memory
+bound.  As a point of comparison, we show an AVX implementation on the
+CPU of the two kernels and we show the best achievable performance in
+the classical SpMV in Figure~\ref{fig:perf_host}. \todo{waiting for
+  complete plot before writing more.}
+
+%% Scaling results as a function of the number of processors, on the MIC
+%% and on the host. The host calculations are performed for reference
+%% only, to bring out the dominance of the MIC due to its advanced
+%% swizzling and channel operations, not available on the host. The
+%% performance on the host is 3x slower than on the MIC. The large cache
+%% of the host alleviates somewhat the lower memory bandwidth, but this
+%% only reduces the cost of transfer {\tt x}.  The cost memory cost for
+%% {\tt A} and {\tt col\_id} is unaffected by the larger cache size.
+
+%% The performance on the MIC scales linearly with the number of
+%% threads. We compare three cases, each with and without bandwidth
+%% reduction using Reverse Cuthil-McKee (RCM). There are $96^3$ RBF
+%% nodes, leading to a matrix with 884,736 rows. The four source vectors
+%% occupy more than 14M bytes, which exceeds the aggregate L2 cache of 31
+%% Mbytes (7.8 Mfloats.)
+
+%% \todo{Is the host architecture described? Problem: I had to change
+%%   hosts a few weeks ago, and it has different characteristics}.
+%% %----------------------
+
+%% Derivatives in an RBFFD formulation are expressed as a sparse
+%% matrix/vector multiplication (SpMV). Using the full 61 nodes of a MIC
+%% processor, and all four threads of each node, we achieve lowly 18
+%% Gflops when calculating a single vector of a single
+%% function. Bandwidth reduction has minimal effect since the cache is
+%% large enough to store both the derivative matrix and the solution
+%% vector. We increase the number of computations per byte transferred by
+%% calculating four different derivatives of four different
+%% functions. With bandwidth reduction, we speed the calculation by a
+%% factor of 8, and achieve upwards of 140 Gflops.
 

### 2013-10-18 ieee_paper.tex
comment
shortening conclusion

@@
 where the elements of $y$ are grouped in sets of 16, which correspond
-to four derivatives applied to four functions, at a single point.
+to four derivatives applied to four functions, at a single point. \todo{this description is innacurate.}
 
@@
       legend to give supercompact first in (a) and compact first in
-      (b)?}}
+      (b)?}}\todo{is there a symbol/color collision?}
   \label{fig:expe_types}
@@
 In the previous sections, we have explored in the practical
-implementation on a MIC architecture of multiple derivative operators
-acting on multiple vectors within the context of RBF-RD. Each
-derivative has an associated sparse matrix with a fixed number of
-nonzeros on each row. While computing a single derivative of multiple
-functions is rather common (vectorization occurs over the vectors), we
-accelerate the algorithm further by considering multiple derivatives
-with corresponding matrices with identical adjacency graphs. We
-specialize to four matrices and four derivatives, with 16 outputs,
-computed as a sum of outer products.
-
-Using a matrix with $64^3$ (or $96^3$ rows, we achieve a speedup of
-xxx relative to the base case of one derivative and one function (1/1
-case), relative to a potential speedup of 16. We have achieved a
-speed of 100 Gflop/s on 60 cores using 240 threads, or a speed of 61\% of
-maximum possible performance. The maximum possible speed for the 1/1
-case is around 40 Gflop/s, of which we achieve 20 Gflop/s. On matrices
-that exceed the cache, a reverse Cuthill-McKee is applied to reduce the
-bandwidth, with an associated speedup of 30\% up to 135 Gflop/s. Our
-optimal implementation makes use of the (as yet unreleased) IMCI MIC instruction set 
-%\todo{that's not AVX on MIC, not sure what the name is}, 
-which includes a number of swizzling and channel swapping operations, for an
-extremely efficient tensor product implementation. A straightforward
-implementation without vectorization runs at around 35 Gflops, or 16\% of peak
-performance. RCM has no effect because the number of floating point
-operations is excessive (the algorithm does not fill the vector
-registers to capacity).
-
-In a future work, we will examine the effect of larger stencil sizes,
-double precision, and the actual cost of computing the derivatives
-within a fluid simulation using RBF-FD.
+implementation on a Intel Xeon Phi card of multiple derivative
+operators acting on multiple vectors within the context of
+RBF-RD. Each derivative has an associated sparse matrix with a fixed
+number of nonzeros on each row. While computing a single derivative of
+multiple functions is rather common (vectorization occurs over the
+vectors), we accelerate the algorithm further by considering multiple
+derivatives with corresponding matrices with identical adjacency
+graphs. We specialize our study to four matrices and four derivatives,
+with 16 outputs, computed as a sum of outer products.
+
+Our implementation makes use of the IMCI MIC instruction set and
+includes a number of swizzling and channel swapping operations, for an
+extremely efficient tensor product implementation. On a $96^3$ 3D
+grid, it reaches a performance of 140 Gflop/s which is 2.7 times
+faster than the best possible performance achievable for a single
+vector and a single derivative and more than 7 times faster than a
+practical implementation.
+
+%% Using a matrix with $96^3$ rows, we achieve a speedup of 7.7 relative
+%% to the base case of one derivative and one function (1/1 case). We
+%% have achieved a speed between 100 Gflop/s to 140 Gflop/s on 60 cores
+%% using 240 threads, or a speed of 47\% to 66\% of the maximum possible
+%% performance of this architecture. As a point of comparison, the
+%% maximum possible speed for the 1/1 case is around 36 Gflop/s, of which
+%% we achieve 18 Gflop/s. Our optimal implementation makes use of the
+%% IMCI MIC instruction set which includes a number of swizzling and
+%% channel swapping operations, for an extremely efficient tensor product
+%% implementation. A straightforward implementation without vectorization
+%% runs at around 35 Gflops, or 16\% of peak performance.
+
+In a future work, we will examine the effect of larger stencil sizes
+and of using double precision. We will also integrate the techniques
+presented in this paper within a fluid simulation using RBF-FD.
 
 \section*{Acknowledgment}
-The first and last authors acknowledge NSF funding under NSF grant DMS-\#0934331 (FSU). 
-NCAR is sponsored by the National Science Foundation. N. Flyer      
-acknowledges support of NSF grant DMS-0934317.
+The first and last authors acknowledge NSF funding under NSF grant
+DMS-\#0934331 (FSU). NCAR is sponsored by the National Science
+Foundation. N. Flyer acknowledges support of NSF grant DMS-0934317.
 

### 2013-10-18 ieee_paper.tex
comment

@@
 \begin{figure*}
-  \centering
-  \subfigure[Performance of $y=Ax$ on the MIC. Squares: base 1/1 case,
-    solid circles: 4/4 case implemented in C++, solid triangles: 4/4
-    case implemented with MIC vector instructions. $-O3$ compilation
-    options. Each symbol has two cases: with (red) and without (blue) 
-    Reverse Cuthill-McKee bandwidth reduction. The grid is $96^3$.]
+  \centering \subfigure[Performance of $y=Ax$ on the MIC. Squares:
+    base 1/1 case, solid circles: 4/4 case implemented in C++, solid
+    triangles: 4/4 case implemented with MIC vector instructions. Each
+    symbol has two cases: with (red) and without (blue) Reverse
+    Cuthill-McKee bandwidth reduction. The grid is $96^3$]
              {\includegraphics[width=.40\linewidth]{figures/mic_performance_nb_threads.pdf}\label{fig:perf_mic}}
@@
   \caption{Vectorization matters in practice. Differences in BW
-    translate in different performance.\todo{this two plots should ``correlate'' }}
+    translate in different performance.\todo{this two plots should ``correlate'' }\todo{legend}}
   \label{fig:comparison_spmv}

### 2013-10-18 ieee_paper.tex
shrinking section 1 2 3

@@
 \label{sec:rbf}
-
-%In the theory of Radial Basis Functions Finite-Difference (RBFFD),
-%derivatives of a function $f(\rvec)$ at node $i$ are expressed as a
-%linear combination of the function values at the stencil center and
-%the nodes connected to node $i$ (Figure~\ref{fig:rbf_stencils}). Thus
-%$y$ is a discrete derivative of the vector $x$.
 
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/rbf_stencils.pdf}
+  \includegraphics[width=.8\linewidth]{figures/rbf_stencils.pdf}
   \caption{RBFFD Stencils. Each node of the stencil is connected to
@@
   \begin{center}
-    \scalebox{.9}{
+    \scalebox{.8}{
       \begin{tabular}{|c|l|}
@@
 
-Figure~\ref{fig:ratio_bytes_flops} shows how these ratios translate
+Figure~\ref{fig:perf_predict} shows how these ratios translate
 into actual performance. This figure provides projected best and worst
@@
 
-\begin{figure*}
+\begin{figure}
   \centering 
-  \subfigure[Absolute performance ]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}}
-%
-  \subfigure[Relative performance to using one matrix and one
-    vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}
-
-  \caption{Estimation of the maximum achievable performance 
-    using a device with a memory to computational units of 150GB/s
-    when varying the number of vectors and matrices. \todo{we could
-      remove the speedup plot if we need space.}}
+  \includegraphics[width=.9\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}
+  \caption{Estimation of the maximum achievable performance using a
+    device with a memory to computational units of 150GB/s when
+    varying the number of vectors and matrices.}
   \label{fig:perf_predict}
-\end{figure*}
+\end{figure}
 

### 2013-10-18 ieee_paper.tex
shrking sect 3

@@
 features of the vectors in the Intel MIC architecture. A vector of 512
-bits is composed of four lanes (sometimes called channels) of 128 bits, 
-which are made of four segments of 32 bits. A swizzling operation allows
-the reorder of the elements within each lane, while a permutation reorders
-entire lanes, without changing their contents. 
-Figure~\ref{fig:swizzling_permutation}
-illustrates these two possibilities. Both permutation and swizzling
-support common reordering and broadcast. In the code, the ``permute''
-function needs to use the {\tt \_mm512\_permute4f128\_epi32}
-instruction which is only defined on integers and therefore in place typecasts
-are necessary to convert to a vector to/from the integer type.
-\todo{I added a comment stating we preferred swizzling.}
-(Note that we prefer swizzling to permutation if possible because we believe 
-it is cheaper. Many instructions have swizzling capability, while permuation
-is its own instruction. \todo{do we need a reference to my note?})
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/swizzling_lanes.pdf}
-  \caption{Swizzling enables the reorder or broadcast of elements within
-    each 128 bit group. Permutation enables to reorder or broadcast
-    whole 128 bit groups. All lanes must be sizzeled identically.
-    \todo{I added to the caption.}}
-  \label{fig:swizzling_permutation}
-\end{figure}
+bits is composed of four lanes (sometimes called channels) of 128
+bits, which are made of four segments of 32 bits. A swizzling
+operation allows the reorder of the elements within each lane (see how
+{\tt mat} is extracted from {\tt mat\_ent} in
+Figure~\ref{code:mat_mul}). While a permutation reorders entire lanes,
+without changing their contents (see for instance how {\tt vect} is
+generated from {\tt all\_vect} in Figure~\ref{code:mat_mul}).  Both
+permutation and swizzling support common reordering and broadcast. In
+the code, the ``permute'' function needs to use the {\tt
+  \_mm512\_permute4f128\_epi32} instruction which is only defined on
+integers and therefore in place typecasts are necessary to convert to
+a vector to/from the integer type.  (Note that we prefer swizzling to
+permutation if possible because we believe it is cheaper. Many
+instructions have swizzling capability, while permuation is its own
+instruction. \todo{do we need a reference to intel Xeon phi instruction set?})
 

### 2013-10-18 ieee_paper.tex
shrinking section 3 more

@@
 OpenCL or CUDA, one can think of these registers as a warp.) The end
-goal is to load and format the nonzero values $a,b,c,d$
-and the vector entries $p,q,r,s$ in two vector registers and 
-apply Fused Multiply-Add on them, which is depicted in
-Figure~\ref{fig:tensor_product}. Bringing the data into the vector
+goal is to load and format the nonzero values $A_h^1,A_h^2,A_h^3,A_h^4$
+and the vector entries $x_m^1,x_m^2,x_m^3,x_m^4$ in two vector registers and 
+apply Fused Multiply-Add on them, which is depicted in the {\tt accu} line of 
+Figure~\ref{code:mat_mul}. Bringing the data into the vector
 register as depicted in the figure is the difficult component of our 
@@
 \end{figure*}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/tensor_product.pdf}
-  \caption{Structure of the vector registers to perform an efficient
-    tensor scalar product. See Figure~\ref{fig:mat_struct} for
-    reference.}
-  \label{fig:tensor_product}
-\end{figure}
 
@@
 Order hint to optimize write traffic on the memory bus.
-\todo{I added "write" to traffic }
 

### 2013-10-18 ieee_paper.tex
cutting

@@
 better understand our results, we first perform some bandwidth
-measurements, followed by the actual SpMV computations.
-
-In this experimental section, as a point of reference we also provide
-some results on a typical CPU architecture. This system is
-...\todo{include a short description of the host.}  All the codes are
-compiled with icc xxx and -O3 optimization. The MIC codes use the IMCI
-instructions while the CPU codes use AVX instructions.
+measurements, followed by the actual SpMV computations. All the codes
+are compiled with the Intel C Compiler version xxx\todo{gordon?} and
+-O3 optimization. The MIC codes use the IMCI instructions exposed by
+the compiler through intrinsics.
 
@@
 
-\begin{figure*}[bth]
+\begin{figure}[bth]
   \centering
-  \subfigure[On MIC with 240 threads, 64-byte memory
-    alignment.]{\includegraphics[width=.40\linewidth]{figures/bandwidth_read_write.pdf}\label{fig:band_rw_MIC}}
-  %
-  \subfigure[On CPU with 32 threads, 32-byte memory
-    alignment.]{\includegraphics[width=.40\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:band_rw_CPU}}
-
-  \caption{Bandwidth performance on the host under idealized
+  \includegraphics[width=.9\linewidth]{figures/bandwidth_read_write.pdf}
+
+  \caption{On MIC with 240 threads, 64-byte memory
+    alignment. Bandwidth performance under idealized
     conditions as a function of matrix row size.  Entries with "cpp"
     denote cases where coding was performed without MIC vector
-    instructions.  \todo{Can we label the x-axis "number of floats" or
+    instructions. \todo{legend}  \todo{Can we label the x-axis "number of floats" or
       "size in MB"?}\todo{Can we use the same symbols in both charts?}
@@
   \label{fig:bandwidth}
-\end{figure*}
+  \label{fig:band_rw_MIC}
+\end{figure}
 
@@
 
-Looking at the results on the host in Figure~\ref{fig:band_rw_CPU}
-show no significant difference between manual vectorization and the
-compiler generated one (there is no store nrngo on the host). It also
-shows quite stable performance with about 70GB/s in the {\tt read}
-benchmark, about 40GB/s in the {\tt write} benchmark and about 49 in
-the {\tt read write} benchmark. The bandwidth achieved on MIC is about
-2 to 3 times faster than the bandwidth achieved on the CPU. The
-kernels we consider being memory bound, we expect MIC to be 2 to 3
-times faster.
-
 Note that SpMV kernels will likely use only simple read and simple
@@
 maximize bandwidth can reach higher
-values~\cite{Saule13-ARXIV}. Comparing to a CPU architecture, we
-expect MIC to be 2 to 3 times faster.
+values~\cite{Saule13-ARXIV}.
 
@@
 
-\begin{figure*}
-  \centering \subfigure[Performance of $y=Ax$ on the MIC. Squares:
+\begin{figure}
+  \centering
+
+  \includegraphics[width=.9\linewidth]{figures/mic_performance_nb_threads.pdf}
+
+  \caption{Performance of $y=Ax$ on the MIC. Squares:
     base 1/1 case, solid circles: 4/4 case implemented in C++, solid
@@
     symbol has two cases: with (red) and without (blue) Reverse
-    Cuthill-McKee bandwidth reduction. The grid is $96^3$]
-             {\includegraphics[width=.40\linewidth]{figures/mic_performance_nb_threads.pdf}\label{fig:perf_mic}}
-%
-\hspace{.4in}
-%
-  \subfigure[Performance on the host on a sparse matrix with $64^3$ rows. Manual
-    vectorization brings significant improvement there as
-    well.]{\includegraphics[width=.40\linewidth]{figures/64x64x64.pdf}\label{fig:perf_host}}
-  \caption{Vectorization matters in practice. Differences in BW
-    translate in different performance.\todo{this two plots should ``correlate'' }\todo{legend}}
+    Cuthill-McKee bandwidth reduction. The grid is $96^3$. Vectorization matters in practice. Differences in BW
+    translate in different performance.\todo{legend} \todo{fix caption}}
   \label{fig:comparison_spmv}
-\end{figure*}
+  \label{fig:perf_mic}
+\end{figure}
 
@@
 4 vectors by 4 matrices using either standard C++ code and our
-optimized implementation. On MIC (Figure~\ref{fig:perf_mic}) the
+optimized implementation. The
 classical SpMV computation reaches 15Gflop/s\todo{check}. The standard
@@
 while our manually vectorized implementation is most likely memory
-bound.  As a point of comparison, we show an AVX implementation on the
-CPU of the two kernels and we show the best achievable performance in
-the classical SpMV in Figure~\ref{fig:perf_host}. \todo{waiting for
-  complete plot before writing more.}
+bound.
 

### 2013-10-18 ieee_paper.tex
description of CPU case

@@
 
+To put things in perspective, we conducted some experiments on a
+classical CPU architecture with 2 Intel Core i7 processor, each has 8
+cores which support AVX instructions. Both the compiler generated and
+the manually written vectorization using AVX instruction reach a
+(read) bandwidth of 70GB/s. This indicates that the performance of the
+classical SpMV can not overcome 16.8Gflop/s. A generic implementation
+of the multiplication of 4 vectors with 4 matrices on the {\tt RBF}
+matrix reaches a performance of 27 Gflop/s; while a manual
+implementation using AVX instruction reaches a performance of 42
+Gflop/s. This confirms that our method is sound and generalizes to
+other architecture than the Intel MIC architecture.
+
 %% Scaling results as a function of the number of processors, on the MIC

### 2013-10-18 ieee_paper.tex
my changes.
Please enter the commit message for your changes. Lines starting

--- 2013-10-18 ieee_paper.tex
@@
 implementation that is independent of dimensionality. RBFs approximate
-a function $f(\xvec)\subset \mathbb{R}^d$ sampled at a set of $N$
+a function $f(\rvec)\subset \mathbb{R}^d$ sampled at a set of $N$
 distinct point locations, $x_j$, by linearly combining translates of a
 single radially symmetric function $\phi(r)$, where $r =
-\|\xvec-\xvec_{j}\|$ denotes the Euclidean distance (e.g., in 2-D
-$\sqrt{(x-x_j)^2+(y-y_j)^2}$) between where the function is evaluated
-$\xvec$ and where the RBF is centered $\xvec_{j}$. That is, the
-interpolant is $s(\xvec) = \sum_{j=1}^{N} w_j
-\phi_i(\|\xvec-\xvec_{j}\|)$. The weights $w_j$ are obtained by
+\|\rvec-\rvec_{j}\|$ denotes the Euclidean distance %(e.g., in 2-D
+%$\sqrt{(x-x_j)^2+(y-y_j)^2}$) 
+between where the function is evaluated
+$\rvec$ and where the RBF is centered $\rvec_{j}$. That is, the
+interpolant is $s(\rvec) = \sum_{j=1}^{N} w_j
+\phi_i(\|\rvec-\rvec_{j}\|)$. The weights $w_j$ are obtained by
 inverting the system
@@
 =
-\parray{c}{ f(\xvec_1) \\ \vdots\\ f(\xvec_N) }. 
+\parray{c}{ f(\rvec_1) \\ \vdots\\ f(\rvec_N) }. 
 \label{eq:rbf}
 \end{equation}
-where $\phi_{ij} = \phi(\|\xvec_i-\xvec_j\|)$. 
+where $\phi_{ij} = \phi(\|\rvec_i-\rvec_j\|)$. 
 %\end{equation}
@@
 desired analytic derivative operator $L$ to the RBF interpolant
-$s(\xvec)$ above and evaluating it at the point locations. For very
+$s(\rvec)$ above and evaluating it at the point locations. For very
 large problems, this is a computationally expensive since the matrix
@@
 every node in the domain to calculate the derivative at a given node
-$\xvec_i$), RBF-generated finite differences (RBF-FD) have been
+$\rvec_i$), RBF-generated finite differences (RBF-FD) have been
 derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBF-FD uses only
-a local set of the $n_z-1$ nearest neighbors to the point $\xvec_i$ to
+a local set of the $n_z-1$ nearest neighbors to the point $\rvec_i$ to
 approximate the derivative. In other words,
-$Lf(\xvec_i)=\sum_{j=1}^{n_z}a_jf(\xvec_j)$. The differentiation
+$Lf(\rvec_i)=\sum_{j=1}^{n_z}a_jf(\rvec_j)$. The differentiation
 weights, $a_j$, are calculated by enforcing that this linear
 combination should be exact for RBFs,
-$\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the
-node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences
+$\{\phi(\|\rvec-\rvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the
+node locations $\{\rvec_j\}_{j=1}^{n_z}$ (classical finite differences
 (FD) would enforce that it be exact for polynomials instead). Similar
@@
 study to four vectors and four matrices. 
-\todo{I have remove reference to $x$, $y$, ... to avoid confusion with what follows.}
+%\todo{I have remove reference to $x$, $y$, ... to avoid confusion with what follows.}
 
-\todo{Add few sentence with discretization. Check Erik}
+5\todo{Add few sentence with discretization. Check Erik}
 In discrete form, Equation (\ref{eq:Q}) takes the form
@@
 large. But the differences are important when the number of vectors is
-low: this highlights that accessing the memory per batch \todo{what do you mean by 
-accessing the memory per batch} is the main
-problem faced when performing a standard SpMV computation.
+low: this highlights that accessing the memory with a granularity of one cache
+line %a cache line basis batch \todo{what do you mean by 
+%accessing the memory per batch} 
+is the main problem faced when performing a standard SpMV computation.
 

### 2013-10-18 ieee_paper.tex
moving the memory layout paragraph

--- 2013-10-18 ieee_paper.tex
@@
 
-In term of memory layout, we adopt a format similar to the ELLPACK
-format to store the matrix since the number of nonzeros per row is
-constant throughout the matrix. The matrix is given in two arrays. The
-first one is the classical {\tt col\_id} array that describes the
-sparsity structure of the matrix. Each row of {\tt col\_id} consists of $n_z$
-entries that label the column index of each nonzero in $A$. The
-other array is the {\tt data} array that gives the values of the 
-nonzeros in the four matrices. The values of the matrices are interleaved
-so that the value $A^l_{col\_id[k]}$ of the $k$th nonzero of the
-$l$th matrix is at index $(k-1)*4+l-1$. The vectors {\tt x} are 
-interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
-consecutive in memory. The vector {\tt y} is similarly
-interleaved. This is schematically shown in Figure~\ref{fig:mat_struct}, 
-where the elements of $y$ are grouped in sets of 16, which correspond
-to four derivatives applied to four functions, at a single point. \todo{this description is innacurate.}
-
 \begin{figure}
@@
 
+In term of memory layout, we adopt a format similar to the ELLPACK
+format to store the matrix since the number of nonzeros per row is
+constant throughout the matrix. The matrix is given in two arrays. The
+first one is the classical {\tt col\_id} array that describes the
+sparsity structure of the matrix. Each row of {\tt col\_id} consists of $n_z$
+entries that label the column index of each nonzero in $A$. The
+other array is the {\tt data} array that gives the values of the 
+nonzeros in the four matrices. The values of the matrices are interleaved
+so that the value $A^l_{col\_id[k]}$ of the $k$th nonzero of the
+$l$th matrix is at index $(k-1)*4+l-1$.  \todo{this description is innacurate.} The vectors {\tt x} are 
+interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
+consecutive in memory. The vector {\tt y} is similarly
+interleaved, grouped in sets of 16, which correspond
+to four derivatives applied to four functions, at a single point.
+
 We discuss how the implementation of the v4m4 kernel works using

### 2013-10-18 ieee_paper.tex
changes.

--- 2013-10-18 ieee_paper.tex
@@
 between where the function is evaluated
-$\rvec$ and where the RBF is centered $\rvec_{j}$. That is, the
+$\rvec$ and where the RBF is centered at $\rvec_{j}$. That is, the
 interpolant is $s(\rvec) = \sum_{j=1}^{N} w_j
@@
 points in the domain, one multiplies the source vector $x$ by the 
-derivative matrix $A_{n_z}$ to obtain the vector $y$ of derivative values, 
+derivative matrix $A$ to obtain the vector $y$ of derivative values, 
 $$
-  y = A_{n_z} x
+  y = A x
 $$
-computation of a single derivative has been reduced to a SpMV, 
+The computation of a single derivative has been reduced to a SpMV, 
 where each row has $n_z$ nonzeros\footnote{We use non-conventional 
@@
 
-5\todo{Add few sentence with discretization. Check Erik}
+%\todo{Add few sentence with discretization. Check Erik}
 In discrete form, Equation (\ref{eq:Q}) takes the form

### 2013-10-18 ieee_paper.tex
new fig 2
temp

@@
   \centering
-  \includegraphics[width=\linewidth]{figures/matrix_not.pdf}
-  \caption{Matrix Structure.}
+
+  \[ \left( \begin{array}{cccc}
+    y^{1,1} & y^{1,2} & y^{1,3} & y^{1,4} \\
+    y^{2,1} & y^{2,2} & y^{2,3} & y^{2,4} \\
+    y^{3,1} & y^{3,2} & y^{3,3} & y^{3,4} \\
+    y^{4,1} & y^{4,2} & y^{4,3} & y^{4,4} 
+  \end{array} \right)
+  = \left(
+  \begin{array}{c}
+    A^1\\
+    A^2\\
+    A^3\\
+    A^4
+  \end{array}\right)
+  \times \left(x^1 x^2 x^3 x^4 \right) \]
+
+%  \includegraphics[width=\linewidth]{figures/matrix_not.pdf}
+  \caption{Matrix of the computation.}
   \label{fig:mat_struct}
@@
 first one is the classical {\tt col\_id} array that describes the
-sparsity structure of the matrix. Each row of {\tt col\_id} consists of $n_z$
-entries that label the column index of each nonzero in $A$. The
-other array is the {\tt data} array that gives the values of the 
-nonzeros in the four matrices. The values of the matrices are interleaved
-so that the value $A^l_{col\_id[k]}$ of the $k$th nonzero of the
-$l$th matrix is at index $(k-1)*4+l-1$.  \todo{this description is innacurate.} The vectors {\tt x} are 
+sparsity structure of the matrix. Each row of {\tt col\_id} consists
+of $n_z$ entries that label the column index of each nonzero in
+$A$. The other array is the {\tt data} array that gives the values of
+the nonzeros in the four matrices. The values of the matrices are
+interleaved by group of four so that the value $A^l_{col\_id[k]}$ of
+the $k$th nonzero of the $l$th matrix is at index $(k-1)*4+l-1$.
+(In other words, the first 4 entries are from $A^1$ and the next 4 are from $A^2$) The vectors {\tt x} are
 interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
-consecutive in memory. The vector {\tt y} is similarly
-interleaved, grouped in sets of 16, which correspond
-to four derivatives applied to four functions, at a single point.
+consecutive in memory. The vector {\tt y} is similarly interleaved,
+grouped in sets of 16, which correspond to four derivatives applied to
+four functions, at a single point.
 

### 2013-10-18 ieee_paper.tex
matrix layout

@@
 \newcommand{\ceil}[1]{\left\lceil#1\right\rceil}
+\newcommand{\floor}[1]{\left\lfloor#1\right\rfloor}
 
@@
 interleaved by group of four so that the value $A^l_{col\_id[k]}$ of
-the $k$th nonzero of the $l$th matrix is at index $(k-1)*4+l-1$.
-(In other words, the first 4 entries are from $A^1$ and the next 4 are from $A^2$) The vectors {\tt x} are
+the $k$th nonzero of the $l$th matrix is at index $\floor{\frac{k-1}{4}}*4*n_m + (l-1)*4+(k-1)\%4$.
+(In other words, the first 4 entries are from $A^{1}$ and the next 4 are from $A^2$.) The vectors {\tt x} are
 interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are

### 2013-10-18 ieee_paper.tex
matrix layout

@@
 interleaved by group of four so that the value $A^l_{col\_id[k]}$ of
-the $k$th nonzero of the $l$th matrix is at index $\floor{\frac{k-1}{4}}*4*n_m + (l-1)*4+(k-1)\%4$.
+the $k$th nonzero of the $l$th matrix is at index $\floor{\frac{k-1}{4}}*4*n_m + (l-1)*4+(k-1)\%4$.\todo{check that formula}
 (In other words, the first 4 entries are from $A^{1}$ and the next 4 are from $A^2$.) The vectors {\tt x} are

### 2013-10-18 ieee_paper.tex
removing some todo

@@
 measurements, followed by the actual SpMV computations. All the codes
-are compiled with the Intel C Compiler version xxx\todo{gordon?} and
+are compiled with the Intel C Compiler version 13 and
 -O3 optimization. The MIC codes use the IMCI instructions exposed by
@@
     denote cases where coding was performed without MIC vector
-    instructions. \todo{legend}  \todo{Can we label the x-axis "number of floats" or
-      "size in MB"?}\todo{Can we use the same symbols in both charts?}
-    \todo{Can we change the color scheme? in (a), I do not see the
-      difference between "read\_cpp" and "read/write cpp"} }
+    instructions. \todo{legend} }
   \label{fig:bandwidth}

### 2013-10-18 ieee_paper.tex
removing some todo

@@
 load. One can load the 16 floating point values for four nonzeroes of
-the matrix at once. The four floats coming from each matrix four 4
-\todo{something wrong with sentence I am in. Did you discuss matrix ordering?}
-consecutive nonzeros are naturally grouped within each lane of the
-SIMD vector. One can use a swizzle operation to broadcast the 4
-elements of a single matrix to take the whole vector. (The broadcast of 
-$A_h^1,A_h^2,A_h^3,A_h^4$ is shown in Figure~\ref{code:mat_mul}.)
+the matrix at once. The four floats coming from each matrix are
+naturally grouped within each lane of the SIMD vector. One can use a
+swizzle operation to broadcast the 4 elements of a single matrix to
+take the whole vector. (The broadcast of $A_h^1,A_h^2,A_h^3,A_h^4$ is
+shown in Figure~\ref{code:mat_mul}.)
 

### 2013-10-18 ieee_paper.tex
GB/s -> ~GB/s; Gflop/s => ~Gflop/s

@@
 expected when computing the 16 multiplications simultaneously and
-reach a total of about 210 Gflop/s. This performance represents
+reach a total of about 210~Gflop/s. This performance represents
 approximatively 10\% of the available flop/s of a Knight Corner
@@
 for analysis purpose and some extracted from and RBF application. A
-performance of 135 Gflop/s is achieved on real scenario. Concluding
+performance of 135~Gflop/s is achieved on real scenario. Concluding
 remarks and perspectives are provided in Section~\ref{sec:ccl}.
@@
 Gflop/s achievable assuming the computation is memory-bound and the
-architecture reaches 150 Gflop/s. Notice that \cite{Saule13-ARXIV}
+architecture reaches 150~Gflop/s. Notice that \cite{Saule13-ARXIV}
 showed a higher peak bandwidth, but we will show in
-Section~\ref{sec:expe} why 150Gflop/s is a better estimate of what one
+Section~\ref{sec:expe} why 150~Gflop/s is a better estimate of what one
 might achieve in these kernels. In single precision, the worse that
@@
 than the best achievable using a classical SpMV computation. The best
-performance reachable is 210Gflop/s: almost six times higher than the
+performance reachable is 210~Gflop/s: almost six times higher than the
 peak of the classical SpMV case.
@@
   \caption{Estimation of the maximum achievable performance using a
-    device with a memory to computational units of 150GB/s when
+    device with a memory to computational units of 150~GB/s when
     varying the number of vectors and matrices.}
@@
 transactions per second. Each has two 32-bit channels, achieving a total
-bandwidth of 320GB/s aggregated across all the memory
+bandwidth of 320~GB/s aggregated across all the memory
 controllers. There are 60 cores clocked at 1.05GHz. Their memory
 interfaces are 32-bit wide with two channels and the total bandwidth
-is 8.4GB/s per core. Thus, the cores should be able to consume 504GB/s
+is 8.4~GB/s per core. Thus, the cores should be able to consume 504~GB/s
 at most. However, the bandwidth between the cores and the memory
@@
 the memory controller. Its precise bandwidth is unknown but it is
-believed to be between 200GB/s and 250GB/s.
+believed to be between 200~GB/s and 250~GB/s.
 
@@
 As explained in Section~\ref{sec:model}, we expect to achieve a performance of
-about 210Gflop/s in single precision in the best case. We will focus
+about 210~Gflop/s in single precision in the best case. We will focus
 on the single precision case with four vectors and four matrices. 
@@
 When using the MIC card (Figure~\ref{fig:band_rw_MIC}) the {\tt
-  read\_cpp} implementation reaches a performance of 160GB/s
+  read\_cpp} implementation reaches a performance of 160~GB/s
 \todo{gordon check exact number} which is better than the {\tt read}
-implementation which only achieves 145GB/s \todo{and that one}. This
+implementation which only achieves 145~GB/s \todo{and that one}. This
 is explained by the compiler adding loop unrolling and software
@@
 not added in the {\tt read} case. On the other hand, the {\tt write}
-case reaches a performance of 155GB/s\todo{and that one} while the
-{\tt write\_cpp} implementation never passes 55GB/s. This difference
+case reaches a performance of 155~GB/s\todo{and that one} while the
+{\tt write\_cpp} implementation never passes 55~GB/s. This difference
 is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction
@@
 useful for some parallel algorithms. The {\tt read/write} benchmark
-performs at 140GB/s while the {\tt read/write cpp} benchmark performs
-at about 90GB/s.
+performs at 140~GB/s while the {\tt read/write cpp} benchmark performs
+at about 90~GB/s.
 
@@
 present the results in Figure~\ref{fig:band_gather}. The {\tt unpack}
-instruction can reach a performance of 140GB/s. (There is no easy way
+instruction can reach a performance of 140~GB/s. (There is no easy way
 to implement the equivalent of the {\tt unpack} operation without
@@
 operations depends in which order the array is read. If the indices
-are ordered ({\tt compact}) then a performance of 140GB/s is
+are ordered ({\tt compact}) then a performance of 140~GB/s is
 reached. If the indices are completely randomized, the performance
-drops to 2GB/s\todo{correct?}. We can see that implementing the
+drops to 2~GB/s\todo{correct?}. We can see that implementing the
 indirection mechanism without vector instructions makes the
 performance drop significantly. (In the {\tt compact} case, it never
-passes 100GB/s and is close to 50GB/s most of the time.)
+passes 100~GB/s and is close to 50~GB/s most of the time.)
 
@@
 in a typical SpMV computation, one can expect a bandwidth between
-140GB/s to 160GB/s. That is why we set in our estimation the best
-achievable bandwidth to 150GB/s despite carefully crafted code to
+140~GB/s to 160~GB/s. That is why we set in our estimation the best
+achievable bandwidth to 150~GB/s despite carefully crafted code to
 maximize bandwidth can reach higher
@@
 computation. The {\tt supercompact} case peaks at
-205Gflop/s\todo{check that} being very close to the predicted peak
-performance of 213Gflop/s. Conversely the performance of the {\tt
-  random} case decreases to 59Gflop/s\todo{check} which is close to
-the lowest predicted performance of 55Gflop/s. One can see that RCM
+205~Gflop/s\todo{check that} being very close to the predicted peak
+performance of 213~Gflop/s. Conversely the performance of the {\tt
+  random} case decreases to 59~Gflop/s\todo{check} which is close to
+the lowest predicted performance of 55~Gflop/s. One can see that RCM
 ordering helps the {\tt random} case but the impact decreases when the
@@
 an {\tt rbf} matrix extracted from a real application. In the {\tt
-  compact} case, the performance can be as high as 195Gflop/s being
+  compact} case, the performance can be as high as 195~Gflop/s being
 within 15\% of the peak predicted performance. The performance of the
-{\tt rbf} case varies between 100Gflop/s and 140Gflop/s. Most of the
+{\tt rbf} case varies between 100~Gflop/s and 140~Gflop/s. Most of the
 time, the RCM ordering provide an improvement which can be as high as
-30Gflop/s.
+30~Gflop/s.
 
@@
 optimized implementation. The
-classical SpMV computation reaches 15Gflop/s\todo{check}. The standard
-C++ implementation reaches a performance of 40Gflop/s while our
-optimized implementation reaches 137Gflop/s\todo{check}. Notice that
+classical SpMV computation reaches 15~Gflop/s\todo{check}. The standard
+C++ implementation reaches a performance of 40~Gflop/s while our
+optimized implementation reaches 137~Gflop/s\todo{check}. Notice that
 RCM ordering has no impact in the standard C++ implementation while it
@@
 the manually written vectorization using AVX instruction reach a
-(read) bandwidth of 70GB/s. This indicates that the performance of the
-classical SpMV can not overcome 16.8Gflop/s. A generic implementation
-of the multiplication of 4 vectors with 4 matrices on the {\tt RBF}
-matrix reaches a performance of 27 Gflop/s; while a manual
-implementation using AVX instruction reaches a performance of 42
-Gflop/s. This confirms that our method is sound and generalizes to
+(read) bandwidth of 70~GB/s. This indicates that the performance of
+the classical SpMV can not overcome 16.8~Gflop/s. A generic
+implementation of the multiplication of 4 vectors with 4 matrices on
+the {\tt RBF} matrix reaches a performance of 27~Gflop/s; while a
+manual implementation using AVX instruction reaches a performance of
+42~Gflop/s. This confirms that our method is sound and generalizes to
 other architecture than the Intel MIC architecture.
@@
 extremely efficient tensor product implementation. On a $96^3$ 3D
-grid, it reaches a performance of 140 Gflop/s which is 2.7 times
+grid, it reaches a performance of 140~Gflop/s which is 2.7 times
 faster than the best possible performance achievable for a single
@@
 practical implementation.
-
-%% Using a matrix with $96^3$ rows, we achieve a speedup of 7.7 relative
-%% to the base case of one derivative and one function (1/1 case). We
-%% have achieved a speed between 100 Gflop/s to 140 Gflop/s on 60 cores
-%% using 240 threads, or a speed of 47\% to 66\% of the maximum possible
-%% performance of this architecture. As a point of comparison, the
-%% maximum possible speed for the 1/1 case is around 36 Gflop/s, of which
-%% we achieve 18 Gflop/s. Our optimal implementation makes use of the
-%% IMCI MIC instruction set which includes a number of swizzling and
-%% channel swapping operations, for an extremely efficient tensor product
-%% implementation. A straightforward implementation without vectorization
-%% runs at around 35 Gflops, or 16\% of peak performance.
 

### 2013-10-18 gflops_speedup.py
Changes

@@ -79,8 +79,8 @@ max_bandwidth = 150
 bandwidths = [150,190]
 nv = [1,4]
 nm = [1,4]
-labels = [(v+m) for v in ["v1","v4"] for m in ["m1","m4"]]
-labels.extend(["v16m1"])
+labels = [(v+','+m) for v in ["nv=1","nv=4"] for m in ["nm=1","nm=4"]]
+labels.extend(["nv=16,nm=1"])
 print labels
 # v loop is slowest
 # m loop is fastest
@@ -122,8 +122,8 @@ print ratios_s_best
 rects2 = plt.bar(index+2*bar_width, ratios_d_worst, bar_width, alpha=1, color='r', label='double precision worst case')
 rects2 = plt.bar(index+3*bar_width, ratios_d_best, bar_width, alpha=.5, color='r', label='double precision best case')
 
-plt.xlabel('Vector/Matrix sizes', fontsize=14)
-plt.ylabel('Peak Gflops', fontsize=14)
+plt.xlabel('Vector/Matrix sizes', fontsize=16)
+plt.ylabel('Peak Gflops', fontsize=16)
 #plt.title('Best possible performance (max application bandwidth: 150 Gbytes/sec')
 plt.xticks(index + bar_width, labels)
 plt.legend(loc=2)  # upper left

### 2013-10-18 ieee_paper.tex
caption

@@
       \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
-    \subfigure[3D, RCM.]{%
+    \subfigure[3D, no RCM.]{%
       \label{fig:rbf3drcm}
@@
   \caption{Different sparsity distributions. In all cases, there are
-    32 nonzeros per line and 300 rows. The last four matrices
-    corresponds to derivative stencils in 2D and 3D RBF-FD
-    calculations.}
+    512 rows and 32 nonzeros per row. The last two matrices
+    corresponds to a derivative stencil in 3D RBF-FD
+    calculations (with and without RCM ordering).}
   \label{fig:spy_plots}
@@
 
-  \caption{Performance of "best code" accross matrix types. The two
-    horizontal lines depict worst and best case performance
-    (Figure~\ref{fig:gflops-peak-perf}). RCM (blue, $\blacktriangle$)
-    and no RCM (red,$\blacktriangledown$).\todo{can we reorder the
-      legend to give supercompact first in (a) and compact first in
-      (b)?}}\todo{is there a symbol/color collision?}
+  \caption{Performance of the manually vectorized code. The two
+    horizontal lines depict the worst and the best predicted performances. }
   \label{fig:expe_types}

### 2013-10-18 ieee_paper.tex

@@
       \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
-    \subfigure[3D, no RCM.]{%
+    \subfigure[3D, RCM.]{%
       \label{fig:rbf3drcm}
@@
   \caption{Different sparsity distributions. In all cases, there are
-    512 rows and 32 nonzeros per row. The last two matrices
-    corresponds to a derivative stencil in 3D RBF-FD
-    calculations (with and without RCM ordering).}
+    32 nonzeros per line and 300 rows. The last four matrices
+    corresponds to derivative stencils in 2D and 3D RBF-FD
+    calculations.}
   \label{fig:spy_plots}

### 2013-10-18 ieee_paper.tex
caption figure 8

--- 2013-10-18 ieee_paper.tex
@@
   \includegraphics[width=\linewidth]{figures/bandwidth_gather_unpack.pdf}
-  \caption{Performance of gather and unpack operations on the MIC.}
+  \caption{Performance of gather and unpack operations.}
   \label{fig:band_gather} 
@@
   \centering
-  \subfigure[Supercompact and Random]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
-  \subfigure[Compact and RBF-FD]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
+  \subfigure[Supercompact and Random.]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
+  \subfigure[Compact and RBF-FD.]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
 

### 2013-10-18 ieee_paper.tex
harmonizing width
captions

@@
 
-%  \includegraphics[width=\linewidth]{figures/matrix_not.pdf}
   \caption{Matrix of the computation.}
@@
 
-  \caption{On MIC with 240 threads, 64-byte memory
-    alignment. Bandwidth performance under idealized
-    conditions as a function of matrix row size.  Entries with "cpp"
-    denote cases where coding was performed without MIC vector
-    instructions. \todo{legend} }
+  \caption{Bandwidth performance under idealized conditions.  Entries
+    with "cpp" (triangles) denote cases where vectorization is
+    compiler generated. \todo{legend} }
   \label{fig:bandwidth}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/bandwidth_gather_unpack.pdf}
+  \includegraphics[width=.9\linewidth]{figures/bandwidth_gather_unpack.pdf}
   \caption{Performance of gather and unpack operations.}
@@
   \centering
-  \subfigure[Supercompact and Random.]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
-  \subfigure[Compact and RBF-FD.]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
+  \subfigure[Supercompact and Random.]{\includegraphics[width=.45\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
+  \subfigure[Compact and RBF-FD.]{\includegraphics[width=.45\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
 
@@
 
-  \caption{Performance of $y=Ax$ on the MIC. Squares:
-    base 1/1 case, solid circles: 4/4 case implemented in C++, solid
-    triangles: 4/4 case implemented with MIC vector instructions. Each
-    symbol has two cases: with (red) and without (blue) Reverse
-    Cuthill-McKee bandwidth reduction. The grid is $96^3$. Vectorization matters in practice. Differences in BW
-    translate in different performance.\todo{legend} \todo{fix caption}}
+  \caption{Performance of $y=Ax$ for a RBF derivative stencil of a
+    $96^3$ 3D grid. Classical SpMV ($n_m=n_v=1$ given in blue) has low
+    performance. The manually vectorized code (red) is 3.5 times
+    faster than the compiler generated one (green). RCM ordering
+    (squares) only improves the performance for our manually
+    vectorized code.}
   \label{fig:comparison_spmv}

### 2013-10-18 ieee_paper.tex
minor changes nad comments

@@
 and its impact on the expected performance of the
-application. Relevant notations are given in Table~\ref{tab:not}.
+application. Relevant notations are given in Figure~\ref{tab:not}.
 
@@
   \end{center}
-  \caption{Notation relative to the matrices (upper section) and to the benchmarks (lower section.)}
+  \caption{Notation relative to the application problem (upper section) and to the benchmarks (lower section.)}
   \label{tab:not}
@@
 
-%% In the special case $n_v=n_m=1$ in single precision
-%% $$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
-%% %$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
-%% $$I_w = \frac{n_z }{2  + (4+cl/2)  n_z} \approx \frac{1}{4+cl/2}$$
-
-%% In the special case $n_v=n_m=4$ in single precision
-%% $$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
-%% %$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
-%% $$I_w = \frac{32 n_z }{64 + (20+cl) n_z} \approx \frac{32}{84} = \frac{8}{21} $$
-%% The above formulas have been applied to MIC, which has a cache size $cl=64$ bytes.
-
-%% \begin{table}
-%%   \centering
-%%   \begin{tabular}{|l|r|r|r|r|}
-%%     \hline
-%%                            & v1m1  & v1m4  & v4m1   & v4m4  \\
-%%     \hline
-%%     Best Single Precision  & 36.36 & 58.18 & 133.33 & 213.33\\
-%%     Worst Single Precision &  4.15 & 14.20 &  16.55 &  55.81\\
-%%     Best Double Precision  & 24.00 & 32.21 &  85.71 & 117.07\\
-%%     Worst Double Precision &  3.93 & 11.88 &  15.58 &  46.15\\
-%%     \hline
-%%   \end{tabular}
-%%   \caption{Estimation of performance in GFlop/s at 150GB/s for a
-%%     matrix of $n_z = 32$}
-%% \end{table}
-
 Figure~\ref{fig:ratio_bytes_flops} presents the flop to byte ratios
@@
     ; and \ref{fig:ratio_worst_nocache} worst case neglecting that the
-    memory transfers are by batch (equivalent to $cl = 1$)}
+    memory transfers are with a grenularity of one cache line (equivalent to $cl = 1$)}
   \label{fig:ratio_bytes_flops}
@@
     device with a memory to computational units of 150~GB/s when
-    varying the number of vectors and matrices.}
+    varying the number of vectors and matrices.\todo{yabel=Peak Performance (Gflop/s)}\todo{no xlabel}}
   \label{fig:perf_predict}
@@
     with "cpp" (triangles) denote cases where vectorization is
-    compiler generated. \todo{legend} }
+    compiler generated. \todo{legend} \todo{ylabel=Bandwidth(GB/s)}}
   \label{fig:bandwidth}
@@
   \includegraphics[width=.9\linewidth]{figures/bandwidth_gather_unpack.pdf}
-  \caption{Performance of gather and unpack operations.}
+  \caption{Performance of gather and unpack operations.\todo{ylabel=Bandwidth(GB/s)}}
   \label{fig:band_gather} 
@@
   \caption{Performance of the manually vectorized code. The two
-    horizontal lines depict the worst and the best predicted performances. }
+    horizontal lines depict the worst and the best predicted performances. \todo{font size} \todo{ylabel=Performance(Gflop/s)}}
   \label{fig:expe_types}
@@
     (squares) only improves the performance for our manually
-    vectorized code.}
+    vectorized code.\todo{ylabel=Performance(Gflop/s)}}
   \label{fig:comparison_spmv}

### 2013-10-18 gflops_speedup.py
figure changes.

@@ -1,184 +1,0 @@
-import sys
-import numpy as np
-import matplotlib.pyplot as plt
-
-#----------------------------------
-def byte_to_float_ratio(nv, nm, bx, nz):
-# no consideration of cache.
-    bi = 4
-    #print nv,nm,nz
-    #rworst = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
-    # in worst case, every vector element brings in an entire cache line (256 bits = 32 floats),
-    # so nv=4 vectors (1/8th cache line), requires a transfer of 32 bytes = 8 floats
-    # 16 vectors = 2 cache lines (, so the formula becomes
-    ratio = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
-    return(ratio)
-#----------------------------------
-#----------------------------------
-def ratio(nv, nm, bx, nz):
-# per row:
-# flops divided by number bytes sent/received
-# nv : number vectors
-# nm : number matrices
-# bx = 4 or 8 (single, double precision)
-# nz = nb nonzeros per row
-# y = A*x
-# y : nv*nm*bx   writes
-# x : nv*bx      loads
-# A : nm*nz*bx   loads
-# col_id : nz*bi loads
-    bi = 4
-
-
-    cl = 64  # size of most caches
-
-    # for each row of A
-
-    flops = 2.*nv*nm*nz
-
-    a_read = bx*nm*nz
-    icol_read = bi*nz
-    y_write = bx*nm*nv
-
-    x_read = nz *  cl*np.ceil((bx*nv)/float(cl))  # assumes small cache, x must be reread
-    rworst = flops/(x_read+a_read+icol_read+y_write)
-
-    x_read = bx*nv   # each element of x is read once
-    rbest = flops/(x_read+a_read+icol_read+y_write)
-    #print "yy: ", [nv,nm,rworst,rbest]
-    return([rworst,rbest])
-
-
-"""
-    cl = 64.
-    #rworst = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
-    # in worst case, every vector element brings in an entire cache line (256 bits = 32 floats),
-    # so nv=4 vectors (1/8th cache line), requires a transfer of 32 bytes = 8 floats
[…]

### 2013-10-18 ieee_paper.tex
minor

@@
   y^{i,j} = A^i x^j 
-$$
-where $j$ ranges over the number of source vectors, $i$ ranges of the number of derivative
-matrices and there are $i*j$ output vectors $y$.
-
-% GORDON I WOULD ELIMINATE BELOW AND END HERE.
-
-%**** PUT ELSEWHERE OR ELIMINATE? ****
-%
-%When solving a one dimensional system of PDEs, one might require an
-%$x$ derivative of multiple functions. For example, the Euler equations
-%require the $x$ derivative of the three components of velocity and
-%pressure. In this case, there are $n_v=4$ vectors $x^k$,
-%$k=0,\cdots,3$.
-%
-%Thus for each vector element $y_i$, we compute $y_i = \sum_j A_{ij}
-%x_j$. If $A_i$ is row $i$ of $A$, $y_i$ is simply the dot product $A_i
-%x$. The next level of generality is to consider $n_v$ vectors $x^k$,
-%$k=0,\cdots,n_v-1$. Whatever the spmv implementation, one achieves
-%improved performance if the matrix formed from the columns $x^k$ are
-%stored in row major order. Thus, $x^0_0,x^1_0,\cdots,x^{n_v-1}_0$, are
-%stored in consecutive memory location. The random access of the
-%elements of $x$ is thus reduced. Maximum efficiency is achieved when
-%$n_v=16$ floats or $8$ doubles, given that cache lines take 64
-%bytes. We will benchmark this case, labeled $Svn$, where $n$ refers to
-%the number of vectors (Iv4 uses four vectors). The $S$ refers to singe
-%precision. A double precision run is labelled $Dvn$.
-%
-%Alternatively, when solving a PDE, one might require derivatives of a
-%given scalar function with respect to coordinate directions $x$, $y$,
-%$z$. Second order operators of often required, such as a second
-%derivative with respect to $x$ or a Laplacian operator. In the RBFFD
-%formulation, on can compute different derivatives using the same
-%stencil, but with different weights. In other words, the adjacency
-%matrix that corresponds to $A$ remains constant, but the matrix
-%elements of $A$ change with the particular derivative.  In this case,
-%label with a superscript $l$ the particular matrix $A^k$. Since the
-%adjacency matrix is assumed invariant, there is only need for a single
-%matrix \ttt{C{ij}}. In Ellpack format, each row is of constant size
-%(the number of nonzeros per row of $A$. $C_{ij}$ is the column number
-%that locates the $j^{th}$ nonzero in row $i$ of $A$.
-%
-%
-%\todo{Must rewrite the above to make it more focused on the
-% application. Remove the architectural/implementation details such as
-%  using ELLPACK. Let's focus on on why this computation is important
-%  and how does the 16 multiplication appears in the equations.}
+$$ where $j$ ranges over the number of source vectors, $i$ ranges of
+  the number of derivative matrices and there are $i*j$ output vectors
+  $y$ (graphically shown in Figure~\ref{fig:struct_comp}).
 
@@
 
-  \caption{Matrix of the computation.}
-  \label{fig:mat_struct}
+  \caption{Structure of the computation.}
+  \label{fig:struct_comp}
 \end{figure}
@@
     ; and \ref{fig:ratio_worst_nocache} worst case neglecting that the
-    memory transfers are with a grenularity of one cache line (equivalent to $cl = 1$)}
+    memory transfers are with a granularity of one cache line (equivalent to $cl = 1$)}
   \label{fig:ratio_bytes_flops}
@@
   \caption{Estimation of the maximum achievable performance using a
-    device with a memory to computational units of 150~GB/s when
+    device with a bandwidth from memory to computational units of 150~GB/s when
     varying the number of vectors and matrices.\todo{yabel=Peak Performance (Gflop/s)}\todo{no xlabel}}

### 2013-10-18 ieee_paper.tex
removing todos

@@
     device with a bandwidth from memory to computational units of 150~GB/s when
-    varying the number of vectors and matrices.\todo{yabel=Peak Performance (Gflop/s)}\todo{no xlabel}}
+    varying the number of vectors and matrices.}
   \label{fig:perf_predict}
@@
     with "cpp" (triangles) denote cases where vectorization is
-    compiler generated. \todo{legend} \todo{ylabel=Bandwidth(GB/s)}}
+    compiler generated.}
   \label{fig:bandwidth}
@@
   \includegraphics[width=.9\linewidth]{figures/bandwidth_gather_unpack.pdf}
-  \caption{Performance of gather and unpack operations.\todo{ylabel=Bandwidth(GB/s)}}
+  \caption{Performance of {\tt gather} and {\tt unpack} operations.}
   \label{fig:band_gather} 
@@
   \caption{Performance of the manually vectorized code. The two
-    horizontal lines depict the worst and the best predicted performances. \todo{font size} \todo{ylabel=Performance(Gflop/s)}}
+    horizontal lines depict the worst and the best predicted performances. \todo{font size}}
   \label{fig:expe_types}
@@
     (squares) only improves the performance for our manually
-    vectorized code.\todo{ylabel=Performance(Gflop/s)}}
+    vectorized code.}
   \label{fig:comparison_spmv}

### 2013-10-18 ieee_paper.tex
Changes.

@@
 \label{code:mat_mul}
-
 \end{figure*}
@@
 we would like to execute a single instruction to load the data for four 
-nonzeros in the matrix, representing 16 floats from the matrix and 16 floats from the vectors
-at once. (The number of nonzeros on a row in our application is
+nonzeros in the matrix, and another instruction to load the data for
+four nonzeros from the vectors.  
+%representing 16 floats from the matrix and 16 floats from the vectors at once. 
+(The number of nonzeros on a row in our application is
 always a multiple of 4; if it weren't, we would add some explicit
 zeroes in the data structure for padding.)
-\todo{Erik, reread paragraph above. single instruction to load 16 floats. Confused.}
+\todo{Erik, I rewrote the paragraph. Please reread.}
 
@@
 
-When using the MIC card (Figure~\ref{fig:band_rw_MIC}) the {\tt
+When using the MIC card (Figure~\ref{fig:band_rw_MIC}), the {\tt
   read\_cpp} implementation reaches a performance of 160~GB/s
-\todo{gordon check exact number} which is better than the {\tt read}
-implementation which only achieves 145~GB/s \todo{and that one}. This
+\todo{gordon check exact number, 168GB/s} which is better than the {\tt read}
+implementation which only achieves 145~GB/s \todo{and that one, 145 is correct}. This
 is explained by the compiler adding loop unrolling and software
-prefetching in the assembly code in the {\tt read\_cpp} case which is
+prefetching in the assembly code in the {\tt read\_cpp} case, which is
 not added in the {\tt read} case. On the other hand, the {\tt write}
-case reaches a performance of 155~GB/s\todo{and that one} while the
-{\tt write\_cpp} implementation never passes 55~GB/s. This difference
-is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction
+case reaches a performance of 158~GB/s\todo{and that one, 158GB} while the
+{\tt write\_cpp} implementation never passes 58~GB/s \todo{I fixed the number}. 
+This difference is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction, 
 which bypasses the Read For Ownership protocol and allows the writes
 to be performed in any order. The compiler can not use this
-instruction on its own because using would break the memory convention
+instruction on its own because this would break the memory convention
 useful for some parallel algorithms. The {\tt read/write} benchmark
@@
 reached. If the indices are completely randomized, the performance
-drops to 2~GB/s\todo{correct?}. We can see that implementing the
+drops to 3~GB/s\todo{correct?. Chaged. It is 3-4}. We can see that implementing the
 indirection mechanism without vector instructions makes the
@@
 in a typical SpMV computation, one can expect a bandwidth between
-140~GB/s to 160~GB/s. That is why we set in our estimation the best
-achievable bandwidth to 150~GB/s despite carefully crafted code to
-maximize bandwidth can reach higher
+140~GB/s to 160~GB/s. That is why we set in our estimation a best
+achievable bandwidth of 150~GB/s despite the fact that a 
+carefully crafted code to maximize bandwidth can reach higher
 values~\cite{Saule13-ARXIV}.
@@
 from memory, and the expense associated with cache misses and gather
-operation is removed from consideration. The cost associated with 
-{\tt A} remains.   \todo{reread the above.}
+operations is removed from consideration. The cost associated with 
+{\tt A} remains.   \todo{reread the above. Reread by Gordon}
 The Compact matrices
@@
 computation. The {\tt supercompact} case peaks at
-205~Gflop/s\todo{check that} being very close to the predicted peak
+208~Gflop/s\todo{check that. changed to 208} being very close to the predicted peak
 performance of 213~Gflop/s. Conversely the performance of the {\tt
-  random} case decreases to 59~Gflop/s\todo{check} which is close to
+  random} case decreases to 56~Gflop/s\todo{check, changed to 56} which is close to
 the lowest predicted performance of 55~Gflop/s. One can see that RCM
 ordering helps the {\tt random} case but the impact decreases when the
-size of the matrix increase.
+size of the matrix increases.
 
 Figure~\ref{fig:perf_realistic} gives the results on the {\tt compact}
-case which represents the best realistic matrix one could find and in
-an {\tt rbf} matrix extracted from a real application. In the {\tt
+case, which represents the best realistic matrix one could find with very
+structured grids and a
+{\tt RBF} matrix extracted from a real application. In the {\tt
   compact} case, the performance can be as high as 195~Gflop/s being
@@
 {\tt rbf} case varies between 100~Gflop/s and 140~Gflop/s. Most of the
-time, the RCM ordering provide an improvement which can be as high as
+time, the RCM ordering provides an improvement which can be as high as
 30~Gflop/s.

### 2013-10-19 ieee_paper.tex
comments

--- 2013-10-18 ieee_paper.tex
@@
 \title{Sparse Matrix Vector Multiplication with Multiple vectors and
-  Multiple Matrices on the MIC Architecture}
+  Multiple Matrices on the MIC Architecture \todo{? more application oriented''}}
 
@@
 \IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
-Email: bollig@gmail.com}}
+Email: bollig@gmail.com}\todo{check affiliation fsu's are strange.}}
 \maketitle

### 2013-10-19 ieee_paper.tex
minor intro

@@
 the MIC architecture, taking into account its advanced swizzling and
-channel interchange features. We present benchmarks that show an
-almost order of magnitude increase in speed compared to efficient
-implementations of a single derivative. We explain the results through
+channel interchange features. We present benchmarks on a real data
+set that show an almost eightfold increase in speed compared to
+efficient implementations of a single derivative, reaching a
+performance of almost 140~Gflop/s. We explain the results through
 consideration of operation count versus memory bandwidth.
@@
 \begin{IEEEkeywords}
-OpenMP; MIC; spmv; sparse matrix; Radial Basis Function;
+MIC; SIMD; SpMV; SpMM; Sparse Matrix; Radial Basis Function;
 \end{IEEEkeywords}
@@
 important kernel in many applied fields such as fluid dynamics 
-\cite{journals/tog/BolzFGS03} 
-to graph analysis (recommender systems~\cite{Brin98}, graph
-drawing~\cite{Koren05}). So improving the performance of SpMV has
-captured the interest of many researcher; including the development of
-various implementations for multi-core
+\cite{journals/tog/BolzFGS03}, recommendation systems~\cite{Brin98} and graph
+drawing~\cite{Koren05}). Naturally, improving the performance of SpMV has
+captured the interest of many researchers; including the development of
+various implementations for 
 CPUs~\cite{Buluc2009SPAA,Williams07} and GPUs~\cite{Bell08,
@@
   journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
-challenge that is faced to obtain good performance for SpMV is that
-the operation are conducted using memory location that are irregular
-and often unpredictible. That makes the kernels mostly memory
+challenge  to obtain good performance for SpMV is that
+the operations are conducted using memory location that are irregular. This makes the kernels mostly memory
 bound and there is a significant instruction overhead per flop.

### 2013-10-19 ieee_paper.tex
minor intro

@@
 with virtually no overhead in all cases thanks to compressed
-representation~\cite{Buluc11}). There is about 8 bytes of the matrix to move
+representation~\cite{Buluc11}). Indeed, there is about 8 bytes of the matrix to move
 in per nonzero (assuming single precision); each nonzero requires two

### 2013-10-19 ieee_paper.tex
intro

@@
 An other solution, and the one we pursue in this paper, is to compute
-multiple SpMV at once on matrices that have the same sparsity
+multiple SpMVs at once on matrices that have the same sparsity
 patterns. Obviously not all the applications have such a property. But
@@
 recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of
-derivatives of Radial Basis Functions(RBF)~\cite{FLBWSC12} can use multiple
+derivatives of Radial Basis Functions (RBF)~\cite{FLBWSC12} can use multiple
 SpMVs simultaneously. In this paper in particular, we investigate the
@@
 In Section~\ref{sec:rbf} we present the computation of RBFs and how it
-can be expressed 16 multiplication of 4 vectors by 4 sparse matrices
-with a common sparsity pattern. Section~\ref{sec:model} presents an
-estimation of the instruction intensity of various form of the
-computations and we show that a 6- fold improvement can be
-expected when computing the 16 multiplications simultaneously and
+can be expressed as sixteen multiplications of four vectors by four sparse matrices
+with identical sparsity patterns. Section~\ref{sec:model} presents an
+estimation of the instruction intensity of various forms of the
+computations and we show that a sixfold improvement can be
+expected when computing the sixteen multiplications simultaneously and
 reach a total of about 210~Gflop/s. This performance represents
 approximatively 10\% of the available flop/s of a Knight Corner
-coprocessor. Therefore, it is necessary to have implementations that
-perform the computation in as little amount of instructions as
+coprocessor. Therefore, it is necessary to have an implementation that
+performs the computation in as few instructions as
 possible. We describe in Section~\ref{sec:impl} the details of the MIC
 architecture and how to use specialized load, store, swizzle and
-permutation instruction to efficiently bring the data in the vector
+permutation instructions to efficiently bring the data in the vector
 registers to be processed. Section~\ref{sec:expe} gives some
-experimental result about the amount of bandwidth that can be achieved
-depending on how the SpMV kernel is written and the actual performance
-of the various kernel on multiple classes of matrices some generated
-for analysis purpose and some extracted from and RBF application. A
+experimental results about the amount of bandwidth that can be achieved
+depending on how the SpMV kernel is written. It also provides the actual performance
+of the various kernel on multiple classes of matrices, some generated
+for analysis purpose, and some extracted from an RBF application. A
 performance of 135~Gflop/s is achieved on real scenario. Concluding

### 2013-10-19 ieee_paper.tex
Went through most of the paper.

--- 2013-10-19 ieee_paper.tex
@@
 \title{Sparse Matrix Vector Multiplication with Multiple vectors and
-  Multiple Matrices on the MIC Architecture \todo{? more application oriented''}}
+  Multiple Matrices on the MIC Architecture}
 
@@
 \IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
-Email: bollig@gmail.com}\todo{check affiliation fsu's are strange.}}
+Email: bollig@gmail.com}}
 \maketitle
@@
 the MIC architecture, taking into account its advanced swizzling and
-channel interchange features. We present benchmarks on a real data
-set that show an almost eightfold increase in speed compared to
-efficient implementations of a single derivative, reaching a
-performance of almost 140~Gflop/s. We explain the results through
+channel interchange features. We present benchmarks that show an
+almost order of magnitude increase in speed compared to efficient
+implementations of a single derivative. We explain the results through
 consideration of operation count versus memory bandwidth.
@@
 \begin{IEEEkeywords}
-MIC; SIMD; SpMV; SpMM; Sparse Matrix; Radial Basis Function;
+OpenMP; MIC; spmv; sparse matrix; Radial Basis Function;
 \end{IEEEkeywords}
@@
 important kernel in many applied fields such as fluid dynamics 
-\cite{journals/tog/BolzFGS03}, recommendation systems~\cite{Brin98} and graph
-drawing~\cite{Koren05}). Naturally, improving the performance of SpMV has
-captured the interest of many researchers; including the development of
-various implementations for 
+\cite{journals/tog/BolzFGS03} 
+to graph analysis (recommender systems~\cite{Brin98}, graph
+drawing~\cite{Koren05}). So improving the performance of SpMV has
+captured the interest of many researcher; including the development of
+various implementations for multi-core
 CPUs~\cite{Buluc2009SPAA,Williams07} and GPUs~\cite{Bell08,
@@
   journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
-challenge  to obtain good performance for SpMV is that
-the operations are conducted using memory location that are irregular. This makes the kernels mostly memory
+challenge that is faced to obtain good performance for SpMV is that
+the operation are conducted using memory location that are irregular
+and often unpredictible. That makes the kernels mostly memory
 bound and there is a significant instruction overhead per flop.
@@
 We now investigate the actual performance that we can obtain when
-multiplying 4 vectors by 4 matrices in single
-precision. Figure~\ref{fig:expe_types} present results for different
+multiplying four vectors by four matrices in single
+precision. Figure~\ref{fig:expe_types} presents results for different
 types of matrices and also shows the minimum and maximum performance
@@
 gives the results for the {\tt supercompact} and {\tt random} matrices
-which represent both the best and the worst case for such a
+that represent both the best and the worst case for such a practical 
 computation. The {\tt supercompact} case peaks at
@@
 within 15\% of the peak predicted performance. The performance of the
-{\tt rbf} case varies between 100~Gflop/s and 140~Gflop/s. Most of the
+{\tt RBF} case varies between 100~Gflop/s and 140~Gflop/s. Most of the
 time, the RCM ordering provides an improvement which can be as high as
@@
 optimized implementation. The
-classical SpMV computation reaches 15~Gflop/s\todo{check}. The standard
-C++ implementation reaches a performance of 40~Gflop/s while our
-optimized implementation reaches 137~Gflop/s\todo{check}. Notice that
+classical SpMV computation reaches 14~Gflop/s\todo{check? 14 on 244 nodes}. 
+The standard C++ implementation reaches a performance of 38~Gflop/s while our
+optimized implementation reaches 140~Gflop/s\todo{check 140 for $64^3$, 135 for $96^3$ on 61 nodes}. 
+Notice that \todo{I do not know that I ran host and mic on the same grid, but your 
+numbers are approximately correct. The numbers are a function of resolution} 
 RCM ordering has no impact in the standard C++ implementation while it
-provides an significant improvement in our manually vectorized
-version. This highlights the C++ implementation is instruction bound
-while our manually vectorized implementation is most likely memory
-bound.
+provides a significant improvement in our manually vectorized
+version. This highlights that the C++ implementation is instruction-bound
+while our manually vectorized implementation is most likely 
+memory-bound.
 
@@
 classical CPU architecture with 2 Intel Core i7 processor, each has 8
-cores which support AVX instructions. Both the compiler generated and
-the manually written vectorization using AVX instruction reach a
+cores that support AVX instructions. Both the compiler-generated and
+the manually-written vectorization using AVX instruction reach a
 (read) bandwidth of 70~GB/s. This indicates that the performance of
-the classical SpMV can not overcome 16.8~Gflop/s. A generic
-implementation of the multiplication of 4 vectors with 4 matrices on
+the classical SpMV can not overcome 17~Gflop/s. A generic
+implementation of the multiplication of four vectors with four matrices on
 the {\tt RBF} matrix reaches a performance of 27~Gflop/s; while a
@@
 42~Gflop/s. This confirms that our method is sound and generalizes to
-other architecture than the Intel MIC architecture.
+architectures than the Intel MIC architecture. \todo{albeit with less 
+impressive results given that swizzling is often not available. Swizzling
+will become available in a new version of the host: enhanced AVX.}
 
@@
 
-In the previous sections, we have explored in the practical
+In the previous sections, we have explored the practical
 implementation on a Intel Xeon Phi card of multiple derivative
@@
 vectors), we accelerate the algorithm further by considering multiple
-derivatives with corresponding matrices with identical adjacency
-graphs. We specialize our study to four matrices and four derivatives,
+matrices (identical sparsity patterns) acting on multiple vectors. 
+We specialize our study to four matrices and four derivatives,
 with 16 outputs, computed as a sum of outer products.
@@
 In a future work, we will examine the effect of larger stencil sizes
-and of using double precision. We will also integrate the techniques
+and double precision. We will also integrate the techniques
 presented in this paper within a fluid simulation using RBF-FD.

### 2013-10-19 ieee_paper.tex
sec 2

@@
 In practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
-three-dimensional flows. These numbers are on a par with what is used in
+three-dimensional flows. These numbers are on par with what is used in
 finite-element codes. Notice that the nonzero elements of the matrix
 correspond to a relation of nearest neighbor. This relationship is not
-necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} which
+necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} \todo{Figure 1 n'est pas nearest neighbor} which
 implies a non-symmetric sparsity pattern in the derivative matrix.

### 2013-10-19 ieee_paper.tex
minor
figure size

@@
 \end{equation}
-where $Q$ is a vector of unknowns %(e.g., 3D Navier-Stokes has
-%three components of velocity, pressure and temperature). 
+where $Q$ is a vector of unknowns.
 For example, in solving a system of equations, it is
@@
   \centering
-  \subfigure[Supercompact and Random.]{\includegraphics[width=.45\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
-  \subfigure[Compact and RBF-FD.]{\includegraphics[width=.45\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
+  \subfigure[Supercompact and Random.]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
+  \subfigure[Compact and RBF-FD.]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
 

### 2013-10-19 ieee_paper.tex
minor

@@
 bandwidth of 320~GB/s aggregated across all the memory
-controllers. There are 60 cores clocked at 1.05GHz. Their memory
+controllers. There are 60 cores clocked at 1.05~GHz. Their memory
 interfaces are 32-bit wide with two channels and the total bandwidth
@@
 The standard C++ implementation reaches a performance of 38~Gflop/s while our
-optimized implementation reaches 140~Gflop/s\todo{check 140 for $64^3$, 135 for $96^3$ on 61 nodes}. 
+optimized implementation reaches 140~Gflop/s\todo{check 140 for $64^3$, 135 for $96^3$ on 61 cores}. 
 Notice that \todo{I do not know that I ran host and mic on the same grid, but your 
@@
 will become available in a new version of the host: enhanced AVX.}
-
-%% Scaling results as a function of the number of processors, on the MIC
-%% and on the host. The host calculations are performed for reference
-%% only, to bring out the dominance of the MIC due to its advanced
-%% swizzling and channel operations, not available on the host. The
-%% performance on the host is 3x slower than on the MIC. The large cache
-%% of the host alleviates somewhat the lower memory bandwidth, but this
-%% only reduces the cost of transfer {\tt x}.  The cost memory cost for
-%% {\tt A} and {\tt col\_id} is unaffected by the larger cache size.
-
-%% The performance on the MIC scales linearly with the number of
-%% threads. We compare three cases, each with and without bandwidth
-%% reduction using Reverse Cuthil-McKee (RCM). There are $96^3$ RBF
-%% nodes, leading to a matrix with 884,736 rows. The four source vectors
-%% occupy more than 14M bytes, which exceeds the aggregate L2 cache of 31
-%% Mbytes (7.8 Mfloats.)
-
-%% \todo{Is the host architecture described? Problem: I had to change
-%%   hosts a few weeks ago, and it has different characteristics}.
-%% %----------------------
-
-%% Derivatives in an RBFFD formulation are expressed as a sparse
-%% matrix/vector multiplication (SpMV). Using the full 61 nodes of a MIC
-%% processor, and all four threads of each node, we achieve lowly 18
-%% Gflops when calculating a single vector of a single
-%% function. Bandwidth reduction has minimal effect since the cache is
-%% large enough to store both the derivative matrix and the solution
-%% vector. We increase the number of computations per byte transferred by
-%% calculating four different derivatives of four different
-%% functions. With bandwidth reduction, we speed the calculation by a
-%% factor of 8, and achieve upwards of 140 Gflops.
 

### 2013-10-19 ieee_paper.tex
removing todo

@@
 computation. The {\tt supercompact} case peaks at
-208~Gflop/s\todo{check that. changed to 208} being very close to the predicted peak
+208~Gflop/s being very close to the predicted peak
 performance of 213~Gflop/s. Conversely the performance of the {\tt
-  random} case decreases to 56~Gflop/s\todo{check, changed to 56} which is close to
+  random} case decreases to 56~Gflop/s which is close to
 the lowest predicted performance of 55~Gflop/s. One can see that RCM
@@
 optimized implementation. The
-classical SpMV computation reaches 14~Gflop/s\todo{check? 14 on 244 nodes}. 
+classical SpMV computation reaches 14~Gflop/s. 
 The standard C++ implementation reaches a performance of 38~Gflop/s while our
-optimized implementation reaches 140~Gflop/s\todo{check 140 for $64^3$, 135 for $96^3$ on 61 cores}. 
+optimized implementation reaches 140~Gflop/s. 
 Notice that \todo{I do not know that I ran host and mic on the same grid, but your 

### 2013-10-19 ieee_paper.tex
changes.

--- 2013-10-19 ieee_paper.tex
@@
 \IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Scientific Computing, 
-Florida State University, Tallahassee, FL 32306-4120\\
-Email: gerlebacher@fsu.edu}
+Florida State University\\ %, Tallahassee %FL 32306-4120\\
+Email: gerlebacher@fsu.edu, bollig@gmail.com}
 \IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte\\
@@
 Email: flyer@ucar.edu}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
-Email: bollig@gmail.com}\todo{check affiliation fsu's are strange.}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
+%Email: bollig@gmail.com}\todo{check affiliation fsu's are strange.}}
 \maketitle
@@
 derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBF-FD uses only
-a local set of the $n_z-1$ nearest neighbors to the point $\rvec_i$ to
+a local set of the $n_z-1$  (often nearest) points in a neighborhood of 
+the point $\rvec_i$ to
 approximate the derivative. In other words,
@@
 In practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
-three-dimensional flows. These numbers are on par with what is used in
+three-dimensional flows. These numbers are on a par with what is used in
 finite-element codes. Notice that the nonzero elements of the matrix
-correspond to a relation of nearest neighbor. This relationship is not
-necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} \todo{Figure 1 n'est pas nearest neighbor} which
+correspond to a relationship of closeness in the physical domain. 
+This relationship is not
+necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} which
 implies a non-symmetric sparsity pattern in the derivative matrix.
@@
 \end{equation}
-where $Q$ is a vector of unknowns.
+where $Q$ is a vector of unknowns %(e.g., 3D Navier-Stokes has
+%three components of velocity, pressure and temperature). 
 For example, in solving a system of equations, it is
@@
 remain unchanged; only the values in the sparse matrix change. 
-change.  Thus, rather than computing a derivative of multiple
+Thus, rather than computing a derivative of multiple
 functions, one can calculate multiple derivatives of a single
@@
 vector processing unit can perform many basic instructions, such as
-addition or division, and mathematical operations, such as sine and
-sqrt, making it possible to reach 8 double precision (or 16 singe precision)
+addition or division, and mathematical operations, such as {\tt sin} and
+{\tt sqrt}, making it possible to reach 8 double precision (or 16 singe precision)
 operations per cycle.  
@@
 When using the MIC card (Figure~\ref{fig:band_rw_MIC}), the {\tt
-  read\_cpp} implementation reaches a performance of 160~GB/s
+  read\_cpp} implementation reaches a performance of 168~GB/s
 \todo{gordon check exact number, 168GB/s} which is better than the {\tt read}
@@
 reached. If the indices are completely randomized, the performance
-drops to 3~GB/s\todo{correct?. Chaged. It is 3-4}. We can see that implementing the
+drops below 5~GB/s\todo{correct?. Chaged. It is 3-4}. We can see that implementing the
 indirection mechanism without vector instructions makes the
@@
 To put things in perspective, we conducted some experiments on a
-classical CPU architecture with 2 Intel Core i7 processor, each has 8
+classical CPU architecture with two Intel Core i7 processors, each has 8
 cores that support AVX instructions. Both the compiler-generated and

### 2013-10-19 ieee_paper.tex
minor

--- 2013-10-19 ieee_paper.tex
@@
 case reaches a performance of 158~GB/s\todo{and that one, 158GB} while the
-{\tt write\_cpp} implementation never passes 58~GB/s \todo{I fixed the number}. 
+{\tt write\_cpp} implementation never passes 58~GB/s. 
 This difference is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction, 

### 2013-10-19 ieee_paper.tex

@@
 finite-element codes. Notice that the nonzero elements of the matrix
-correspond to a relation of nearest neighbor. This relationship is not
-necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} \todo{Figure 1 n'est pas nearest neighbor} which
+correspond to a relationship of closeness in the physical domain. 
+This relationship is not
+necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} which
 implies a non-symmetric sparsity pattern in the derivative matrix.

### 2013-10-19 ieee_paper.tex
affiliation

@@
 Florida State University\\ %, Tallahassee %FL 32306-4120\\
-Email: gerlebacher@fsu.edu, bollig@gmail.com}
+}
 \IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte\\
-Email: esaule@uncc.edu}
+}
 \IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR \\
-Email: flyer@ucar.edu}
+}
+Email: gerlebacher@fsu.edu, esaule@uncc.edu, bollig@gmail.com, flyer@ucar.edu}
 %\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\

### 2013-10-19 ieee_paper.tex
email list

@@
 }
-Email: gerlebacher@fsu.edu, esaule@uncc.edu, bollig@gmail.com, flyer@ucar.edu}
+Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com }
 %\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\

### 2013-10-19 ieee_paper.tex
removing todos

@@
 Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com }
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
-%Email: bollig@gmail.com}\todo{check affiliation fsu's are strange.}}
 \maketitle
@@
 study to four vectors and four matrices.
-%\todo{I have remove reference to $x$, $y$, ... to avoid confusion with what follows.}
 
-%\todo{Add few sentence with discretization. Check Erik}
 In discrete form, Equation (\ref{eq:Q}) takes the form
@@
 low: this highlights that accessing the memory with a granularity of one cache
-line %a cache line basis batch \todo{what do you mean by 
-%accessing the memory per batch} 
+line
 is the main problem faced when performing a standard SpMV computation.
@@
   read\_cpp} implementation reaches a performance of 168~GB/s
-\todo{gordon check exact number, 168GB/s} which is better than the {\tt read}
-implementation which only achieves 145~GB/s \todo{and that one, 145 is correct}. This
+ which is better than the {\tt read}
+implementation which only achieves 145~GB/s. This
 is explained by the compiler adding loop unrolling and software
@@
 not added in the {\tt read} case. On the other hand, the {\tt write}
-case reaches a performance of 158~GB/s\todo{and that one, 158GB} while the
+case reaches a performance of 158~GB/s while the
 {\tt write\_cpp} implementation never passes 58~GB/s. 
@@
 reached. If the indices are completely randomized, the performance
-drops below 5~GB/s\todo{correct?. Chaged. It is 3-4}. We can see that implementing the
+drops below 5~GB/s. We can see that implementing the
 indirection mechanism without vector instructions makes the

### 2013-10-19 ieee_paper.tex
section 3 and figure placement

@@
 
-\begin{figure}[bth]
+\begin{figure}[tbh]
   \begin{center}
@@
 
+\begin{figure*}[tbh]
+  \centering
+  \subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
+  \subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
+  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}
+
+  \caption{Ratio of flops to bytes in single precision:  
+    \ref{fig:ratio_best} best case; \ref{fig:ratio_worst} worst case when $cl=64$ 
+    ; and \ref{fig:ratio_worst_nocache} worst case neglecting that the
+    memory transfers are with a granularity of one cache line (equivalent to $cl = 1$)}
+  \label{fig:ratio_bytes_flops}
+\end{figure*}
+
 If we assume an algorithm where the rows are processed one after
@@
 
-The computation intensity is the amount of computations performed per
+The computation intensity is the amount of floating point operations performed per
 byte transfered. In the worst case and in the best case, we have
@@
 can easily see the potential improvement in the computational
-intensity when the number of matrices of vectors increase. There is a
+intensity when the number of matrices or vectors increases. There is a
 significant difference between the best and the worst case: there is a
-eightfold difference in the one matrix -- one vector (1/1) case but that gap closes
+eightfold difference in the one matrix -- one vector ($n_m=n_v=1$) case but that gap closes
 with the increase in the number of vectors and matrices to a fourfold
-difference in the four4matrices and four vectors (4/4) case and 1.6 fold in the 
-16/16 case. 
+difference in the four matrices and four vectors ($n_m=n_v=4$) case and 1.6 fold in the 
+$n_m=n_v=16$ case. 
 The ratios computed with $cl = 1$ are mostly
@@
 
-\begin{figure*}
-  \centering
-  \subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
-  \subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
-  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}
-
-  \caption{Ratio of flops to bytes in single precision:  
-    \ref{fig:ratio_best} best case; \ref{fig:ratio_worst} worst case when $cl=64$ 
-    ; and \ref{fig:ratio_worst_nocache} worst case neglecting that the
-    memory transfers are with a granularity of one cache line (equivalent to $cl = 1$)}
-  \label{fig:ratio_bytes_flops}
-\end{figure*}
-
 Figure~\ref{fig:perf_predict} shows how these ratios translate
@@
 Gflop/s achievable assuming the computation is memory-bound and the
-architecture reaches 150~Gflop/s. Notice that \cite{Saule13-ARXIV}
+architecture reaches 150~GB/s. Notice that \cite{Saule13-ARXIV}
 showed a higher peak bandwidth, but we will show in
-Section~\ref{sec:expe} why 150~Gflop/s is a better estimate of what one
+Section~\ref{sec:expe} why 150~GB/s is a better estimate of what one
 might achieve in these kernels. In single precision, the worse that
-one could achieve by using 4 vectors and 4 matrices is much higher
+one could achieve by using four vectors and four matrices is much higher
 than the best achievable using a classical SpMV computation. The best

### 2013-10-19 ieee_paper.tex
figure fonts/lines/sizes mostly uniform.

--- 2013-10-19 ieee_paper.tex
@@
 Florida State University\\ %, Tallahassee %FL 32306-4120\\
-}
+Email: gerlebacher@fsu.edu, bollig@gmail.com}
 \IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte\\
-}
+Email: esaule@uncc.edu}
 \IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR \\
-}
-Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com }
+Email: flyer@ucar.edu}}
 %\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
@@
 case reaches a performance of 158~GB/s\todo{and that one, 158GB} while the
-{\tt write\_cpp} implementation never passes 58~GB/s. 
+{\tt write\_cpp} implementation never passes 58~GB/s \todo{I fixed the number}. 
 This difference is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction, 
@@
 computation. The {\tt supercompact} case peaks at
-208~Gflop/s being very close to the predicted peak
+208~Gflop/s\todo{check that. changed to 208} being very close to the predicted peak
 performance of 213~Gflop/s. Conversely the performance of the {\tt
-  random} case decreases to 56~Gflop/s which is close to
+  random} case decreases to 56~Gflop/s\todo{check, changed to 56} which is close to
 the lowest predicted performance of 55~Gflop/s. One can see that RCM
@@
 optimized implementation. The
-classical SpMV computation reaches 14~Gflop/s. 
+classical SpMV computation reaches 14~Gflop/s\todo{check? 14 on 244 nodes}. 
 The standard C++ implementation reaches a performance of 38~Gflop/s while our
-optimized implementation reaches 140~Gflop/s. 
+optimized implementation reaches 140~Gflop/s\todo{check 140 for $64^3$, 135 for $96^3$ on 61 nodes}. 
 Notice that \todo{I do not know that I ran host and mic on the same grid, but your 
@@
 will become available in a new version of the host: enhanced AVX.}
+
+%% Scaling results as a function of the number of processors, on the MIC
+%% and on the host. The host calculations are performed for reference
+%% only, to bring out the dominance of the MIC due to its advanced
+%% swizzling and channel operations, not available on the host. The
+%% performance on the host is 3x slower than on the MIC. The large cache
+%% of the host alleviates somewhat the lower memory bandwidth, but this
+%% only reduces the cost of transfer {\tt x}.  The cost memory cost for
+%% {\tt A} and {\tt col\_id} is unaffected by the larger cache size.
+
+%% The performance on the MIC scales linearly with the number of
+%% threads. We compare three cases, each with and without bandwidth
+%% reduction using Reverse Cuthil-McKee (RCM). There are $96^3$ RBF
+%% nodes, leading to a matrix with 884,736 rows. The four source vectors
+%% occupy more than 14M bytes, which exceeds the aggregate L2 cache of 31
+%% Mbytes (7.8 Mfloats.)
+
+%% \todo{Is the host architecture described? Problem: I had to change
+%%   hosts a few weeks ago, and it has different characteristics}.
+%% %----------------------
+
+%% Derivatives in an RBFFD formulation are expressed as a sparse
+%% matrix/vector multiplication (SpMV). Using the full 61 nodes of a MIC
+%% processor, and all four threads of each node, we achieve lowly 18
+%% Gflops when calculating a single vector of a single
+%% function. Bandwidth reduction has minimal effect since the cache is
+%% large enough to store both the derivative matrix and the solution
+%% vector. We increase the number of computations per byte transferred by
+%% calculating four different derivatives of four different
+%% functions. With bandwidth reduction, we speed the calculation by a
+%% factor of 8, and achieve upwards of 140 Gflops.
 

### 2013-10-19 ieee_paper.tex
MIC

--- 2013-10-19 ieee_paper.tex
@@
 instruction cache, and a 512kB L2 cache. The architecture of a core is
-based the traditional Pentium architecture, which has been extended to
+based on the traditional Pentium architecture, which has been extended to
 64-bit. A core can hold four hardware contexts at any given time. At each
@@
 addition or division, and mathematical operations, such as {\tt sin} and
-{\tt sqrt}, making it possible to reach 8 double precision (or 16 singe precision)
+{\tt sqrt}, making it possible to reach 16 single precision (or 8 double precision)
 operations per cycle.  

### 2013-10-19 ieee_paper.tex

--- 2013-10-19 ieee_paper.tex
@@
 Florida State University\\ %, Tallahassee %FL 32306-4120\\
+Email: gerlebacher@fsu.edu, bollig@gmail.com}
 }
 \IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte\\
-}
+Email: erik.saule@uncc.edu}
 \IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR \\
-}
-Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com }
+Email: flyer@ucar.edu}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
+%Email: bollig@gmail.com}\todo{check affiliation fsu's are strange.}}
 \maketitle

### 2013-10-19 ieee_paper.tex
Figures done. Perhaps make legends smaller?

@@
 Email: gerlebacher@fsu.edu, bollig@gmail.com}
-}
 \IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte\\

### 2013-10-19 ieee_paper.tex
minor

@@
 Florida State University\\ %, Tallahassee %FL 32306-4120\\
-Email: gerlebacher@fsu.edu, bollig@gmail.com}
+}
 \IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte\\
-Email: erik.saule@uncc.edu}
+}
 \IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR \\
-Email: flyer@ucar.edu}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
-%Email: bollig@gmail.com}\todo{check affiliation fsu's are strange.}}
+}
+Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com }
 \maketitle
@@
 at most. However, the bandwidth between the cores and the memory
-controllers is limited by the ring network that connect the cores and
+controllers is limited by the ring network that connects the cores and
 the memory controller. Its precise bandwidth is unknown but it is

### 2013-10-19 ieee_paper.tex
email

@@
 \IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Scientific Computing, 
-Florida State University\\ %, Tallahassee %FL 32306-4120\\
-}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte\\
-}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR \\
-}
-Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com }
+Florida State University}
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte}
+\IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR}
+Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com}
 \maketitle

### 2013-10-19 ieee_paper.tex
formula checked

@@
 performance represents 10\% of the peak performance of the
-architecture which is impossible without an efficient vectorization of
+architecture which is impossible to reach without an efficient vectorization of
 the kernel. Indeed, to reach such a performance, a Fused Multiply-Add
@@
 interleaved by group of four so that the value $A^l_{col\_id[k]}$ of
-the $k$th nonzero of the $l$th matrix is at index $\floor{\frac{k-1}{4}}*4*n_m + (l-1)*4+(k-1)\%4$.\todo{check that formula}
+the $k$th nonzero of the $l$th matrix is at index $\floor{\frac{k-1}{4}}*4*n_m + (l-1)*4+(k-1)\%4$.
 (In other words, the first 4 entries are from $A^{1}$ and the next 4 are from $A^2$.) The vectors {\tt x} are

### 2013-10-19 ieee_paper.tex
todo removed

@@
 based on the traditional Pentium architecture, which has been extended to
-64-bit. A core can hold four hardware contexts at any given time. At each
+64-bit. A core can hold four hardware contexts at any given time (240 threads in total). At each
 clock cycle, instructions from a single thread are executed. Due to
@@
 register as depicted in the figure is the difficult component of our 
-proposed algorith, which we proceed to explain further. 
+proposed algorithm, which we proceed to explain further. 
 A thread will perform the multiplications one row at a time,
-and parallelism is achieved by giving blocks of rows to each core
+and parallelism is achieved by giving blocks of rows to each thread
 using an OpenMP construct.
@@
 four nonzeros from the vectors.  
-%representing 16 floats from the matrix and 16 floats from the vectors at once. 
 (The number of nonzeros on a row in our application is
@@
 zeroes in the data structure for padding.)
-\todo{Erik, I rewrote the paragraph. Please reread.}
 

### 2013-10-19 ieee_paper.tex
section 4

@@
 lane of the vector register using an {\tt unpack}
-operation\footnote{Notice that there are two calls to unpack for the
-  LSB and MSB. Both are mandatory even if one is known not to be
-  useful according the documentation of the hardware}. (The column
+operation\footnote{Notice that there are two calls to {\tt unpack} for the
+  LSB and MSB. Both are mandatory even if one is known to be
+  nilpotent according the documentation of the hardware}. (The column
 pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They

### 2013-10-19 ieee_paper.tex
changes.

--- 2013-10-19 ieee_paper.tex
@@
 \begin{document}
-\title{Sparse Matrix Vector Multiplication with Multiple vectors and
-  Multiple Matrices on the MIC Architecture \todo{? more application oriented''}}
+\title{Acceleration of Derivative Calculations with Application to Radial Basis Function --
+Finite-Differences on the MIC Architecture}
+%\title{Sparse Matrix Vector Multiplication with Multiple vectors and
+  %Multiple Matrices on the MIC Architecture \todo{? more application oriented''}}
 
@@
 \IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Scientific Computing, 
-Florida State University}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR}
-Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com}
+Florida State University\\ %, Tallahassee %FL 32306-4120\\
+Email: gerlebacher@fsu.edu, bollig@gmail.com}
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte\\
+Email: erik.saule@uncc.edu}
+\IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR \\
+Email: flyer@ucar.edu}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{4}Department of Scientific Computing, Florida State University\\
+%Email: bollig@gmail.com}\todo{check affiliation fsu's are strange.}}
 \maketitle
@@
 at most. However, the bandwidth between the cores and the memory
-controllers is limited by the ring network that connects the cores and
+controllers is limited by the ring network that connect the cores and
 the memory controller. Its precise bandwidth is unknown but it is
@@
 performance represents 10\% of the peak performance of the
-architecture which is impossible to reach without an efficient vectorization of
+architecture which is impossible without an efficient vectorization of
 the kernel. Indeed, to reach such a performance, a Fused Multiply-Add
@@
 instructions have swizzling capability, while permuation is its own
-instruction. \todo{do we need a reference to intel Xeon phi instruction set?})
+instruction. %\todo{do we need a reference to intel Xeon phi instruction set?})
 

### 2013-10-19 ieee_paper.tex

--- 2013-10-19 ieee_paper.tex
@@
 based on the traditional Pentium architecture, which has been extended to
-64-bit. A core can hold four hardware contexts at any given time (240 threads in total). At each
+64-bit. A core can hold four hardware contexts at any given time. At each
 clock cycle, instructions from a single thread are executed. Due to
@@
 register as depicted in the figure is the difficult component of our 
-proposed algorithm, which we proceed to explain further. 
+proposed algorith, which we proceed to explain further. 
 A thread will perform the multiplications one row at a time,
-and parallelism is achieved by giving blocks of rows to each thread
+and parallelism is achieved by giving blocks of rows to each core
 using an OpenMP construct.
@@
 four nonzeros from the vectors.  
+%representing 16 floats from the matrix and 16 floats from the vectors at once. 
 (The number of nonzeros on a row in our application is
@@
 zeroes in the data structure for padding.)
+\todo{Erik, I rewrote the paragraph. Please reread.}
 

### 2013-10-19 ieee_paper.tex
minor
briging code one page early

--- 2013-10-19 ieee_paper.tex
@@
 
+\begin{figure*}
+  \centering
+\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{code.pdf}}%
+\subfigure[Content of the vector registers]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
+
+\caption{Code snippet of the multiplication of 4 vectors with 4
+  matrices. When a line changes the content of a vector register its
+  new content is shown on the right. The effects of swizzling and
+  permutations are highlighted using a color code.}
+\label{code:mat_mul}
+\end{figure*}
+
 As explained in Section~\ref{sec:model}, we expect to achieve a performance of
@@
 using an OpenMP construct.
-
-\begin{figure*}
-  \centering
-\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{code.pdf}}%
-\subfigure[Content of the vector registers]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
-
-\caption{Code snippet of the multiplication of 4 vectors with 4
-  matrices. When a line changes the content of a vector register its
-  new content is shown on the right. The effects of swizzling and
-  permutations are highlighted using a color code.}
-\label{code:mat_mul}
-\end{figure*}
 
@@
 are compiled with the Intel C Compiler version 13 and
--O3 optimization. The MIC codes use the IMCI instructions exposed by
+-O3 optimization. The codes use the IMCI instructions exposed by
 the compiler through intrinsics.
@@
 Figure~\ref{fig:bandwidth} gives some benchmark that compute the
-bandwidth achieved when reading and/or writing large arrays on both
-the MIC card and the CPU architecture. The read benchmark performs a
-simple sum of the array. The write benchmark sets whole the elemtns of
+bandwidth achieved when reading and/or writing large arrays. The read benchmark performs a
+simple sum of the array. The write benchmark sets all the elements of
 the array to 0. While the read/write benchmarks copies the array into

### 2013-10-19 ieee_paper.tex
no need for instruction set
Fixed grammar. Changed 4 to four.

--- 2013-10-19 ieee_paper.tex
@@
 the nonzeros in the four matrices. The values of the matrices are
-interleaved by group of four so that the value $A^l_{col\_id[k]}$ of
+interleaved by groups of four so that the value $A^l_{col\_id[k]}$ of
 the $k$th nonzero of the $l$th matrix is at index $\floor{\frac{k-1}{4}}*4*n_m + (l-1)*4+(k-1)\%4$.
@@
 
-Figure~\ref{fig:bandwidth} gives some benchmark that compute the
+Figure~\ref{fig:bandwidth} gives some benchmark that computes the
 bandwidth achieved when reading and/or writing large arrays on both
 the MIC card and the CPU architecture. The read benchmark performs a
-simple sum of the array. The write benchmark sets whole the elemtns of
-the array to 0. While the read/write benchmarks copies the array into
+simple sum of the array. The write benchmark sets all the elemnts of
+the array to zero, while the read/write benchmarks copies the array into
 another one. We also investigate the difference between a straight C++
 implementation (compiled with -O3 with proper tagging to inform the
-compiler of memory alignments) and a simple implementation using
-vector register explicitly.
+compiler of memory alignments) and a simple implementation with
+explicit use of the vector registers.
 
@@
 case reaches a performance of 158~GB/s while the
-{\tt write\_cpp} implementation never passes 58~GB/s. 
+{\tt write\_cpp} implementation never exceeds 58~GB/s. 
 This difference is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction, 
@@
 drops below 5~GB/s. We can see that implementing the
-indirection mechanism without vector instructions makes the
-performance drop significantly. (In the {\tt compact} case, it never
-passes 100~GB/s and is close to 50~GB/s most of the time.)
+indirection mechanism without vector instructions lowers the
+performance significantly. (In the {\tt compact} case, it never
+surpasses 100~GB/s and is close to 50~GB/s most of the time.)
 
@@
 
-To summarize, using vector instruction seems mandatory to reach the
-highest bandwidth on the MIC architecture. For the instructions used
+To summarize, using vector instructions appears necessary in order
+toreach the highest bandwidth on the MIC architecture. For the instructions used
 in a typical SpMV computation, one can expect a bandwidth between
-140~GB/s to 160~GB/s. That is why we set in our estimation a best
+140~GB/s to 160~GB/s. That is why we use in our estimations a best
 achievable bandwidth of 150~GB/s despite the fact that a 
@@
 structured grids and a
-{\tt RBF} matrix extracted from a real application. In the {\tt
-  compact} case, the performance can be as high as 195~Gflop/s being
+{\tt RBF} derivative matrix extracted from a real 3D application. In the {\tt
+  compact} case, the performance can be as high as 195~Gflop/s, which is
 within 15\% of the peak predicted performance. The performance of the
@@
 actual performance by transforming classical SpMV computation into the
-multiplication of 4 vectors by 4 matrices for the computation of the
+multiplication of four vectors by four matrices for the computation of the
 derivative of RBFs? Does using manual vectorization improve actual
@@
 performance achieved by a classical SpMV and by the multiplication of
-4 vectors by 4 matrices using either standard C++ code and our
+four vectors by four matrices using either standard C++ code and our
 optimized implementation. The

### 2013-10-19 ieee_paper.tex
typo

--- 2013-10-19 ieee_paper.tex
@@
 instructions have swizzling capability, while permuation is its own
-instruction. \todo{do we need a reference to intel Xeon phi instruction set?})
+instruction.)
 
@@
 
-When using the MIC card (Figure~\ref{fig:band_rw_MIC}), the {\tt
-  read\_cpp} implementation reaches a performance of 168~GB/s
+The {\tt
+  read, cpp} implementation reaches a performance of 168~GB/s
  which is better than the {\tt read}
@@
 is explained by the compiler adding loop unrolling and software
-prefetching in the assembly code in the {\tt read\_cpp} case, which is
+prefetching in the assembly code in the {\tt read, cpp} case, which is
 not added in the {\tt read} case. On the other hand, the {\tt write}
 case reaches a performance of 158~GB/s while the
-{\tt write\_cpp} implementation never passes 58~GB/s. 
+{\tt write, cpp} implementation never passes 58~GB/s. 
 This difference is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction, 
@@
 useful for some parallel algorithms. The {\tt read/write} benchmark
-performs at 140~GB/s while the {\tt read/write cpp} benchmark performs
+performs at 140~GB/s while the {\tt read/write, cpp} benchmark performs
 at about 90~GB/s.
 
-Note that SpMV kernels will likely use only simple read and simple
-writes on the MIC architecture. We also benchmark the performance we
+Note that SpMV kernels are not likely to use only simple read and simple
+writes. We also benchmark the performance we
 can achieve using {\tt gather} and {\tt unpack} operations and we
@@
 achievable bandwidth of 150~GB/s despite the fact that a 
-carefully crafted code to maximize bandwidth can reach higher
+ code carefully crafted to maximize bandwidth can reach higher
 values~\cite{Saule13-ARXIV}.

### 2013-10-19 ieee_paper.tex

--- 2013-10-19 ieee_paper.tex
@@
 \title{Acceleration of Derivative Calculations with Application to Radial Basis Function --
-Finite-Differences on the MIC Architecture}
+Finite-Differences on the Intel MIC Architecture}
 %\title{Sparse Matrix Vector Multiplication with Multiple vectors and
@@
 or 16 different derivatives. We demonstrate a novel implementation on
-the MIC architecture, taking into account its advanced swizzling and
+the Intel MIC architecture, taking into account its advanced swizzling and
 channel interchange features. We present benchmarks on a real data
@@
 \subsection{Bringing the data into the vector register}
+
+\begin{figure*}
+  \centering
+\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{code.pdf}}%
+\subfigure[Content of the vector registers]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
+
+\caption{Code snippet of the multiplication of 4 vectors with 4
+  matrices. When a line changes the content of a vector register its
+  new content is shown on the right. The effects of swizzling and
+  permutations are highlighted using a color code.}
+\label{code:mat_mul}
+\end{figure*}
 
@@
 register as depicted in the figure is the difficult component of our 
-proposed algorith, which we proceed to explain further. 
+proposed algorithm, which we proceed to explain further. 
 A thread will perform the multiplications one row at a time,
-and parallelism is achieved by giving blocks of rows to each core
+and parallelism is achieved by giving blocks of rows to each thread
 using an OpenMP construct.
-
-\begin{figure*}
-  \centering
-\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{code.pdf}}%
-\subfigure[Content of the vector registers]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
-
-\caption{Code snippet of the multiplication of 4 vectors with 4
-  matrices. When a line changes the content of a vector register its
-  new content is shown on the right. The effects of swizzling and
-  permutations are highlighted using a color code.}
-\label{code:mat_mul}
-\end{figure*}
 
@@
 instructions have swizzling capability, while permuation is its own
-instruction. %\todo{do we need a reference to intel Xeon phi instruction set?})
+instruction.
 
@@
 
-Figure~\ref{fig:bandwidth} gives some benchmark that computes the
+Figure~\ref{fig:bandwidth} shows a benchmark that computes the
 bandwidth achieved when reading and/or writing large arrays on both
@@
 is explained by the compiler adding loop unrolling and software
-prefetching in the assembly code in the {\tt read\_cpp} case, which is
+prefetching in the assembly code in the {\tt read, cpp} case, which is
 not added in the {\tt read} case. On the other hand, the {\tt write}
 case reaches a performance of 158~GB/s while the
-{\tt write\_cpp} implementation never exceeds 58~GB/s. 
+{\tt write, cpp} implementation never exceeds 58~GB/s. 
 This difference is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction, 
@@
 the matrix. As a result, only 32 values of {\tt x} are read
-from memory, and the expense associated with cache misses and gather
-operations is removed from consideration. The cost associated with 
-{\tt A} remains.   \todo{reread the above. Reread by Gordon}
-The Compact matrices
+from memory, and the expense associated with cache misses is removed from consideration. The cost associated with 
+{\tt A} remains. The Compact matrices
 (Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
@@
 reutilisation is the lowest from one row to the next.
+The other type of matrix is used in derivative stencils in 3D
+RBF-FD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-points
+stencil of a 3D $8^3$ grid).
 
-The other two types of matrices are used in derivative stencils in 3D
-RBF-FD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-points
-stencil of a 3D $8^3$ grid). We also apply a Cuthill-Mckee reordering to
-all the matrices.  This ordering technique is designed to reduce the
+We also apply a Reverse Cuthill-Mckee reordering to
+all the matrices.  This ordering technique aims to reduce the
 distance between the nonzeros and the diagonal, hopefully allowing a
@@
   \caption{Performance of the manually vectorized code. The two
-    horizontal lines depict the worst and the best predicted performances. \todo{font size}}
+    horizontal lines depict the worst and the best predicted performances.}
   \label{fig:expe_types}
@@
 computation. The {\tt supercompact} case peaks at
-208~Gflop/s being very close to the predicted peak
+208~Gflop/s which is very close to the predicted peak
 performance of 213~Gflop/s. Conversely the performance of the {\tt
@@
   compact} case, the performance can be as high as 195~Gflop/s, which is
-within 15\% of the peak predicted performance. The performance of the
+within 15\% of the predicted  peak performance. The performance of the
 {\tt RBF} case varies between 100~Gflop/s and 140~Gflop/s. Most of the
@@
 The standard C++ implementation reaches a performance of 38~Gflop/s while our
-optimized implementation reaches 140~Gflop/s. 
-Notice that \todo{I do not know that I ran host and mic on the same grid, but your 
-numbers are approximately correct. The numbers are a function of resolution} 
+optimized implementation almost reaches 140~Gflop/s. 
+Notice that 
 RCM ordering has no impact in the standard C++ implementation while it
@@
 42~Gflop/s. This confirms that our method is sound and generalizes to
-architectures than the Intel MIC architecture. \todo{albeit with less 
-impressive results given that swizzling is often not available. Swizzling
-will become available in a new version of the host: enhanced AVX.}
+architectures other than the Intel MIC architecture.
 

### 2013-10-19 ieee_paper.tex
shrinking text

@@
   y} vectors, which lead to the size of the {\tt x} and {\tt y} vectors:
-$$s_x = n_v b_x n_r$$ $$s_y = n_v n_m b_x n_r$$
+$$s_x = n_v b_x n_r\mbox{, ~}s_y = n_v n_m b_x n_r$$
 
@@
 controllers is limited by the ring network that connects the cores and
-the memory controller. Its precise bandwidth is unknown but it is
+the memory controller. Its precise bandwidth 
 believed to be between 200~GB/s and 250~GB/s.
@@
 64-bit values or as a vector of 16 32-bit values, respectively. The
-vector processing unit can perform many basic instructions, such as
+vector processing unit can perform basic instructions, such as
 addition or division, and mathematical operations, such as {\tt sin} and
@@
 on the single precision case with four vectors and four matrices. 
-Similar techniques apply for other matrix/vector combinations. This target
+Similar techniques apply for other combinations. This target
 performance represents 10\% of the peak performance of the
-architecture which is impossible to reach without an efficient vectorization of
-the kernel. Indeed, to reach such a performance, a Fused Multiply-Add
+architecture which cannot be reached without an efficient vectorization. Indeed, to reach such a performance, a Fused Multiply-Add
 instruction on fully loaded registers must be executed at most every
@@
 using a swizzle operation. Then each value is multiplied by 4 
-(because there are 4 vectors) 
+(because $n_v=4$) 
 and the offset vector $(3,2,1,0)$ is added to each lane to obtain the
@@
 performed to bring the 16 entries of the {\tt x} vector into the SIMD
-register. At this point we have a SIMD register where each lane is the
+register. At this point we have a register where each lane is the
 four vector entries for a nonzero element. Using a broadcast
@@
 \label{sec:expe}
+
+\begin{figure}[tbh]
+  \centering
+  \includegraphics[width=.9\linewidth]{figures/bandwidth_read_write.pdf}
+
+  \caption{Bandwidth performance under idealized conditions.  Entries
+    with "cpp" (triangles) denote cases where vectorization is
+    compiler generated.}
+  \label{fig:bandwidth}
+  \label{fig:band_rw_MIC}
+\end{figure}
+
 In this section, we describe a series of numerical experiments to
@@
 \subsection{Bandwidth}
-
-\begin{figure}[bth]
-  \centering
-  \includegraphics[width=.9\linewidth]{figures/bandwidth_read_write.pdf}
-
-  \caption{Bandwidth performance under idealized conditions.  Entries
-    with "cpp" (triangles) denote cases where vectorization is
-    compiler generated.}
-  \label{fig:bandwidth}
-  \label{fig:band_rw_MIC}
-\end{figure}
 

### 2013-10-19 ieee_paper.tex
ccl
harmonizing code and text

@@
 constant throughout the matrix. The matrix is given in two arrays. The
-first one is the classical {\tt col\_id} array that describes the
+first one is {\tt col\_id} that describes the
 sparsity structure of the matrix. Each row of {\tt col\_id} consists
 of $n_z$ entries that label the column index of each nonzero in
-$A$. The other array is the {\tt data} array that gives the values of
+$A$. The other array is  {\tt data} that gives the values of
 the nonzeros in the four matrices. The values of the matrices are
@@
 
-In a future work, we will examine the effect of larger stencil sizes
-and double precision. We will also integrate the techniques
+In a future work, we will examine the effect of larger stencil sizes, double precision and port common techniques from SpMV. We will also integrate the techniques
 presented in this paper within a fluid simulation using RBF-FD.

### 2013-10-19 ieee_paper.tex
sentence hacking

@@
 interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
-consecutive in memory. The vector {\tt y} is similarly interleaved,
+consecutive in memory. The vector {\tt y} is similarly 
 grouped in sets of 16, which correspond to four derivatives applied to
-four functions, at a single point.
+four functions, at a single point.\looseness=-1
 

### 2013-10-19 ieee_paper.tex
sentence hacking

@@
 reach a total of about 210~Gflop/s. This performance represents
-approximatively 10\% of the available flop/s of a Knight Corner
+approximatively 10\% of the available flop/s of a Xeon Phi
 coprocessor. Therefore, it is necessary to have an implementation that
@@
 performance of 135~Gflop/s is achieved on real scenario. Concluding
-remarks and perspectives are provided in Section~\ref{sec:ccl}.
+remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1
 
@@
 
-\subsection{Knights Corner}
+\subsection{Intel Xeon Phi}
 

### 2013-10-19 ieee_paper.tex
ccl

@@
 
-In a future work, we will examine the effect of larger stencil sizes, double precision and port common techniques from SpMV. We will also integrate the techniques
-presented in this paper within a fluid simulation using RBF-FD.
+In a future work, we will examine the effect of larger stencil sizes,
+double precision. The presented performance do not reach the predicted
+peak and we believe that common techniques from SpMV such as cache
+partitioning can be successfully applied. We will also integrate the
+techniques presented in this paper within a fluid simulation using
+RBF-FD.
 

### 2013-10-19 ieee_paper.tex
spell check

@@
 reach a total of about 210~Gflop/s. This performance represents
-approximatively 10\% of the available flop/s of a Xeon Phi
+approximately 10\% of the available flop/s of a Xeon Phi
 coprocessor. Therefore, it is necessary to have an implementation that
@@
 where $\phi_{ij} = \phi(\|\rvec_i-\rvec_j\|)$.
-%\end{equation}
-%\begin{equation}
-%\begin{bmatrix}
-%\phi(\|\xvec_{1}-\xvec_{1}\|) & \phi(\|\xvec_{1}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{1}-\xvec_{N}\|) \\
-%\vdots & \ddots & \vdots  \\
-%\phi (\|\xvec_{n}-\xvec_{1}\|) & \phi(\|\xvec_{n}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{N}-\xvec_{N}\|)
-%\end{bmatrix}
 
@@
 On the other hand, in the worst case, every time the {\tt x} vector is
-accessed, the value needs to be transfered from memory again. So in
+accessed, the value needs to be transferred from memory again. So in
 total, there are as many transfers as the number of nonzeros in the
@@
 single byte at a time. Instead, a minimum number of bytes, equal to
-the size of a cacheline $cl$, are transfered at once.  When there are
+the size of a cacheline $cl$, are transferred at once.  When there are
 multiple vectors, each nonzero element uses $n_v$ consecutive
-entries. In the worst case, the number of bytes read and transfered is
+entries. In the worst case, the number of bytes read and transferred is
 $$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}} $$ 
@@
 The computation intensity is the amount of floating point operations performed per
-byte transfered. In the worst case and in the best case, we have
+byte transferred. In the worst case and in the best case, we have
 $$I_b = \frac{O}{b_T} = \frac{2 n_v n_m}{ (b_i + b_x n_m) + n_v n_m b_x n_z^{-1} + n_v b_x n_z^{-1} }$$
@@
 We discuss how the implementation of the v4m4 kernel works using
-vectorial instructions. The code is given for reference in
+vector instructions. The code is given for reference in
 Figure~\ref{code:mat_mul}. (We will show in Section~\ref{sec:expe}
@@
 permutation if possible because we believe it is cheaper. Many
-instructions have swizzling capability, while permuation is its own
+instructions have swizzling capability, while permutation is its own
 instruction.
@@
 The values from the nonzero elements in the matrix are the simplest to
-load. One can load the 16 floating point values for four nonzeroes of
+load. One can load the 16 floating point values for four nonzeros of
 the matrix at once. The four floats coming from each matrix are

### 2013-10-19 ieee_paper.tex
typos
reordered figures
figure placement
Changes.

@@
 sought.  The derivative at all points takes the form of a sparse
-matrix/vector multiplication (spmv).
+matrix/vector multiplication (SpMV).
 
-In this paper, we consider the case of local stencils with the number
+In this paper, we consider the case of local stencils with a fixed number
 of nodes at each point and encode the sparse matrix in ELLPACK
@@
 provide some useful yet limited improvements). Register blocking~\cite{conf/ppsc/Toledo97} does
-not apply well to most the matrix at hand (despite it can be applied
-with virtually no overhead in all cases thanks to compressed
-representation~\cite{Buluc11}). Indeed, there is about 8 bytes of the matrix to move
-in per nonzero (assuming single precision); each nonzero requires two
+not work well to many of the matrices in general use (although it can be applied
+with virtually no overhead thanks to compressed
+representations~\cite{Buluc11}.) Indeed, there are about 8 bytes of the matrix to transfer
+from memory 
+per nonzero in single precision; each nonzero requires two
 floating point operations leading to a flop-to-byte ratio of at most
 $\frac{1}{4}$. This limits the obtained performance to at most a
-quarter of the bandwidth of the architecture wasting a lot of
+quarter of the bandwidth of the architecture, wasting a lot of
 potentially useful cycles. The commonly used techniques are mostly
@@
 Fortunately that fate is not inevitable. One solution would be to
-schedule a more instruction intensive kernel simultaneously with the
+schedule a more instruction-intensive kernel simultaneously with the
 execution of SpMV, relying on some hardware threading capabilities,
-such as HyperThreading, to reduce the cycle wastage. However, most ot
-the applications that use SpMV do not typically have an instruction
-intensive kernel to run simultaneously.
+such as HyperThreading, to reduce the cycle wastage. However, most of 
+the applications that use SpMV do not typically have an 
+instruction-intensive kernel to run simultaneously.
 
@@
 every node in the domain to calculate the derivative at a given node
-$\rvec_i$), RBF-generated finite differences (RBF-FD) have been
-derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBF-FD uses only
+$\rvec_i$), RBF-generated finite differences (RBFFD) have been
+derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBFFD uses only
 a local set of the $n_z-1$  (often nearest) points in a neighborhood of 
@@
 each of size $n_z \times n_z$. Each linear solve produces a row of the
-RBF-FD differentiation matrix $A_{n_z}$, resulting in a $N \times N$
+RBFFD differentiation matrix $A_{n_z}$, resulting in a $N \times N$
 matrix with $n_zN$ nonzero entries. To evaluate the derivative at all
@@
 
-\begin{figure}[tbh]
+\begin{figure}[t]
   \begin{center}
@@
 
-\begin{figure}[tbh]
+\begin{figure}[t]
   \centering
@@
 \subsection{Bandwidth}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=.9\linewidth]{figures/bandwidth_gather_unpack.pdf}
+  \caption{Performance of {\tt gather} and {\tt unpack} operations.}
+  \label{fig:band_gather} 
+\end{figure}
 
@@
 surpasses 100~GB/s and is close to 50~GB/s most of the time.)
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.9\linewidth]{figures/bandwidth_gather_unpack.pdf}
-  \caption{Performance of {\tt gather} and {\tt unpack} operations.}
-  \label{fig:band_gather} 
-\end{figure}
-%
-
-To summarize, using vector instructions appears necessary in order
-toreach the highest bandwidth on the MIC architecture. For the instructions used
-in a typical SpMV computation, one can expect a bandwidth between
-140~GB/s to 160~GB/s. That is why we use in our estimations a best
-achievable bandwidth of 150~GB/s despite the fact that a 
- code carefully crafted to maximize bandwidth can reach higher
-values~\cite{Saule13-ARXIV}.
-
-\subsection{Instances}
-
 \def\ww{.19\textwidth}
-\begin{figure*}[tbh]
+\begin{figure*}[t]
   \begin{center}
@@
     512 rows and 32 nonzeros per row. The last two matrices
-    corresponds to a derivative stencil in 3D RBF-FD
+    corresponds to a derivative stencil in 3D RBFFD
     calculations (with and without RCM ordering).}
@@
 \end{figure*}
+
+To summarize, using vector instructions appears necessary in order
+to reach the highest bandwidth on the MIC architecture. For the instructions used
+in a typical SpMV computation, one can expect a bandwidth between
+140~GB/s to 160~GB/s. That is why we use in our estimations a best
+achievable bandwidth of 150~GB/s despite the fact that a 
+ code carefully crafted to maximize bandwidth can reach higher
+values~\cite{Saule13-ARXIV}.
+
+\subsection{Instances}
 
@@
 The other type of matrix is used in derivative stencils in 3D
-RBF-FD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-points
+RBFFD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-points
 stencil of a 3D $8^3$ grid).
@@
 
+\vspace{-0.3em}
 \subsection{Computations}
@@
   \subfigure[Supercompact and Random.]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
-  \subfigure[Compact and RBF-FD.]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
+  \subfigure[Compact and RBFFD.]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
 
@@
 
-\begin{figure}
+\begin{figure}[t]
   \centering
@@
 
+\vspace{-0.5em}
 \section{Conclusion}
@@
 techniques presented in this paper within a fluid simulation using
-RBF-FD.
+RBFFD.
 

### 2013-10-19 ieee_paper.tex
changes.

@@
 case of the derivative of RBFs where four derivatives of four
-different functions is expressed as the multiplication of four vectors
+different functions are expressed as the multiplication of four vectors
 by four matrices with identical sparsity patterns leading to the
@@
 at a time has been investigated before to densify the
-computations~\cite{Im01} but this is the first time that the
+computations~\cite{Im01} but this is the first time, to the 
+best of our knowledge, that the
 densification comes from adding both vectors and matrices.
@@
 To perform our analysis, we focus our attention on the improvement
-that we can achieve on the Intel Xeon Phi processor. It follow the
+that can be achieved on the Intel Xeon Phi processor. It follows the
 Many Integrated Core (MIC) architecture, which has a significant
@@
 for sparse linear algebra compared to more classical CPU or GPU
-architecture~\cite{Saule13-ARXIV, Liu:2013:ESM:2464996.2465013, cramer2012openmp}.
+architectures~\cite{Saule13-ARXIV, Liu:2013:ESM:2464996.2465013, cramer2012openmp}.
 
-In Section~\ref{sec:rbf} we present the computation of RBFs and how it
+In Section~\ref{sec:rbf}, we present the computation of RBFs and how it
 can be expressed as sixteen multiplications of four vectors by four sparse matrices
@@
 estimation of the instruction intensity of various forms of the
-computations and we show that a sixfold improvement can be
-expected when computing the sixteen multiplications simultaneously and
+computations. We show that a sixfold improvement can be
+expected when computing the sixteen multiplications simultaneously to
 reach a total of about 210~Gflop/s. This performance represents
@@
 depending on how the SpMV kernel is written. It also provides the actual performance
-of the various kernel on multiple classes of matrices, some generated
-for analysis purpose, and some extracted from an RBF application. A
-performance of 135~Gflop/s is achieved on real scenario. Concluding
+of the various kernels on multiple classes of matrices, some generated
+for purpose of analysis, and some extracted from an RBF application. A
+performance of 135~Gflop/s is achieved using RBF-generated matrices. Concluding
 remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1
@@
 %$\sqrt{(x-x_j)^2+(y-y_j)^2}$) 
-between where the function is evaluated
-$\rvec$ and where the RBF is centered at $\rvec_{j}$. That is, the
+between $\rvec$, where the function is evaluated
+and and $\rvec_j$,  where the RBF is centered. That is, the
 interpolant is $s(\rvec) = \sum_{j=1}^{N} w_j
@@
 in (\ref{eq:rbf}) is full and inversion requires O$(N^3)$
-operations. To alleviate the cost of this global approach (i.e. using
+operations. To alleviate the cost of this global approach (i.e., using
 every node in the domain to calculate the derivative at a given node
@@
 $\{\phi(\|\rvec-\rvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the
-node locations $\{\rvec_j\}_{j=1}^{n_z}$ (classical finite differences
-(FD) would enforce that it be exact for polynomials instead). Similar
+node locations $\{\rvec_j\}_{j=1}^{n_z}$ (classical finite-differences
+(FD) would enforce that it be exact for polynomials instead.) Similar
 to FD, as the stencil size $n_z$ increases so does the order of the
@@
 where each row has $n_z$ nonzeros\footnote{We use non-conventional 
-symbology for differentiation matrix, source and destination vectors
+symbology for the differentiation matrix, source and destination vectors, 
 anticipating what follows.}. 
@@
 This relationship is not
-necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} which
+necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils}, which
 implies a non-symmetric sparsity pattern in the derivative matrix.
@@
 have the same sparsity distribution as long as the derivative stencils
-remain unchanged; only the values in the sparse matrix change. 
+remain unchanged; only the values in the sparse matrix vary. 
 Thus, rather than computing a derivative of multiple
@@
 leading to an overall performance benefit. In practice however, the number of 
-vectors whose derivatives are required is limited, limiting the 
+vectors whose derivatives are required is limited, capping the 
 performance of the matrix/vector multiplication. Similarly, the number
-of different derivatives in a particular computation is limited. It is 
+of different derivatives in a particular computation is finite. It is 
 for this reason that we seek to combine the benefits of multiple
@@
 
-In discrete form, Equation (\ref{eq:Q}) takes the form
+In discrete form, Equation (\ref{eq:Q}) becomes
 $$
@@
 and its impact on the expected performance of the
-application. Relevant notations are given in Figure~\ref{tab:not}.
+application. Relevant notation are given in Figure~\ref{tab:not}.
 
@@
 cores. \cite{Saule13-ARXIV} showed that there is very little cache trashing
-in practice, but it showed that having elements of the vectors used by
+in practice; however having elements of the vectors used by
 multiple cores can have a significant impact on the performance
@@
 for single precision computation in the best and the worst case on a
-classical cache based architecture ($cl = 64$) and assuming $cl=1$. We
+classical cache-based architecture ($cl = 64$) and assuming $cl=1$. We
 can easily see the potential improvement in the computational
@@
 low: this highlights that accessing the memory with a granularity of one cache
-line
-is the main problem faced when performing a standard SpMV computation.
+line is the main problem faced when performing a standard SpMV computation.
 
@@
 controllers is limited by the ring network that connects the cores and
-the memory controller. Its precise bandwidth 
+the memory controller. Its precise bandwidth is
 believed to be between 200~GB/s and 250~GB/s.
@@
 
-In a future work, we will examine the effect of larger stencil sizes,
+In future work, we will examine the effect of larger stencil sizes and
 double precision. The presented performance do not reach the predicted

### 2013-10-19 ieee_paper.tex
changes.

@@
 
-Each core in the architecture has a 32kB L1 data cache, a 32kB L1
-instruction cache, and a 512kB L2 cache. The architecture of a core is
+Each core in the architecture has a 32~kB L1 data cache, a 32~kB L1
+instruction cache, and a 512~kB L2 cache. The architecture of a core is
 based on the traditional Pentium architecture, which has been extended to
@@
 cycle. However, only one vector or floating point instruction can be
-executed at each cycle.
+executed per cycle.
 
@@
 processing unit. Each core has 32 512-bit SIMD registers that can be
-used for double or single precision, that is, either as a vector of 8
+used for double or single precision, that is, either as a vector of eight
 64-bit values or as a vector of 16 32-bit values, respectively. The
 vector processing unit can perform basic instructions, such as
-addition or division, and mathematical operations, such as {\tt sin} and
+addition or division, and mathematical operations, such as {\tt sin} or
 {\tt sqrt}, making it possible to reach 16 single precision (or 8 double precision)
@@
 The unit also supports Fused Multiply-Add (FMA)
-operations, which are typically counted as two operations for
+operations, which are typically counted as two operations 
 when benchmarking. Therefore, the peak performance of the 5110P
-card is 1 Tflop/s in double precision and 2 Tflop/s in single
+card is one Tflop/s in double precision and two Tflop/s in single
 precision. If FMA cannot be used, only half of these rates can be
@@
 performance represents 10\% of the peak performance of the
-architecture which cannot be reached without an efficient vectorization. Indeed, to reach such a performance, a Fused Multiply-Add
+architecture, which cannot be reached without an efficient vectorization. 
+Indeed, to reach such a performance, a Fused Multiply-Add
 instruction on fully loaded registers must be executed at most every
@@
 constant throughout the matrix. The matrix is given in two arrays. The
-first one is {\tt col\_id} that describes the
+first array is {\tt col\_id} that describes the
 sparsity structure of the matrix. Each row of {\tt col\_id} consists
@@
 the $k$th nonzero of the $l$th matrix is at index $\floor{\frac{k-1}{4}}*4*n_m + (l-1)*4+(k-1)\%4$.
-(In other words, the first 4 entries are from $A^{1}$ and the next 4 are from $A^2$.) The vectors {\tt x} are
+(In other words, the first four entries are from $A^{1}$ and the 
+next four are from $A^2$.) The vectors {\tt x} are
 interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
@@
 goal is to load and format the nonzero values $A_h^1,A_h^2,A_h^3,A_h^4$
-and the vector entries $x_m^1,x_m^2,x_m^3,x_m^4$ in two vector registers and 
+and the vector entries $x_m^1,x_m^2,x_m^3,x_m^4$ into two vector registers and 
 apply Fused Multiply-Add on them, which is depicted in the {\tt accu} line of 
 Figure~\ref{code:mat_mul}. Bringing the data into the vector
-register as depicted in the figure is the difficult component of our 
+register as shown in the figure is the difficult component of our 
 proposed algorithm, which we proceed to explain further. 
@@
 {\tt mat} is extracted from {\tt mat\_ent} in
-Figure~\ref{code:mat_mul}). While a permutation reorders entire lanes,
+Figure~\ref{code:mat_mul}). On the other hand, a permutation reorders entire lanes,
 without changing their contents (see for instance how {\tt vect} is
@@
   \_mm512\_permute4f128\_epi32} instruction which is only defined on
-integers and therefore in place typecasts are necessary to convert to
-a vector to/from the integer type.  (Note that we prefer swizzling to
+integers and therefore in place typecasts are necessary to convert
+a vector to and from integer type.  (Note that we prefer swizzling to
 permutation if possible because we believe it is cheaper. Many
 instructions have swizzling capability, while permutation is its own
-instruction.
+instruction.)
 
@@
 in the matrix will cause 16 operations to be performed, but it only
-uses 4 floats from the matrix and 4 floats from the {\tt x} vector. So
+uses four floats from the matrix and four floats from the {\tt x} vector. So
 we would like to execute a single instruction to load the data for four 
@@
 (The number of nonzeros on a row in our application is
-always a multiple of 4; if it weren't, we would add some explicit
-zeroes in the data structure for padding.)
+always a multiple of four; if it weren't, we would add some explicit
+zeros in the data structure for padding.)
 
@@
 the matrix at once. The four floats coming from each matrix are
-naturally grouped within each lane of the SIMD vector. One can use a
-swizzle operation to broadcast the 4 elements of a single matrix to
-take the whole vector. (The broadcast of $A_h^1,A_h^2,A_h^3,A_h^4$ is
+naturally grouped within each lane of the SIMD register. One can use a
+swizzle operation to broadcast the four elements of a single matrix to
+fill the whole register. (The broadcast of $A_h^1,A_h^2,A_h^3,A_h^4$ is
 shown in Figure~\ref{code:mat_mul}.)
@@
 Loading the entries from the {\tt x} vector is a little bit more
-involved. First, the columns pointers of four different nonzeros are
+involved. First, the column pointers of four different nonzeros are
 loaded into a vector register and the values are distributed one per
@@
 LSB and MSB. Both are mandatory even if one is known to be
-  nilpotent according the documentation of the hardware}. (The column
+nilpotent according the documentation of the hardware.} (The column
 pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They
-are replicated so that each value fills an single vector lane 
-using a swizzle operation. Then each value is multiplied by 4 
+are replicated so that each value fills a single vector lane 
+using a swizzle operation. Then each value is multiplied by four
 (because $n_v=4$) 
@@
 correct index of the elements of the {\tt x} vector within each lane
-using a Fused Multiply-Add operation. Then a {\tt gather} operation is
+using a Fused Multiply-Add operation. A {\tt gather} operation is then
 performed to bring the 16 entries of the {\tt x} vector into the SIMD

### 2013-10-19 ieee_paper.tex
fixing wrong hypenation
simple change to reach 10 pages
changes.

@@
 
-The {\tt
-  read, cpp} implementation reaches a performance of 168~GB/s
+The {\tt read, cpp} implementation reaches a performance of 168~GB/s
  which is better than the {\tt read}
@@
 randomly (uniformly) distributed in the matrix; they represent the
-worst case scenario for a cache based architecture where the cache
+worst case scenario for a cache-based architecture where the cache
 reutilisation is the lowest from one row to the next.
@@
 all the matrices.  This ordering technique aims to reduce the
-distance between the nonzeros and the diagonal, hopefully allowing a
-better cache reutilisation. The reordered version of the matrix can be
+distance between the nonzeros and the diagonal, hopefully increasing 
+the cache hit ratio.  The reordered version of the matrix can be
 seen in Figure~\ref{fig:rbf3drcm}.

### 2013-10-19 ieee_paper.tex
another mbox to prevent hyphenation

@@
 
-Loading the entries from the {\tt x} vector is a little bit more
+Loading the entries from the {\tt x} vector is  more
 involved. First, the column pointers of four different nonzeros are
@@
 are compiled with the Intel C Compiler version 13 and
--O3 optimization. The codes use the IMCI instructions exposed by
+\mbox{-O3} optimization. The codes use the IMCI instructions exposed by
 the compiler through intrinsics.
@@
 another one. We also investigate the difference between a straight C++
-implementation (compiled with -O3 with proper tagging to inform the
+implementation (compiled with \mbox{-O3} with proper tagging to inform the
 compiler of memory alignments) and a simple implementation with

### 2013-10-19 ieee_paper.tex
Final changes from Gordon last reading (with Erik).

--- 2013-10-19 ieee_paper.tex
@@
 To put things in perspective, we conducted some experiments on a
-classical CPU architecture with two Intel Core i7 processors, each has 8
+classical CPU architecture with two Intel Core i7 processors, each has eight
 cores that support AVX instructions. Both the compiler-generated and
-the manually-written vectorization using AVX instruction reach a
+the manually-written vectorization using AVX instructions reach a
 (read) bandwidth of 70~GB/s. This indicates that the performance of
@@
 the {\tt RBF} matrix reaches a performance of 27~Gflop/s; while a
-manual implementation using AVX instruction reaches a performance of
+manual implementation using AVX instructions reaches a performance of
 42~Gflop/s. This confirms that our method is sound and generalizes to
@@
 In the previous sections, we have explored the practical
-implementation on a Intel Xeon Phi card of multiple derivative
+implementation on an Intel Xeon Phi card of multiple derivative
 operators acting on multiple vectors within the context of
-RBF-RD. Each derivative has an associated sparse matrix with a fixed
+RBFFD. Each derivative has an associated sparse matrix with a fixed
 number of nonzeros on each row. While computing a single derivative of
-multiple functions is rather common (vectorization occurs over the
-vectors), we accelerate the algorithm further by considering multiple
-matrices (identical sparsity patterns) acting on multiple vectors. 
+multiple functions is rather common, we accelerate the algorithm further 
+by considering multiple
+matrices, with identical sparsity patterns, acting on multiple vectors. 
 We specialize our study to four matrices and four derivatives,
@@
 In future work, we will examine the effect of larger stencil sizes and
-double precision. The presented performance do not reach the predicted
+double precision. The presented performance does not reach the predicted
 peak and we believe that common techniques from SpMV such as cache

### 2013-10-19 figure_files/figure10/performance.py
Added data and python scripts necessary to generate the figures.

@@ -1,0 +1,119 @@
+
+#ell_kd-tree_rcm_sym_1_x_weights_direct__no_hv_stsize_32_3d_32x_32y_32z.out:method_8a_multi, threads: 240, Max Gflops: 162.098709, min time: 0.207000 (ms)
+
+#Create secondary file with
+#nx ny nz rcm/norcm gflops
+
+str = "#ell_kd-tree_rcm_sym_1_x_weights_direct__no_hv_stsize_32_3d_32x_32y_32z.out:method_8a_multi, threads: 240, Max Gflops: 162.098709, min time: 0.  207000 (ms)"
+
+import matplotlib.pyplot as plt
+import numpy as np
+import myutil
+import csv
+
+def readData(filename):
+    x1 = []
+    y1 = []
+    y2 = []
+    with open(filename, "rb") as infile:
+        reader = csv.reader(infile)
+        for i, line in enumerate(reader):
+            x1.extend([int(line[0])])
+            y1.extend([float(line[1])])
+            y2.extend([float(line[2])])
+    print [x1,y1,y2]
+    return [x1,y1,y2]
+
+# in the file: x, norcm, rcm
+[nx,n2,n3] = readData("none1.data")
+[cx,c2,c3] = readData("compact.data")
+[sx,s2,s3] = readData("supercompact.data")
+[rx,r2,r3] = readData("random.data")
+
+
+#-------------------------
+plt.clf()
+plt.plot(sx,s2,'rs-',label='supercompact, no rcm', ms=10, lw=2)
+plt.plot(sx,s3,'r^-',label='supercompact, rcm', ms=10, lw=2)
+plt.plot(rx,r2,'bs-',label='random, no rcm', ms=10, lw=2)
+plt.plot(rx,r3,'b^-',label='random, rcm', ms=10, lw=2)
+plt.ylim([50,220])
+plt.ylabel('Performance (Gflop/s)', fontsize=18)
+plt.xlabel('number rows / 1000', fontsize=18)
+plt.legend(loc=7, fontsize=18)  # upper left
+plt.rc('xtick', labelsize=16)
+plt.rc('ytick', labelsize=16)
+myutil.horizontal_lines()
+plt.savefig.format = "pdf"
+plt.savefig("random_supercompact.pdf")
+#-------------------------
+plt.clf()
+plt.plot(cx,c2,'rs-',label='compact, no rcm', ms=10, lw=2)
[…]

### 2013-10-19 figure_files/figure10/performance_compact_rbf.py

@@ -1,0 +1,121 @@
+
+#ell_kd-tree_rcm_sym_1_x_weights_direct__no_hv_stsize_32_3d_32x_32y_32z.out:method_8a_multi, threads: 240, Max Gflops: 162.098709, min time: 0.207000 (ms)
+
+#Create secondary file with
+#nx ny nz rcm/norcm gflops
+
+str = "#ell_kd-tree_rcm_sym_1_x_weights_direct__no_hv_stsize_32_3d_32x_32y_32z.out:method_8a_multi, threads: 240, Max Gflops: 162.098709, min time: 0.  207000 (ms)"
+
+import matplotlib.pyplot as plt
+import numpy as np
+import myutil
+import csv
+
+def readData(filename):
+    x1 = []
+    y1 = []
+    y2 = []
+    with open(filename, "rb") as infile:
+        reader = csv.reader(infile)
+        for i, line in enumerate(reader):
+            x1.extend([int(line[0])])
+            y1.extend([float(line[1])])
+            y2.extend([float(line[2])])
+    print [x1,y1,y2]
+    return [x1,y1,y2]
+
+# in the file: x, norcm, rcm
+[nx,n2,n3] = readData("none1.data")
+[cx,c2,c3] = readData("compact.data")
+[sx,s2,s3] = readData("supercompact.data")
+[rx,r2,r3] = readData("random.data")
+
+
+#-------------------------
+"""
+plt.clf()
+plt.plot(sx,s2,'rs-',label='supercompact, no rcm', ms=10, lw=2)
+plt.plot(sx,s3,'r^-',label='supercompact, rcm', ms=10, lw=2)
+plt.plot(rx,r2,'bs-',label='random, no rcm', ms=10, lw=2)
+plt.plot(rx,r3,'b^-',label='random, rcm', ms=10, lw=2)
+plt.ylim([50,220])
+plt.ylabel('Performance (Gflop/s)', fontsize=18)
+plt.xlabel('number rows / 1000', fontsize=18)
+plt.legend(loc=7, fontsize=18)  # upper left
+plt.rc('xtick', labelsize=16)
+plt.rc('ytick', labelsize=16)
+myutil.horizontal_lines()
+plt.savefig.format = "pdf"
+plt.savefig("random_supercompact.pdf")
+"""
+#-------------------------
+plt.clf()
+plt.plot(cx,c2,'rs-',label='compact, no rcm', ms=10, lw=2)
[…]

### 2013-10-19 figure_files/figure10/performance_super_random.py

@@ -1,0 +1,121 @@
+
+#ell_kd-tree_rcm_sym_1_x_weights_direct__no_hv_stsize_32_3d_32x_32y_32z.out:method_8a_multi, threads: 240, Max Gflops: 162.098709, min time: 0.207000 (ms)
+
+#Create secondary file with
+#nx ny nz rcm/norcm gflops
+
+str = "#ell_kd-tree_rcm_sym_1_x_weights_direct__no_hv_stsize_32_3d_32x_32y_32z.out:method_8a_multi, threads: 240, Max Gflops: 162.098709, min time: 0.  207000 (ms)"
+
+import matplotlib.pyplot as plt
+import numpy as np
+import myutil
+import csv
+
+def readData(filename):
+    x1 = []
+    y1 = []
+    y2 = []
+    with open(filename, "rb") as infile:
+        reader = csv.reader(infile)
+        for i, line in enumerate(reader):
+            x1.extend([int(line[0])])
+            y1.extend([float(line[1])])
+            y2.extend([float(line[2])])
+    print [x1,y1,y2]
+    return [x1,y1,y2]
+
+# in the file: x, norcm, rcm
+[nx,n2,n3] = readData("none1.data")
+[cx,c2,c3] = readData("compact.data")
+[sx,s2,s3] = readData("supercompact.data")
+[rx,r2,r3] = readData("random.data")
+
+
+#-------------------------
+plt.clf()
+plt.plot(sx,s2,'rs-',label='supercompact, no rcm', ms=10, lw=2)
+plt.plot(sx,s3,'r^-',label='supercompact, rcm', ms=10, lw=2)
+plt.plot(rx,r2,'bs-',label='random, no rcm', ms=10, lw=2)
+plt.plot(rx,r3,'b^-',label='random, rcm', ms=10, lw=2)
+plt.ylim([50,220])
+plt.ylabel('Performance (Gflop/s)', fontsize=18)
+plt.xlabel('number rows / 1000', fontsize=18)
+plt.legend(loc=7, fontsize=18)  # upper left
+plt.rc('xtick', labelsize=16)
+plt.rc('ytick', labelsize=16)
+myutil.horizontal_lines()
+plt.savefig.format = "pdf"
+plt.savefig("random_supercompact.pdf")
+#-------------------------
+"""
+plt.clf()
+plt.plot(cx,c2,'rs-',label='compact, no rcm', ms=10, lw=2)
[…]

### 2013-10-19 figure_files/figure11/m.py

@@ -1,0 +1,54 @@
+#!/usr/bin/python
+# Run on the mic, by launching from the host.
+
+import os
+
+#-------------------------------------------------------
+# not clear the threads command is working
+THREADS= "export OMP_NUM_THREADS=244"
+SCHEDULE= "export OMP_SCHEDULE=static,64"
+KMC= "export KMP_AFFINITY=granularity=fine,scatter"
+KMC= "export KMP_AFFINITY=scatter"
+KMC= "export KMP_AFFINITY=granularity=fine,compact"
+
+SCHEDULE= "export OMP_SCHEDULE=dynamic,64"
+KMC= "export KMP_AFFINITY=compact"
+LIB = "export LD_LIBRARY_PATH=/opt/intel/composer_xe_2013/lib/mic/"
+F="matrix/ell_ell_sym_1_x_weights_direct__no_hv_stsize_32_3d_8x_8y_8z.bmtx"
+F= "matrix/ell_ell_x_weights_direct__no_hv_stsize_32_3d_64x_64y_64z.bmtx"
+F="matrix/ell_kd-tree_x_weights_direct__no_hv_stsize_32_3d_128x_128y_128z.bmtx"
+F="matrix/ell_kd-tree_x_weights_direct__no_hv_stsize_32_3d_192x_192y_192z.bmtx"
+F="matrix/ell_kd-tree_rcm_sym_1_x_weights_direct__no_hv_stsize_32_3d_192x_192y_192z.bmtx"
+F="matrix/ell_kd-tree_rcm_sym_1_x_weights_direct__no_hv_stsize_32_3d_128x_128y_128z.bmtx"
+F="matrix/ell_x_weights_direct__no_hv_stsize_32_3d_64x_64y_64z.bmtx"
+F="matrix/ell_kd-tree_rcm_sym_1_x_weights_direct__no_hv_stsize_32_3d_96x_96y_96z.bmtx"
+F="matrix/ell_kd-tree_x_weights_direct__no_hv_stsize_32_3d_96x_96y_96z.bmtx"
+F="matrix/ell_kd-tree_rcm_sym_1_x_weights_direct__no_hv_stsize_32_3d_64x_64y_64z.bmtx"
+F="matrix/ell_kd-tree_x_weights_direct__no_hv_stsize_32_3d_64x_64y_64z.bmtx"
+
+
+EXEC= "./linux/release/spmv_openmp_mic %s 1" % F
[…]

### 2013-10-19 figure_files/figure11/performance.py

@@ -1,0 +1,77 @@
+
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+import csv
+
+#----------------------------------
+
+# 3 columns: filenmame, rows, bandwidth
+def plot_file(infile,outfile):
+    dict = {}
+    dict['64_norcm'] = {}
+    dict['96_norcm'] = {}
+    dict['64_rcm'] = {}
+    dict['96_rcm'] = {}
+    with open(infile, 'r') as csvfile:
+        csvfile.seek(0)
+        reader = csv.reader(csvfile, delimiter=' ')
+        for row in reader:
+            if row == []: break
+            if row[0][0] == '#': continue
+            rfile = row[0]
+            method = row[1]
+            r1 = float(row[2])
+            r2 = float(row[3])
+
+            try:
+                dict[rfile][method].append([r1,r2])
+            except:
+                print(rfile)
+                dict[rfile][method] = []
+                dict[rfile][method].append([r1,r2])
+
+    for f in dict.keys():
+        for m in dict[f].keys():
+            dict[f][m] = zip(*dict[f][m])  # * means transpose
+
+
+#### BEGINNING OF CHART
+
+    def plot_file(xlist, ylist, col, sym, leg):
+        lines = ['-','-','-','-']
+        count = 0
+        x = xlist
+        y = ylist
+        p = plt.plot(x,y, col+sym+'-', label=leg)
+        lines.append(p)
+        count += 1
+
+
+    cols = ['b','r','m','c','k','g']
+    syms = ['o','^','s','v','h','x']
+    fi = 0
+    for f in dict.keys():
+        col = cols[fi]; fi+=1; mi = 0
+        for m in dict[f].keys():
+            sym = syms[mi]; mi+=1
+            leg = f + "_" + m
+            plot_file(dict[f][m][0], dict[f][m][1], col, sym, leg)
+
+    plt.xlabel('Nb threads')
+    plt.ylabel('Gflops')
+    plt.title('Performance RBF-FD derivative MIC, 96^3')
+    plt.legend(loc=2, fontsize=10)  # upper left
+    plt.grid(True)
+    #plt.ylim(0,100)
+
+    plt.tight_layout()
+    plt.savefig.format = "pdf"
+    plt.savefig(outfile)
+#plt.show()
+#### END OF CHART
+#----------------------------------------------------------------------
+
+filename = "raw2"
+plot_file(filename+".dat",filename+".pdf")
+

### 2013-10-19 figure_files/figure11/performance_96cube.py

@@ -1,0 +1,77 @@
+
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+import csv
+
+#----------------------------------
+
+# 3 columns: filenmame, rows, bandwidth
+def plot_file(infile,outfile):
+    dict = {}
+    dict['64_norcm'] = {}
+    dict['96_norcm'] = {}
+    dict['64_rcm'] = {}
+    dict['96_rcm'] = {}
+    with open(infile, 'r') as csvfile:
+        csvfile.seek(0)
+        reader = csv.reader(csvfile, delimiter=' ')
+        for row in reader:
+            if row == []: break
+            if row[0][0] == '#': continue
+            rfile = row[0]
+            method = row[1]
+            r1 = float(row[2])
+            r2 = float(row[3])
+
+            try:
+                dict[rfile][method].append([r1,r2])
+            except:
+                print(rfile)
+                dict[rfile][method] = []
+                dict[rfile][method].append([r1,r2])
+
+    for f in dict.keys():
+        for m in dict[f].keys():
+            dict[f][m] = zip(*dict[f][m])  # * means transpose
+
+
+#### BEGINNING OF CHART
+
+    def plot_file(xlist, ylist, col, sym, leg):
+        lines = ['-','-','-','-']
+        count = 0
+        x = xlist
+        y = ylist
+        p = plt.plot(x,y, col+sym+'-', label=leg)
+        lines.append(p)
+        count += 1
+
+
+    cols = ['b','r','m','c','k','g']
+    syms = ['o','^','s','v','h','x']
+    fi = 0
+    for f in dict.keys():
+        col = cols[fi]; fi+=1; mi = 0
+        for m in dict[f].keys():
+            sym = syms[mi]; mi+=1
+            leg = f + "_" + m
+            plot_file(dict[f][m][0], dict[f][m][1], col, sym, leg)
+
+    plt.xlabel('Nb threads')
+    plt.ylabel('Gflops')
+    #plt.title('Performance RBF-FD derivative MIC, 96^3')
+    plt.legend(loc=2, fontsize=10)  # upper left
+    plt.grid(True)
+    #plt.ylim(0,100)
+
+    plt.tight_layout()
+    plt.savefig.format = "pdf"
+    plt.savefig(outfile)
+#plt.show()
+#### END OF CHART
+#----------------------------------------------------------------------
+
+filename = "raw2"
+plot_file(filename+".dat",filename+".pdf")
+

### 2013-10-19 figure_files/figure11/plot_for_natasha_ncar_end_of_year.py

@@ -1,0 +1,82 @@
+
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+import csv
+
+#----------------------------------
+
+# 3 columns: filenmame, rows, bandwidth
+def plot_file(infile,outfile):
+    dict = {}
+    #dict['64_norcm'] = {}
+    dict['96_norcm'] = {}
+    #dict['64_rcm'] = {}
+    dict['96_rcm'] = {}
+    with open(infile, 'r') as csvfile:
+        csvfile.seek(0)
+        reader = csv.reader(csvfile, delimiter=' ')
+        for row in reader:
+            if row == []: break
+            if row[0][0] == '#': continue
+            rfile = row[0]
+            method = row[1]
+            print method
+            r1 = float(row[2])
+            r2 = float(row[3])
+
+            try:
+                if method != "method_8a_multi_cpp":
+                    dict[rfile][method].append([r1,r2])
+            except:
+                try:
+                    print(rfile)
+                    dict[rfile][method] = []
+                    dict[rfile][method].append([r1,r2])
+                except:
+                    pass
+
+    for f in dict.keys():
+        for m in dict[f].keys():
+            dict[f][m] = zip(*dict[f][m])  # * means transpose
+
+
+#### BEGINNING OF CHART
+
+    def plot_file(xlist, ylist, col, sym, leg):
+        lines = ['-','-','-','-']
+        count = 0
+        x = xlist
+        y = ylist
+        p = plt.plot(x,y, col+sym+'-', label=leg)
+        lines.append(p)
+        count += 1
+
+
+    cols = ['b','r','m','c','k','g']
+    syms = ['o','^','s','v','h','x']
+    fi = 0
+    for f in dict.keys():
+        col = cols[fi]; fi+=1; mi = 0
+        for m in dict[f].keys():
+            sym = syms[mi]; mi+=1
+            leg = f + "_" + m
+            plot_file(dict[f][m][0], dict[f][m][1], col, sym, leg)
+
+    plt.xlabel('Nb threads')
+    plt.ylabel('Performance (Gflop/s)')
+    plt.title('Performance RBF-FD derivative MIC, 96^3')
+    plt.legend(loc=2, fontsize=14)  # upper left
+    plt.grid(True)
+    #plt.ylim(0,100)
+
+    plt.tight_layout()
+    plt.savefig.format = "pdf"
+    plt.savefig(outfile)
+#plt.show()
+#### END OF CHART
[…]

### 2013-10-19 figure_files/figure4/matrix_stencil_32.py

@@ -1,0 +1,185 @@
+"""
+Show how to modify the coordinate formatter to report the image "z"
+value of the nearest pixel given x and y
+"""
+
+#Docs example: http://matplotlib.org/examples/pylab_examples/multi_image.html
+
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+from matplotlib.cm import colors
+
+#----------------------------------------------------------------------
+def byte_to_float_ratio(nv, nm, bx, nz):
+# no consideration of cache.
+    bi = 4
+    #print nv,nm,nz
+    #rworst = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
+    # in worst case, every vector element brings in an entire cache line (256 bits = 32 floats),
+    # so nv=4 vectors (1/8th cache line), requires a transfer of 32 bytes = 8 floats
+    # 16 vectors = 2 cache lines (, so the formula becomes
+    # Assume infinite cache size, and cache line size = 1 float = 4 bytes. So there is no bandwidth waste
+    cl = 4 # bytes
+    flops = 2.*nv*nm
+    # count in flops
+    x_read = bx*nv
+    a_read = bx*nm
+    icol_read = bi
+    y_write = bx*nm*nv/float(nz)
+    ratio = (flops)/(x_read+a_read+icol_read+y_write)
+    #ratio = (2.*nv*nm)/(bx*nv+bx*nm+bx*nm*nv/float(nz))+bi)
+    # assumes cl=1
+    cl = 64.  # size of most caches in bytes
+    x_read = cl*np.ceil((bx*nv)/float(cl))  # assumes small cache, x must be reread
+    rworst = flops/(x_read+a_read+icol_read+y_write)
+    #rworst = (2.*nv*nm)/((cl*np.ceil((bx*nv)/cl)+nm+nm*nv/float(nz))+bi)
+    x_read = bx*nv/float(nz)   # each element of x is read once
+    rbest = flops/(x_read+a_read+icol_read+y_write)
+    print "xx: ", [nv,nm,rworst,rbest]
+    return(ratio)
+#----------------------------------------------------------------------
+def ratio_cache(nv, nm, bx, nz):
+# byte to float taking cache effect into account
+# per row:
[…]

### 2013-10-19 figure_files/figure5/gflops_speedup_figure5.py

@@ -1,0 +1,184 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+
+#----------------------------------
+def byte_to_float_ratio(nv, nm, bx, nz):
+# no consideration of cache.
+    bi = 4
+    #print nv,nm,nz
+    #rworst = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
+    # in worst case, every vector element brings in an entire cache line (256 bits = 32 floats),
+    # so nv=4 vectors (1/8th cache line), requires a transfer of 32 bytes = 8 floats
+    # 16 vectors = 2 cache lines (, so the formula becomes
+    ratio = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
+    return(ratio)
+#----------------------------------
+#----------------------------------
+def ratio(nv, nm, bx, nz):
+# per row:
+# flops divided by number bytes sent/received
+# nv : number vectors
+# nm : number matrices
+# bx = 4 or 8 (single, double precision)
+# nz = nb nonzeros per row
+# y = A*x
+# y : nv*nm*bx   writes
+# x : nv*bx      loads
+# A : nm*nz*bx   loads
+# col_id : nz*bi loads
+    bi = 4
+
+
+    cl = 64  # size of most caches
+
+    # for each row of A
+
+    flops = 2.*nv*nm*nz
+
+    a_read = bx*nm*nz
+    icol_read = bi*nz
+    y_write = bx*nm*nv
+
+    x_read = nz *  cl*np.ceil((bx*nv)/float(cl))  # assumes small cache, x must be reread
+    rworst = flops/(x_read+a_read+icol_read+y_write)
+
+    x_read = bx*nv   # each element of x is read once
+    rbest = flops/(x_read+a_read+icol_read+y_write)
+    #print "yy: ", [nv,nm,rworst,rbest]
+    return([rworst,rbest])
+
+
+"""
+    cl = 64.
+    #rworst = (2.*nv*nm)/(bx*(nv+nm+nm*nv/float(nz))+bi)
+    # in worst case, every vector element brings in an entire cache line (256 bits = 32 floats),
+    # so nv=4 vectors (1/8th cache line), requires a transfer of 32 bytes = 8 floats
[…]

### 2013-10-19 figure_files/figure7-8/bandwidth_tests.cpp

@@ -1,0 +1,521 @@
+#include "bandwidth_tests.h"
+#include "timer_eb.h"
+#include "util.h"
+#include <vector>
+#include <algorithm>
+
+namespace spmv {
+
+void MemoryBandwidth::evictFromCache()
+{
+// evict data from cache
+#pragma omp parallel
+{
+    evict_array_from_cache(&vec_vt[0], sizeof(float)*nb_rows);
+    evict_array_from_cache(&result_vt[0], sizeof(float)*nb_rows);
+    evict_array_from_cache(&col_id_t[0], sizeof(int)*nb_rows);
+}
+}
+//----------------------------------------------------------------------
+MemoryBandwidth::MemoryBandwidth()
+{
+    max_gflops = 0.;
+    max_bandwidth = 0.;
+    elapsed = 0.; 
+    min_elapsed = 0.; 
+
+    int offset = 3; // ignore first offset measurements
+    tm["spmv"] = new EB::Timer("memory benchmarks", offset);
+    tm["spmv1"] = new EB::Timer("readWriteAlone", offset);
+
+    pj = new ProjectSettings("./test.conf");
+ 
+	// Parse file created by python script
+	// Uncomment if not using a script
+	bool use_script = REQUIRED<bool>("USE_PYTHON_SCRIPT");
+    printf("use_script: %d\n", (int) use_script);
+
+	if (use_script) {
+    	pj->ParseFile("bench.conf");
+	}
+
+    //pj = new ProjectSettings("/mnt/global/LCSE/gerlebacher/mic/test.conf");
+    experiment_s  = REQUIRED<std::string>("bandwidth_experiment");
+    col_id_type_s = REQUIRED<std::string>("col_id_type");
+    std::string rr = REQUIRED<std::string>("nb_rows");
+    sscanf(rr.c_str(), "%d", &this->nb_rows);
+    printf("constructor: nb_rows= %d\n", this->nb_rows);
+    delete pj;
+}
+//----------------------------------------------------------------------
+void MemoryBandwidth::free()
+{
+    _mm_free(vec_vt);
+    _mm_free(result_vt);
+    _mm_free(col_id_t);
+}
+//----------------------------------------------------------------------
+void MemoryBandwidth::initialize()
+{
+    printf("initialize, entered\n");
+    printf("initialize: nb rows: %d\n", nb_rows);
[…]

### 2013-10-19 figure_files/figure7-8/figure.py

@@ -1,0 +1,114 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+import csv
+
+def parse(infile):
+    rows = []
+    nb_rows = []
+    all_nb_rows = []
+    all_bw = []
+    with open(infile, 'r') as csvfile:
+        csvfile.seek(0)
+        reader = csv.reader(csvfile, delimiter=' ')
+        for i,row in enumerate(reader):
+# if bandwidth in row, continue
+            try:
+                if row[5] == 'bandwidth=':
+                    rows.append(row)
+                else:
+                    continue
+            except:
+                pass
+    #print rows[0]
+    #print rows[1]
+    #print rows[2]
+
+    for r in rows:
+        #print r
+        #print r[1], r[6]
+        bandwidth = float(r[6])
+        nb_rows = int(r[1])
+        #print r
+        print bandwidth, nb_rows
+        all_nb_rows.append(nb_rows)
+        all_bw.append(bandwidth)
+
+    #print rows
+    #print bw
+    return [all_nb_rows, all_bw]
+
+#----------------------------------------------------------------------
+plt.clf()
+[rows, bw] = parse("bench=read_cpp_col=compact_rows=00.out")
+p1, = plt.plot(rows, bw, '-b^', label='read, cpp', ms=6, lw=2)
+
+[rows, bw] = parse("bench=write_col=compact_rows=00.out")
+p2, = plt.plot(rows, bw, '-rs', label='write', ms=6, lw=2)
+
+[rows, bw] = parse("bench=read_col=compact_rows=00.out")
+p3, = plt.plot(rows, bw, '-bs', label='read', ms=6, lw=2)
+
+[rows, bw] = parse("bench=read_write_col=compact_rows=00.out")
+p4, = plt.plot(rows, bw, '-gs', label='read/write', ms=6, lw=2)
+
+[rows, bw] = parse("bench=read_write_cpp_col=compact_rows=00.out")
+p5, = plt.plot(rows, bw, '-g^', label='read/write, cpp', ms=6, lw=2)
+
+[rows, bw] = parse("bench=write_cpp_col=compact_rows=00.out")
[…]

### 2013-10-19 figure_files/figure7-8/myutil.py

@@ -1,0 +1,25 @@
+
+import matplotlib.pyplot as plt
+
+#-----------------------------------------------
+def parse_line(line):
+    in_file,method = line[0].split(':')
+    rcm = len(in_file.split('rcm'))
+    rcm = rcm-1    # 0 if no RCM, 1 if rcm
+    threads = line[1].split(',')[0].split(':')[1]
+    max_flops = line[2].split(',')[0].split(':')[1]
+    numbs = in_file.split('3d_')[1].split('_')
+    nx = int( numbs[0].split('x')[0] )
+    ny = int( numbs[1].split('y')[0] )
+    nz = int( numbs[2].split('z')[0].split('z')[0] )
+    return [in_file,max_flops,nx,ny,nz,rcm]
+#-----------------------------------------------
+def horizontal_lines():
+    # worst and best cases single precision
+    horizontal_line(55.8, 0, 1600)
+    horizontal_line(213.3, 0, 1600)
+
+def horizontal_line(ylev, xmin=0, xmax=1):
+    x = [xmin, xmax]
+    y = [ylev, ylev]
+    plt.plot(x, y, 'k-',lw=2)

### 2013-10-19 figure_files/figure9/sparse.py

@@ -1,0 +1,51 @@
+import csv
+import numpy as np
+import random as rnd
+import scipy.sparse as sps
+import matplotlib.pyplot as plt
+#plt.spy(a, marker='.',markersize=2)
+#plt.show()
+
+
+def plot_sparse(mat, filename):
+    plt.spy(mat, markersize=1)
+    plt.tight_layout()
+    #plt.rc('xtick',labelsize=16)
+    #plt.rc('ytick',labelsize=16)
+    plt.savefig.format = "pdf"
+    plt.savefig(filename)
+    plt.clf()
+
+#--------------------------------
+nrows = 512
+#--------------------------------
+# compact matrix.
+mat = sps.csr_matrix((nrows, nrows))
+ones = np.ones(2*nrows)
+mat.setdiag(ones, k=0)
+for i in range(1,16):
+    mat.setdiag(ones, k=i)
+    mat.setdiag(ones, k=-i)
+plot_sparse(mat, "compact_matrix.pdf")
+#--------------------------------
+# supercompact matrix
+ones = np.ones((nrows,32))
+xones = sps.csr_matrix(ones)
+xsparse = sps.csr_matrix((nrows,nrows-32))
+mat1 = sps.hstack([xones,xsparse])
+plot_sparse(mat1, "supercompact_matrix.pdf")
+#--------------------------------
+# random matrix
+rg = range(nrows)
+col_id = np.zeros((nrows,32),int)
+for i in range(nrows):
+    a = rnd.sample(rg, 32)
+    col_id[i,:] = np.sort(a)
+
+col_id = np.ravel(col_id)
+values = np.ravel(np.ones((nrows,32),float))
+ptrs = range(0,32*(nrows+1),32)
+mat2 = sps.csr_matrix((values,col_id,ptrs),shape=(nrows,nrows))
+plot_sparse(mat2, "random_matrix.pdf")
+#--------------------------------
+#--------------------------------

### 2014-01-14 acm/RBF_Section.tex
Paper for ACM.

%\documentclass{article}

%\usepackage{amsmath, amssymb, graphicx, multirow}

%\begin{document}

%\section{Introduction}

\section{Introduction and Motivation for RBF-FD}
\label{sec:rbf}

RBFs approximate a function $f(\mathbf{x}) \in \mathbb{R}^d$ sampled at $N$ distinct node locations by linearly combining translates of a single radially symmetric function, e.g. $\phi(r) = e^{-\varepsilon^2r^2}$, where $r=\|\mathbf{x}-\mathbf{x}_{i}\|$ denotes the Euclidean distance between where the function $f(\mathbf{x})$ is evaluated, $\mathbf{x}$, and where the RBF, $\phi$, is centered $\mathbf{x}_{i}$. The parameter $\varepsilon$ controls the shape of the RBF. Note that the argument of an RBF is simply a scalar distance $r$, independent of any coordinate system or dimension. As a result, nodes can be scattered as desired across complex physical domains with the implementation of the method being independent of dimensionality. Thus, no mesh generation is needed and algorithmic complexity does not increase with dimension.

To obtain a RBF derivative approximation at a node location that will result in a sparse differentiation matrix (DM), the RBF-FD method has been developed over the last decade \cite{TAI1,TAI2,SDY02,WrFo06}. RBF-FD are conceptually similar to standard finite differences (FD). However, the differentiation weights for approximating a derivative at a given node (forming one row of the RBF-FD DM) enforce that the linear combination of function values at the $n$ nearest node locations (a stencil) with $n \in N$ be exact for RBFs centered at each of the $n$ nodes rather than polynomials. The result is a sparse matrix with $nN$ nonzero entries, where $n<<N$. As in FD methods, the order of the method increases with stencil size. An illustration of an RBF stencil on a sphere is given in Figure \ref{fig:stencil}.

\begin{figure}%[htb]
\centering
\includegraphics[width=0.5\textwidth]{figures/RBFStencil}\label{fig:stencil}
\caption{An example of an RBF-FD stencil of size $n=21$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square. In nonsparse format, the DM  is $1024\times1024$ with 21 nonzero entries per row.}
\end{figure}

Due to the above attributes and its simplicity of implementation, the RBF-FD is gaining popularity in modeling systems of PDEs in a variety of fields in fluid dynamics (e.g., geosciences, combustion, aerodynamics \cite{Bayona13,CDNT,FoL11,FLBWSC12,SPLM}). For example, the full shallow water equations in a 3D Cartesian coordinate system for a rotating fluid are as follows:
\begin{eqnarray}
\dfrac{\partial u}{\partial t} &=& - \left( u\dfrac{\partial u}{\partial x} + v\dfrac{\partial u}{\partial y} + w\dfrac{\partial u}{\partial z} + f(yw - zv) + g\dfrac{\partial h}{\partial x} \right)
\\
\dfrac{\partial v}{\partial t} &=& - \left( u\dfrac{\partial v}{\partial x} + v\dfrac{\partial v}{\partial y} + w\dfrac{\partial v}{\partial z} + f(zu - xw) + g\dfrac{\partial h}{\partial x} \right)
\\
\dfrac{\partial w}{\partial t} &=& - \left( u\dfrac{\partial w}{\partial x} + v\dfrac{\partial w}{\partial y} + w\dfrac{\partial w}{\partial z} + f(xv - yu) + g\dfrac{\partial h}{\partial x} \right)
\\
\frac{\partial h}{\partial t}& =&-\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right) , \label{height}
\end{eqnarray}
where $f$ is the Coriolis force, $\{u,v,w\}$ are the components of the velocity vector in the respective $\{x,y,z\}$ directions, and $h$ is the geopotential height (analogous to pressure). Thus, we have four variables $\{u,v,w,h\}$ and three DMs, $D_x, D_y, D_z$ representing $\dfrac{\partial}{\partial x}, \dfrac{\partial}{\partial y},\dfrac{\partial w}{\partial z}$, respectively. However, in order to stably time step the equations with an explicit time-stepping scheme, most numerical methods (including RBF-FD) require a fourth operator, hyperviscosity, to be added to the each equation. Hyperviscosity takes the form of $\triangle^p$ ($\triangle$ denotes the Laplacian operator), where $p\geq3$, which is approximated by the matrix $D_{hyp}$. Therefore, in order to solve this system of PDEs, the application of 4 DM to 4 unknowns, i.e. 16 SpMV is required every time step. Due to the fact that RBFs are independent of a coordinate system, taking into account only distances to nearest neighbors, the DMs have the following properties:
\begin{itemize}
\item For a given $n$, regardless of what operator is approximated, e.g. $\dfrac{\partial}{\partial x}$ or the Laplacian, RBF-FD DMs will have the same number of nonzeros per row.
\item Also, for a given $n$ and $N$, regardless of what operator is approximated, the DMs will have the same sparsity pattern. Examples of an RBF-FD DM can be seen in Figure \ref{fig:spy_plots}d and e.
\end{itemize}

Multiple variables transform a SpMV into a SpMM (Sparse Matrix/dense Matrix multiplication), which improves register utilization and decreases cache
misses by vectorizing over the multiple source vectors. Furthermore, due to the properties just mentioned, multiple derivatives of a single function can be calculated rather than computing a derivative of multiple functions. Using the PDE system above as an example, the block structure of the SpMM for computing all derivatives needed can be seen in Figure \ref{fig:struct_comp}.
\begin{figure}
  \centering
  \[ \left( \begin{array}{cccc}
    \underline{u}_x     & \underline{v}_x     & \underline{w}_x     & \underline{h}_x \\
    \underline{u}_y     & \underline{v}_y     & \underline{w}_y     & \underline{h}_y \\
    \underline{u}_z     & \underline{v}_z     & \underline{w}_z     & \underline{h}_z \\
    \underline{u}_{hyp} & \underline{v}_{hyp} & \underline{w}_{hyp} & \underline{h}_{hyp}
  \end{array} \right)
  = \left(
  \begin{array}{c}
    D_{x}\\
    D_{y}\\
    D_{z}\\
    D_{hyp}
  \end{array}\right)
  \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) \]
  \caption{Block structure of the computation.}
  \label{fig:struct_comp}
\end{figure}
where $\{\underline{u},\underline{v}, \underline{w}, \underline{h}\}$ are the functions values at all nodes of the corresponding variables with the left hand side being the resulting derivative approximations. The increased memory bandwidth due to an increase in the number of derivative matrices is offset by better cache utilization, leading to an overall performance benefit. In practice however, the number of vectors whose derivatives are required is limited, capping the performance of the matrix/vector multiplication. Similarly, the number of different derivatives in a particular computation is finite. It is for this reason that we seek to combine the benefits of multiple vectors and multiple matrices simultaneously. We limit this initial study to four vectors and four matrices, which is a common number found in fluid dynamic calculations.

%\end{document}

### 2014-01-14 acm/aa.tex
Paper for ACM.

\begin{figure}
\begin{center}
	\centering
	\begin{subfigure}[t]{0.75in}
	\centering
	\includegraphics[width=1.0\textwidth]{../figures/appendices/direct_vs_indirect_weights/compare_weight_generation/xsfc_vs_xsfc_alt_on_sph32_times_sine_20x/sph32_times_sin20x-eps-converted-to.pdf}
	\caption{Test function:  \\ $Y_{3}^{2} \sin(20 x) $.  }
		\label{fig:direct_vs_indirect_manufactured_base}
	\end{subfigure}

	\begin{subfigure}[h]{0.75in}
		\centering
	\includegraphics[width=1.0\textwidth]{../figures/appendices/direct_vs_indirect_weights/compare_weight_generation/xsfc_vs_xsfc_alt_on_sph32_times_sine_20x/pdx_sph32_times_sin20x-eps-converted-to.pdf}
	\caption{$\mathbf{p}_{x} \cdot \nabla ( Y_{3}^{2} \sin(20 x))$ }
			\label{fig:direct_vs_indirect_manufactured_xsfc}
	\end{subfigure}
	\begin{subfigure}[h]{.75in}
		\centering
	\includegraphics[width=1.0\textwidth]{../figures/appendices/direct_vs_indirect_weights/compare_weight_generation/lsfc_vs_px_grad_dot_px_grad/lsfc_sph32_times_sin20x-eps-converted-to.pdf}
	\caption{Surface Laplacian: \\ $\LaplaceBeltrami ( Y_{3}^{2} \sin(20 x) )$  }
				\label{fig:direct_vs_indirect_manufactured_lsfc}
	\end{subfigure}
	\caption{Test function and its projected derivatives on the surface of the unit sphere. }
				\label{fig:direct_vs_indirect_manufactured_solution}
	\end{center}
\end{figure}

### 2014-01-14 acm/analysis.tex
Paper for ACM.

\section{Analysis}
In this section, we perform some simple analysis.

Under the best scenarios, the peak floating point performance is 2 Tflops and the peak measured bandwidth, (read: 
!80 GB/sec, write: 160 GB/sec. We seek to answer the question as to whether our algorithm is limited by the bandwidth or the floating point j
performance.

Under the best of circumstances, read occurs in $n_{rT}10^{-6}/180$ ms, writes occur in $n_{wT}10^{-6}/160$ sec. 
The calculation takes $n_F 10^{-9}/2$ ms. Many more bytes are read in than are written. Thus, we ignore the difference between the read and write 
performance.

Consider stencils of size 32, 4 vectors and 4 matrices,

The ratio of times taken by floating point operations to memory transfers is
$$
R = \frac{n_F}{n_{wT}+n_{rT}} = \frac{8}{b_x+1} \frac{10^{-6}}{160} \; \frac{2}{10^{-9}}
$$
$$
R = \frac{8}{b_x+1}  \; 10^3 80
$$

Assume $N=64$, stencils of size 32. Thus there are $250000*2*32\times 16=256 \times 10^6$ operations. At a speed of $2\times 10^{12}$ operations per sec, 
the calculation should take a minimum of $128\times 10^{-6}$ sec = 0.128 ms. Reading/writing all the data requires a transfer of 
$250000\times (32 \times 4 + 4 + 4) = 34\times 10^6$ bytes, which takes $\frac{34\times 10^6}{200\times 10^9}=0.17$ ms. 
Thus, calculation and computation take on the same order. 
Rato of calculation to transfer time is $.128/0.17=0.75 = 1/1.33$.

Now, consider the 1 vector and 1 matrix case. There are $16\times 10^6$ flops taking $8\times 10^{-6}$ sec, and a transfer of 
$250000\times (32+1+1) =8.5\times 10^6$ bytes, which takes $\frac{8.5\times 10^6}{200\times 10^9}=4.25=0.04$ ms. 
Ratio of calculation to transfer time is $.008/.04=0.2=1/5$.

Conclusion: a gain of a factor of 3.7 flops per byte, approximately 4.

If instead, one had 32 vectors, the gain would be much higher. I should implement a BENCHMARK to calculate a matrix/matrix multiplication, 
and execute it 10 times, AND make sure all elements stay in cache, to measure an idealized flop rate.

Consider A * (v,v,v,v,v,....) . Measure the performance as a function of the number of cores. The Gflop rate will measure optimum performance. 
In reality, one must take into account the cost of loop index calculations, loop indexing, etc. The hope of course, is that this cost is 
insignificant. It can be measured via loop unrolling.

TODO: need practical Gflop rate.
Use supercompact. Given 64^3 grid, as #threads increases, available cache increases, so performance shoudl go up. With low number of processors, 
I should not have enough cache (at least if using random). Using supercompact, and in addition, eliinating gather and load_epi32 so I am executing 
summation of $v_i^2$, I should get good measurement of GFlops rate. Cache will be sufficient. All I want is to avoid loads.

\subsection{measure theoretical speedup}
\subsection{measure measured speedup}
\begin{itemize}
\item register density effect: would affect flop rate
\end{itemize}

### 2014-01-14 acm/code.tex
Paper for ACM.

--- 2013-10-18 code.tex
@@
  for (int n=0; n < nz; n+=4) {
-   float* a = &dom.col_id_t[0]+n+nz*r;
+   float* a = &dom.col_id[0]+n+nz*r;
    __m512i vect_i = _mm512_setzero_ps();
@@
 
-   mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));
+   mat_ent = _mm512_load_ps(dom.data + nb_mat*(n + r*nz));
 

### 2014-01-14 acm/ieee_paper.tex
Paper for ACM.

--- 2013-10-19 ieee_paper.tex
@@
-\documentclass[10pt,conference,compsocconf]{IEEEtran}
+%\documentclass[10pt,conference,compsocconf]{IEEEtran}
+%\documentclass{acm_proc_article-sp}
+\documentclass{sig-alternate}
 \input{support/macros} % color is defined in macros or misc_mac
@@
 
-\usepackage[pdftex]{graphicx}
+\usepackage{graphicx}
+%%\usepackage[pdftex]{graphicx}
 \usepackage{subfigure}
 %\usepackage{fixltx2e}
-\usepackage{url}
-\hyphenation{op-tical net-works semi-conduc-tor}
+%%\usepackage{url}
+%%\hyphenation{op-tical net-works semi-conduc-tor}
 
@@
 
-\author{\IEEEauthorblockN{Gordon Erlebacher\IEEEauthorrefmark{1},
-Erik Saule\IEEEauthorrefmark{2}, Natasha Flyer\IEEEauthorrefmark{3}, 
-and Evan Bollig\IEEEauthorrefmark{1}}
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Scientific Computing, 
-Florida State University}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR}
-Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com}
+\subtitle{[Extended Abstract]}
+
+%\numberofauthors{4}
+
+%\author{
+%\alignauthor
+%Gordon Erlebacher\\ %\titlenote{Dr.~Trovato insisted his name be first.}\\
+%       \affaddr{Department of Scientific Computing}\\
+%       \affaddr{Florida State University}\\
+%       \affaddr{Tallahassee, FL 32306-4120}\\
+%       \email{gordon.erlebach@gmail.com}
+%\alignauthor
+%Erik Saule
+%        \affaddr{Department of Computer Science}\\
+%        \affaddr{University of North Carolina at Charlotte}\\
+%        \email{esaule@uncc.edu}
+%\and
+%\alignauthor
+%Natasha Flyer
+%        \affaddr{Computational and Information System Laboratory}\\
+%        \affaddr{UCAR}\\
+%        \email{flyer@ucar.edu}
+%\alignauthor
+%Evan Bollig
+%        \affaddr{Minnesota Supercomputer Institute}\\
+%        \affaddr{University of Minnesota}\\
+%        \affaddr{Minneapolis, MN}\\
+%        \email{bollig@gmail.com}
+%}
 \maketitle
+
+%\author{\IEEEauthorblockN{Gordon Erlebacher\IEEEauthorrefmark{1},
+%Erik Saule\IEEEauthorrefmark{2}, Natasha Flyer\IEEEauthorrefmark{3}, 
+%and Evan Bollig\IEEEauthorrefmark{1}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Scientific Computing, 
+%Florida State University}
+%\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte}
+%\IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR}
+%Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com}
+%\maketitle
 
@@
 
-\begin{IEEEkeywords}
-MIC; SIMD; SpMV; SpMM; Sparse Matrix; Radial Basis Function;
-\end{IEEEkeywords}
+% A category with the (minimum) three required fields
+\category{H.4}{Information Systems Applications}{Miscellaneous}
+%A category including the fourth, optional field follows...
+\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
 
-\IEEEpeerreviewmaketitle
+\terms{Theory}
+
+\keywords{MIC, SIMD, SpMV, Sparse Matrix, Radial Basis Function}
+
+%\begin{IEEEkeywords}
+%MIC; SIMD; SpMV; SpMM; Sparse Matrix; Radial Basis Function;
+%\end{IEEEkeywords}
+
+%\IEEEpeerreviewmaketitle
 
@@
 
-\section{Derivatives of Radial Basis Functions}
-\label{sec:rbf}
-
-In this paper, we propose a new idea that is applicable to radial
-basis functions (RBFs). Their strength is the ability to randomly
-distribute points across complex physical domains, and have an
-implementation that is independent of dimensionality. RBFs approximate
-a function $f(\rvec)\subset \mathbb{R}^d$ sampled at a set of $N$
-distinct point locations, $x_j$, by linearly combining translates of a
-single radially symmetric function $\phi(r)$, where $r =
-\|\rvec-\rvec_{j}\|$ denotes the Euclidean distance %(e.g., in 2-D
-%$\sqrt{(x-x_j)^2+(y-y_j)^2}$) 
-between $\rvec$, where the function is evaluated
-and and $\rvec_j$,  where the RBF is centered. That is, the
-interpolant is $s(\rvec) = \sum_{j=1}^{N} w_j
-\phi_i(\|\rvec-\rvec_{j}\|)$. The weights $w_j$ are obtained by
-inverting the system
-\begin{equation}
-\parray{lccr}{
-\phi_{11} & \phi_{12} & \cdots & \phi_{1N} \\
-\vdots & \ddots & \vdots & \vdots \\
-\phi_{n1} & \phi_{n2} & \cdots & \phi_{NN} 
-}
-\parray{c}{ w_{1} \\ \vdots \\ w_{N} }
-=
-\parray{c}{ f(\rvec_1) \\ \vdots\\ f(\rvec_N) }. 
-\label{eq:rbf}
-\end{equation}
-where $\phi_{ij} = \phi(\|\rvec_i-\rvec_j\|)$.
-
-The RBF differentiation matrix, $D_N$, is derived by applying the
-desired analytic derivative operator $L$ to the RBF interpolant
-$s(\rvec)$ above and evaluating it at the point locations. For very
-large problems, this is a computationally expensive since the matrix
-in (\ref{eq:rbf}) is full and inversion requires O$(N^3)$
-operations. To alleviate the cost of this global approach (i.e., using
-every node in the domain to calculate the derivative at a given node
-$\rvec_i$), RBF-generated finite differences (RBFFD) have been
-derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBFFD uses only
-a local set of the $n_z-1$  (often nearest) points in a neighborhood of 
-the point $\rvec_i$ to
-approximate the derivative. In other words,
-$Lf(\rvec_i)=\sum_{j=1}^{n_z}a_jf(\rvec_j)$. The differentiation
-weights, $a_j$, are calculated by enforcing that this linear
-combination should be exact for RBFs,
-$\{\phi(\|\rvec-\rvec_{j}\|)\}_{j=1}^{n_z}$, centered at each of the
-node locations $\{\rvec_j\}_{j=1}^{n_z}$ (classical finite-differences
-(FD) would enforce that it be exact for polynomials instead.) Similar
-to FD, as the stencil size $n_z$ increases so does the order of the
-method.
-
-For a total of $N$ points, there will be $N$ linear systems to solve,
-each of size $n_z \times n_z$. Each linear solve produces a row of the
-RBFFD differentiation matrix $A_{n_z}$, resulting in a $N \times N$
-matrix with $n_zN$ nonzero entries. To evaluate the derivative at all
-points in the domain, one multiplies the source vector $x$ by the 
-derivative matrix $A$ to obtain the vector $y$ of derivative values, 
-$$
-  y = A x
-$$
-The computation of a single derivative has been reduced to a SpMV, 
-where each row has $n_z$ nonzeros\footnote{We use non-conventional 
-symbology for the differentiation matrix, source and destination vectors, 
-anticipating what follows.}. 
-In practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
-three-dimensional flows. These numbers are on par with what is used in
-finite-element codes. Notice that the nonzero elements of the matrix
-correspond to a relationship of closeness in the physical domain. 
-This relationship is not
-necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils}, which
-implies a non-symmetric sparsity pattern in the derivative matrix.
-
-\begin{figure}[tbh]
-  \centering
-  \includegraphics[width=.8\linewidth]{figures/rbf_stencils.pdf}
-  \caption{RBFFD Stencils. Each node of the stencil is connected to
-    $n_z-1$ stencil nodes in addition to itself. In the figure, node
-    $A$ is connected to $B$, but $B$ is {\em not\/} connected to
-    $A$. Thus adjacency graph of $A$ is not-symmetric.}
-  \label{fig:rbf_stencils}
-\end{figure}
-
-Many problems in fluid dynamics and in the geosciences require the
-solution to transport equations of the form
-\begin{equation}
-\pf{Q}{t} = f(Q,\grad{Q},\Laplacian{Q})  \label{eq:Q}
-\end{equation}
-where $Q$ is a vector of unknowns.
-For example, in solving a system of equations, it is
-often necessary to compute the derivative of multiple functions along
-the same coordinate direction,
-typically four for the Euler equations, or five for the Navier-Stokes
-equations. Multiple right-hand sides transform a SpMV into a SpMM
-(Sparse Matrix/dense Matrix multiplication), which improves register
-utilization and decreases cache misses by vectorizing over the
-multiple source vectors. Further improvements are possible by
-recognizing that different derivative matrices (e.g., gradient and 
-Laplacian operators), 
-have the same sparsity distribution as long as the derivative stencils
-remain unchanged; only the values in the sparse matrix vary. 
-Thus, rather than computing a derivative of multiple
-functions, one can calculate multiple derivatives of a single
-function. The increased memory bandwidth due to an increase in the
-number of derivative matrices is offset by better cache utilization,
-leading to an overall performance benefit. In practice however, the number of 
-vectors whose derivatives are required is limited, capping the 
-performance of the matrix/vector multiplication. Similarly, the number
-of different derivatives in a particular computation is finite. It is 
-for this reason that we seek to combine the benefits of multiple
-vectors and multiple matrices simultaneously. We limit this initial 
-study to four vectors and four matrices.
-
-In discrete form, Equation (\ref{eq:Q}) becomes
-$$
-  y^{i,j} = A^i x^j 
-$$ where $j$ ranges over the number of source vectors, $i$ ranges of
-  the number of derivative matrices and there are $i*j$ output vectors
-  $y$ (graphically shown in Figure~\ref{fig:struct_comp}).
-
-\begin{figure}
-  \centering
-
-  \[ \left( \begin{array}{cccc}
-    y^{1,1} & y^{1,2} & y^{1,3} & y^{1,4} \\
-    y^{2,1} & y^{2,2} & y^{2,3} & y^{2,4} \\
-    y^{3,1} & y^{3,2} & y^{3,3} & y^{3,4} \\
-    y^{4,1} & y^{4,2} & y^{4,3} & y^{4,4} 
-  \end{array} \right)
-  = \left(
-  \begin{array}{c}
-    A^1\\
-    A^2\\
-    A^3\\
-    A^4
-  \end{array}\right)
-  \times \left(x^1 x^2 x^3 x^4 \right) \]
-
-  \caption{Structure of the computation.}
-  \label{fig:struct_comp}
-\end{figure}
+\input{RBF_Section}
+%\input{old_RBF_Section}
 
@@
   \centering
-\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{code.pdf}}%
+\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{figures/code.pdf}}%
 \subfigure[Content of the vector registers]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
@@
 \section*{Acknowledgment}
-The first and last authors acknowledge NSF funding under NSF grant
-DMS-\#0934331 (FSU). NCAR is sponsored by the National Science
-Foundation. N. Flyer acknowledges support of NSF grant DMS-0934317.
+Erlebacher acknowledges funding from NSF grant
+DMS-\#0934331 (FSU), and the use of the MIC at FSU and at LCSE/U. Minnesota (provided by Prof. D. Yuen). NCAR is sponsored by NSF. 
+Flyer acknowledges support of NSF grant DMS-0934317.
 

### 2014-01-14 acm/old_RBF_Section.tex
Paper for ACM.

--- 2013-10-19 ieee_paper.tex
@@
-\documentclass[10pt,conference,compsocconf]{IEEEtran}
-\input{support/macros} % color is defined in macros or misc_mac
-\input{support/misc_mac}
-\input{support/setupicase}
-
-\def\red#1{\textbf{\textcolor{red}{#1}}}
-\def\blue#1{\textbf{\textcolor{blue}{#1}}}
-\def\qes#1{{\blue{*** For Erik: #1 ***}}}
-\def\es#1{{\blue{*** For Erik: #1 ***}}}
-\def\ge#1{{\red{*** For Gordon: #1 ***}}}
-\def\ttt#1{{\tt #1}}
-\def\bold#1{{\bf #1}}
-
-\usepackage{soul}
-\usepackage{xspace}
-\usepackage{color} 
-\definecolor{darkgreen}{rgb}{0,0.5,0}
-\usepackage[colorlinks=true,% 
-  linkcolor=red,% 
-  citecolor=darkgreen,%
-  urlcolor=blue]{hyperref}
-
-\usepackage{listings}
-
-\newcommand{\todo}[1]{{\color{red}\textbf{\hl{#1}}\xspace}}
-
-\def\qes#1{}
-\def\es#1{}
-%\def\ge#1{}
-%\usepackage{morefloats}
-
-\newcommand{\ceil}[1]{\left\lceil#1\right\rceil}
-\newcommand{\floor}[1]{\left\lfloor#1\right\rfloor}
-
-\usepackage{amsmath}
-\usepackage{amssymb}
-
-\usepackage[pdftex]{graphicx}
-\usepackage{subfigure}
-%\usepackage{fixltx2e}
-\usepackage{url}
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\begin{document}
-\title{Acceleration of Derivative Calculations with Application to Radial Basis Function --
-Finite-Differences on the Intel MIC Architecture}
-%\title{Sparse Matrix Vector Multiplication with Multiple vectors and
-  %Multiple Matrices on the MIC Architecture \todo{? more application oriented''}}
-
-\author{\IEEEauthorblockN{Gordon Erlebacher\IEEEauthorrefmark{1},
-Erik Saule\IEEEauthorrefmark{2}, Natasha Flyer\IEEEauthorrefmark{3}, 
-and Evan Bollig\IEEEauthorrefmark{1}}
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Scientific Computing, 
-Florida State University}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR}
-Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com}
-\maketitle
-
-\begin{abstract}
-In this paper, we develop an efficient scheme for the calculation of
-derivatives within the context of Radial Basis Function
-Finite-Difference (RBFFD). RBF methods express functions as a linear
-combination of radial basis functions on an arbitrary set of
-nodes. The Finite-Difference component expresses this combination over
-a local set of nodes neighboring the point where the derivative is
-sought.  The derivative at all points takes the form of a sparse
-matrix/vector multiplication (SpMV).
-
-In this paper, we consider the case of local stencils with a fixed number
-of nodes at each point and encode the sparse matrix in ELLPACK
-format. We increase the number of operations relative to memory
-bandwidth by calculating four derivatives of four different functions,
-or 16 different derivatives. We demonstrate a novel implementation on
-the Intel MIC architecture, taking into account its advanced swizzling and
-channel interchange features. We present benchmarks on a real data
-set that show an almost eightfold increase in speed compared to
-efficient implementations of a single derivative, reaching a
-performance of almost 140~Gflop/s. We explain the results through
-consideration of operation count versus memory bandwidth.
-\end{abstract}
-
-\begin{IEEEkeywords}
-MIC; SIMD; SpMV; SpMM; Sparse Matrix; Radial Basis Function;
-\end{IEEEkeywords}
-
-\IEEEpeerreviewmaketitle
-
-\section{Introduction}
-The multiplication of sparse matrix by a dense vector (SpMV) is an
-important kernel in many applied fields such as fluid dynamics 
-\cite{journals/tog/BolzFGS03}, recommendation systems~\cite{Brin98} and graph
-drawing~\cite{Koren05}). Naturally, improving the performance of SpMV has
-captured the interest of many researchers; including the development of
-various implementations for 
-CPUs~\cite{Buluc2009SPAA,Williams07} and GPUs~\cite{Bell08,
-  conf/ipps/KreutzerHWFBB12,
-  journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
-challenge  to obtain good performance for SpMV is that
-the operations are conducted using memory location that are irregular. This makes the kernels mostly memory
-bound and there is a significant instruction overhead per flop.
-
-Common improvement techniques such as register blocking, bandwidth
-reduction (matrix reordering~\cite{Cuthill69}), partitioning to fit in
-cache or TLB~\cite{Nishtala07,Temam:1992:CBS:147877.148091,conf/ppsc/Toledo97},
-unrolling~\cite{Mellor-Crummey04} have impacts which are very
-dependent on the matrix and overall do not lead to dramatic
-improvement. (The state-of-the-art techniques in OSKI~\cite{Vuduc05}
-provide some useful yet limited improvements). Register blocking~\cite{conf/ppsc/Toledo97} does
-not work well to many of the matrices in general use (although it can be applied
-with virtually no overhead thanks to compressed
-representations~\cite{Buluc11}.) Indeed, there are about 8 bytes of the matrix to transfer
-from memory 
-per nonzero in single precision; each nonzero requires two
-floating point operations leading to a flop-to-byte ratio of at most
-$\frac{1}{4}$. This limits the obtained performance to at most a
-quarter of the bandwidth of the architecture, wasting a lot of
-potentially useful cycles. The commonly used techniques are mostly
-designed to reach that bound rather than overcome it.
-
-Fortunately that fate is not inevitable. One solution would be to
-schedule a more instruction-intensive kernel simultaneously with the
-execution of SpMV, relying on some hardware threading capabilities,
-such as HyperThreading, to reduce the cycle wastage. However, most of 
-the applications that use SpMV do not typically have an 
-instruction-intensive kernel to run simultaneously.
-
-An other solution, and the one we pursue in this paper, is to compute
-multiple SpMVs at once on matrices that have the same sparsity
-patterns. Obviously not all the applications have such a property. But
-important classes of applications such as graph
-recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of
-derivatives of Radial Basis Functions (RBF)~\cite{FLBWSC12} can use multiple
-SpMVs simultaneously. In this paper in particular, we investigate the
-case of the derivative of RBFs where four derivatives of four
-different functions are expressed as the multiplication of four vectors
-by four matrices with identical sparsity patterns leading to the
-simultaneous execution of 16 SpMVs at a time. Using multiple vectors
-at a time has been investigated before to densify the
-computations~\cite{Im01} but this is the first time, to the 
-best of our knowledge, that the
-densification comes from adding both vectors and matrices.
-
-To perform our analysis, we focus our attention on the improvement
-that can be achieved on the Intel Xeon Phi processor. It follows the
-Many Integrated Core (MIC) architecture, which has a significant
-memory bandwidth and peak flop throughput thanks to its 512-bit large
-SIMD registers. The Xeon Phi processor has been shown to be promising
-for sparse linear algebra compared to more classical CPU or GPU
-architectures~\cite{Saule13-ARXIV, Liu:2013:ESM:2464996.2465013, cramer2012openmp}.
-
-In Section~\ref{sec:rbf}, we present the computation of RBFs and how it
-can be expressed as sixteen multiplications of four vectors by four sparse matrices
-with identical sparsity patterns. Section~\ref{sec:model} presents an
-estimation of the instruction intensity of various forms of the
-computations. We show that a sixfold improvement can be
-expected when computing the sixteen multiplications simultaneously to
-reach a total of about 210~Gflop/s. This performance represents
-approximately 10\% of the available flop/s of a Xeon Phi
-coprocessor. Therefore, it is necessary to have an implementation that
-performs the computation in as few instructions as
-possible. We describe in Section~\ref{sec:impl} the details of the MIC
-architecture and how to use specialized load, store, swizzle and
-permutation instructions to efficiently bring the data in the vector
-registers to be processed. Section~\ref{sec:expe} gives some
-experimental results about the amount of bandwidth that can be achieved
-depending on how the SpMV kernel is written. It also provides the actual performance
-of the various kernels on multiple classes of matrices, some generated
-for purpose of analysis, and some extracted from an RBF application. A
-performance of 135~Gflop/s is achieved using RBF-generated matrices. Concluding
-remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1
-
 \section{Derivatives of Radial Basis Functions}
@@
 \end{figure}
-
-\section{Modelization of the Potential Improvements}
-\label{sec:model}
-
-We saw in the previous section that one can express the RBF problem as
-a multiplication of four matrices by four vectors. We present here an
-estimation of the variation on the flop intensity of the computation
-and its impact on the expected performance of the
-application. Relevant notation are given in Figure~\ref{tab:not}.
-
-\begin{figure}[t]
-  \begin{center}
-    \scalebox{.8}{
-      \begin{tabular}{|c|l|}
-        %\hline
-        %& & \\
-        \hline
-        $b_i$ & number of bytes per index \\
-        $b_x$ & number of bytes per value \\
-        $n_z$ & number of nonzeros per row of $A$ \\
-        $n_r$ & number of column/rows of $A$ \\
-        $n_c$ & total number of nonzeros\\
-        $n_v$ & number of {\tt x} vectors \\
-        $n_m$ & number of matrices \\
-        $s_M$ & size of the $n_m$ matrices in bytes\\
-        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
-        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
-        \hline
-        $cl$    & size of a cache line in bytes\\
-        $b_{wT}$ & number of bytes written to memory  \\
-        $b_{rT}$ & minimum number of bytes read from memory  \\
-        $b_T$   & minimum number of bytes transferred \\
-        $B_{rT}$ & maximum number of bytes read from memory \\
-        $B_T$   & maximum number of bytes transferred \\
-        $O$     & number of floating point operations \\
-        $I_b$   & maximum computational intensity\\
-        $I_w$   & minimum computational intensity\\        
-        \hline
-      \end{tabular}
-    }
-  \end{center}
-  \caption{Notation relative to the application problem (upper section) and to the benchmarks (lower section.)}
-  \label{tab:not}
-\end{figure}
-
-Each vector in the problem is of dimension $n_r$ and each entry
-takes $b_x$ bytes. There are $n_v$ {\tt x} vectors and $n_v n_m$ {\tt
-  y} vectors, which lead to the size of the {\tt x} and {\tt y} vectors:
-$$s_x = n_v b_x n_r\mbox{, ~}s_y = n_v n_m b_x n_r$$
-
-The matrix is composed of $n_r$ rows and columns with $n_z$ nonzeros
-per row leading to a total of $$n_c = n_r n_z$$ nonzero entries in
-the matrix. Each of these nonzero entries has one index of size $b_i$
-and $n_m$ values of size $b_x$. The matrices have a total size
-of $$s_M = n_c (b_i + b_x n_m) = n_r n_z (b_i + b_x n_m)$$
-
-\begin{figure*}[tbh]
-  \centering
-  \subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
-  \subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
-  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}
-
-  \caption{Ratio of flops to bytes in single precision:  
-    \ref{fig:ratio_best} best case; \ref{fig:ratio_worst} worst case when $cl=64$ 
-    ; and \ref{fig:ratio_worst_nocache} worst case neglecting that the
-    memory transfers are with a granularity of one cache line (equivalent to $cl = 1$)}
-  \label{fig:ratio_bytes_flops}
-\end{figure*}
-
-If we assume an algorithm where the rows are processed one after
-the other, the amount of memory written is precisely the
-size of the {\tt y} vector. (This assumption removes the possibility of
-blocking or cache partitioning techniques.) $$b_{wT} = s_y = n_v n_m b_x n_r$$
-
-The amount of data read from memory depends highly on both the
-algorithm's execution path, and on how the matrix is structured. But
-in the best case both the matrix $A$ and the source vector {\tt x} are
-read once from the main memory. (We assume that all the elements of
-$\tt x$ are involved in the SpMV.)  Thus, $$b_{rT} = s_M + s_x = n_r
-n_z (b_i + b_x n_m) + n_v b_x n_r$$
- $$b_T = b_{rT} + b_{wT} =  n_r n_z (b_i + b_x n_m) + n_v b_x n_r (1 + n_m)$$
-
-Notice that there is no reason for a piece of the matrix to be read
-multiple times. But assuming that each element of the {\tt x} vector
-is read a single time is a strong assumption. If using a single core,
-it assumes that either the cache of the architecture can store the full
-{\tt x} vector or that the matrix is sufficiently well structured 
-to cause no cache trashing. If using multiple cores, this assumes that
-no element of the {\tt x} vectors will be used by multiple
-cores. \cite{Saule13-ARXIV} showed that there is very little cache trashing
-in practice; however having elements of the vectors used by
-multiple cores can have a significant impact on the performance
-(growing with $n_v$).
-
-On the other hand, in the worst case, every time the {\tt x} vector is
-accessed, the value needs to be transferred from memory again. So in
-total, there are as many transfers as the number of nonzeros in the
-matrix. Note however that most architectures cannot read memory a
-single byte at a time. Instead, a minimum number of bytes, equal to
-the size of a cacheline $cl$, are transferred at once.  When there are
-multiple vectors, each nonzero element uses $n_v$ consecutive
-entries. In the worst case, the number of bytes read and transferred is
-$$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}} $$ 
-$$B_T = n_v n_m b_x n_r + n_r n_z \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)$$
-
-In SpMV, each nonzero of the matrix requires two floating point
-operations: one for performing the multiplication and one for
-accumulating the result row-wise. Here we are dealing with $n_v n_m$
-simultaneous SpMVs and the number of floating point operations is
-$$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$$
-
-The computation intensity is the amount of floating point operations performed per
-byte transferred. In the worst case and in the best case, we have
-$$I_b = \frac{O}{b_T} = \frac{2 n_v n_m}{ (b_i + b_x n_m) + n_v n_m b_x n_z^{-1} + n_v b_x n_z^{-1} }$$
-%$$I_w = \frac{O}{B_T} = \frac{2 n_v n_m}{n_v n_m b_x n_z^{-1} + \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)}$$
-$$I_w = \frac{O}{B_T} = \frac{2 n_v n_m}{(b_i+b_x n_m) + n_v n_m b_x n_z^{-1} + cl \ceil{\frac{n_vb_x}{cl}} }$$
-
-Figure~\ref{fig:ratio_bytes_flops} presents the flop to byte ratios
-for single precision computation in the best and the worst case on a
-classical cache-based architecture ($cl = 64$) and assuming $cl=1$. We
-can easily see the potential improvement in the computational
-intensity when the number of matrices or vectors increases. There is a
-significant difference between the best and the worst case: there is a
-eightfold difference in the one matrix -- one vector ($n_m=n_v=1$) case but that gap closes
-with the increase in the number of vectors and matrices to a fourfold
-difference in the four matrices and four vectors ($n_m=n_v=4$) case and 1.6 fold in the 
-$n_m=n_v=16$ case. 
-The ratios computed with $cl = 1$ are mostly
-similar to the one with $cl=64$ when the number of vectors is
-large. But the differences are important when the number of vectors is
-low: this highlights that accessing the memory with a granularity of one cache
-line is the main problem faced when performing a standard SpMV computation.
-
-Figure~\ref{fig:perf_predict} shows how these ratios translate
-into actual performance. This figure provides projected best and worst
-Gflop/s achievable assuming the computation is memory-bound and the
-architecture reaches 150~GB/s. Notice that \cite{Saule13-ARXIV}
-showed a higher peak bandwidth, but we will show in
-Section~\ref{sec:expe} why 150~GB/s is a better estimate of what one
-might achieve in these kernels. In single precision, the worse that
-one could achieve by using four vectors and four matrices is much higher
-than the best achievable using a classical SpMV computation. The best
-performance reachable is 210~Gflop/s: almost six times higher than the
-peak of the classical SpMV case.
-
-\begin{figure}
-  \centering 
-  \includegraphics[width=.9\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}
-  \caption{Estimation of the maximum achievable performance using a
-    device with a bandwidth from memory to computational units of 150~GB/s when
-    varying the number of vectors and matrices.}
-  \label{fig:perf_predict}
-\end{figure}
-
-\section{Efficient Implementation on the Intel Xeon Phi processor}
-\label{sec:impl}
-
-\subsection{Intel Xeon Phi}
-
-In this work, we use an Intel Xeon Phi 5110P coprocessor. This card
-has 8 memory controllers that can each execute 5 billion
-transactions per second. Each has two 32-bit channels, achieving a total
-bandwidth of 320~GB/s aggregated across all the memory
-controllers. There are 60 cores clocked at 1.05~GHz. Their memory
-interfaces are 32-bit wide with two channels and the total bandwidth
-is 8.4~GB/s per core. Thus, the cores should be able to consume 504~GB/s
-at most. However, the bandwidth between the cores and the memory
-controllers is limited by the ring network that connects the cores and
-the memory controller. Its precise bandwidth is
-believed to be between 200~GB/s and 250~GB/s.
-
-Each core in the architecture has a 32~kB L1 data cache, a 32~kB L1
-instruction cache, and a 512~kB L2 cache. The architecture of a core is
-based on the traditional Pentium architecture, which has been extended to
-64-bit. A core can hold four hardware contexts at any given time (240 threads in total). At each
-clock cycle, instructions from a single thread are executed. Due to
-some hardware constraints, two hardware contexts must be used to reach
-the peak instruction throughput in that architecture. Similar to the
-Pentium architecture, a core has two different concurrent instruction
-pipelines that allow the execution of two instructions per
-cycle. However, only one vector or floating point instruction can be
-executed per cycle.
-
-Most of the performance of the architecture comes from the vector
-processing unit. Each core has 32 512-bit SIMD registers that can be
-used for double or single precision, that is, either as a vector of eight
-64-bit values or as a vector of 16 32-bit values, respectively. The
-vector processing unit can perform basic instructions, such as
-addition or division, and mathematical operations, such as {\tt sin} or
-{\tt sqrt}, making it possible to reach 16 single precision (or 8 double precision)
-operations per cycle.  
-The unit also supports Fused Multiply-Add (FMA)
-operations, which are typically counted as two operations 
-when benchmarking. Therefore, the peak performance of the 5110P
-card is one Tflop/s in double precision and two Tflop/s in single
-precision. If FMA cannot be used, only half of these rates can be
-achieved.
-
-\subsection{Bringing the data into the vector register}
-
-\begin{figure*}
-  \centering
-\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{code.pdf}}%
-\subfigure[Content of the vector registers]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
-
-\caption{Code snippet of the multiplication of 4 vectors with 4
-  matrices. When a line changes the content of a vector register its
-  new content is shown on the right. The effects of swizzling and
-  permutations are highlighted using a color code.}
-\label{code:mat_mul}
-\end{figure*}
-
-As explained in Section~\ref{sec:model}, we expect to achieve a performance of
-about 210~Gflop/s in single precision in the best case. We will focus
-on the single precision case with four vectors and four matrices. 
-Similar techniques apply for other combinations. This target
-performance represents 10\% of the peak performance of the
-architecture, which cannot be reached without an efficient vectorization. 
-Indeed, to reach such a performance, a Fused Multiply-Add
-instruction on fully loaded registers must be executed at most every
-10 cycles.
-
-In term of memory layout, we adopt a format similar to the ELLPACK
-format to store the matrix since the number of nonzeros per row is
-constant throughout the matrix. The matrix is given in two arrays. The
-first array is {\tt col\_id} that describes the
-sparsity structure of the matrix. Each row of {\tt col\_id} consists
-of $n_z$ entries that label the column index of each nonzero in
-$A$. The other array is  {\tt data} that gives the values of
-the nonzeros in the four matrices. The values of the matrices are
-interleaved by groups of four so that the value $A^l_{col\_id[k]}$ of
-the $k$th nonzero of the $l$th matrix is at index $\floor{\frac{k-1}{4}}*4*n_m + (l-1)*4+(k-1)\%4$.
-(In other words, the first four entries are from $A^{1}$ and the 
-next four are from $A^2$.) The vectors {\tt x} are
-interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
-consecutive in memory. The vector {\tt y} is similarly 
-grouped in sets of 16, which correspond to four derivatives applied to
-four functions, at a single point.\looseness=-1
-
-We discuss how the implementation of the v4m4 kernel works using
-vector instructions. The code is given for reference in
-Figure~\ref{code:mat_mul}. (We will show in Section~\ref{sec:expe}
-that an implementation that relies on compiler vectorization does not lead
-to desirable performance, which confirms the results
-of~\cite{Saule13-ARXIV}.) The registers in the MIC architecture are
-512 bits wide and can store 16 floats or integers. (In comparison to
-OpenCL or CUDA, one can think of these registers as a warp.) The end
-goal is to load and format the nonzero values $A_h^1,A_h^2,A_h^3,A_h^4$
-and the vector entries $x_m^1,x_m^2,x_m^3,x_m^4$ into two vector registers and 
-apply Fused Multiply-Add on them, which is depicted in the {\tt accu} line of 
-Figure~\ref{code:mat_mul}. Bringing the data into the vector
-register as shown in the figure is the difficult component of our 
-proposed algorithm, which we proceed to explain further. 
-A thread will perform the multiplications one row at a time,
-and parallelism is achieved by giving blocks of rows to each thread
-using an OpenMP construct.
-
-The core of the technique is the use of swizzling and permutation
-features of the vectors in the Intel MIC architecture. A vector of 512
-bits is composed of four lanes (sometimes called channels) of 128
-bits, which are made of four segments of 32 bits. A swizzling
-operation allows the reorder of the elements within each lane (see how
-{\tt mat} is extracted from {\tt mat\_ent} in
-Figure~\ref{code:mat_mul}). On the other hand, a permutation reorders entire lanes,
-without changing their contents (see for instance how {\tt vect} is
-generated from {\tt all\_vect} in Figure~\ref{code:mat_mul}).  Both
-permutation and swizzling support common reordering and broadcast. In
-the code, the ``permute'' function needs to use the {\tt
-  \_mm512\_permute4f128\_epi32} instruction which is only defined on
-integers and therefore in place typecasts are necessary to convert
-a vector to and from integer type.  (Note that we prefer swizzling to
-permutation if possible because we believe it is cheaper. Many
-instructions have swizzling capability, while permutation is its own
-instruction.)
-
-Every operation that loads data from memory into a register
-transfers 512 bits, or 16 floats at a time. A single nonzero
-in the matrix will cause 16 operations to be performed, but it only
-uses four floats from the matrix and four floats from the {\tt x} vector. So
-we would like to execute a single instruction to load the data for four 
-nonzeros in the matrix, and another instruction to load the data for
-four nonzeros from the vectors.  
-(The number of nonzeros on a row in our application is
-always a multiple of four; if it weren't, we would add some explicit
-zeros in the data structure for padding.)
-
-The values from the nonzero elements in the matrix are the simplest to
-load. One can load the 16 floating point values for four nonzeros of
-the matrix at once. The four floats coming from each matrix are
-naturally grouped within each lane of the SIMD register. One can use a
-swizzle operation to broadcast the four elements of a single matrix to
-fill the whole register. (The broadcast of $A_h^1,A_h^2,A_h^3,A_h^4$ is
-shown in Figure~\ref{code:mat_mul}.)
-
-Loading the entries from the {\tt x} vector is  more
-involved. First, the column pointers of four different nonzeros are
-loaded into a vector register and the values are distributed one per
-lane of the vector register using an {\tt unpack}
-operation\footnote{Notice that there are two calls to {\tt unpack} for the
-LSB and MSB. Both are mandatory even if one is known to be
-nilpotent according the documentation of the hardware.} (The column
-pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They
-are replicated so that each value fills a single vector lane 
-using a swizzle operation. Then each value is multiplied by four
-(because $n_v=4$) 
-and the offset vector $(3,2,1,0)$ is added to each lane to obtain the
-correct index of the elements of the {\tt x} vector within each lane
-using a Fused Multiply-Add operation. A {\tt gather} operation is then
-performed to bring the 16 entries of the {\tt x} vector into the SIMD
-register. At this point we have a register where each lane is the
-four vector entries for a nonzero element. Using a broadcast
-permutation, one can replicate an entry of the four vectors across
-each lane. (The broadcast of $x_m^1,x_m^2,x_m^3,x_m^4$ is shown in
-Figure~\ref{code:mat_mul}.)
-
-Now that $A_h^1,A_h^2,A_h^3,A_h^4$ and $x_m^1,x_m^2,x_m^3,x_m^4$ are
-properly laid out in the vectors, a partial value for $y_i$ can be
-computed. Once all the nonzeros of the row have been processed, the
-exact $y_i$ value can be sent to memory using a the No Read No Global
-Order hint to optimize write traffic on the memory bus.
-
-\section{Experimental Validation}
-\label{sec:expe}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=.9\linewidth]{figures/bandwidth_read_write.pdf}
-
-  \caption{Bandwidth performance under idealized conditions.  Entries
-    with "cpp" (triangles) denote cases where vectorization is
-    compiler generated.}
-  \label{fig:bandwidth}
-  \label{fig:band_rw_MIC}
-\end{figure}
-
-In this section, we describe a series of numerical experiments to
-confirm the performance predictions of Section~\ref{sec:model}. To
-better understand our results, we first perform some bandwidth
-measurements, followed by the actual SpMV computations. All the codes
-are compiled with the Intel C Compiler version 13 and
-\mbox{-O3} optimization. The codes use the IMCI instructions exposed by
-the compiler through intrinsics.
-
-\subsection{Bandwidth}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=.9\linewidth]{figures/bandwidth_gather_unpack.pdf}
-  \caption{Performance of {\tt gather} and {\tt unpack} operations.}
-  \label{fig:band_gather} 
-\end{figure}
-
-Figure~\ref{fig:bandwidth} shows a benchmark that computes the
-bandwidth achieved when reading and/or writing large arrays on both
-the MIC card and the CPU architecture. The read benchmark performs a
-simple sum of the array. The write benchmark sets all the elemnts of
-the array to zero, while the read/write benchmarks copies the array into
-another one. We also investigate the difference between a straight C++
-implementation (compiled with -O3 with proper tagging to inform the
-compiler of memory alignments) and a simple implementation with
-explicit use of the vector registers.
-
-The {\tt read, cpp} implementation reaches a performance of 168~GB/s
- which is better than the {\tt read}
-implementation which only achieves 145~GB/s. This
-is explained by the compiler adding loop unrolling and software
-prefetching in the assembly code in the {\tt read, cpp} case, which is
-not added in the {\tt read} case. On the other hand, the {\tt write}
-case reaches a performance of 158~GB/s while the
-{\tt write, cpp} implementation never exceeds 58~GB/s. 
-This difference is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction, 
-which bypasses the Read For Ownership protocol and allows the writes
-to be performed in any order. The compiler can not use this
-instruction on its own because this would break the memory convention
-useful for some parallel algorithms. The {\tt read/write} benchmark
-performs at 140~GB/s while the {\tt read/write, cpp} benchmark performs
-at about 90~GB/s.
-
-Note that SpMV kernels are not likely to use only simple read and simple
-writes. We also benchmark the performance we
-can achieve using {\tt gather} and {\tt unpack} operations and we
-present the results in Figure~\ref{fig:band_gather}. The {\tt unpack}
-instruction can reach a performance of 140~GB/s. (There is no easy way
-to implement the equivalent of the {\tt unpack} operation without
-using explicit registers.) The performance of the {\tt gather}
-operations depends in which order the array is read. If the indices
-are ordered ({\tt compact}) then a performance of 140~GB/s is
-reached. If the indices are completely randomized, the performance
-drops below 5~GB/s. We can see that implementing the
-indirection mechanism without vector instructions lowers the
-performance significantly. (In the {\tt compact} case, it never
-surpasses 100~GB/s and is close to 50~GB/s most of the time.)
-\def\ww{.19\textwidth}
-\begin{figure*}[t]
-  \begin{center}
-    \subfigure[Supercompact.]{%
-      \label{fig:supercompact}
-      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.png}}%
-    \subfigure[Compact.]{%
-      \label{fig:compact}
-      \includegraphics[width=\ww]{figures/compact_matrix-crop.png}}%
-    \subfigure[Random.]{%
-      \label{fig:random}
-      \includegraphics[width=\ww]{figures/random_matrix-crop.png}}%
-    %% \subfigure[2D, no RCM.]{%
-    %%   \label{fig:rbf2dnorcm}
-    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.png}}%
-    %% \subfigure[2D, RCM.]{%
-    %%   \label{fig:rbf2drcm}
-    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
-    \subfigure[3D, RCM.]{%
-      \label{fig:rbf3dnorcm} 
-      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
-    \subfigure[3D, no RCM.]{%
-      \label{fig:rbf3drcm}
-      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.png}}%
-  \end{center}
-
-  \caption{Different sparsity distributions. In all cases, there are
-    512 rows and 32 nonzeros per row. The last two matrices
-    corresponds to a derivative stencil in 3D RBFFD
-    calculations (with and without RCM ordering).}
-  \label{fig:spy_plots}
-\end{figure*}
-
-To summarize, using vector instructions appears necessary in order
-to reach the highest bandwidth on the MIC architecture. For the instructions used
-in a typical SpMV computation, one can expect a bandwidth between
-140~GB/s to 160~GB/s. That is why we use in our estimations a best
-achievable bandwidth of 150~GB/s despite the fact that a 
- code carefully crafted to maximize bandwidth can reach higher
-values~\cite{Saule13-ARXIV}.
-
-\subsection{Instances}
-
-The different matrices used to perform our experiments and analysis
-are presented in Figure~\ref{fig:spy_plots}. The
-Supercompact matrices (Figure~\ref{fig:supercompact}) have been
-generated to only have nonzero elements in the first 32 columns of
-the matrix. As a result, only 32 values of {\tt x} are read
-from memory, and the expense associated with cache misses is removed from consideration. The cost associated with 
-{\tt A} remains. The Compact matrices
-(Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
-centered around the diagonal and it represents the ideal case for many
-applications that rely on sparse matrix vector multiplication. The nonzero
-elements of the Random
-matrices (Figure~\ref{fig:compact}) see their nonzero elements
-randomly (uniformly) distributed in the matrix; they represent the
-worst case scenario for a cache-based architecture where the cache
-reutilisation is the lowest from one row to the next.
-The other type of matrix is used in derivative stencils in 3D
-RBFFD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-points
-stencil of a 3D $8^3$ grid).
-
-We also apply a Reverse Cuthill-Mckee reordering to
-all the matrices.  This ordering technique aims to reduce the
-distance between the nonzeros and the diagonal, hopefully increasing 
-the cache hit ratio.  The reordered version of the matrix can be
-seen in Figure~\ref{fig:rbf3drcm}.
-
-\vspace{-0.3em}
-\subsection{Computations}
-
-\begin{figure*}[t]
-  \centering
-  \subfigure[Supercompact and Random.]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
-  \subfigure[Compact and RBFFD.]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
-
-  \caption{Performance of the manually vectorized code. The two
-    horizontal lines depict the worst and the best predicted performances.}
-  \label{fig:expe_types}
-\end{figure*}
-
-We now investigate the actual performance that we can obtain when
-multiplying four vectors by four matrices in single
-precision. Figure~\ref{fig:expe_types} presents results for different
-types of matrices and also shows the minimum and maximum performance
-predicted in Section~\ref{sec:model}. Figure~\ref{fig:perf_ideal}
-gives the results for the {\tt supercompact} and {\tt random} matrices
-that represent both the best and the worst case for such a practical 
-computation. The {\tt supercompact} case peaks at
-208~Gflop/s which is very close to the predicted peak
-performance of 213~Gflop/s. Conversely the performance of the {\tt
-  random} case decreases to 56~Gflop/s which is close to
-the lowest predicted performance of 55~Gflop/s. One can see that RCM
-ordering helps the {\tt random} case but the impact decreases when the
-size of the matrix increases.
-
-Figure~\ref{fig:perf_realistic} gives the results on the {\tt compact}
-case, which represents the best realistic matrix one could find with very
-structured grids and a
-{\tt RBF} derivative matrix extracted from a real 3D application. In the {\tt
-  compact} case, the performance can be as high as 195~Gflop/s, which is
-within 15\% of the predicted  peak performance. The performance of the
-{\tt RBF} case varies between 100~Gflop/s and 140~Gflop/s. Most of the
-time, the RCM ordering provides an improvement which can be as high as
-30~Gflop/s.
-
-\begin{figure}[t]
-  \centering
-
-  \includegraphics[width=.9\linewidth]{figures/mic_performance_nb_threads.pdf}
-
-  \caption{Performance of $y=Ax$ for a RBF derivative stencil of a
-    $96^3$ 3D grid. Classical SpMV ($n_m=n_v=1$ given in blue) has low
-    performance. The manually vectorized code (red) is 3.5 times
-    faster than the compiler generated one (green). RCM ordering
-    (squares) only improves the performance for our manually
-    vectorized code.}
-  \label{fig:comparison_spmv}
-  \label{fig:perf_mic}
-\end{figure}
-
-We finally investigate the central questions of this paper. Do we gain
-actual performance by transforming classical SpMV computation into the
-multiplication of four vectors by four matrices for the computation of the
-derivative of RBFs? Does using manual vectorization improve actual
-application performance? Figure~\ref{fig:comparison_spmv} compares the
-performance achieved by a classical SpMV and by the multiplication of
-four vectors by four matrices using either standard C++ code and our
-optimized implementation. The
-classical SpMV computation reaches 14~Gflop/s. 
-The standard C++ implementation reaches a performance of 38~Gflop/s while our
-optimized implementation almost reaches 140~Gflop/s. 
-Notice that 
-RCM ordering has no impact in the standard C++ implementation while it
-provides a significant improvement in our manually vectorized
-version. This highlights that the C++ implementation is instruction-bound
-while our manually vectorized implementation is most likely 
-memory-bound.
-
-To put things in perspective, we conducted some experiments on a
-classical CPU architecture with two Intel Core i7 processors, each has eight
-cores that support AVX instructions. Both the compiler-generated and
-the manually-written vectorization using AVX instructions reach a
-(read) bandwidth of 70~GB/s. This indicates that the performance of
-the classical SpMV can not overcome 17~Gflop/s. A generic
-implementation of the multiplication of four vectors with four matrices on
-the {\tt RBF} matrix reaches a performance of 27~Gflop/s; while a
-manual implementation using AVX instructions reaches a performance of
-42~Gflop/s. This confirms that our method is sound and generalizes to
-architectures other than the Intel MIC architecture.
-
-\vspace{-0.5em}
-\section{Conclusion}
-\label{sec:ccl}
-
-In the previous sections, we have explored the practical
-implementation on an Intel Xeon Phi card of multiple derivative
-operators acting on multiple vectors within the context of
-RBFFD. Each derivative has an associated sparse matrix with a fixed
-number of nonzeros on each row. While computing a single derivative of
-multiple functions is rather common, we accelerate the algorithm further 
-by considering multiple
-matrices, with identical sparsity patterns, acting on multiple vectors. 
-We specialize our study to four matrices and four derivatives,
-with 16 outputs, computed as a sum of outer products.
-
-Our implementation makes use of the IMCI MIC instruction set and
-includes a number of swizzling and channel swapping operations, for an
-extremely efficient tensor product implementation. On a $96^3$ 3D
-grid, it reaches a performance of 140~Gflop/s which is 2.7 times
-faster than the best possible performance achievable for a single
-vector and a single derivative and more than 7 times faster than a
-practical implementation.
-
-In future work, we will examine the effect of larger stencil sizes and
-double precision. The presented performance does not reach the predicted
-peak and we believe that common techniques from SpMV such as cache
-partitioning can be successfully applied. We will also integrate the
-techniques presented in this paper within a fluid simulation using
-RBFFD.
-
-\section*{Acknowledgment}
-The first and last authors acknowledge NSF funding under NSF grant
-DMS-\#0934331 (FSU). NCAR is sponsored by the National Science
-Foundation. N. Flyer acknowledges support of NSF grant DMS-0934317.
-
-\bibliographystyle{plain} %{IEEEabrv,paper.bib}
-\bibliography{paper, bollig}
-\end{document}

### 2014-01-14 acm/paper_submission.tex
Paper for ACM.

%\documentclass[10pt,conference,compsocconf]{IEEEtran}
\documentclass{acm_proc_article-sp}
\input{support/macros} % color is defined in macros or misc_mac
\input{support/misc_mac}
%\input{support/setupicase}
\usepackage{amsmath}  %NATASHA

\def\red#1{\textbf{\textcolor{red}{#1}}}
\def\blue#1{\textbf{\textcolor{blue}{#1}}}
\def\qes#1{{\blue{*** For Erik: #1 ***}}}
\def\es#1{{\blue{*** For Erik: #1 ***}}}
\def\ge#1{{\red{*** For Gordon: #1 ***}}}
\def\ttt#1{{\tt #1}}
\def\bold#1{{\bf #1}}

\usepackage{soul}
\usepackage{xspace}
\usepackage{color}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\usepackage[colorlinks=true,%
  linkcolor=red,%
  citecolor=darkgreen,%
  urlcolor=blue]{hyperref}

\newcommand{\todo}[1]{{\color{red}\textbf{\hl{#1}}\xspace}}

\def\qes#1{}
\def\es#1{}
%\def\ge#1{}
%\usepackage{morefloats}

\newcommand{\ceil}[1]{\left\lceil#1\right\rceil}

\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
%%\usepackage{subfigure}
%\usepackage{fixltx2e}
%%\usepackage{url}
%%\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\title{Sparse Matrix Vector Multiplication with Multiple vectors and
  Multiple Matrices on the MIC Architecture}

%\begin{abstract}
\subtitle{[Extended Abstract]}

\numberofauthors{4}

\author{
\alignauthor
Gordon Erlebacher\\ %\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Department of Scientific Computing}\\
       \affaddr{Florida State University}\\
       \affaddr{Tallahassee, FL 32306-4120}\\
       \email{gordon.erlebach@gmail.com}
\alignauthor
Erik Saule
        \affaddr{Department of Computer Science}\\
        \affaddr{University of North Carolina at Charlotte}\\
        \email{esaule@uncc.edu}
\alignauthor
Natasha Flyer
        \affaddr{Computational and Information System Laboratory}\\
        \affaddr{UCAR}\\
        \email{flyer@ucar.edu}
\alignauthor
Evan Bollig
        \affaddr{Minnesota Supercomputer Institute}\\
        \affaddr{University of Minnesota}\\
        \affaddr{Minneapolis, MN}\\
        \email{bollig@gmail.com}
}
\maketitle

\begin{abstract}
In this paper, we develop an efficient scheme for the calculation of
derivatives within the context of Radial Basis Function
Finite-Difference (RBFFD). RBF methods express functions as a linear
combination of radial basis functions on an arbitrary set of
nodes. The Finite-Difference component expresses this combination over
a local set of nodes neighboring the point where the derivative is
sought.  The derivative at all points takes the form of a sparse
matrix/vector multiplication (spmv).

In this paper, we consider the case of local stencils with the number
of nodes at each point and encode the sparse matrix in ELLPACK
format. We increase the number of operations relative to memory
bandwidth by calculating four derivatives of four different functions,
or 16 different derivatives. We demonstrate a novel implementation on
the MIC architecture, taking into account its advanced swizzling and
channel interchange features. We present benchmarks that show an
almost order of magnitude increase in speed compared to efficient
implementations of a single derivative. We explain the results through
consideration of operation count versus memory bandwidth.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{OpenMP, MIC, spmv, sparse matrix, Radial Basis Function

\section{Introduction}
\cite{Bell08, Vuduc05, Nishtala07, Stock12-TACC, Cuthill69, cramer2012openmp, Buluc2009_SPAA, Buluc11, Im01, Mellor-Crummey04, Nishtala07, Saad94sparskit, Williams07}
\cite{Temam:1992:CBS:147877.148091}
\cite{DBLP:journals/ijhpca/ShantharamCR11,
conf/ppsc/Toledo97, Liu:2013:ESM:2464996.2465013, Molka:2009:MPC:1636712.1637764,%
DBLP:journals/corr/abs-1101-0091, conf/ipps/BulucWOD11, conf/ipps/KreutzerHWFBB12,%
kumar2012accelerating, journals/concurrency/VazquezFG11}

SpMV is an important kernel for lots of stuff. So improving the
performance of SpMV has caputred the interest of many researcher.

The main challenge that is faced to improve good performance for SpMV
is that the operation are conducted using memory location that are
irregular and often unpredictible. That make that the kernels are
mostly memory bound and there is a significant instruction overhead
per flop.

Common improvement techniques such as register blocking, bandiwdth
reduction (matrix reordering), partitioning to fit in cache or TLB
have impacts which are very dependent on the matrix and overall do not
lead to dramatic improvement. Assuming register blocking does not
apply well to the matrix at hand (which is true for most matrices),
there is about 8 bytes of the matrix to move in per non zero (assuming
single precision); each nonzero requires two floating point operations
leading to a flop-to-byte ratio of at most $\frac{1}{4}$. This limits
the obtained performance to at most a quarter of the bandwidth of the
architecture wasting a lot of potentially useful cycles. The commonly
used techniques are mostly designed to reach that bound rather than
overcome it.

Fortunately that fate is not inevitable. One solution would be to pair
multiple component of an application to schedule a more instruction
intensive kernel simultaneously with SpMV, relying on some hardware
threading capabilities, such as HyperThreading, to reduce the cycle
wastage. However most ot the applications that use SpMV do not
typically have an instruction intensive kernel to run simultaneously.

An other solution, and the one we pursue in this paper, is to compute
multiple SpMV at once on matrices that have the same sparsity
patterns. Obviously not all the applications have such a property. But
important classes of applications such as graph
recommendation~\cite{}, eigensolving~\cite{} and the computation of
derivative of Radial Basis Functions(RBF)~\cite{} can use multiple
SpMVs simultaneously. In this paper in particular, we investigate the
case of the derivative of RBFs where four derivatives of four
different function is expressed as the multiplication of four vectors
by four matrices with identical sparsity patterns leading to the
simultaneous execution of 16 SpMVs at a time.

To perform our analysis, we focus our attention on the improvement
that we can achieve on the Intel Xeon Phi processor. It follow the
Many Integrated Core (MIC) architecture, which has a significant
memory bandwidth and peak flop throughput thanks to its 512-bit large
SIMD registers. The Xeon Phi processor has been shown to be promising
for sparse linear algebra compared to more classical CPU or GPU
architecture~\cite{}.

In Section~\ref{sec:rbf} we present the computation of RBFs and how it
can be expressed 16 multiplication of 4 vectors by 4 sparse matrices
with a common sparsity pattern. Section~\ref{sec:model} presents an
estimation of the instruction intensity of various form of the
computations and we show that a 5- to 6- fold improvement can be
expected when computing the 16 multiplications simultaneously and
reach a total of about 200 Gflop/s. This performance represents
approximatively 10\% of the available flop/s of a Knight Corner
coprocessor. Therefore, it is necessary to have implementations that
perform the computation in as little amount of instructions as
possible. We describe in Section~\ref{sec:impl} the details of the MIC
architecture and how to use specialized load, store, swizzle and
permutation instruction to efficiently bring the data in the vector
registers to be processed. Section~\ref{sec:expe} gives some
experimental result about the amount of bandwidth that can be achieved
depending on how the spmv kernel is written and the actual performance
of the various kernel on multiple classes of matrices some generated
for analysis purpose and some extracted from and RBF application. A
performance of xxx GFlop/s is achieved on real scenario. Concluding
remarks and perspectives are provided in Section~\ref{sec:ccl}.

\section{Derivatives of Radial Basis Functions}
\label{sec:rbf}

%In the theory of Radial Basis Functions Finite-Difference (RBFFD),
%derivatives of a function $f(\rvec)$ at node $i$ are expressed as a
%linear combination of the function values at the stencil center and
%the nodes connected to node $i$ (Figure~\ref{fig:rbf_stencils}). Thus
%$y$ is a discrete derivative of the vector $x$.

In this paper, we propose yet a new idea that is applicable to radial basis functions (RBFs). Their strength is the ability to randomly distribute points across complex physical domains, and have an implementation that is independent of dimensionality. RBFs approximate a function $f(\xvec)\subset \mathbb{R}^d$ sampled at a set of $N$ distinct point locations, $x_j$, by linearly combining translates of a single radially symmetric function $\phi(\ep r)$, where  $r = \|\xvec-\xvec_{j}\|$ denotes the Euclidean distance (e.g., in 2-D $\sqrt{(x-x_j)^2+(y-y_j)^2}$) between where the function is evaluated $\xvec$ and where the RBF is centered $\xvec_{j}$. That is, the interpolant is $s(\xvec) =  \sum_{j=1}^{N} w_j \phi_i(\|\xvec-\xvec_{j}\|)$. The weights $w_j$ are obtained by inverting the system
\begin{equation}
\begin{bmatrix}
\phi(\|\xvec_{1}-\xvec_{1}\|) & \phi(\|\xvec_{1}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{1}-\xvec_{N}\|) \\
\vdots & \ddots & \vdots  \\
\phi (\|\xvec_{n}-\xvec_{1}\|) & \phi(\|\xvec_{n}-\xvec_{2}\|) & \cdots & \phi(\|\xvec_{N}-\xvec_{N}\|)
\end{bmatrix}
\begin{bmatrix}
w_{1} \\
\vdots \\
w_{N}
\end{bmatrix}
=
\begin{bmatrix}
f(\xvec_1) \\
\vdots\\
f(xvec_N)
\end{bmatrix}. \label{eq:rbf}
\end{equation}

The RBF differentiation matrix, $D_N$, is derived by applying the desired analytic derivative operator $L$ to the RBF interpolant $s(\xvec)$ above and evaluating it at the point locations. For very large problems, this is an computationally expensive since the matrix in (\ref{eq:rbf}) is full and inversion requires O$(N^3)$ operations. To alleviate the cost of this global approach (i.e. using every node in the domain to calculate the derivative at a given node $\xvec_i$), RBF-generated finite differences (RBF-FD) have been derived \cite{TAI1,TAI2,SDY02,WrFo06,FoL11,FLBWSC12}. RBF-FD use only a local set of the $n_z-1$ nearest neighbors to the point $\xvec_i$ to approximate the derivative. In other words, $Lf(\xvec_i)=\sum_{j=1}^{n_z}a_jf(\xvec_j)$. The differentiation weights, $a_j$, are calculated by enforcing that this linear combination should be exact for RBFs, $\{\phi(\|\xvec-\xvec_{j}\|)\}_{j=1}^n_z$, centered at each of the node locations $\{\xvec_j\}_{j=1}^{n_z}$ (classical finite differences (FD) would enforce that it be exact for polynomials instead). Similar to FD, as the stencil size $n_z$ increases so does the order of the method.

For a total of $N$ points, there will be $N$ linear systems to solve, each of size $n_z \times n_z$. Each linear solve produces a row of the RBF-FD differentiation matrix $D_{n_z}$, resulting in a $N \ times N$ matrix with $n_zN$ nonzero entries. To evaluate the derivative at all points in the domain, takes the form
$$
  \mathbf{g} = D_{n_z} \mathbf{f}
$$
where $f$ is the source vector of function values and $g$ is the resulting vector of derivative values. The computation of a single derivative has been reduced to a SpMV, where each row has $n_z$ nonzeros. Since the sparsity of each row is constant, ELLPACK~\cite{} becomes the most appropriate matrix compression scheme\cite{}. In practice, $n_z=32$ in two-dimensional flows and 64 or 100 for three-dimensional flows. These numbers are similar to what is used in finite-element codes.

Many problems in fluid dynamics and in the geosciences require the solution to transport equations of the form
$$
\pf{Q}{t} = f(Q,Q_x, Q_y, Q_z, \Laplacian{Q})
$$
where $Q$ is a vector of unknowns (3 components of velocity,
pressure, temperature). For example, in solving a system of equations, it is often necessary to compute the $x$
derivative of multiple functions, typically four for the Euler
equations, or five for the Navier-Stokes equations. Multiple
right-hand sides transform a SpMV into a SpMM (Sparse Matrix/dense
Matrix multiplication), which improves register utilization and
decreases cache misses by vectorizing over the multiple source
vectors. Further improvements are possible by recognizing that
different derivative matrices (x,y,z and Laplacian for example), have
the same sparsity distribution; only the weight change.  Thus,
alternative to computing a derivative of multiple functions, we can
calculate multiple derivatives of a single functions. The increased
memory bandwidth due to an increase in the number of derivative
matrices is offset by better cache utilization, leading to an overall
benefit.

% GORDON I WOULD ELIMINATE BELOW AND END HERE.

**** PUT ELSEWHERE OR ELIMINATE? ****

When solving a one dimensional system of PDEs, one might require an
$x$ derivative of multiple functions. For example, the Euler equations
require the $x$ derivative of the three components of velocity and
pressure. In this case, there are $n_v=4$ vectors $x^k$,
$k=0,\cdots,3$.

Thus for each vector element $y_i$, we compute $y_i = \sum_j A_{ij}
x_j$. If $A_i$ is row $i$ of $A$, $y_i$ is simply the dot product $A_i
x$. The next level of generality is to consider $n_v$ vectors $x^k$,
$k=0,\cdots,n_v-1$. Whatever the spmv implementation, one achieves
improved performance if the matrix formed from the columns $x^k$ are
stored in row major order. Thus, $x^0_0,x^1_0,\cdots,x^{n_v-1}_0$, are
stored in consecutive memory location. The random access of the
elements of $x$ is thus reduced. Maximum efficiency is achieved when
$n_v=16$ floats or $8$ doubles, given that cache lines take 64
bytes. We will benchmark this case, labeled $Svn$, where $n$ refers to
the number of vectors (Iv4 uses four vectors). The $S$ refers to singe
precision. A double precision run is labelled $Dvn$.

Alternatively, when solving a PDE, one might require derivatives of a
given scalar function with respect to coordinate directions $x$, $y$,
$z$. Second order operators of often required, such as a second
derivative with respect to $x$ or a Laplacian operator. In the RBFFD
formulation, on can compute different derivatives using the same
stencil, but with different weights. In other words, the adjacency
matrix that corresponds to $A$ remains constant, but the matrix
elements of $A$ change with the particular derivative.  In this case,
label with a superscript $l$ the particular matrix $A^k$. Since the
adjacency matrix is assumed invariant, there is only need for a single
matrix \ttt{C{ij}}. In Ellpack format, each row is of constant size
(the number of nonzeros per row of $A$. $C_{ij}$ is the column number
that locates the $j^{th}$ nonzero in row $i$ of $A$.

\todo{Must rewrite the above to make it more focused on the
  application. Remove the architectural/implementation details such as
  using ELLPACK. Let's focus on on why this computation is important
  and how does the 16 multiplication appears in the equations.}

\def\wide{2.5in}
\newfig{figures/rbf_stencils.pdf}{\wide}{RBFFD Stencils. Each node of the stencil is connected to $n_z-1$ stencil nodes in addition to itself. In the figure, node $A$ is connected to $B$, but $B$ is {\em not\/} connected to $A$. Thus adjacency graph of $A$ is not-symmetric. This must be corrected when reducing the bandwidth of $A$ with a Cuthill\-McKee algorithm, which assumes symmetry.}{fig:rbf_stencils}

\newfig{figures/matrix_structure.pdf}{\wide}{Matrix Structure.}{fig:mat_struct}

\section{Modelization of the Potential Improvements}
\label{sec:model}

We saw in the previous section that one can express the RBF problem as
a multiplication of four matrices by four vectors. We present here an
estimation of the variation on the flop intensity of the computation
and its impact on the expected performance of the
application. Relevant notations are given in Table~\ref{tab:not}.

\begin{figure}
  \begin{center}
    \scalebox{.9}{
      \begin{tabular}{|c|l|}
        %\hline
        %& & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of non-zero\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the matrice(s) in bytes\\
        $s_x$ & size of the {\tt x} vector(s) in bytes\\
        $s_y$ & size of the {\tt y} vector(s) in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred  \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred  \\
        $O$     & number of floating point operations \\
        $I_b$   & maximal computational intensity\\
        $I_w$   & minimum computational intensity\\
        \hline
      \end{tabular}
    }
  \end{center}
  \caption{Notations}
  \label{tab:not}
\end{figure}

Each vector in the problem is of dimension $n_r$ and each dimension
tasks $b_x$ bytes. There are $n_v$ {\tt x} vectors and $n_v n_m$ {\tt
  y} vectors, which lead to the size of the {\tt x} and {\tt y} vectors:
$$s_x = n_v b_x n_r$$ $$s_y = n_v n_m b_x n_r$$

The matrix is composed of $n_r$ rows and columns with $n_z$ non-zeros
per row leading to a total of $$n_c = n_r n_z$$ non-zero entries in
the matrix. Each of these non-zero entries has one index of size $b_i$
and $n_m$ values of size $b_x$. The matrices have a total size
of $$s_M = n_c (b_i + b_x n_m) = n_r n_z (b_i + b_x n_m)$$

If we assume an algorithm where the rows are processed one after
the other, the amount of memory written is precisely the
size of the {\tt y} vector. (This assumption removes the possibility of
cache partitioning techniques.) $$b_{wT} = s_y = n_v n_m b_x n_r$$

The amount of data read from memory depends highly on both the algorithm's
execution path, and on %algorithm execute,
how the matrix is structured. But in the best case
both the matrix $A$ %is read a single time
and the source vector {\tt x} are read once
%read a single time
from the main memory. (We assume that all the elements of $\tt x$ are involved
in the SpMV.)  %is used is reasonnable since only oddly shaped matrix will
%left a dimension of the vector unused.)
Thus, $$b_{rT} = s_M + s_x = n_r n_z
(b_i + b_x n_m) + n_v b_x n_r$$
 $$b_T = b_{rT} + b_{wT} =  n_r n_z (b_i + b_x n_m) + n_v b_x n_r (1 + n_m)$$

Notice that there is no reason for a piece of the matrix to be read
multiple times. But assuming that each element of the {\tt x} vector
is read a single time is a strong assumption. If using a single core,
it assumes that either the cache of the architecture can store the full
{\tt x} vector or that the matrix is sufficiently well structured
causing no cache trashing. If using multiple core, this assumes that
no element of the {\tt x} vectors will be used by multiple
cores. \cite{Saule12} showed that there is very little cache trashing
in practice, but it showed that having elements of the vectors used by
multiple cores can have a significant impact on the performance
(growing with $n_v$).

On the other hand, in the worst case, every time the {\tt x} vector is
accessed, the value needs to be transfered from memory again. So in
total, there are as many transfers as the number of non-zeros in the
matrix. Note however that most architectures cannot read memory a single
byte at a time. Instead, a minimum number of bytes,equal to the size o a cacheline,
are transfered at once.
%multiple bytes, with each chunk equal to the data do not allow to read the memory
%one byte at a time but always in a set of multiple bytes. On most
%cached architecture, that
%the minimum size corresponds to the size $cl$ of
%a cacheline (typically 64 bytes).
When there are multiple
vectors, each non-zero element uses $n_v$ consecutive entries. The
worst case number of bytes read and transfered are
$$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}} $$
$$B_T = n_v n_m b_x n_r + n_r n_z \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)$$

In SpMV, each non-zero of the matrix requires two floating point
operations: one for performing the multiplication and one for
accumulating the result row-wise. Here we are dealing with $n_v n_m$
simultaneous SpMVs and the number of floating point operations is
$$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$$

The computation intensity is the amount of computations performed per
byte transfered. In the worst case and in the best case, we have
$$I_b = \frac{O}{b_T} = \frac{2 n_v n_m n_z}{n_z (b_i + b_x n_m) + n_v b_x (1 + n_m)}$$
$$I_w = \frac{O}{B_T} = \frac{2 n_v n_m n_z}{n_v n_m b_x + n_z \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)}$$

In the special case $n_v=n_m=1$ in single precision
$$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
%$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
$$I_w = \frac{n_z }{2  + (4+cl/2)  n_z} \approx \frac{1}{4+cl/2}$$

In the special case $n_v=n_m=4$ in single precision
$$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
%$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
$$I_w = \frac{32 n_z }{64 + (20+cl) n_z} \approx \frac{32}{84} = \frac{8}{21} $$
The above formulas have been applied to MIC, which has a cache size $cl=64$ bytes.

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|r|r|}
    \hline
                           & v1m1  & v1m4  & v4m1   & v4m4  \\
    \hline
    Best Single Precision  & 36.36 & 58.18 & 133.33 & 213.33\\
    Worst Single Precision &  4.15 & 14.20 &  16.55 &  55.81\\
    Best Double Precision  & 24.00 & 32.21 &  85.71 & 117.07\\
    Worst Double Precision &  3.93 & 11.88 &  15.58 &  46.15\\
    \hline
  \end{tabular}
  \caption{Estimation of performance in GFlop/s at 150GB/s for a
    matrix of $n_z = 32$}
\end{table}

\begin{figure*}
  \centering
  \subfigure[Absolute performance ]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops_peak_perf}}
%
  \subfigure[Relative performance to base case using one matrix and one
    vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}

  \caption{Estimation of the reachable performance one can
      achieve using a device with a memory to computational units of
      150GB/s when varying the number of vectors and matrices}
  \label{fig:perf_predict}
\end{figure*}

\todo{conclude on the evaluation using the figure/table. Note that we
  base prediction on 150GB/s which is not the peak of saule12. Note
  that if you are in best in v1m1 you might go to bad v4m4, but if you
  are in bad v1m1, you will essentially stay there.}

\section{Efficient Implementation on the Intel Xeon Phi processor}
\label{sec:impl}

\subsection{Knights Corner}

In this work, we use a pre-release KNC card SE10P. The card has 8
memory controllers where each of them can execute 5.5 billion
transactions per second and has two 32-bit channels. That is the
architecture can achieve a total bandwidth of 352GB/s aggregated
across all the memory controllers. There are 61 cores clocked at
1.05GHz. The cores’ memory interface are 32-bit wide with two channels
and the total bandwidth is 8.4GB/s per core. Thus, the cores should be
able to consume 512.4GB/s at most. However, the bandwidth between
the cores and the memory controllers is limited by the ring network
which connects them and theoretically supports at most 220GB/s.

Each core in the architecture has a 32kB L1 data cache, a 32kB L1
instruction cache, and a 512kB L2 cache. The architecture of a core is
based on the Pentium architecture: though its design has been updated
to 64-bit. A core can hold 4 hardware contexts at any time. And at
each clock cycle, instructions from a single thread are executed. Due
to the hardware constraints and to overlap latency, a core never
executes two instructions from the same hardware context
consecutively. In other words, if a program only uses one thread, half
of the clock cycles are wasted. Since there are 4 hardware contexts
available, the instructions from a single thread are executed
in-order. As in the Pentium architecture, a core has two di↵erent
concur- rent instruction pipelines (called U-pipe and V-pipe) which
allow the execution of two instructions per cycle. However, some
instructions are not available on both pipelines: only one vector or
floating point instruction can be executed at each cycle, but two ALU
instructions can be executed in the same cycle.

Most of the performance of the architecture comes from the vector
processing unit. Each of Intel Xeon Phi’s cores has 32 512-bit SIMD
registers which can be used for double or single precision, that is,
either as a vector of 8 64-bit values or as a vector of 16 32-bit
values, respectively. The vector processing unit can perform many
basic instructions, such as addition or division, and mathematical
operations, such as sine and sqrt, allowing to reach 8 double
precision operations per cycle (16 single precision). The unit also
sup- ports Fused Multiply-Add (FMA) operations which are typically
counted as two operations for benchmarking purposes. Therefore, the
peak performance of the SE10P card is 1.0248 Tflop/s in double
precision (2.0496 Tflop/s in single precision) and half without FMA.

\todo{rewrite that to make sure it fits the description of the card
  used in the experiments. This text with copy/pasted from the PPAM
  paper}

\subsection{Bringing the data in vector register}

Registers on the MIC are 512 bits wide, and can store 16 floats or 8
doubles. It is through these registers that the \#SIMD pragma is able
to achieve vectorization. (In comparison to OpenCL or CUDA, one can
think of these registers as a warp.)  We use the following vector
operations in our code, which we'll explain in the text.

\def\loadps{\ttt{\_mm512\_load\_ps}}
\def\loadds{\ttt{\_mm512\_load\_ds}}
\def\fmadps{\ttt{\_mm512\_fmad\_ps}}
\def\gatherps{\ttt{\_mm512\_i32gather\_ps}}
\def\swizzleps{\ttt{\_mm512\_swizzle\_ps}}
\def\storenrngops{\ttt{\_mm512\_storenrngo\_ps}}
\def\castsi{\ttt{\_mm512\_castsi512\_ps}}
\def\permute{\ttt{\_mm512\_permute4f128\_epi32}}
\def\intmask{\ttt{\_mm512\_int2mask}}
\def\loadunpack{\ttt{\_mm512\_mask\_loadunpacklo\_epi32}}
\def\castitops{\ttt{\_mm512\_castsi512\_ps}}
\def\castpstoi{\ttt{\_mm512\_castps\_si512\_ps}}

\begin{figure*}
  %
  \begin{center}
    \begin{tabular}{|l|l|}
      \hline
      \loadps &  load 16 consecutive floats to register\\
      \fmadps &  multiply/add of 16 floats\\
      \gatherps &  gather 16, possibly disconnected floats from memory\\
      \swizzleps &  reorder a channel, and duplicate across channels\\
      \storenrngops &  store 16 floats to memory without register rewrite or reordering\\
      \castpstoi & reinterpret 16 floats as integers\\
      \castitops & reinterpret 16 integers as floats\\
      \permute &  permute channels; do not change individual channels\\
      \intmask &  \\
      \loadunpack &  \\
      \hline
    \end{tabular}
  \end{center}
  %
  \caption{List of MIC vectorial instructions used}
\end{figure*}
In the vector instructions, \ttt{ps} refers to a float (4 bytes),
while \ttt{epi32} refers to a 4-byte integer. Although we do not
discuss double precision in this paper, \ttt{ds} refers to an 8 byte
double precision real number (i.e., \loadds).  Each vector register is
broken up into four channels of 128 bits each. It is possible to
interchange these channels at a cost of a "few" \qes{exact numbers?}
cycles, and it is also possible to execute swizzle operations within a
channel. For example, if the 128 bytes of each channel are labelled as
$ABCD$, the vector instruction \ttt{\_mm512\_swizzle\_ps(v,
  \_MM\_SWIZ\_REG\_AAAA)} replaces each channel by $AAAA$.

Here is a function that reads in four floats (a,b,c,d) and creates the 16-float vector dddd,cccc,bbbb,aaaa:
\begin{verbatim}
__m512 read_aaaa(float* a)
{
    int int_mask_lo = (1 << 0) + (1 << 4) + (1 << 8) + (1 << 12);
    __mmask16 mask_lo = _mm512_int2mask(int_mask_lo);
    __m512 v1_old;
    v1_old = _mm512_setzero_ps();
    v1_old = _mm512_mask_loadunpacklo_ps(v1_old, mask_lo, a);
    v1_old = _mm512_mask_loadunpackhi_ps(v1_old, mask_lo, a);
    v1_old = _mm512_swizzle_ps(v1_old, _MM_SWIZ_REG_AAAA);
    return v1_old;
}
\end{verbatim}

Here is a function that takes four floats from float register \ttt{v1}, and
places them in each of the four lanes. Because the permulation function only
operates on 4-byte integers, it is necessary to convert (in place) each float
to an integer (the bit structure is not changed), permute the lanes and cast
the integers back to floats (with no modification of the bit structure)
\begin{verbatim}
__m512 permute(__m512 v1, _MM_PERM_ENUM perm)
{
    __m512i vi = _mm512_castps_si512(v1);
    vi = _mm512_permute4f128_epi32(vi, perm);
    v1 = _mm512_castsi512_ps(vi);
    return v1;
}
\end{verbatim}

\newfig{figures/swizzling.pdf}{\wide}{Duplication of the first four
  bytes of each channel across the entire channel.}{fig:swizzling}

\newfig{figures/channel_permutation.pdf}{\wide}{Channel Permutations
  on the MIC}{fig:permutation}

\newfig{figures/tensor_product.pdf}{\wide}{Tensor scalar product using
  channels and swizzling.}{fig:tensor_product}

\section{Experimental Validation}
\label{sec:expe}

\subsection{Instances}

\todo{How were instances generated. Detail the different type of
  instances and the rational for trying them. Detail how the realistic
  instances were generated. Explain matrix ordering (Kutill McKee).}

\subsection{Bandwidth}

\begin{figure*}

  \subfigure[Bandwidth performance under idealized conditions as a
    function of matrix row size. Entries with "cpp" denote cases where
    coding was performed without MIC vector
    instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_readwrite.pdf}\label{fig:band_rw}}
%
  \subfigure[Bandwidth performance under
  idealized conditions as a function of matrix row size. Entries with
  "cpp" denote cases where coding was performed without MIC vector
  instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_gather.pdf}\label{fig:band_gather}}

  \subfigure[Bandwidth performance under idealized conditions as a
    function of matrix row size. Entries with "cpp" denote cases where
    coding was performed without MIC vector instructions. The greater
    speed of the cpp version is obtained throught the use of
    \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}. All memory
    is aligned on 64
    bytes.]{\includegraphics[width=.49\linewidth]{figures/test3_gather.pdf}\label{fig:band_gather_ivdep}}
%
  \subfigure[Bandwidth performance on the host under idealized
    conditions as a function of matrix row size. Entries with "cpp"
    denote cases where coding was performed without AVX vector
    instructions. The speed on the CPU matches the speed with AVX
    instructions. All memory is aligned on 32
    bytes.]{\includegraphics[width=.49\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:read_write}}

\caption{Bandwidth Analysis.\todo{not 100\% sure what this plots are.}
  \todo{can we harmozie y-axis?} \todo{ylabel should read "Bandwidth (GB/s)".}
  \todo{remove title.}}

\end{figure*}

\subsection{Computations}

\begin{figure*}
  \centering

  \subfigure[Performance of $y=Ax$ on the MIC. Squares: base 1/1 case,
    solid circles: 4/4 case implemented in C++, solid triangles: 4/4
    case implemented with MIC vector instructions. $-O3$ compilation
    options. Each group consists of four cases: grids of $64^3$ and
    $96^3$ with and without Reverse Cuthill
    McKee.\todo{I don't think this is supercompact but RBF 3D.}]{\includegraphics[width=.49\linewidth]{figures/mic_performance_nb_threads.pdf}\label{}}
%
  \subfigure[performance on the host on a 64x64x64 matrix. Manual
    vectoization brings significant improvement there as
    well.]{\includegraphics[width=.49\linewidth]{figures/64x64x64.pdf}\label{}}

  \caption{Vectorization matters in practice. Differences in BW
    translate in different performance.\todo{this two plots should ``correlate'' }}
\end{figure*}

\begin{figure*}[t]
  \centering
  \subfigure[Supercompact\todo{is the title? this looks like super compact.}]{\includegraphics[width=.49\linewidth]{figures/supercompact_max_perf.pdf}}
  \subfigure[Compact]{\includegraphics[width=.49\linewidth]{figures/compact_max_perf.pdf}}

  \subfigure[Random]{\includegraphics[width=.49\linewidth]{figures/random_max_perf.pdf}}
  \subfigure[RBF 3D]{\includegraphics[width=.49\linewidth]{figures/rbf_max_perf_run1.pdf}}
  \caption{Performance of "best code" accross instance types. \todo{Why stop at 1.6M
      rows. Is that because of memory limitation? I guess that's what
      it is}\todo{can we remove the title?}\todo{can we harmonize y-axis?}}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}

  \caption{Performance on a MIC of derivative computation for RBFFD
    with 884,736 nodes distributed quasi-randomly in a cube. Single
    derivative of a single function (base case, triangle) and four
    derivatives of four functions (multi case, circles). No bandwidth
    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
\end{figure}

Derivatives in an RBFFD formulation are expressed as a sparse
matrix/vector multiplication (SpMV). Using the full 61 nodes of a MIC
processor, and all four threads of each node, we achieve lowly 18
Gflops when calculating a single vector of a single
function. Bandwidth reduction has minimal effect since the cache is
large enough to store both the derivative matrix and the solution
vector. We increase the number of computations per byte transferred by
calculating four different derivatives of four different
functions. With bandwidth reduction, we speed the calculation by a
factor of 8, and achieve upwards of 140 Gflops.

\section{Conclusion}
\label{sec:ccl}
In the previous sections, we have explored in the practical implementation on a MIC architecture of multiple derivative operators acting on multiple vectors within the context of RBF-RD. Each derivative has an associated sparse matrix with a fixed number
of non-zeros on each row. While computing a single derivative of multiple functions is rather common (vectorization occurs over the vectors), we accelerate the algorithm further by considering multiple derivatives with corresponding
matrices with identical adjacency graphs. We specialize to four matrices and four derivatives, with 16 outputs,
computed as a sum of outer products.

Using a matrix with $64^3$ (or $96^3$ rows, we achieve a speedup of xxx relative to the base case of one derivative and one function (1/1/ case), releative to a potential speedup of 16. We have achieved a speed 100 Gflops on 60 nodes using 240 threads, or a speed of 61\% of maximum possible performance. The maximum possible speed for the 1/1 case is around 40 Gflops, of which we achieve 20 Gflops. On matrices that exceed cache, a reverse Cuthill-McKee is applied
to reduce the bandwidth, with an associated speedup of 30\% up to 135 Gflops. Our optimal implementation makes
use of the AVX instruction set, which makes use of swizzling and channel swapping operations for an extremely
efficient tensor product implementation. A straightforward implementation without AVX run at around 35 Gflops,
or 16\% of peak performance. RCM has no effect because the number of floating point operations is excessive (the
algorithm does not fill the vector registers to capacity).

In a future work, we will examine the effect of larger stencil sizes, double precision, and the actual cost
of computing the derivatives within a fluid simulation using RBF-FD.

\section*{Acknowledgment}
The first and last authors acknowledge NSF funding under NSF grant DMS-\#0934331 (FSU). NCAR is sponsored by the National Science Foundation. N. Flyer acknowledges support of NSF grants DMS-0934317.

\bibliographystyle{abbrv}
%\bibliographystyle{plain} %{IEEEabrv,paper.bib}
\bibliography{paper, bollig}
\end{document}

### 2014-01-15 acm/RBF_Section.tex
New figure from Natasha + intro text from Natasha.

@@
 
-\begin{figure}%[htb]
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/RBFStencil}\label{fig:stencil}
-\caption{An example of an RBF-FD stencil of size $n=21$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square.
+\includegraphics[width=0.8\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
+\caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square.
 In nonsparse format, the DM  is $1024\times1024$ with 21 nonzero entries per row.}

### 2014-01-15 acm/ieee_paper.tex
New figure from Natasha + intro text from Natasha.

@@
 
-An other solution, and the one we pursue in this paper, is to compute
-multiple SpMVs at once on matrices that have the same sparsity
-patterns. Obviously not all the applications have such a property. But
-important classes of applications such as graph
-recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of
-derivatives of Radial Basis Functions (RBF)~\cite{FLBWSC12} can use multiple
-SpMVs simultaneously. In this paper in particular, we investigate the
-case of the derivative of RBFs where four derivatives of four
-different functions are expressed as the multiplication of four vectors
-by four matrices with identical sparsity patterns leading to the
-simultaneous execution of 16 SpMVs at a time. Using multiple vectors
-at a time has been investigated before to densify the
-computations~\cite{Im01} but this is the first time, to the 
-best of our knowledge, that the
-densification comes from adding both vectors and matrices.
+Another solution, pursued in this paper, is to compute multiple SpMVs at once on matrices that have the same sparsity pattern.
+Obviously not all the applications have such a property.
+However, important classes of applications such as graph recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of derivatives for solving systems of PDEs using Radial Basis Function-generated Finite Differences (RBF-FD)~\cite{FLBWSC12} can use multiple SpMVs simultaneously.
+Since RBF-FD is a mesh-less method, easily handling irregular geometries and local refinement with algorithmic complexity not increasing with dimension as well as high-order derivative approximations, they are rapidly gaining ground in science and engineering modeling communities \cite{Bayona13,CDNT,FoL11,FLBWSC12,SPLM}).
+As a result, it is of interest to develop an efficient implementation on novel computer platforms for the calculation of derivatives within the context of RBF-FD.
+This paper investigates the case of approximating four different derivatives, using RBF-FD, for four different functions (a common scenario in 3D fluid dynamics modeling).
+Since the RBF-FD differentiation matrices approximating the four different derivative operators have an identical sparsity pattern, this leads to the simultaneous execution of 16 SpMVs at a time.
+Using multiple vectors at a time has been investigated before to densify the computations~\cite{Im01} but, to the best of the authors’ knowledge, this is the first time that the densification comes from adding both vectors and matrices.
 
@@
 
-In Section~\ref{sec:rbf}, we present the computation of RBFs and how it
-can be expressed as sixteen multiplications of four vectors by four sparse matrices
-with identical sparsity patterns. Section~\ref{sec:model} presents an
+In Section~\ref{sec:rbf}, we introduce and further motivate the RBF-FD method, giving an example of how calculating the derivatives for a common system of PDEs in fluid dynamics can be expressed as sixteen multiplications of four vectors by four sparse matrices with identical sparsity patterns.
+Section~\ref{sec:model} presents an
 estimation of the instruction intensity of various forms of the
@@
 experimental results about the amount of bandwidth that can be achieved
-depending on how the SpMV kernel is written. It also provides the actual performance
-of the various kernels on multiple classes of matrices, some generated
-for purpose of analysis, and some extracted from an RBF application. A
-performance of 135~Gflop/s is achieved using RBF-generated matrices. Concluding
-remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1
+depending on how the SpMV kernel is written. 
+It also provides the actual performance of the various kernels on multiple classes of matrices, some generated for purpose of analysis, and some extracted from an application of RBF-FD.
+A performance of 135~Gflop/s is achieved using RBF-FD differentiation matrices.
+Concluding remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1
 

### 2014-01-15 acm/RBF_Section.tex
Fixed formatting issues.

@@
 For example, the full shallow water equations in a 3D Cartesian coordinate system for a rotating fluid are as follows:
+{
+\small
 \begin{eqnarray}
 \dfrac{\partial u}{\partial t} &=& - \left( u\dfrac{\partial u}{\partial x} + v\dfrac{\partial u}{\partial y} + w\dfrac{\partial u}{\partial z} + f(yw - zv) + g\dfrac{\partial h}{\partial x} \right) 
-\\
+\nonumber \\
 \dfrac{\partial v}{\partial t} &=& - \left( u\dfrac{\partial v}{\partial x} + v\dfrac{\partial v}{\partial y} + w\dfrac{\partial v}{\partial z} + f(zu - xw) + g\dfrac{\partial h}{\partial x} \right)
-\\
+\nonumber\\
 \dfrac{\partial w}{\partial t} &=& - \left( u\dfrac{\partial w}{\partial x} + v\dfrac{\partial w}{\partial y} + w\dfrac{\partial w}{\partial z} + f(xv - yu) + g\dfrac{\partial h}{\partial x} \right)
-\\
-\frac{\partial h}{\partial t}& =&-\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right) , \label{height}
+\nonumber\\
+\frac{\partial h}{\partial t}& =&-\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right) , \label{height} \nonumber
 \end{eqnarray}
-where $f$ is the Coriolis force, $\{u,v,w\}$ are the components of the velocity vector in the respective $\{x,y,z\}$ directions, and $h$ is the geopotential height (analogous to pressure).
+} where $f$ is the Coriolis force, $\{u,v,w\}$ are the components of the velocity vector in the respective $\{x,y,z\}$ directions, and $h$ is the geopotential height (analogous to pressure).
 Thus, we have four variables $\{u,v,w,h\}$ and three DMs, $D_x, D_y, D_z$ representing $\dfrac{\partial}{\partial x}, \dfrac{\partial}{\partial y},\dfrac{\partial w}{\partial z}$, respectively.
@@
 Furthermore, due to the properties just mentioned, multiple derivatives of a single function can be calculated rather than computing a derivative of multiple functions.
-Using the PDE system above as an example, the block structure of the SpMM for computing all derivatives needed can be seen in Figure \ref{fig:struct_comp}.
-\begin{figure}
+Using the PDE system above as an example, the block structure of the SpMM for computing all derivatives needed can be seen in Figure \ref{fig:struct_comp},
+\begin{figure}[h]
   \centering
+ {\small
   \[ \left( \begin{array}{cccc}
@@
   \begin{array}{c}
-    D_{x}\\
-    D_{y}\\
-    D_{z}\\
-    D_{hyp}
+    D_{x}\\ D_{y}\\ D_{z}\\ D_{hyp}
   \end{array}\right)
   \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) \]
+ }
+  \vspace{-1ex}
   \caption{Block structure of the computation.}

### 2014-01-15 acm/ieee_paper.tex
WIth new updates.

@@
 \def\es#1{{\blue{*** For Erik: #1 ***}}}
-\def\ge#1{{\red{*** For Gordon: #1 ***}}}
+\def\ge#1{{\red{#1}}}
+\def\NOTE#1{{\red{*** NOTEGordon: #1 ***}}}
 \def\ttt#1{{\tt #1}}
@@
 Finite-Difference (RBFFD). RBF methods express functions as a linear
-combination of radial basis functions on an arbitrary set of
+combination of \ge{spherically} symmetric basis functions on an arbitrary set of
 nodes. The Finite-Difference component expresses this combination over
@@
 format. We increase the number of operations relative to memory
-bandwidth by calculating four derivatives of four different functions,
+bandwidth by \ge{interleaving} the calculation of  four derivatives 
+of four different functions,
 or 16 different derivatives. We demonstrate a novel implementation on
@@
 \section{Introduction}
-The multiplication of sparse matrix by a dense vector (SpMV) is an
+The multiplication of a sparse matrix by a dense vector (SpMV) is an
 important kernel in many applied fields such as fluid dynamics 
@@
   conf/ipps/KreutzerHWFBB12,
-  journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
-challenge  to obtain good performance for SpMV is that
-the operations are conducted using memory location that are irregular. This makes the kernels mostly memory
-bound and there is a significant instruction overhead per flop.
+  journals/concurrency/VazquezFG11,kumar2012accelerating}. \ge{The main
+challenge  to obtain good performance for matrix vector multiplication in general, and 
+sparse matrix vector multiplication in particular, is the low ratio of floating point operations
+to memory bandwidth. When the matrix is not dense, the problem is exacerbated due to non-uniform 
+access patterns.}
+%{SpMV is that
+%the operations are conducted using memory locations that are irregularly spaced. 
+%This makes the kernels mostly memory bound
+% and there is a significant instruction overhead per flop.
+
+% REMOVED by GE
+%This makes the kernels strongly memory bound.  There is also a 
+%significant instruction overhead per flop.
+\NOTE{I rewrote last few sentences: write more precisely? Blas 2 is Matrix/vector, and that is still somewhat memory bound.}
 
@@
 cache or TLB~\cite{Nishtala07,Temam:1992:CBS:147877.148091,conf/ppsc/Toledo97},
-unrolling~\cite{Mellor-Crummey04} have impacts which are very
+unrolling~\cite{Mellor-Crummey04} have impacts that are very
 dependent on the matrix and overall do not lead to dramatic
@@
 
-Another solution, pursued in this paper, is to compute multiple SpMVs at once on matrices that have the same sparsity pattern.
-Obviously not all the applications have such a property.
-However, important classes of applications such as graph recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of derivatives for solving systems of PDEs using Radial Basis Function-generated Finite Differences (RBF-FD)~\cite{FLBWSC12} can use multiple SpMVs simultaneously.
-Since RBF-FD is a mesh-less method, easily handling irregular geometries and local refinement with algorithmic complexity not increasing with dimension as well as high-order derivative approximations, they are rapidly gaining ground in science and engineering modeling communities \cite{Bayona13,CDNT,FoL11,FLBWSC12,SPLM}).
-As a result, it is of interest to develop an efficient implementation on novel computer platforms for the calculation of derivatives within the context of RBF-FD.
-This paper investigates the case of approximating four different derivatives, using RBF-FD, for four different functions (a common scenario in 3D fluid dynamics modeling).
-Since the RBF-FD differentiation matrices approximating the four different derivative operators have an identical sparsity pattern, this leads to the simultaneous execution of 16 SpMVs at a time.
+Another solution, pursued in this paper, \ge{is to compute multiple SpMVs using matrices with identical spasity patterns, but with different matrix elements.}  Obviously not all the applications have such a property.
+However, important classes of applications such as graph recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of derivatives for solving systems of PDEs using Radial Basis Function-generated Finite Differences (RBF-FD)~\cite{FLBWSC12} \ge{fall into the category of applications that require multiple SpMVs simultaneously.} RBF-FD is a meshless method, which easily handles irregular geometries and local refinement with algorithmic complexity independent of dimensionality,
+and which can produce high-order derivative approximations. These methods are rapidly gaining 
+ground in science and engineering modeling communities \cite{Bayona13,CDNT,FoL11,FLBWSC12,SPLM}). 
+As a result, it is of interest to develop an efficient implementation on novel computer plat
+forms for the calculation of derivatives within the context of RBF-FD, which thus far are responsible 
+for the bulk of computer resources. \ge{Typical equations solved by RBF-FD depend on multiple derivatives
+of multiple functions. Rather than compute each derivative individually, this paper investigates 
+how to compute all derivatives simultaneously in an interleaved fashion. Specifically, 
+we calculate four different derivatives (corresponding to four different sparse matrices with identical 
+sparsity pattern) of four different functions
+%is an interThis paper investigates the possibility  to increase the ratio  case of approximating 
+%four different derivatives, using RBF-FD, for four different functions 
+(a common scenario in 3D fluid dynamics modeling) for a total of 16 derivatives. }
+%Since the RBF-FD differentiation matrices approximating the four different derivative operators have an identical sparsity pattern, this leads to the simultaneous execution of 16 SpMVs at a time.
 Using multiple vectors at a time has been investigated before to densify the computations~\cite{Im01} but, to the best of the authors’ knowledge, this is the first time that the densification comes from adding both vectors and matrices.
@@
 
-In Section~\ref{sec:rbf}, we introduce and further motivate the RBF-FD method, giving an example of how calculating the derivatives for a common system of PDEs in fluid dynamics can be expressed as sixteen multiplications of four vectors by four sparse matrices with identical sparsity patterns.
+In Section~\ref{sec:rbf}, we introduce and further motivate the RBF-FD method, giving 
+an example of how calculating the derivatives for a common system of PDEs in fluid dynamics 
+can be expressed as sixteen multiplications of four vectors by four sparse matrices with 
+identical sparsity patterns.
+
 Section~\ref{sec:model} presents an
@@
 approximately 10\% of the available flop/s of a Xeon Phi
-coprocessor. Therefore, it is necessary to have an implementation that
+coprocessor. \ge{Therefore, it is necessary to have an implementation that
 performs the computation in as few instructions as
-possible. We describe in Section~\ref{sec:impl} the details of the MIC
+possible.} \NOTE{I DO NOT FOLLOW the Therefore in the previous sentence.}
+We describe in Section~\ref{sec:impl} the details of the MIC
 architecture and how to use specialized load, store, swizzle and
-permutation instructions to efficiently bring the data in the vector
+permutation instructions to efficiently bring the data from memory into the vector
 registers to be processed. Section~\ref{sec:expe} gives some
@@
 depending on how the SpMV kernel is written. 
-It also provides the actual performance of the various kernels on multiple classes of matrices, some generated for purpose of analysis, and some extracted from an application of RBF-FD.
-A performance of 135~Gflop/s is achieved using RBF-FD differentiation matrices.
-Concluding remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1
+It also provides the actual performance of the various kernels on multiple classes of matrices, 
+some generated for purpose of analysis, and some extracted from an application of RBF-FD. 
+A performance of 135~Gflop/s is achieved using RBF-FD differentiation matrices. Concluding remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1
 

### 2014-01-15 acm/ieee_paper.tex
More changes.

@@
 \def\es#1{{\blue{*** For Erik: #1 ***}}}
-\def\ge#1{{\red{#1}}}
-\def\NOTE#1{{\red{*** NOTEGordon: #1 ***}}}
+\def\ge#1{{\blue{#1}}}
+\def\NOTE#1{{\red{NOTE: #1}}}
 \def\ttt#1{{\tt #1}}
@@
 possible.} \NOTE{I DO NOT FOLLOW the Therefore in the previous sentence.}
-We describe in Section~\ref{sec:impl} the details of the MIC
+\ge{We describe in Section~\ref{sec:impl} the details of the MIC
 architecture and how to use specialized load, store, swizzle and
 permutation instructions to efficiently bring the data from memory into the vector
-registers to be processed. Section~\ref{sec:expe} gives some
-experimental results about the amount of bandwidth that can be achieved
-depending on how the SpMV kernel is written. 
+registers to be processed. Section~\ref{sec:expe} presents results on 
+memory access and computational speed and their relation to choices made in the 
+SpMV kernels.}
+%experimental results  discussing the relationship between the structure of $A$ and 
+%the ; specifcally on the amount of bandwidth that can be achieved
+%depending on how the SpMV kernel is written. 
 It also provides the actual performance of the various kernels on multiple classes of matrices, 

### 2014-01-15 acm/RBF_Section.tex
More changes.

@@
 To obtain a RBF derivative approximation at a node location that will result in a sparse differentiation matrix (DM), the RBF-FD method has been developed over the last decade \cite{TAI1,TAI2,SDY02,WrFo06}.
-RBF-FD are conceptually similar to standard finite differences (FD).
-However, the differentiation weights for approximating a derivative at a given node (forming one row of the RBF-FD DM) enforce that the linear combination of function values at the $n$ nearest node locations (a stencil) with $n \in N$ be exact for RBFs centered at each of the $n$ nodes rather than polynomials.
-The result is a sparse matrix with $nN$ nonzero entries, where $n<<N$.
+RBF-FD is conceptually similar to standard finite differences (FD).
+However, the differentiation weights for approximating a derivative at a given node (forming one row of the RBF-FD derivative matrix) are derived from a function approxiation within the stencil formed by a linear combination of RBFs centered at the different points of the stencil.
+In general, $n<<N$.
+%that the linear combination of function values at the $n$ nearest node locations (a stencil) with $n \in N$ be exact for RBFs centered at each of the $n$ nodes rather than polynomials. 
+The resulting derivative matrix (DM) is sparse matrix with $nN$ nonzero entries.
 As in FD methods, the order of the method increases with stencil size.
-An illustration of an RBF stencil on a sphere is given in Figure \ref{fig:stencil}.
+An illustration of an RBF 32 node stencil on a sphere,  is given in Figure \ref{fig:stencil}.
 
@@
 Thus, we have four variables $\{u,v,w,h\}$ and three DMs, $D_x, D_y, D_z$ representing $\dfrac{\partial}{\partial x}, \dfrac{\partial}{\partial y},\dfrac{\partial w}{\partial z}$, respectively.
-However, in order to stably time step the equations with an explicit time-stepping scheme, most numerical methods (including RBF-FD) require a fourth operator, hyperviscosity, to be added to the each equation.
+Moreover, in order to stably time step the equations with an explicit time-stepping scheme, most numerical methods (including RBF-FD) require a fourth operator, hyperviscosity, to be added to the each equation.
 Hyperviscosity takes the form of $\triangle^p$ ($\triangle$ denotes the Laplacian operator), where $p\geq3$, which is approximated by the matrix $D_{hyp}$.
 Therefore, in order to solve this system of PDEs, the application of 4 DM to 4 unknowns, i.e.
-16 SpMV is required every time step.
+16 SpMV is required at every time step.
 Due to the fact that RBFs are independent of a coordinate system, taking into account only distances to nearest neighbors, the DMs have the following properties:
 \begin{itemize}
-\item For a given $n$, regardless of what operator is approximated, e.g. $\dfrac{\partial}{\partial x}$ or the Laplacian, RBF-FD DMs will have the same number of nonzeros per row.
-\item Also, for a given $n$ and $N$, regardless of what operator is approximated, the DMs will have the same sparsity pattern.
-Examples of an RBF-FD DM can be seen in Figure \ref{fig:spy_plots}d and e.
+\item For a given $n$, regardless of the operator approximated, e.g. $\dfrac{\partial}{\partial x}$ or the Laplacian, RBF-FD DMs will have the same number of nonzeros per row.
+\item For a given $n$ and $N$, regardless of the operator approximated, the DMs will have the same sparsity pattern. Examples of an RBF-FD DM can be seen in Figures \ref{fig:spy_plots}d and e.
 \end{itemize}
@@
 misses by vectorizing over the multiple source vectors.
-Furthermore, due to the properties just mentioned, multiple derivatives of a single function can be calculated rather than computing a derivative of multiple functions.
+Furthermore, due to the properties just mentioned, multiple derivatives of a single function can be calculated rather than computing a single derivative of multiple functions.
 Using the PDE system above as an example, the block structure of the SpMM for computing all derivatives needed can be seen in Figure \ref{fig:struct_comp},
@@
 where $\{\underline{u},\underline{v}, \underline{w}, \underline{h}\}$ are the functions values at all nodes of the corresponding variables with the left hand side being the resulting derivative approximations.
-The increased memory bandwidth due to an increase in the number of derivative matrices is offset by better cache utilization, leading to an overall performance benefit.
+The increased memory bandwidth due to an increase in the number of derivative matrices is offset by improved cache utilization, leading to an overall performance benefit.
 In practice however, the number of vectors whose derivatives are required is limited, capping the performance of the matrix/vector multiplication.
@@
 
+\ge{We also note that the DMs are time-independent. They are computed once and used repeatedly to update the solution 
+in time. Thus, the cost to calculate the derivatives, although high, is amortized by the very long numerical simulations.}
+
 %\end{document}

### 2014-01-16 acm/RBF_Section.tex
small correction in RBF section.

@@
 RBF-FD is conceptually similar to standard finite differences (FD).
-However, the differentiation weights for approximating a derivative at a given node (forming one row of the RBF-FD derivative matrix) are derived from a function approxiation within the stencil formed by a linear combination of RBFs centered at the different points of the stencil.
-In general, $n<<N$.
-%that the linear combination of function values at the $n$ nearest node locations (a stencil) with $n \in N$ be exact for RBFs centered at each of the $n$ nodes rather than polynomials. 
-The resulting derivative matrix (DM) is sparse matrix with $nN$ nonzero entries.
-As in FD methods, the order of the method increases with stencil size.
-An illustration of an RBF 32 node stencil on a sphere,  is given in Figure \ref{fig:stencil}.
+However, the differentiation weights for approximating a derivative at a given node (forming one row of the RBF-FD derivative matrix (DM)) enforce that the linear combination of function values at the $n << N$ nearest node locations (a stencil) be exact for RBFs centered at each of the $n$ nodes rather than polynomials.
+The result is a sparse matrix with only $n$ entries in each of its $N$ rows, resulting in a total of $nN$ nonzero entries.
 
@@
 \caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square.
-In nonsparse format, the DM  is $1024\times1024$ with 21 nonzero entries per row.}
+In nonsparse format, the DM  is $1024\times1024$ with 32 nonzero entries per row.}
 \end{figure}

### 2014-01-17 acm/ieee_paper.tex
Incorporated Evan's changes. Blue is stuff I change (most of the stuff I cahnged). Red are questions.

@@
 channel interchange features. We present benchmarks on a real data
-set that show an almost eightfold increase in speed compared to
+set that show an almost \ge{sevenfold} \note{eightfold; there were inconsistencies in the ppaer.} increase in speed compared to
 efficient implementations of a single derivative, reaching a
@@
 ground in science and engineering modeling communities \cite{Bayona13,CDNT,FoL11,FLBWSC12,SPLM}). 
-As a result, it is of interest to develop an efficient implementation on novel computer plat
-forms for the calculation of derivatives within the context of RBF-FD, which thus far are responsible 
-for the bulk of computer resources. \ge{Typical equations solved by RBF-FD depend on multiple derivatives
+As a result, it is of interest to develop an efficient implementation on novel computer platforms for the calculation of derivatives within the context of RBF-FD; derivative calculations that account for the bulk of computer resources when running a numerical simulation implemented with RBF-FD.
+\ge{Typical equations solved by RBF-FD depend on multiple derivatives
 of multiple functions. Rather than compute each derivative individually, this paper investigates 
@@
 estimation of the instruction intensity of various forms of the
-computations. We show that a sixfold improvement can be
+computations. We show that a \ge{sevenfold} \NOTE{sixfold was inconsistent with eighfold in the abstract.} improvement can be
 expected when computing the sixteen multiplications simultaneously to
@@
 performs the computation in as few instructions as
-possible.} \NOTE{I DO NOT FOLLOW the Therefore in the previous sentence.}
+possible.} \NOTE{I DO NOT FOLLOW the Therefore in the previous sentence. Evan has the same concern.}
 \ge{We describe in Section~\ref{sec:impl} the details of the MIC
@@
 based on the traditional Pentium architecture, which has been extended to
-64-bit. A core can hold four hardware contexts at any given time (240 threads in total). At each
+64-bit with hyperthreading. A core can hold four hardware contexts at any given time (240 threads in total). At each
 clock cycle, instructions from a single thread are executed. Due to
@@
 \subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{figures/code.pdf}}%
-\subfigure[Content of the vector registers]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
+\subfigure[Content of the vector registers \ge{in the source code in (a)}.]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
 
@@
 computed. Once all the nonzeros of the row have been processed, the
-exact $y_i$ value can be sent to memory using a the No Read No Global
-Order hint to optimize write traffic on the memory bus.
+exact $y_i$ value can be sent to memory using a the "No Read No Global
+Order" hint to optimize write traffic on the memory bus.
 
@@
 This difference is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction, 
-which bypasses the Read For Ownership protocol and allows the writes
+which bypasses the \ge{"Read For Ownership"} protocol and allows the writes
 to be performed in any order. The compiler can not use this
@@
 are presented in Figure~\ref{fig:spy_plots}. The
-Supercompact matrices (Figure~\ref{fig:supercompact}) have been
+\ttt{Supercompact} matrices (Figure~\ref{fig:supercompact}) have been
 generated to only have nonzero elements in the first 32 columns of
@@
 from memory, and the expense associated with cache misses is removed from consideration. The cost associated with 
-{\tt A} remains. The Compact matrices
+{\tt A} remains. The \ttt{Compact} matrices
 (Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
@@
 applications that rely on sparse matrix vector multiplication. The nonzero
-elements of the Random
+elements of the \ttt{Random}
 matrices (Figure~\ref{fig:compact}) see their nonzero elements
@@
 The other type of matrix is used in derivative stencils in 3D
-RBFFD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-points
+RBFFD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-point
 stencil of a 3D $8^3$ grid).
@@
   \caption{Performance of $y=Ax$ for a RBF derivative stencil of a
-    $96^3$ 3D grid. Classical SpMV ($n_m=n_v=1$ given in blue) has low
+    $96^3$ 3D grid. Classical "one-matrix one-vector" SpMV ($n_m=n_v=1$ given in blue) has low
     performance. The manually vectorized code (red) is 3.5 times
@@
 
-To put things in perspective, we conducted some experiments on a
+To put things in perspective, we conducted similar experiments on a
 classical CPU architecture with two Intel Core i7 processors, each has eight
@@
 includes a number of swizzling and channel swapping operations, for an
-extremely efficient tensor product implementation. On a $96^3$ 3D
+extremely efficient tensor product implementation. \NOTE{From Evan: this is the first time $96^3$ is mentionned. Might wish to mention it earlier.} On a $96^3$ 3D
 grid, it reaches a performance of 140~Gflop/s which is 2.7 times

### 2014-01-17 acm/RBF_Section.tex

@@
 The parameter $\varepsilon$ controls the shape of the RBF.
-Note that the argument of an RBF is simply a scalar distance $r$, independent of any coordinate system or dimension.
+Note that the argument of an RBF is simply a scalar distance, $r$, independent of any coordinate system or dimension.
 As a result, nodes can be scattered as desired across complex physical domains with the implementation of the method being independent of dimensionality.
@@
 Thus, we have four variables $\{u,v,w,h\}$ and three DMs, $D_x, D_y, D_z$ representing $\dfrac{\partial}{\partial x}, \dfrac{\partial}{\partial y},\dfrac{\partial w}{\partial z}$, respectively.
-Moreover, in order to stably time step the equations with an explicit time-stepping scheme, most numerical methods (including RBF-FD) require a fourth operator, hyperviscosity, to be added to the each equation.
+Moreover, in order to stably time step the equations with an explicit time-stepping scheme, most numerical methods (including RBF-FD) require a fourth operator, hyperviscosity, to be added to each equation.
 Hyperviscosity takes the form of $\triangle^p$ ($\triangle$ denotes the Laplacian operator), where $p\geq3$, which is approximated by the matrix $D_{hyp}$.
@@
 Furthermore, due to the properties just mentioned, multiple derivatives of a single function can be calculated rather than computing a single derivative of multiple functions.
-Using the PDE system above as an example, the block structure of the SpMM for computing all derivatives needed can be seen in Figure \ref{fig:struct_comp},
-\begin{figure}[h]
-  \centering
- {\small
-  \[ \left( \begin{array}{cccc}
+Using the PDE system above as an example, the block structure of the SpMM for computing all derivatives needed is %can be seen in Equation \ref{fig:struct_comp},
+%\begin{figure}[h]
+\begin{equation}
+  %\centering
+ %{\small
+   \left( \begin{array}{cccc}
     \underline{u}_x     & \underline{v}_x     & \underline{w}_x     & \underline{h}_x \\
@@
   \end{array}\right)
-  \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) \]
- }
+  \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) 
   \vspace{-1ex}
-  \caption{Block structure of the computation.}
+  %\caption{Block structure of the computation.}
   \label{fig:struct_comp}
-\end{figure}
+%\end{figure}
+\end{equation}
 where $\{\underline{u},\underline{v}, \underline{w}, \underline{h}\}$ are the functions values at all nodes of the corresponding variables with the left hand side being the resulting derivative approximations.

### 2014-01-17 acm/RBF_Section.tex
sec 1 and 2

@@
 Thus, no mesh generation is needed and algorithmic complexity does not increase with dimension.
+\NOTE{While reading that paragraph, I was under the impression that the physical domain needs to be a sphere.}
 
@@
 
-\ge{We also note that the DMs are time-independent. They are computed once and used repeatedly to update the solution 
-in time. Thus, the cost to calculate the derivatives, although high, is amortized by the very long numerical simulations.}
+\ge{We also note that the DMs are time-independent. They are computed once \ee{at the beginning of the simulation} and used repeatedly to update the solution 
+in time \ee{(for each iteration of the simulation)}. Thus, the cost to calculate the derivatives, although high, is amortized by the very long numerical simulations.}
 

### 2014-01-17 acm/ieee_paper.tex
sec 1 and 2

@@
 \def\blue#1{\textbf{\textcolor{blue}{#1}}}
+\def\green#1{\textbf{\textcolor{green}{#1}}}
 \def\qes#1{{\blue{*** For Erik: #1 ***}}}
 \def\es#1{{\blue{*** For Erik: #1 ***}}}
+\def\ee#1{{\green{#1}}}
 \def\ge#1{{\blue{#1}}}
@@
 efficient implementations of a single derivative, reaching a
-performance of almost 140~Gflop/s. We explain the results through
+performance of almost 140~Gflop/s \ee{in single precision}. We explain the results through
 consideration of operation count versus memory bandwidth.
@@
 Another solution, pursued in this paper, \ge{is to compute multiple SpMVs using matrices with identical spasity patterns, but with different matrix elements.}  Obviously not all the applications have such a property.
-However, important classes of applications such as graph recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of derivatives for solving systems of PDEs using Radial Basis Function-generated Finite Differences (RBF-FD)~\cite{FLBWSC12} \ge{fall into the category of applications that require multiple SpMVs simultaneously.} RBF-FD is a meshless method, which easily handles irregular geometries and local refinement with algorithmic complexity independent of dimensionality,
+However, important classes of applications such as graph recommendation~\cite{Kucuktunc13-SNAM}, \NOTE{eigensolving does not match that description} and the computation of derivatives for solving systems of PDEs using Radial Basis Function-generated Finite Differences (RBF-FD)~\cite{FLBWSC12} \ge{fall into the category of applications that require multiple SpMVs simultaneously.} RBF-FD is a meshless method, which easily handles irregular geometries and local refinement with algorithmic complexity independent of dimensionality,
 and which can produce high-order derivative approximations. These methods are rapidly gaining 
-ground in science and engineering modeling communities \cite{Bayona13,CDNT,FoL11,FLBWSC12,SPLM}). 
+ground in science and engineering modeling communities \cite{Bayona13,CDNT,FoL11,FLBWSC12,SPLM}. 
 As a result, it is of interest to develop an efficient implementation on novel computer platforms for the calculation of derivatives within the context of RBF-FD; derivative calculations that account for the bulk of computer resources when running a numerical simulation implemented with RBF-FD.
@@
 approximately 10\% of the available flop/s of a Xeon Phi
-coprocessor. \ge{Therefore, it is necessary to have an implementation that
-performs the computation in as few instructions as
-possible.} \NOTE{I DO NOT FOLLOW the Therefore in the previous sentence. Evan has the same concern.}
+coprocessor.
+\ee{Provided the computation is mostly irregular, it is necessary to have an implementation that organizes the data in such a way it can perform a fully efficient Fused Multiply-Add every 10 cycles.} \NOTE{Rewrote, is it more clear?}
 \ge{We describe in Section~\ref{sec:impl} the details of the MIC
@@
 some generated for purpose of analysis, and some extracted from an application of RBF-FD. 
-A performance of 135~Gflop/s is achieved using RBF-FD differentiation matrices. Concluding remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1
+A performance of 135~Gflop/s in single precision is achieved using RBF-FD differentiation matrices, which is 3.75 imes better than the theoretical peak performance of an SpMV operation.
+Concluding remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1
 

### 2014-01-17 acm/ieee_paper.tex
sectino 4

@@
 low: this highlights that accessing the memory with a granularity of one cache
-line is the main problem faced when performing a standard SpMV computation.
+line is a main problem faced when performing a standard SpMV computation.
 
@@
 vector instructions. The code is given for reference in
-Figure~\ref{code:mat_mul}. (We will show in Section~\ref{sec:expe}
+Figure~\ref{code:mat_mul} along with a graphical depiction of the content of the registers. (We will show in Section~\ref{sec:expe}
 that an implementation that relies on compiler vectorization does not lead
@@
 
-The core of the technique is the use of swizzling and permutation
-features of the vectors in the Intel MIC architecture. A vector of 512
-bits is composed of four lanes (sometimes called channels) of 128
-bits, which are made of four segments of 32 bits. A swizzling
-operation allows the reorder of the elements within each lane (see how
-{\tt mat} is extracted from {\tt mat\_ent} in
-Figure~\ref{code:mat_mul}). On the other hand, a permutation reorders entire lanes,
-without changing their contents (see for instance how {\tt vect} is
-generated from {\tt all\_vect} in Figure~\ref{code:mat_mul}).  Both
-permutation and swizzling support common reordering and broadcast. In
-the code, the ``permute'' function needs to use the {\tt
+\ee{The core of the technique is to load the maximum amount of data
+  into large register and format the data for processing within these
+  registers. That way the amount of interaction with the memory is
+  minimized. The Intel MIC is more for these operations than classical
+  CPUs because it features larger vector registers and more elaborate
+  vector reorganization instructions than namely swizzling,
+  permutations and broadcasts. (The AVX instruction sets on classical
+  CPUs is more primitive in that aspect.)} A vector of 512 bits is
+composed of four lanes (sometimes called channels) of 128 bits, which
+are made of four segments of 32 bits. A swizzling operation allows the
+reorder of the elements within each lane (see how {\tt mat} is
+extracted from {\tt mat\_ent} in Figure~\ref{code:mat_mul}). On the
+other hand, a permutation reorders entire lanes, without changing
+their contents (see for instance how {\tt vect} is generated from {\tt
+  all\_vect} in Figure~\ref{code:mat_mul}).  Both permutation and
+swizzling support common reordering and broadcast. In the code, the
+``permute'' function needs to use the {\tt
   \_mm512\_permute4f128\_epi32} instruction which is only defined on
-integers and therefore in place typecasts are necessary to convert
-a vector to and from integer type.  (Note that we prefer swizzling to
+integers and therefore in place typecasts are necessary to convert a
+vector to and from integer type.  (Note that we prefer swizzling to
 permutation if possible because we believe it is cheaper. Many
@@
 Order" hint to optimize write traffic on the memory bus.
+
+\ee{The modelization of Section~\ref{sec:model} shows that the memory
+  transfer caps the computation at 210GFlop/s, which is 10\% of the
+  peak single precision of the architecture. To reach this value, the
+  cores need to execute one "useful" Fused Multiply-Add operation
+  every ten cycles. The inner loop of the kernel spends 7 vector
+  instructions to layout the data and then 3 per tensor product for a
+  total of 12 vector instructions. Most of these instruction have a
+  one cycle throughput cost except the gather operations which might
+  need to perform up to 16 memory accesses. There are also some scalar
+  operations but there are few of them and are executed on a different
+  pipeline than the vector operations. Overall, the inner loop can
+  emit the 4 Fused Multiply-Add in less than 40 cycles.}
 

### 2014-01-17 acm/ieee_paper.tex
edits are done

@@
 Figure~\ref{fig:bandwidth} shows a benchmark that computes the
-bandwidth achieved when reading and/or writing large arrays on both
-the MIC card and the CPU architecture. The read benchmark performs a
+bandwidth achieved when reading and/or writing large arrays on 
+the MIC card. The read benchmark performs a
 simple sum of the array. The write benchmark sets all the elemnts of
@@
 implementation (compiled with \mbox{-O3} with proper tagging to inform the
-compiler of memory alignments) and a simple implementation with
+compiler of memory alignments) and an implementation with
 explicit use of the vector registers.
@@
 
-To put things in perspective, we conducted similar experiments on a
-classical CPU architecture with two Intel Core i7 processors, each has eight
-cores that support AVX instructions. Both the compiler-generated and
-the manually-written vectorization using AVX instructions reach a
-(read) bandwidth of 70~GB/s. This indicates that the performance of
-the classical SpMV can not overcome 17~Gflop/s. A generic
-implementation of the multiplication of four vectors with four matrices on
-the {\tt RBF} matrix reaches a performance of 27~Gflop/s; while a
-manual implementation using AVX instructions reaches a performance of
-42~Gflop/s. This confirms that our method is sound and generalizes to
-architectures other than the Intel MIC architecture.
+\ee{The main idea of this paper, computing 16 derivatives at a time,
+  can be applied on different architectures such as multiprocessor
+  multicore or GPU. To illustrate this fact, we conducted similar
+  experiments on a classical CPU architecture with two Intel Core i7
+  (SandyBridge) processors, each has eight cores that support AVX
+  instructions.}  Both the compiler-generated and the manually-written
+vectorization using AVX instructions reach a (read) bandwidth of
+70~GB/s. This indicates that the performance of the classical SpMV can
+not overcome 17~Gflop/s. A generic implementation of the
+multiplication of four vectors with four matrices on the {\tt RBF}
+matrix reaches a performance of 27~Gflop/s. We also implemented an AVX
+vectorized version which is less compact than the one on MIC because
+of smaller vector registers and the abscence of a gather instruction
+(AVX2 on Haswell processor has that feature). It reaches a
+performance of 42~Gflop/s. This confirms that our method is sound and
+generalizes to architectures other than the Intel MIC architecture.
 
@@
 In future work, we will examine the effect of larger stencil sizes and
-double precision. The presented performance does not reach the predicted
-peak and we believe that common techniques from SpMV such as cache
-partitioning can be successfully applied. We will also integrate the
-techniques presented in this paper within a fluid simulation using
-RBFFD.
+double precision. The presented performance does not reach the
+predicted peak and we believe that common techniques from SpMV such as
+cache partitioning can be successfully applied. We will also integrate
+the techniques presented in this paper within a fluid simulation using
+RBFFD. We will also look developping a similar technique on more
+commonly available devices such as GPUs.
 

### 2014-01-17 acm/ieee_paper.tex
edits

@@
 multiplication of four vectors with four matrices on the {\tt RBF}
-matrix reaches a performance of 27~Gflop/s. We also implemented an AVX
+matrix reaches a performance of 27~Gflop/s. \ee{We also implemented an AVX
 vectorized version which is less compact than the one on MIC because
 of smaller vector registers and the abscence of a gather instruction
-(AVX2 on Haswell processor has that feature). It reaches a
+(Haswell processors support AVX2 which have that feature).} It reaches a
 performance of 42~Gflop/s. This confirms that our method is sound and
@@
 the techniques presented in this paper within a fluid simulation using
-RBFFD. We will also look developping a similar technique on more
-commonly available devices such as GPUs.
+RBFFD. \ee{We will also look developping a similar technique on more
+commonly available devices such as GPUs.}
 

### 2014-01-17 acm/ieee_paper.tex
minor

@@
 some generated for purpose of analysis, and some extracted from an application of RBF-FD. 
-A performance of 135~Gflop/s in single precision is achieved using RBF-FD differentiation matrices, which is 3.75 imes better than the theoretical peak performance of an SpMV operation.
-Concluding remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1
+A performance of 135~Gflop/s in single precision is achieved using RBF-FD differentiation matrices, which is 3.75 times better than the theoretical peak performance of an SpMV operation.
+Concluding remarks and perspectives are provided in Section~\ref{sec:ccl}.%\looseness=-1
 

### 2014-01-17 acm/ieee_paper.tex
scaling figure

@@
 \def\ge#1{{\blue{#1}}}
-\def\NOTE#1{{\red{NOTE: #1}}}
+%\def\NOTE#1{{\red{NOTE: #1}}}
+\def\NOTE#1{{}}
 \def\ttt#1{{\tt #1}}
@@
   \centering
+\scalebox{.8}{
 \subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{figures/code.pdf}}%
 \subfigure[Content of the vector registers \ge{in the source code in (a)}.]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
+}
 

### 2014-01-17 acm/ieee_paper.tex
line numbers

@@
 \subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{figures/code.pdf}}%
-\subfigure[Content of the vector registers \ge{in the source code in (a)}.]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}
+\subfigure[Content of the vector registers \ge{in the source code in (a)}.]{\hspace{-0.07\linewidth}\includegraphics[width=.55\linewidth]{figures/code-annote.pdf}}
 }

### 2014-01-17 acm/ieee_paper.tex
line number

@@
 \def\ge#1{{\blue{#1}}}
-%\def\NOTE#1{{\red{NOTE: #1}}}
-\def\NOTE#1{{}}
+\def\NOTE#1{{\red{NOTE: #1}}}
+%\def\NOTE#1{{}}
 \def\ttt#1{{\tt #1}}
@@
 reorder of the elements within each lane (see how {\tt mat} is
-extracted from {\tt mat\_ent} in Figure~\ref{code:mat_mul}). On the
+extracted from {\tt mat\_ent} in Figure~\ref{code:mat_mul} VIII to X). On the
 other hand, a permutation reorders entire lanes, without changing
 their contents (see for instance how {\tt vect} is generated from {\tt
-  all\_vect} in Figure~\ref{code:mat_mul}).  Both permutation and
+  all\_vect} in Figure~\ref{code:mat_mul} VII to IX).  Both permutation and
 swizzling support common reordering and broadcast. In the code, the
@@
 lane of the vector register using an {\tt unpack}
-operation\footnote{Notice that there are two calls to {\tt unpack} for the
+operation in line IV\footnote{Notice that there are two calls to {\tt unpack} for the
 LSB and MSB. Both are mandatory even if one is known to be
@@
 are replicated so that each value fills a single vector lane 
-using a swizzle operation. Then each value is multiplied by four
+using a swizzle operation (in line V). Then each value is multiplied by four
 (because $n_v=4$) 
@@
 correct index of the elements of the {\tt x} vector within each lane
-using a Fused Multiply-Add operation. A {\tt gather} operation is then
-performed to bring the 16 entries of the {\tt x} vector into the SIMD
+using a Fused Multiply-Add operation in line VI. A {\tt gather} operation is then
+performed in line VII to bring the 16 entries of the {\tt x} vector into the SIMD
 register. At this point we have a register where each lane is the
@@
 each lane. (The broadcast of $x_m^1,x_m^2,x_m^3,x_m^4$ is shown in
-Figure~\ref{code:mat_mul}.)
+Figure~\ref{code:mat_mul} in line IX.)
 
@@
 properly laid out in the vectors, a partial value for $y_i$ can be
-computed. Once all the nonzeros of the row have been processed, the
+computed in line XI. Once all the nonzeros of the row have been processed, the
 exact $y_i$ value can be sent to memory using a the "No Read No Global

### 2014-01-17 acm/ieee_paper.tex
line number

@@
 \def\es#1{{\blue{*** For Erik: #1 ***}}}
-\def\ee#1{{\green{#1}}}
-\def\ge#1{{\blue{#1}}}
-\def\NOTE#1{{\red{NOTE: #1}}}
-%\def\NOTE#1{{}}
+%\def\ee#1{{\green{#1}}}
+\def\ee#1{{#1}}
+%\def\ge#1{{\blue{#1}}}
+\def\ge#1{{#1}}
+%\def\NOTE#1{{\red{NOTE: #1}}}
+\def\NOTE#1{{}}
 \def\ttt#1{{\tt #1}}
@@
 channel interchange features. We present benchmarks on a real data
-set that show an almost \ge{sevenfold} \note{eightfold; there were inconsistencies in the ppaer.} increase in speed compared to
+set that show an almost \ge{sevenfold} \NOTE{eightfold; there were inconsistencies in the paper.} increase in speed compared to
 efficient implementations of a single derivative, reaching a
@@
 reorder of the elements within each lane (see how {\tt mat} is
-extracted from {\tt mat\_ent} in Figure~\ref{code:mat_mul} VIII to X). On the
+extracted from {\tt mat\_ent} in Figure~\ref{code:mat_mul} line VIII to X). On the
 other hand, a permutation reorders entire lanes, without changing
 their contents (see for instance how {\tt vect} is generated from {\tt
-  all\_vect} in Figure~\ref{code:mat_mul} VII to IX).  Both permutation and
+  all\_vect} in Figure~\ref{code:mat_mul} line VII to IX).  Both permutation and
 swizzling support common reordering and broadcast. In the code, the

### 2014-01-18 acm/RBF_Section.tex
minor
scaling figure
Modified figure 11 and added to the caption.

@@
 The result is a sparse matrix with only $n$ entries in each of its $N$ rows, resulting in a total of $nN$ nonzero entries.
+An example stencil is shown in
+%Figure~\ref{fig:stencil}.  % Latex is making this figure 2. Do not know why. 
+Figure~1.
 
-\begin{figure}[t]
+\begin{figure}[tbh]
 \centering
-\includegraphics[width=0.8\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
+\includegraphics[width=0.7\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
 \caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square.
@@
 For example, the full shallow water equations in a 3D Cartesian coordinate system for a rotating fluid are as follows:
-{
-\small
+{ \small
 \begin{eqnarray}
@@
 \end{eqnarray}
-} where $f$ is the Coriolis force, $\{u,v,w\}$ are the components of the velocity vector in the respective $\{x,y,z\}$ directions, and $h$ is the geopotential height (analogous to pressure).
-Thus, we have four variables $\{u,v,w,h\}$ and three DMs, $D_x, D_y, D_z$ representing $\dfrac{\partial}{\partial x}, \dfrac{\partial}{\partial y},\dfrac{\partial w}{\partial z}$, respectively.
+}
+where $f$ is the Coriolis force, $\{u,v,w\}$ are the components of the velocity vector in the respective $\{x,y,z\}$ directions, and $h$ is the geopotential height (analogous to pressure).
+Thus, we have four variables $\{u,v,w,h\}$ and three DMs, $D_x, D_y, D_z$ representing
+\begingroup
+\everymath{\footnotesize}
+\scriptsize
+$\dfrac{\partial}{\partial x}, \dfrac{\partial}{\partial y},\dfrac{\partial}{\partial z}$, 
+\endgroup
+respectively.
 Moreover, in order to stably time step the equations with an explicit time-stepping scheme, most numerical methods (including RBF-FD) require a fourth operator, hyperviscosity, to be added to each equation.
@@
 \begin{itemize}
-\item For a given $n$, regardless of the operator approximated, e.g. $\dfrac{\partial}{\partial x}$ or the Laplacian, RBF-FD DMs will have the same number of nonzeros per row.
+\item For a given $n$, regardless of the operator approximated, e.g.,
+$\partial_x$, 
+%$\dfrac{\partial}{\partial x}$ 
+or the Laplacian, RBF-FD DMs will have the same number of nonzeros per row.
 \item For a given $n$ and $N$, regardless of the operator approximated, the DMs will have the same sparsity pattern. Examples of an RBF-FD DM can be seen in Figures \ref{fig:spy_plots}d and e.
@@
 %\begin{figure}[h]
+\begingroup
+\everymath{\footnotesize}
+\scriptsize
 \begin{equation}
@@
 \end{equation}
+\endgroup
 where $\{\underline{u},\underline{v}, \underline{w}, \underline{h}\}$ are the functions values at all nodes of the corresponding variables with the left hand side being the resulting derivative approximations.

### 2014-01-18 acm/ieee_paper.tex
no need for "extended abstract"
Modified figure 11 and added to the caption.

@@
 
-\section{Modelization of the Potential Improvements}
+\section{Modelization of the \\Potential Improvements}
 \label{sec:model}
@@
 within 15\% of the predicted  peak performance. The performance of the
-{\tt RBF} case varies between 100~Gflop/s and 140~Gflop/s. Most of the
+{\tt RBF} case varies between 100 and 140~Gflop/s. Most of the
 time, the RCM ordering provides an improvement which can be as high as
@@
     (squares) only improves the performance for our manually
-    vectorized code.}
+    vectorized code.
+    We also show peak theoretical performance for the base case and the 4 vector/1 matrix case. Not shown is the peak theoretical performance of 210 Gflops for the 4/4 case. 
+    }
   \label{fig:comparison_spmv}
@@
 provides a significant improvement in our manually vectorized
-version. This highlights that the C++ implementation is instruction-bound
+version. Thus, the C++ implementation is likely instruction-bound
 while our manually vectorized implementation is most likely 

### 2014-01-18 figure_files/figure11/plot_for_paper_figure11.py

@@ -3,6 +3,7 @@ import sys
 import numpy as np
 import matplotlib.pyplot as plt
 import csv
+import myutil
 
 #----------------------------------
 
@@ -94,6 +95,11 @@ def plot_file(infile,outfile):
     plt.grid(True)
     #plt.ylim(0,100)
     plt.xlim([0,250])
+
+    myutil.horizontal_line(35,  xmin=205, xmax=250, col='b', lw=4)
+    myutil.horizontal_line(135, xmin=205, xmax=250, col='g', lw=4)
+    plt.text(130,32.5,'peak: 35 Gflop/s', fontsize=16)
+    plt.text(130,132.5,'peak: 135 Gflop/s', fontsize=16)
 
     plt.tight_layout()
     plt.savefig.format = "pdf"

### 2014-01-18 acm/ieee_paper.tex
Final paper submitted to ACM Conference. http://www.lrr.in.tum.de/~gerndt/ICS2014/

@@
 % A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
+%\category{I.6}{Simulation and Modeling}{Applications} %\break
+\category{F.2}{Analysis of Algorithms and Problem Complexity}{Numerical Algorithms and Problems}
 %A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
 
-\terms{Theory}
+%\terms{Theory}
 
@@
 \ee{The core of the technique is to load the maximum amount of data
-  into large register and format the data for processing within these
-  registers. That way the amount of interaction with the memory is
-  minimized. The Intel MIC is more for these operations than classical
+  into large the registers and format the data for processing within these
+  registers, minimizing the amount of interaction with memory. 
+  The Intel MIC is more for these operations than classical
   CPUs because it features larger vector registers and more elaborate
-  vector reorganization instructions than namely swizzling,
+  vector reorganization instructions, such as swizzling,
   permutations and broadcasts. (The AVX instruction sets on classical
-  CPUs is more primitive in that aspect.)} A vector of 512 bits is
+  CPUs are more primitive in that aspect.)} A vector of 512 bits is
 composed of four lanes (sometimes called channels) of 128 bits, which
@@
 
-\ee{The modelization of Section~\ref{sec:model} shows that the memory
-  transfer caps the computation at 210GFlop/s, which is 10\% of the
+\ee{The modelization of Section~\ref{sec:model} establishes that the 
+  memory
+  transfer caps the computation at 210~GFlop/s, which is 10\% of the
   peak single precision of the architecture. To reach this value, the
@@
     vectorized code.
-    We also show peak theoretical performance for the base case and the 4 vector/1 matrix case. Not shown is the peak theoretical performance of 210 Gflops for the 4/4 case. 
+    We also show peak theoretical performance for v1m1 and v4m1 on the MIC. 
+    Not shown is the peak theoretical performance of 210 Gflops for v4m4. 
     }
@@
 the techniques presented in this paper within a fluid simulation using
-RBFFD. \ee{We will also look developping a similar technique on more
+RBFFD. \ee{We will also look into developing a similar technique on more
 commonly available devices such as GPUs.}

### 2014-01-18 figure_files/figure11/plot_for_paper_figure11.py

@@ -96,10 +96,10 @@ def plot_file(infile,outfile):
     #plt.ylim(0,100)
     plt.xlim([0,250])
 
-    myutil.horizontal_line(35,  xmin=205, xmax=250, col='b', lw=4)
-    myutil.horizontal_line(135, xmin=205, xmax=250, col='g', lw=4)
-    plt.text(130,32.5,'peak: 35 Gflop/s', fontsize=16)
-    plt.text(130,132.5,'peak: 135 Gflop/s', fontsize=16)
+    myutil.horizontal_line(35,  xmin=210, xmax=250, col='b', lw=4)
+    myutil.horizontal_line(135, xmin=210, xmax=250, col='g', lw=4)
+    plt.text(120,33.8,'v1m1 peak: 35 Gflop/s', fontsize=14)
+    plt.text(120,133.8,'v4m1 peak: 135 Gflop/s', fontsize=14)
 
     plt.tight_layout()
     plt.savefig.format = "pdf"

### 2014-04-15 acm/answers_to_reviews_500words.txt
ICS paper.

Review 1
The cost of computing the derivatives requires the inversion of a nxn system of equations for each node. For stencil sizes of 64 or 128, this can be expensive. But the calculation of derivative weights is done only once. Our method is useful when solving time-dependent PDEs or stationary PDEs via iterative methods, and the derivatives are to be computed thousands if not millions of times, thus amortizing the original cost.

Regarding the time to compute the derivation matrix. This operation is typically not very optimized because the same matrix is often dumped to disk and reused later to perform many analyses each composed of many tens of thousands 
of spmv operations. To give a point of reference, the cost of computing the matrix has an order of magnitude of a thousand spmvs. If one is performing a single simulation, then the cost of generating the matrix is important, but when perforing a parametric study on a single grid, the cost of the derivative matrix calculation is negligible.

Effectively, we could go to n_m=n_v to 16 based on figure 3. However, as mentionned in the introduction, the transport equations for Euler's equations require $x,y,z$ derivatives of $u,v,w,p$ and the calculation of a special hyperviscosity operator. Thus, there are four derivatives applied to four functions. If we had more functions and more derivatives, we could certainly increase the values of n_m and n_v.

We use 1D arrays for all operations. If the data is contiguous, a 1D array is equivalent to a 2D array, with our own control using pointers.  Low level programming, assembler, CUDA, etc. all work better with 1D arrays. Finally, we would like have a certain degree of control of where our data goes in memory, and thus it is easier to think about our arrays as one-dimensional.

Review 3
The front end to the MIC was an Intel(R) Xeon(R) CPU E5-2650 running at 2.0GHZ, with a total of 16 cores (8 per socket), with hyperthreading enabled and 64 GB of RAM running at 1600 MHz.

Review 4
Given the short space provided, we were only brief in our description of the RBF-FD, and the implementation on the MIC, although we provided source code for those interested in the details.  The stencil size is the number of non-zeros in each row of the matrix $A$ used in the matrix multiplication y = A*x.   Equation (1) in the paper represents the product y = A x, where y is the computed derivative, A is the derivative matrix, and x is the function whose derivative is being computed.  The size of the different matrices is clearly shown in Equation (1).

From Erik: 
[3/31/14, 11:44:59 AM] erik saule: I think we should first give the full name of the CPU. If you know on which machine this was run, you should have the information dans /proc/cpuinfo
[3/31/14, 11:45:44 AM] erik saule: there is indeed little difference between Xeon E5 architecture and core i7. The most different is in the support for ECC memory and higher core count
[3/31/14, 11:46:12 AM] erik saule: but let's say explicitely it is a Xeon E5-2650 at 2Ghz

### 2014-04-15 acm/bak/ieee_paper_final.tex
ICS paper.

--- 2014-01-18 acm/ieee_paper.tex
@@
 \input{support/setupicase}
-
-\def\red#1{\textbf{\textcolor{red}{#1}}}
-\def\blue#1{\textbf{\textcolor{blue}{#1}}}
-\def\green#1{\textbf{\textcolor{green}{#1}}}
-\def\qes#1{{\blue{*** For Erik: #1 ***}}}
-\def\es#1{{\blue{*** For Erik: #1 ***}}}
-%\def\ee#1{{\green{#1}}}
-\def\ee#1{{#1}}
-%\def\ge#1{{\blue{#1}}}
-\def\ge#1{{#1}}
-%\def\NOTE#1{{\red{NOTE: #1}}}
-\def\NOTE#1{{}}
-\def\ttt#1{{\tt #1}}
-\def\bold#1{{\bf #1}}
 
@@
   urlcolor=blue]{hyperref}
+
+%% \def\red#1{\textbf{\textcolor{red}{#1}}}
+%% \def\blue#1{\textbf{\textcolor{blue}{#1}}}
+%% \def\green#1{\textbf{\textcolor{green}{#1}}}
+%% \def\qes#1{{\blue{*** For Erik: #1 ***}}}
+%% \def\es#1{{\blue{*** For Erik: #1 ***}}}
+%% %\def\ee#1{{\green{#1}}}
+%% \def\ee#1{{#1}}
+%% %\def\ge#1{{\blue{#1}}}
+%% %\def\gee#1{}
+%% \def\ge#1{{#1}}
+%% %\def\NOTE#1{{\red{NOTE: #1}}}
+%% \def\NOTE#1{{}}
+%% \def\ttt#1{{\tt #1}}
+%% \def\bold#1{{\bf #1}}
 
@@
 
+% deal with figures that span two columns
+\usepackage{dblfloatfix}
+
 \usepackage{graphicx}
@@
 \begin{document}
+\end{document}
 \title{Acceleration of Derivative Calculations with Application to Radial Basis Function --
@@
 
-\subtitle{[Extended Abstract]}
+%\subtitle{[Extended Abstract]}
 
@@
 
-\begin{figure*}
-  \centering
-\scalebox{.8}{
-\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{figures/code.pdf}}%
-\subfigure[Content of the vector registers \ge{in the source code in (a)}.]{\hspace{-0.07\linewidth}\includegraphics[width=.55\linewidth]{figures/code-annote.pdf}}
-}
-
-\caption{Code snippet of the multiplication of 4 vectors with 4
-  matrices. When a line changes the content of a vector register its
-  new content is shown on the right. The effects of swizzling and
-  permutations are highlighted using a color code.}
-\label{code:mat_mul}
-\end{figure*}
+%\begin{figure*}
+%  \centering
+%\scalebox{.8}{
+%\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{figures/code.pdf}}%
+%\subfigure[Content of the vector registers {in the source code}.]{\hspace{0.07\linewidth}\includegraphics[width=.55\linewidth]{figures/code-annote.pdf}}
+%%\subfigure[Content of the vector registers \ge{in the source code}.]{\hspace{-0.07\linewidth}\includegraphics[width=.55\linewidth]{figures/code-annote.pdf}}
+%}
+%
+%\caption{Code snippet of the multiplication of 4 vectors with 4
+%  matrices. When a line changes the content of a vector register its
+%  new content is shown on the right. The effects of swizzling and
+%  permutations are highlighted using a color code.}
+%\label{code:mat_mul}
+%\end{figure*}
 
@@
 surpasses 100~GB/s and is close to 50~GB/s most of the time.)
+
 \def\ww{.19\textwidth}
-\begin{figure*}[t]
+\begin{figure*}[b]
+% Figure 8
   \begin{center}
@@
     %%   \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
-    \subfigure[3D, RCM.]{%
+    \subfigure[3D, \ge{no RCM}.]{%
       \label{fig:rbf3dnorcm} 
       \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
-    \subfigure[3D, no RCM.]{%
+    \subfigure[3D, \ge{RCM}.]{%
       \label{fig:rbf3drcm}
@@
   \label{fig:spy_plots}
+\end{figure*}
+
+\begin{figure*}
+% Figure 9
+  \centering
+% was width=.48
+  \subfigure[Supercompact and Random.]{\includegraphics[width=.4\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
+% was width=.48
+  \subfigure[Compact and RBFFD.]{\includegraphics[width=.4\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
+
+  \caption{Performance of the manually vectorized code. The two
+    horizontal lines depict the worst and the best predicted performances.}
+  \label{fig:expe_types}
 \end{figure*}
@@
 
-\begin{figure*}[t]
-  \centering
-  \subfigure[Supercompact and Random.]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
-  \subfigure[Compact and RBFFD.]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
-
-  \caption{Performance of the manually vectorized code. The two
-    horizontal lines depict the worst and the best predicted performances.}
-  \label{fig:expe_types}
-\end{figure*}
-
 We now investigate the actual performance that we can obtain when
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
+% Figure 10
   \centering

### 2014-04-15 acm/ieee_paper_final.tex
ICS paper.

--- 2014-01-18 acm/ieee_paper.tex
@@
   urlcolor=blue]{hyperref}
+
+\def\red#1{\textbf{\textcolor{red}{#1}}}
+\def\blue#1{\textbf{\textcolor{blue}{#1}}}
+\def\green#1{\textbf{\textcolor{green}{#1}}}
+\def\qes#1{{\blue{*** For Erik: #1 ***}}}
+\def\es#1{{\blue{*** For Erik: #1 ***}}}
+%\def\ee#1{{\green{#1}}}
+\def\ee#1{{#1}}
+\def\FIXED{{\red{FIXED}}}
+%\def\NOTE#1{{\red{NOTE: #1}}}
+\def\NOTE#1{{}}
+\def\ttt#1{{\tt #1}}
+\def\bold#1{{\bf #1}}
+
+%\def\ge#1{{\blue{#1}}}  % answer to first review
+\def\ge#1{{#1}}
+
+\def\gee#1{}
+%\def\ger#1{{\red{#1}}}  % answer to second review
+\def\ger#1{#1}
 
@@
 
-Common improvement techniques such as register blocking, bandwidth
-reduction (matrix reordering~\cite{Cuthill69}), partitioning to fit in
+Common improvement techniques such as  bandwidth
+reduction (matrix reordering~\cite{Cuthill69}), register blocking, partitioning to fit in
 cache or TLB~\cite{Nishtala07,Temam:1992:CBS:147877.148091,conf/ppsc/Toledo97},
@@
 Concluding remarks and perspectives are provided in Section~\ref{sec:ccl}.%\looseness=-1
+%--------------------------------------------
 
-\input{RBF_Section}
+%\input{RBF_Section}
 %\input{old_RBF_Section}
 
+\section{Introduction and Motivation for RBF-FD}
+\label{sec:rbf}
+
+RBFs approximate a function $f(\mathbf{x}) \in \mathbb{R}^d$ sampled at $N$ distinct node locations by linearly combining translates of a single radially symmetric function, e.g.
+$\phi(r) = e^{-\varepsilon^2r^2}$, where $r=\|\mathbf{x}-\mathbf{x}_{i}\|$ denotes the Euclidean distance between where the function $f(\mathbf{x})$ is evaluated, $\mathbf{x}$, and where the RBF, $\phi$, is centered $\mathbf{x}_{i}$.
+The parameter $\varepsilon$ controls the shape of the RBF.
+Note that the argument of an RBF is simply a scalar distance, $r$, independent of any coordinate system or dimension.
+As a result, nodes can be scattered as desired across complex physical domains with the implementation of the method being independent of dimensionality.
+Thus, no mesh generation is needed and algorithmic complexity does not increase with dimension.
+\NOTE{While reading that paragraph, I was under the impression that the physical domain needs to be a sphere.}
+
+To obtain a RBF derivative approximation at a node location that will result in a sparse differentiation matrix (DM), the RBF-FD method has been developed over the last decade \cite{TAI1,TAI2,SDY02,WrFo06}.
+RBF-FD is conceptually similar to standard finite differences (FD).
+However, the differentiation weights for approximating a derivative at a given node (forming one row of the RBF-FD derivative matrix (DM)) enforce that the linear combination of function values at the $n << N$ nearest node locations (a stencil) be exact for RBFs centered at each of the $n$ nodes rather than polynomials.
+The result is a sparse matrix with only $n$ entries in each of its $N$ rows, resulting in a total of $nN$ nonzero entries.
+An example stencil is shown in
+%Figure~\ref{fig:stencil}.  % Latex is making this figure 2. Do not know why. 
+Figure~1.
+
+\begin{figure}[tbh]
+\centering
+\includegraphics[width=0.7\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
+\caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square.
+In nonsparse format, the DM  is $1024\times1024$ with 32 nonzero entries per row.}
+\end{figure}
+
+Due to the above attributes and its simplicity of implementation, the RBF-FD is gaining popularity in modeling systems of PDEs in a variety of fields in fluid dynamics (e.g., geosciences, combustion, aerodynamics \cite{Bayona13,CDNT,FoL11,FLBWSC12,SPLM}).
+For example, the full shallow water equations in a 3D Cartesian coordinate system for a rotating fluid are as follows:
+{ \small
+\begin{eqnarray}
+\dfrac{\partial u}{\partial t} &=& - \left( u\dfrac{\partial u}{\partial x} + v\dfrac{\partial u}{\partial y} + w\dfrac{\partial u}{\partial z} + f(yw - zv) + g\dfrac{\partial h}{\partial x} \right) 
+\nonumber \\
+\dfrac{\partial v}{\partial t} &=& - \left( u\dfrac{\partial v}{\partial x} + v\dfrac{\partial v}{\partial y} + w\dfrac{\partial v}{\partial z} + f(zu - xw) + g\dfrac{\partial h}{\partial x} \right)
+\nonumber\\
+\dfrac{\partial w}{\partial t} &=& - \left( u\dfrac{\partial w}{\partial x} + v\dfrac{\partial w}{\partial y} + w\dfrac{\partial w}{\partial z} + f(xv - yu) + g\dfrac{\partial h}{\partial x} \right)
+\nonumber\\
+\frac{\partial h}{\partial t}& =&-\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right) , \label{height} \nonumber
+\end{eqnarray}
+}
+where $f$ is the Coriolis force, $\{u,v,w\}$ are the components of the velocity vector in the respective $\{x,y,z\}$ directions, and $h$ is the geopotential height (analogous to pressure).
+Thus, we have four variables $\{u,v,w,h\}$ and three DMs, $D_x, D_y, D_z$ representing  $\partial_x$, $\partial_y$, $\partial_z$,
+%\begingroup
+%\everymath{\footnotesize}
+%\scriptsize
+%$\dfrac{\partial}{\partial x}, \dfrac{\partial}{\partial y},\dfrac{\partial}{\partial z}$, 
+%\endgroup
+respectively.
+Moreover, in order to stably time step the equations with an explicit time-stepping scheme, most numerical methods (including RBF-FD) require a fourth operator, hyperviscosity, to be added to each equation.
+Hyperviscosity takes the form of $\triangle^p$ ($\triangle$ denotes the Laplacian operator), where $p\geq3$, which is approximated by the matrix $D_{hyp}$.
+Therefore, in order to solve this system of PDEs, the application of 4 DM to 4 unknowns, i.e.
+16 SpMV is required at every time step.
+Due to the fact that RBFs are independent of a coordinate system, taking into account only distances to nearest neighbors, the DMs have the following properties:
+\begin{itemize}
+\item For a given $n$, regardless of the operator approximated, e.g.,
+$\partial_x$, 
+%$\dfrac{\partial}{\partial x}$ 
+or the Laplacian, RBF-FD DMs will have the same number of nonzeros per row.
+\item For a given $n$ and $N$, regardless of the operator approximated, the DMs will have the same sparsity pattern. Examples of an RBF-FD DM can be seen in Figures \ref{fig:spy_plots}d-e.
+\end{itemize}
+
+Multiple variables transform a SpMV into a SpMM (Sparse Matrix/dense Matrix multiplication), which improves register utilization and decreases cache
+misses by vectorizing over the multiple source vectors.
+Furthermore, due to the properties just mentioned, multiple derivatives of a single function can be calculated rather than computing a single derivative of multiple functions.
+Using the PDE system above as an example, the block structure of the SpMM for computing all derivatives needed is %can be seen in Equation \ref{fig:struct_comp},
+%\begin{figure}[h]
+\begingroup
+\everymath{\footnotesize}
+\scriptsize
+\begin{equation}
+  %\centering
+ %{\small
+   \left( \begin{array}{cccc}
+    \underline{u}_x     & \underline{v}_x     & \underline{w}_x     & \underline{h}_x \\
+    \underline{u}_y     & \underline{v}_y     & \underline{w}_y     & \underline{h}_y \\
+    \underline{u}_z     & \underline{v}_z     & \underline{w}_z     & \underline{h}_z \\
+    \underline{u}_{hyp} & \underline{v}_{hyp} & \underline{w}_{hyp} & \underline{h}_{hyp}
+  \end{array} \right)
+  = \left(
+  \begin{array}{c}
+    D_{x}\\ D_{y}\\ D_{z}\\ D_{hyp}
+  \end{array}\right)
+  \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) 
+  \vspace{-1ex}
+  %\caption{Block structure of the computation.}
+  \label{fig:struct_comp}
+%\end{figure}
+\end{equation}
+\endgroup
+where $\{\underline{u},\underline{v}, \underline{w}, \underline{h}\}$ are the functions values at all nodes of the corresponding variables with the left hand side being the resulting derivative approximations.
+The increased memory bandwidth due to an increase in the number of derivative matrices is offset by improved cache utilization, leading to an overall performance benefit.
+In practice however, the number of vectors whose derivatives are required is limited, capping the performance of the matrix/vector multiplication.
+Similarly, the number of different derivatives in a particular computation is finite.
+It is for this reason that we seek to combine the benefits of multiple vectors and multiple matrices simultaneously.
+We limit this initial study to four vectors and four matrices, which is a common number found in fluid dynamic calculations.
+Equation~\ref{fig:struct_comp} is of the form $Y=AX$, where $Y$ encodes the $x,y,z$ discrete derivatives (and hyperviscosity operator) of $u,v,w,h$, $X$ encodes the variables $(u,v,w,h)$ and $A$ represents the 16 discrete differentiation operators.
+
+We also note that the DMs are time-independent. They are computed once \ee{at the beginning of the simulation} and used repeatedly to update the solution 
+in time (for each iteration of the simulation). Thus, the cost to calculate the derivatives, although high \ger{(on the order of 1000 SpMVs)}, 
+is amortized by the very long numerical simulations.
+\ger{The startup cost is eliminated when performing parameter studies on a given node distribution.
+Furthermore, the computation of the derivative matrices are done serially, but are trivially parallelizable.}
+
+%--------------------------------------------
 \section{Modelization of the \\Potential Improvements}
@@
 used for double or single precision, that is, either as a vector of eight
-64-bit values or as a vector of 16 32-bit values, respectively. The
-vector processing unit can perform basic instructions, such as
-addition or division, and mathematical operations, such as {\tt sin} or
-{\tt sqrt}, making it possible to reach 16 single precision (or 8 double precision)
-operations per cycle.  
+64-bit values or as a vector of 16 32-bit values, respectively. 
+\ger{The vector processing unit can perform arithmetic (addition and multiplication) 
+instructions making it possible to reach 16 single precision 
+(or 8 double precision) operations per cycle. 
 The unit also supports Fused Multiply-Add (FMA)
@@
 precision. If FMA cannot be used, only half of these rates can be
-achieved.
+achieved.}
+
+%addition or %division, and mathematical operations, such as {\tt sin} or
+%{\tt sqrt}, making it possible to reach 16 single precision (or 8 double precision)
+%operations per cycle.
 
@@
 \subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{figures/code.pdf}}%
-\subfigure[Content of the vector registers \ge{in the source code in (a)}.]{\hspace{-0.07\linewidth}\includegraphics[width=.55\linewidth]{figures/code-annote.pdf}}
+\subfigure[Content of the vector registers.]{\hspace{0.07\linewidth}\includegraphics[width=.55\linewidth]{figures/code-annote.pdf}}
+%%\subfigure[Content of the vector registers \ge{in the source code}.]{\hspace{-0.07\linewidth}\includegraphics[width=.55\linewidth]{figures/code-annote.pdf}}
 }
@@
 of $n_z$ entries that label the column index of each nonzero in
-$A$. The other array is  {\tt data} that gives the values of
+$A$. %***
+The other array is  {\tt data} that gives the values of
 the nonzeros in the four matrices. The values of the matrices are
@@
 grouped in sets of 16, which correspond to four derivatives applied to
-four functions, at a single point.\looseness=-1
+four functions, at a single point. %\ger{All matrices are stored as one-dimensional vectors and pointers to individual rows and elements are explicitely computed.}
+\looseness=-1
 
@@
 
-The {\tt read, cpp} implementation reaches a performance of 168~GB/s
+The {\tt read\_cpp} implementation reaches a performance of 168~GB/s
  which is better than the {\tt read}
@@
 is explained by the compiler adding loop unrolling and software
-prefetching in the assembly code in the {\tt read, cpp} case, which is
+prefetching in the assembly code in the {\tt read\_cpp} case, which is
 not added in the {\tt read} case. On the other hand, the {\tt write}
 case reaches a performance of 158~GB/s while the
-{\tt write, cpp} implementation never exceeds 58~GB/s. 
-This difference is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction, 
-which bypasses the \ge{"Read For Ownership"} protocol and allows the writes
+{\tt write\_cpp} implementation never exceeds 58~GB/s. 
+This difference is the result of {\tt \_mm512\_store\_nrngo\_ps}, 
+an instruction, which bypasses the \ge{"Read For Ownership"} protocol 
+and allows the writes
 to be performed in any order. The compiler can not use this
@@
 useful for some parallel algorithms. The {\tt read/write} benchmark
-performs at 140~GB/s while the {\tt read/write, cpp} benchmark performs
+performs at 140~GB/s while the {\tt read/write\_cpp} benchmark performs
 at about 90~GB/s.
@@
 elements of the \ttt{Random}
-matrices (Figure~\ref{fig:compact}) see their nonzero elements
+matrices (Figure~\ref{fig:random}) see their nonzero elements
 randomly (uniformly) distributed in the matrix; they represent the
@@
 \subsection{Computations}
-
-\begin{figure*}[t]
-  \centering
-  \subfigure[Supercompact and Random.]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
-  \subfigure[Compact and RBFFD.]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
-
-  \caption{Performance of the manually vectorized code. The two
-    horizontal lines depict the worst and the best predicted performances.}
-  \label{fig:expe_types}
-\end{figure*}
 
@@
 
-\ee{The main idea of this paper, computing 16 derivatives at a time,
+\ger{The main idea of this paper, computing 16 derivatives at a time,
 can be applied on different architectures such as multiprocessor
-  multicore or GPU. To illustrate this fact, we conducted similar
-  experiments on a classical CPU architecture with two Intel Core i7
-  (SandyBridge) processors, each has eight cores that support AVX
-  instructions.}  Both the compiler-generated and the manually-written
-vectorization using AVX instructions reach a (read) bandwidth of
-70~GB/s. This indicates that the performance of the classical SpMV can
-not overcome 17~Gflop/s. A generic implementation of the
-multiplication of four vectors with four matrices on the {\tt RBF}
-matrix reaches a performance of 27~Gflop/s. \ee{We also implemented an AVX
-vectorized version which is less compact than the one on MIC because
-of smaller vector registers and the abscence of a gather instruction
-(Haswell processors support AVX2 which have that feature).} It reaches a
-performance of 42~Gflop/s. This confirms that our method is sound and
-generalizes to architectures other than the Intel MIC architecture.
+multicore or GPU. To illustrate this fact, we conducted similar experiments
+on Cascade at the Minnesota Supercomputing Institute
+using two XEON E5 2670 nodes (8 cores, 2 threads and 20 Mbytes of L2
+cache per core, with hyperthreading disabled.) The cores support the AVX instruction set. 
+Both the compiler-generated code and the code written with the 
+help of the AVX instruction set reach a read bandwidth of
+85~GB/s. Using cache lines of 256 bits, we find a best/worst speeds of 4/20\; Gflops/s and 
+50/120\; Gflops/s, respectively for the 1vec/1mat and 4vec/4mat cases. 
+Our implementation on Cascade reaches 45 Gflops for the 4/4\, case, or 41\% of peak theoretical 
+performance on this particular processor setup. For references, we achieve 
+140/210=66\% of peak theoretical performance on the MIC. Note that peak performance 
+is not that of the processor, but of the best possible algorithm on the processor. 
+The gain in the 4/4 case with respect to 4x 1/1 case is respectively 4x on the MIC 
+and 2.5x on the front end.  
+}
 
@@
 includes a number of swizzling and channel swapping operations, for an
-extremely efficient tensor product implementation. \NOTE{From Evan: this is the first time $96^3$ is mentionned. Might wish to mention it earlier.} On a $96^3$ 3D
+extremely efficient tensor product implementation. 
+\NOTE{From Evan: this is the first time $96^3$ is mentionned. Might wish to mention 
+it earlier.} On a $96^3$ 3D
 grid, it reaches a performance of 140~Gflop/s which is 2.7 times
@@
 the techniques presented in this paper within a fluid simulation using
-RBFFD. \ee{We will also look into developing a similar technique on more
-commonly available devices such as GPUs.}
+RBFFD. \ge{We will also look into developing a similar technique on the Kepler
+family of GPUs, which offers swizzling and channeling operations.}
 

### 2014-04-15 acm/ieee_paper_final.tex
adding author affiliations and replacing figures
typo

@@
 
-%\numberofauthors{4}
+\numberofauthors{2}
 
-%\author{
-%\alignauthor
-%Gordon Erlebacher\\ %\titlenote{Dr.~Trovato insisted his name be first.}\\
-%       \affaddr{Department of Scientific Computing}\\
-%       \affaddr{Florida State University}\\
+\author{
+\alignauthor Gordon Erlebacher\\ 
+      \affaddr{Department of Scientific Computing}\\
+      \affaddr{Florida State University}\\
 %      \affaddr{Tallahassee, FL 32306-4120}\\
-%       \email{gordon.erlebach@gmail.com}
-%\alignauthor
-%Erik Saule
-%        \affaddr{Department of Computer Science}\\
-%        \affaddr{University of North Carolina at Charlotte}\\
-%        \email{esaule@uncc.edu}
-%\and
-%\alignauthor
-%Natasha Flyer
-%        \affaddr{Computational and Information System Laboratory}\\
-%        \affaddr{UCAR}\\
-%        \email{flyer@ucar.edu}
-%\alignauthor
-%Evan Bollig
-%        \affaddr{Minnesota Supercomputer Institute}\\
-%        \affaddr{University of Minnesota}\\
+      \email{gordon.erlebach@gmail.com}
+\alignauthor Erik Saule\\
+       \affaddr{Department of Computer Science}\\
+       \affaddr{University of North Carolina at Charlotte}\\
+       \email{esaule@uncc.edu}
+\and
+\alignauthor Natasha Flyer\\
+       \affaddr{Computational and Information Systems Laboratory}\\
+       \affaddr{UCAR}\\
+       \email{flyer@ucar.edu}
+\alignauthor Evan Bollig\\
+       \affaddr{Minnesota Supercomputer Institute}\\
+       \affaddr{University of Minnesota}\\
 %       \affaddr{Minneapolis, MN}\\
-%        \email{bollig@gmail.com}
-%}
+       \email{bollig@gmail.com}
+}
 \maketitle
@@
 
-\begin{figure}[tbh]
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.7\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
+\includegraphics[width=0.65\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
 \caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square.
@@
 %--------------------------------------------
-\section{Modelization of the \\Potential Improvements}
+\section{Modeling the\\ Potential Improvements}
 \label{sec:model}
@@
 
-\begin{figure}[t]
+\begin{figure}[tbh]
   \begin{center}
@@
 
-\begin{figure}
+\begin{figure}[t]
   \centering 
@@
 on Cascade at the Minnesota Supercomputing Institute
-using two XEON E5 2670 nodes (8 cores, 2 threads and 20 Mbytes of L2
+using two Intel Xeon E5 2670 nodes (8 cores, 2 threads and 20 Mbytes of L2
 cache per core, with hyperthreading disabled.) The cores support the AVX instruction set. 

### 2014-04-16 acm/ieee_paper_final.tex
Change section prior to conclusion. The numbers to refect the fact that the cache of Cascade, E5-3670 has 512 bit cache line (64 bytes). .

@@
 \centering
-\includegraphics[width=0.65\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
+\includegraphics[width=0.60\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
 \caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square.
@@
   \centering
-  \subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
-  \subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
-  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}
+  \subfigure[Best case.]{\includegraphics[width=.28\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
+  \subfigure[Worst case.]{\includegraphics[width=.28\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
+  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.28\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}
 
@@
 % was width=.48
-  \subfigure[Supercompact and Random.]{\includegraphics[width=.4\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
+  \subfigure[Supercompact and Random.]{\includegraphics[width=.35\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
 % was width=.48
-  \subfigure[Compact and RBFFD.]{\includegraphics[width=.4\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
+  \subfigure[Compact and RBFFD.]{\includegraphics[width=.35\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
 
@@
 
-  \includegraphics[width=.9\linewidth]{figures/mic_performance_nb_threads.pdf}
+  \includegraphics[width=.8\linewidth]{figures/mic_performance_nb_threads.pdf}
 
@@
 help of the AVX instruction set reach a read bandwidth of
-85~GB/s. Using cache lines of 256 bits, we find a best/worst speeds of 4/20\; Gflops/s and 
-50/120\; Gflops/s, respectively for the 1vec/1mat and 4vec/4mat cases. 
-Our implementation on Cascade reaches 45 Gflops for the 4/4\, case, or 41\% of peak theoretical 
+85~GB/s. Using cache lines of 512 bits, we find a best/worst speeds of 2.4/20\; Gflops/s and 
+32/120\; Gflops/s, respectively for the 1vec/1mat and 4vec/4mat cases. 
+Our implementation on Cascade reaches 45 Gflops for the 4/4\, case, or 45/120=37\% of peak theoretical 
 performance on this particular processor setup. For references, we achieve 
@@
 }
+
+% cascade
+%[  2.35701906   8.04733728   9.37931034  31.62790698  36.75675676]
+%[  20.60606061   32.96969697   75.55555556  120.88888889  226.66666667]
 
@@
 vector and a single derivative and more than 7 times faster than a
-practical implementation.
+best case practical 1/1 implementation.
 

### 2014-04-16 figure_files/figure5/gflops_speedup_figure5.py

@@ -118,6 +118,9 @@ bar_width = 0.35/2
 rects1 = plt.bar(index, ratios_s_worst, bar_width, alpha=1, color='b', label='single precision worst case')
 rects1 = plt.bar(index+bar_width, ratios_s_best, bar_width, alpha=.5, color='b', label='single precision best case')
 #print ratios_s_best
+print("worst, best")
+print(ratios_s_worst)
+print(ratios_s_best)
 
 rects2 = plt.bar(index+2*bar_width, ratios_d_worst, bar_width, alpha=1, color='r', label='double precision worst case')
 rects2 = plt.bar(index+3*bar_width, ratios_d_best, bar_width, alpha=.5, color='r', label='double precision best case')

### 2014-04-16 acm/ieee_paper_final.tex
minor

@@
 derivatives within the context of Radial Basis Function
-Finite-Difference (RBFFD). RBF methods express functions as a linear
+Finite-Difference (RBF-FD). RBF methods express functions as a linear
 combination of \ge{spherically} symmetric basis functions on an arbitrary set of
@@
 
-\ee{The modelization of Section~\ref{sec:model} establishes that the 
+\ee{The model of Section~\ref{sec:model} establishes that the 
   memory
@@
     512 rows and 32 nonzeros per row. The last two matrices
-    corresponds to a derivative stencil in 3D RBFFD
+    corresponds to a derivative stencil in 3D RBF-FD
     calculations (with and without RCM ordering).}
@@
 % was width=.48
-  \subfigure[Compact and RBFFD.]{\includegraphics[width=.35\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
+  \subfigure[Compact and RBF-FD.]{\includegraphics[width=.35\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
 
@@
 The other type of matrix is used in derivative stencils in 3D
-RBFFD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-point
+RBF-FD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-point
 stencil of a 3D $8^3$ grid).
@@
 operators acting on multiple vectors within the context of
-RBFFD. Each derivative has an associated sparse matrix with a fixed
+RBF-FD. Each derivative has an associated sparse matrix with a fixed
 number of nonzeros on each row. While computing a single derivative of
@@
 the techniques presented in this paper within a fluid simulation using
-RBFFD. \ge{We will also look into developing a similar technique on the Kepler
+RBF-FD. \ge{We will also look into developing a similar technique on the Kepler
 family of GPUs, which offers swizzling and channeling operations.}

### 2014-04-16 acm/acm_paper_submitted.tex
acm_paper_final_truetype.pdf and acm_paper_submitted.tex are the sources submitted ot the ACM conference (camera ready). (and dummy.ps)

--- 2014-04-16 acm/ieee_paper_final.tex
@@
 \centering
-\includegraphics[width=0.60\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
+\includegraphics[width=0.60\linewidth]{figures/RBFStencil_n32_a.pdf}\label{fig:stencil}
 \caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square.
@@
 Our implementation on Cascade reaches 45 Gflops for the 4/4\, case, or 45/120=37\% of peak theoretical 
-performance on this particular processor setup. For references, we achieve 
+performance on this particular processor setup. For reference, we achieve 
 140/210=66\% of peak theoretical performance on the MIC. Note that peak performance 
 is not that of the processor, but of the best possible algorithm on the processor. 
-The gain in the 4/4 case with respect to 4x 1/1 case is respectively 4x on the MIC 
-and 2.5x on the front end.  
+The gain in the 4/4 case with respect to 4x 1/1 case is 4x on the MIC 
+and 2.5x on the front end, respectively.  
 }

### 2014-04-17 acm/acm_paper_submitted.tex
Updated paper. Slightly larger figures, added some space below some figures. Column width check.
pdfltx acm_paper_submitted.tex

@@
 \input{support/misc_mac}
-\input{support/setupicase}
+% resets margins. Bad. 
+%\input{support/setupicase}
 
@@
 
-\section{Introduction and Motivation for RBF-FD}
+\pagebreak
+\section{Introduction and Motivation \\for RBF-FD}
 \label{sec:rbf}
@@
 \centering
-\includegraphics[width=0.60\linewidth]{figures/RBFStencil_n32_a.pdf}\label{fig:stencil}
+\includegraphics[width=0.80\linewidth]{figures/RBFStencil_n32_a.pdf}\label{fig:stencil}
 \caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square.
 In nonsparse format, the DM  is $1024\times1024$ with 32 nonzero entries per row.}
+\vspace{3ex}
 \end{figure}
@@
   \centering
-  \subfigure[Best case.]{\includegraphics[width=.28\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
-  \subfigure[Worst case.]{\includegraphics[width=.28\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
-  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.28\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}
+  \subfigure[Best case.]{\includegraphics[width=.31\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
+  \subfigure[Worst case.]{\includegraphics[width=.31\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
+  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.31\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}
 
@@
   \centering 
-  \includegraphics[width=.9\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}
+  \includegraphics[width=.95\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}
   \caption{Estimation of the maximum achievable performance using a
@@
   \label{fig:perf_predict}
+  \vspace{3ex}
 \end{figure}
@@
 four functions, at a single point. %\ger{All matrices are stored as one-dimensional vectors and pointers to individual rows and elements are explicitely computed.}
-\looseness=-1
+%GE \looseness=-1
 
@@
   \label{fig:band_rw_MIC}
+  \vspace{2ex}
 \end{figure}
@@
   \label{fig:band_gather} 
+  \vspace{2ex}
 \end{figure}
@@
 % was width=.48
-  \subfigure[Supercompact and Random.]{\includegraphics[width=.35\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
+  \subfigure[Supercompact and Random.]{\includegraphics[width=.45\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
 % was width=.48
-  \subfigure[Compact and RBF-FD.]{\includegraphics[width=.35\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
+  \subfigure[Compact and RBF-FD.]{\includegraphics[width=.45\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}
 
@@
 
-  \includegraphics[width=.8\linewidth]{figures/mic_performance_nb_threads.pdf}
+  \includegraphics[width=.9\linewidth]{figures/mic_performance_nb_threads.pdf}
 
@@
     We also show peak theoretical performance for v1m1 and v4m1 on the MIC. 
-    Not shown is the peak theoretical performance of 210 Gflops for v4m4. 
-    }
+    Not shown is the peak theoretical performance of 210 Gflops for v4m4.  }
   \label{fig:comparison_spmv}
   \label{fig:perf_mic}
+  \vspace{2ex}
 \end{figure}
@@
 \section*{Acknowledgment}
-Erlebacher acknowledges funding from NSF grant
+Erlebacher was supported under NSF grant
 DMS-\#0934331 (FSU), and the use of the MIC at FSU and at LCSE/U. Minnesota (provided by Prof. D. Yuen). NCAR is sponsored by NSF. 

### 2014-06-09 slides/2014_ics/slides-ics.tex
draft of slides

\documentclass{beamer}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{epsfig}
%\usepackage{figlatex}
\usepackage{verbatim}
%\usepackage{listings}

\usepackage{subfigure}

%\usepackage{algorithm}
%\usepackage[noend]{algpseudocode}

\usepackage[ruled,vlined]{algorithm2e}

\usepackage{float}

\usetheme{Madrid}
 \setbeamertemplate{navigation symbols}{}
 \setbeamercovered{transparent}

\newcommand{\card}[1]{\ensuremath{|#1|}}
\newcommand{\union}{\ensuremath{\cup}}

\newcommand{\ceil}[1]{\left\lceil#1\right\rceil}
\newcommand{\floor}[1]{\lfloor#1\rfloor}
\newcommand{\todo}[1]{{\bf TODO: #1}}

\usealerttemplate{\color{red}\bf}{}

\graphicspath{{fig/}}

\newcommand{\MIC}{Intel MIC\xspace}

\title[RBF-FD on \MIC]{Acceleration of Derivative Calculations with
  Application to Radial Basis Function ­ Finite-Differences on the
  Intel MIC Architecture}

\date{ICS 2014}

\author[Erik Saule]{Gordon Erlebacher, {\bf Erik Saule}, Natasha Flyer, Evan
  Bollig}

\institute[UNCC]{Florida State University\\{\bf University of North Carolina at Charlotte}\\UCAR\\University of Minnesota}

\begin{document}

\maketitle

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[hidesubsection,hidesubsubsection]
\end{frame}

\AtBeginSection[]
{
  \begin{frame}<beamer>
    \frametitle{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

\section{Introduction}

\subsection{RBF-FD}

\begin{frame}
  \frametitle{Derivative as Linear operators}
\end{frame}

\begin{frame}
  \frametitle{RBF-FD methods}

  \todo{Add definition of RBF}

  \begin{columns}
    \column{.45\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/RBFStencil_n32_a}

    An example of an RBF-FD stencil of size $n=32$ on a sphere of
    $N=1024$ nodes to approximate any derivative operator at the point
    in the square.

    \column{.55\linewidth}

    \begin{block}{Shallow waters equations of a rotating fluid}
      \tiny
      \begin{eqnarray}
        \dfrac{\partial u}{\partial t} &=& - \left( u\dfrac{\partial u}{\partial x} + v\dfrac{\partial u}{\partial y} + w\dfrac{\partial u}{\partial z} + f(yw - zv) + g\dfrac{\partial h}{\partial x} \right) 
        \nonumber \\
        \dfrac{\partial v}{\partial t} &=& - \left( u\dfrac{\partial v}{\partial x} + v\dfrac{\partial v}{\partial y} + w\dfrac{\partial v}{\partial z} + f(zu - xw) + g\dfrac{\partial h}{\partial x} \right)
        \nonumber\\
        \dfrac{\partial w}{\partial t} &=& - \left( u\dfrac{\partial w}{\partial x} + v\dfrac{\partial w}{\partial y} + w\dfrac{\partial w}{\partial z} + f(xv - yu) + g\dfrac{\partial h}{\partial x} \right)
        \nonumber\\
        \frac{\partial h}{\partial t}& =&-\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right) , \label{height} \nonumber
      \end{eqnarray}
    \end{block}

    How to compute one simple derivative: sparse matrix
    multiplication. Bound by (roughly) 25GFlop/s. Can we do better?
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Compute all the simple derivative at once}

%  \begin{columns}
%    \column{.5\linewidth}
    \begin{block}{Idea}
      \scriptsize
      \begin{equation}
        \left( \begin{array}{cccc}
            \underline{u}_x     & \underline{v}_x     & \underline{w}_x     & \underline{h}_x \\
            \underline{u}_y     & \underline{v}_y     & \underline{w}_y     & \underline{h}_y \\
            \underline{u}_z     & \underline{v}_z     & \underline{w}_z     & \underline{h}_z \\
            \underline{u}_{hyp} & \underline{v}_{hyp} & \underline{w}_{hyp} & \underline{h}_{hyp}
          \end{array} \right)
        = \left(
          \begin{array}{c}
            D_{x}\\ D_{y}\\ D_{z}\\ D_{hyp}
          \end{array}\right)
        \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) 
        \vspace{-1ex}
      \end{equation}
    \end{block}  
%    \column{.5\linewidth}

    \begin{block}{Why would it work}
      All the derivative matrices have the same sparsity pattern. In
      other words, the non zeros are at the same positions in the
      matrix, only the values of the non zero change.
    \end{block}

%  \end{columns}
\end{frame}

\subsection{MIC}

\begin{frame}
  \frametitle{What is Intel MIC}
  Intel Many integrated Core (MIC) is Intel's response to GPUs
  becoming popular in High Performance Computing.

  \begin{block}{What GPUs do well?}
    \begin{itemize}
      \item Get a lot of GFlops by using hundreds of cores
      \item Each core has large SIMD-like abilities
      \item Hide memory latency by using fast cycle context switch
    \end{itemize}
  \end{block}

  \begin{block}{What GPUs do not do well?}
    \begin{itemize}
      \item Alien to program
      \item Poor support for legacy applications
      \item Inter thread communications
      \item Branching
    \end{itemize}
  \end{block}

  Goal of \MIC : do all of it well!
\end{frame}

\begin{frame}
  \frametitle{Overall Architecture}
  \begin{columns}
    \column{18em}    
    \includegraphics[width=18em]{slides-figures/MIC-overview.pdf}
    \column{17em}

    \begin{itemize}
      \item 8 memory controllers
        \begin{itemize}
          \item GDDR5
          \item 2 channels (32-bit)
          \item 5.5GT/s
          \item 352GB/s aggregated peak
            \begin{itemize}
              \item twice the GPU's
              \item but you typically get half
            \end{itemize}
        \end{itemize}
      \item 50+ cores
        \begin{itemize}
          \item 32KB of L1 cache
          \item 512KB of L2 cache
          \item LRU
          \item 8-way associative
        \end{itemize}
      \item 1 PCI-e controller
        \begin{itemize}
          \item to the host (2GB/s guaranteed to memory)
        \end{itemize}
    \end{itemize}
  \end{columns}

\end{frame}

\begin{frame}
  \frametitle{Core architecture}
  \begin{columns}
    \column{20em}
    \includegraphics[width=20em,height=20em]{slides-figures/MIC-core.png}%from intelsdg  
    \column{12em}
    \begin{itemize}
      \small
    \item Clocked at 1Ghz
    \item 64-bit
    \item 4 hardware threads
      \begin{itemize}
        \footnotesize
        \item no context switching
        \item no 2 instructions for the same thread
      \end{itemize}
    \item A vectorial unit
      \begin{itemize}
        \footnotesize
        \item (more details on the next slides)
      \end{itemize}
    \item Two instruction pipes:
      \begin{itemize}
        \footnotesize
        \item 2 ALU ops
        \item ALU + MEM ops
        \item ALU + VFP ops
        \item VFP + MEM ops
      \end{itemize}
    \item In-order execution
    \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Vectorial Unit (SIMD) : The classics}

  \begin{columns}
    \column{15em}

    \includegraphics[width=15em]{slides-figures/vect-basic.png}%from intelsdg

    \column{15em}

    {\small 32 512-bits registers (16x32, 8x64)}

    \begin{block}{Binary operators (1 cycle throughput)}
      \small mul, add, sub, and, xor, or, fma
    \end{block}

    \begin{block}{Short Vector Math Library}
      sqrt, rsqrt, log, exp, div
    \end{block}

  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{When Should I Use \MIC ?}

  \begin{block}{Key points of SE10P}
    \begin{itemize}
      \item Large memory bandwidth (peak: 220GB/s)
      \item 61 cores with mandatory use of hardware threading
      \item 512-bit wide SIMD registers: 
        \begin{itemize} 
          \item FMA: up to 2x16 SP Flop/c (2x8 DP Flop/c)
          \item otherwise: up to 16 SP Flop/c (8 DP Flop/c)
        \end{itemize}
      \item On a 61-core configuration at 1.05Ghz:
        \begin{itemize} 
        \item FMA: 2x16x61x1.05Ghz = 2.048 TFlop/s SP (1.024TFlop/s DP)
        \item otherwise: 16x61x1.05Ghz = 1.024 GFlop/s SP (512GFlop/s DP)
        \end{itemize}
    \end{itemize}
  \end{block}

  Lots of bandwidth? Fused Multiply Add? Large vector registers?

  Sounds like the perfect system for Sparse Matrix operations!
\end{frame}

\begin{frame}
  \frametitle{Getting actual bandwidth}

  \begin{columns}
    \column{.49\linewidth}

    \begin{center}
      \includegraphics[width=\linewidth,page=1]{slides-figures/PPAM-data/bandwidth.pdf}

      Read
    \end{center}

    \column{.49\linewidth}

    \begin{center}
      \includegraphics[width=\linewidth,page=2]{slides-figures/PPAM-data/bandwidth.pdf}

      Write
    \end{center}

  \end{columns}

  \begin{itemize}
  \item Using the appropriate vectorial instructions gives significant
    improvements.
  \item Read Peak: 183GB/s.
  \item Write Peak: 160GB/s.
  \end{itemize}

\end{frame}

\section{Performance Prediction}

\begin{frame}
  \frametitle{Is there hope in this technique? Let's model it.}

  \begin{columns}
    \column{.4\linewidth}
    \scalebox{.55}{
      \begin{tabular}{|c|l|}
        % \hline
        % & & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row of $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of nonzeros\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the $n_m$ matrices in bytes\\
        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred \\
        $O$     & number of floating point operations \\
        $I_b$   & maximum computational intensity\\
        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }
    \column{.6\linewidth}

    Size of the input vector:\\$s_x = n_v b_x n_r$

    Size of the matrix:\\$s_M = n_c (b_i + b_x n_m) = n_r n_z (b_i + b_x n_m)$

    Size of the write (output vector):\\$b_{wT} = s_y = n_v n_m b_x n_r$

    Data Read (best case):\\$b_{rT} = s_M + s_x$

    Data Read (worst case):\\$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}}$

    Number of floating point operation:\\$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$
  \end{columns}

  \begin{center}
    Intensity (Best case):\\
    $I_b = \frac{O}{b_T+ b_{wT}} = \frac{2 n_v n_m}{ (b_i + b_x n_m) + n_v n_m b_x n_z^{-1} + n_v b_x n_z^{-1} }$

    Intensity (Worst case):\\
    $I_w = \frac{O}{B_T+ b_{wT}} = \frac{2 n_v n_m}{(b_i+b_x n_m) + n_v n_m b_x n_z^{-1} + cl \ceil{\frac{n_vb_x}{cl}} }$
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Ok... So will it work? Flop to byte ratio}

  \begin{columns}
    \column{.33\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_best-crop.pdf}

    \begin{center}
      Best case\\~
    \end{center}

    \column{.33\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_worst-crop.pdf}
    \begin{center}
      Worst case\\~
    \end{center}

    \column{.33\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_no_cache-crop.pdf}
    \begin{center}
      Worst case \\no cacheline
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{In terms of flops?}

  \includegraphics[width=.9\linewidth]{slides-figures/ICS-figures/gflops_peak.pdf}

\end{frame}

\section{Implementation}

\begin{frame}
  \frametitle{Vectorial Unit (SIMD) : Rearranging}

  \begin{columns}
    \column{.45\linewidth}

    \includegraphics[width=15em]{slides-figures/vect-swizzling.png}%from intelsdg

    \column{.45\linewidth}

    Registers are organized in 4 lanes.

    \begin{block}{Permutation}
      Permutation instructions allow to reorder lanes according to
      given patterns.
    \end{block}

    \begin{block}{Swizzling}
      Almost all instructions support recombining elements within a
      lane. As it is part of the instruction, swizzling comes with no
      throughput penalty (but certainly a high latency). These also
      typically follow a given set of patterns.
    \end{block}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Vectorial Unit (SIMD) : Unaligned loads/stores}

  \begin{columns}
    \column{.45\linewidth}

    \begin{block}{Gather}
      Allows to load data from memory at arbitrary offsets. Takes one
      cycle per cache line accessed.
    \end{block}

    \includegraphics[width=\linewidth]{slides-figures/gather.pdf}

    \begin{block}{Scatter}
      Same principle as gather but does store operations.
    \end{block}

    \column{.45\linewidth}

    \begin{block}{Unpack}
      Allows to load packed data from memory using a bit mask write.
    \end{block}

    \includegraphics[width=\linewidth]{slides-figures/gather.pdf}

    \begin{block}{Pack}
      Same principle as unpack but does store operations.
    \end{block}    
  \end{columns}  
\end{frame}

\begin{frame}
  \frametitle{How to do it? (4 vectors, 4 matrices)}

  \begin{block}{Observations}
    \begin{itemize}
    \item We aim at over 200Gflop/s in single precision. (\~{}  10\% of peak.)
    \item One instruction in ten must be a Fused Multiply Add.
    \end{itemize}
  \end{block}

  \begin{figure}
    \centering
    \scalebox{.85}{
      \subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{slides-figures/ICS-figures/code.pdf}}%
      \subfigure[Content of the vector registers.]{\hspace{-0.07\linewidth}\includegraphics[width=.55\linewidth]{slides-figures/ICS-figures/code-annote.pdf}}
    }
  \end{figure}  
\end{frame}

\section{Experiments}

\begin{frame}
  \frametitle{Results : Extreme cases}
  \begin{columns}
    \column{.4\linewidth}

    \begin{block}{Supercompact}
      \centering
      \includegraphics[width=.6\linewidth]{slides-figures/ICS-figures/supercompact_matrix-crop.png}
    \end{block}

    \begin{block}{Random}
      \centering
      \includegraphics[width=.6\linewidth]{slides-figures/ICS-figures/random_matrix-crop.png}
    \end{block}

    \column{.6\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/random_supercompact.pdf}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Results : Real cases}
  \begin{columns}
    \column{.4\linewidth}
    \begin{block}{Compact}
      \centering
      \includegraphics[width=.6\linewidth]{slides-figures/ICS-figures/compact_matrix-crop.png}
    \end{block}

    \begin{block}{RBF-FD}
      \centering
      \includegraphics[width=.6\linewidth]{slides-figures/ICS-figures/kd-tree-3d-norcm-crop.png}
    \end{block}

    \column{.6\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/rbf_compact.pdf}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Does it help?}
  \begin{center}
    \includegraphics[width=.7\linewidth]{slides-figures/ICS-figures/mic_performance_nb_threads.pdf}
  \end{center}

  Performance of $y=Ax$ for a RBF derivative stencil of a $96^3$ 3D
  grid. Shown: The peaks of common methods
\end{frame}

\section{Conclusion}

\begin{frame}
  \frame{Conclusion}

  \begin{block}{Conclusion}
  \end{block}

  \begin{block}{Future Works}
    \begin{itemize}
    \item More than 1 Xeon Phi?
    \item More than 1 node?
    \item Perform a complete simulation.
    \item GPU?
    \end{itemize}
  \end{block}
\end{frame}

\bibliographystyle{alpha}
\bibliography{slides}

\end{document}

### 2014-06-09 slides/2014_ics/slides-figures/ICS-figures/plot_for_paper_figure11.py

@@ -1,0 +1,108 @@
+
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+import csv
+
+#----------------------------------
+
+# 3 columns: filenmame, rows, bandwidth
+def plot_file(infile,outfile):
+    dict = {}
+    #dict['64_norcm'] = {}
+    dict['96_norcm'] = {}
+    #dict['64_rcm'] = {}
+    dict['96_rcm'] = {}
+    with open(infile, 'r') as csvfile:
+        csvfile.seek(0)
+        reader = csv.reader(csvfile, delimiter=' ')
+        for row in reader:
+            if row == []: break
+            if row[0][0] == '#': continue
+            rfile = row[0]
+            method = row[1]
+            #print method
+            r1 = float(row[2])
+            r2 = float(row[3])
+
+            try:
+                #if method != "method_8a_multi_cpp":
+                    dict[rfile][method].append([r1,r2])
+            except:
+                try:
+                    #print(rfile)
+                    dict[rfile][method] = []
+                    dict[rfile][method].append([r1,r2])
+                except:
+                    pass
+
+    for f in dict.keys():
+        for m in dict[f].keys():
+            dict[f][m] = zip(*dict[f][m])  # * means transpose
+
+
+#### BEGINNING OF CHART
+
+    def plot_file(xlist, ylist, col, sym, leg):
+        lines = ['-','-','-','-']
+        count = 0
+        x = xlist
+        y = ylist
+        p = plt.plot(x,y, col+sym+'-', label=leg, ms=6, lw=2)
+        lines.append(p)
+        count += 1
+
+
+    cols = ['b','r','m','c','k','g']
+    syms = ['o','^','s','v','h','x']
+    fi = 0
+    #for f in dict.keys():
+        #print "** key= ", f
+
+    """
+    for f in dict.keys():
+        print "** key= ", f
+        col = cols[fi]; fi+=1; mi = 0
+        for m in dict[f].keys():
+            print "  ++ key= ", m
+            sym = syms[mi]; mi+=1
+            leg = f + "_" + m
+            plot_file(dict[f][m][0], dict[f][m][1], col, sym, leg)
+    """
+
+    d = dict['96_rcm']['method_8a_multi']
[…]

### 2014-06-10 slides/2014_ics/slides-ics.tex
progress on slides

@@
 
-\begin{frame}
-  \frametitle{What is Intel MIC}
-  Intel Many integrated Core (MIC) is Intel's response to GPUs
-  becoming popular in High Performance Computing.
+% \begin{frame}
+%   \frametitle{What is Intel MIC}
+%   Intel Many integrated Core (MIC) is Intel's response to GPUs
+%   becoming popular in High Performance Computing.
 
-  \begin{block}{What GPUs do well?}
-    \begin{itemize}
-      \item Get a lot of GFlops by using hundreds of cores
-      \item Each core has large SIMD-like abilities
-      \item Hide memory latency by using fast cycle context switch
-    \end{itemize}
-  \end{block}
+%   \begin{block}{What GPUs do well?}
+%     \begin{itemize}
+%       \item Get a lot of GFlops by using hundreds of cores
+%       \item Each core has large SIMD-like abilities
+%       \item Hide memory latency by using fast cycle context switch
+%     \end{itemize}
+%   \end{block}
 
-  \begin{block}{What GPUs do not do well?}
-    \begin{itemize}
-      \item Alien to program
-      \item Poor support for legacy applications
-      \item Inter thread communications
-      \item Branching
-    \end{itemize}
-  \end{block}
+%   \begin{block}{What GPUs do not do well?}
+%     \begin{itemize}
+%       \item Alien to program
+%       \item Poor support for legacy applications
+%       \item Inter thread communications
+%       \item Branching
+%     \end{itemize}
+%   \end{block}
 
-  Goal of \MIC : do all of it well!
-\end{frame}
+%   Goal of \MIC : do all of it well!
+% \end{frame}
 
 \begin{frame}
-  \frametitle{Overall Architecture}
+  \frametitle{Intel MIC: Overall Architecture}
   \begin{columns}
@@
             \begin{itemize}
-              \item twice the GPU's
+              \item about same as a K40 
               \item but you typically get half
@@
         \footnotesize
-        \item (more details on the next slides)
+        \item 512-bit registers
+        \item 16 SP floats
+        \item 8 DP floats
+        \item support FMA
       \end{itemize}
@@
 
-\begin{frame}
-  \frametitle{Vectorial Unit (SIMD) : The classics}
+% \begin{frame}
+%   \frametitle{Vectorial Unit (SIMD) : The classics}
 
-  \begin{columns}
-    \column{15em}
+%   \begin{columns}
+%     \column{15em}
 
-    \includegraphics[width=15em]{slides-figures/vect-basic.png}%from intelsdg
+%     \includegraphics[width=15em]{slides-figures/vect-basic.png}%from intelsdg
 
-    \column{15em}
+%     \column{15em}
 
-    {\small 32 512-bits registers (16x32, 8x64)}
+%     {\small 32 512-bits registers (16x32, 8x64)}
 
-    \begin{block}{Binary operators (1 cycle throughput)}
-      \small mul, add, sub, and, xor, or, fma
-    \end{block}
+%     \begin{block}{Binary operators (1 cycle throughput)}
+%       \small mul, add, sub, and, xor, or, fma
+%     \end{block}
 
-    \begin{block}{Short Vector Math Library}
-      sqrt, rsqrt, log, exp, div
-    \end{block}
+%     \begin{block}{Short Vector Math Library}
+%       sqrt, rsqrt, log, exp, div
+%     \end{block}
 
-  \end{columns}
-\end{frame}
+%   \end{columns}
+% \end{frame}
 
@@
 \begin{frame}
-  \frametitle{Getting actual bandwidth}
+  \frametitle{Have to use vector operations to get high bandwidth}
 
@@
 
+  \tiny Source: PPAM 2013
 \end{frame}
@@
   \begin{columns}
-    \column{.33\linewidth}
+    \column{.49\linewidth}
     \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_best-crop.pdf}
@@
 
-    \column{.33\linewidth}
+    \column{.49\linewidth}
     \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_worst-crop.pdf}
@@
 
-    \column{.33\linewidth}
-    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_no_cache-crop.pdf}
-    \begin{center}
-      Worst case \\no cacheline
-    \end{center}
+    % \column{.33\linewidth}
+    % \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_no_cache-crop.pdf}
+    % \begin{center}
+    %   Worst case \\no cacheline
+    % \end{center}
   \end{columns}
+
+  \tiny Also, taking cacheline into account makes a large difference (see paper for details).
 \end{frame}
@@
 \begin{frame}
-  \frametitle{In terms of flops?}
+  \frametitle{In terms of flops? (Assuming memory bandwidth bound)}
 
@@
 
-\section{Implementation}
+\section{Implementation (4 vectors, 4 matrices)}
+
+\begin{frame}
+  \frametitle{Observation on the 4 vectors, 4 matrices case}
+
+  \begin{itemize}
+  \item We aim at over 200Gflop/s in single precision. (\~{}  10\% of peak.)
+  \item One instruction in ten must be a Fused Multiply Add.
+  \item The kernel must be fairly instruction efficient
+  \end{itemize}
+\end{frame}
 
@@
 
-    \includegraphics[width=15em]{slides-figures/vect-swizzling.png}%from intelsdg
+    \hspace*{-.1\linewidth}\includegraphics[width=1.2\linewidth]{slides-figures/vect-swizzling.png}%from intelsdg
 
@@
 \begin{frame}
-  \frametitle{Vectorial Unit (SIMD) : Unaligned loads/stores}
+  \frametitle{Vectorial Unit (SIMD) : smarter loads}
 
@@
     \end{block}
-
     \includegraphics[width=\linewidth]{slides-figures/gather.pdf}
-
-    \begin{block}{Scatter}
-      Same principle as gather but does store operations.
-    \end{block}
 
@@
 
-    \includegraphics[width=\linewidth]{slides-figures/gather.pdf}
+    \includegraphics[width=\linewidth]{slides-figures/unpack.pdf}
 
-    \begin{block}{Pack}
-      Same principle as unpack but does store operations.
-    \end{block}    
   \end{columns}
@@
 
-  \begin{block}{Observations}
-    \begin{itemize}
-    \item We aim at over 200Gflop/s in single precision. (\~{}  10\% of peak.)
-    \item One instruction in ten must be a Fused Multiply Add.
-    \end{itemize}
-  \end{block}
-
   \begin{figure}
     \centering
-    \scalebox{.85}{
+    \hspace*{-1em}\scalebox{.95}{
       \subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{slides-figures/ICS-figures/code.pdf}}%

### 2014-06-10 slides/2014_ics/slides-ics.tex
ccl

@@
 \begin{frame}
-  \frame{Conclusion}
+  \frametitle{Conclusion}
 
   \begin{block}{Conclusion}
+    \begin{itemize}
+    \item RBF-FD needs to compute different derivative of multiple functions.
+    \item The different derivatives have the same sparsity structure.
+    \item Idea: perform all the derivatives simultaneously
+    \item The flop-to-byte ratio increases significantly
+    \item Vector reordering operations of Xeon Phi enables efficient execution
+    \item Performs efficiently: 
+      \begin{itemize}
+      \item 3.5 times faster than peak of traditional SpMV  
+      \item faster than peak of SpMM (v=4)
+      \end{itemize}
+    \end{itemize}
   \end{block}

### 2014-06-10 slides/2014_ics/slides-ics.tex
first slide on derivative

@@
 
-\begin{frame}
-  \frametitle{Derivative as Linear operators}
+\begin{frame}[fragile]
+  \frametitle{Derivative as linear algebra operators}
+
+  \begin{columns}
+    \column{.5\linewidth}
+    \begin{block}{2nd derivative, 1D}
+
+      \includegraphics[width=\linewidth]{slides-figures/1d.pdf}
+
+      $$u''_i = \frac{u_{i+1}-2u_i+u_{i-1}}{h^2}$$
+      \scalebox{.8}{$A = \begin{pmatrix}
+        \dots & \dots & \dots & \dots & \dots & \dots & \dots \\
+        \dots & 1 & -2 & 1 & 0 & 0 & \dots  \\
+        \dots & 0 & 1 & -2 & 1 & 0 & \dots \\
+        \dots & 0 & 0 & 1 & -2 & 1 & \dots  \\
+        \dots & \dots & \dots & \dots & \dots & \dots & \dots
+      \end{pmatrix}$}
+      $$u'' = Au$$
+    \end{block}
+
+    \column{.49\linewidth}
+    \begin{block}{2nd derivative, 2D}
+      \begin{center}
+        \includegraphics[width=.6\linewidth]{slides-figures/2d.pdf}
+      \end{center}
+
+        $L_{i,j} = \bigtriangledown^2 u_{i,j} = \frac{1}{h^2} (u_{i+1,j}  +u_{i-1,j}$
+
+        $~~~~~~~~~~~~~~~~+u_{i,j-1}+u_{i,j+1}-4u_i)$
+    \end{block}
+
+  \end{columns}
 \end{frame}
@@
   \end{columns}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Compute all the simple derivative at once}
-
-%  \begin{columns}
-%    \column{.5\linewidth}
-    \begin{block}{Idea}
-      \scriptsize
-      \begin{equation}
-        \left( \begin{array}{cccc}
-            \underline{u}_x     & \underline{v}_x     & \underline{w}_x     & \underline{h}_x \\
-            \underline{u}_y     & \underline{v}_y     & \underline{w}_y     & \underline{h}_y \\
-            \underline{u}_z     & \underline{v}_z     & \underline{w}_z     & \underline{h}_z \\
-            \underline{u}_{hyp} & \underline{v}_{hyp} & \underline{w}_{hyp} & \underline{h}_{hyp}
-          \end{array} \right)
-        = \left(
-          \begin{array}{c}
-            D_{x}\\ D_{y}\\ D_{z}\\ D_{hyp}
-          \end{array}\right)
-        \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) 
-        \vspace{-1ex}
-      \end{equation}
-    \end{block}  
-%    \column{.5\linewidth}
-
-    \begin{block}{Why would it work}
-      All the derivative matrices have the same sparsity pattern. In
-      other words, the non zeros are at the same positions in the
-      matrix, only the values of the non zero change.
-    \end{block}
-
-%  \end{columns}
 \end{frame}
@@
 
+% \begin{frame}
+%   \frametitle{Have to use vector operations to get high bandwidth}
+
+%   \begin{columns}
+%     \column{.49\linewidth}
+
+%     \begin{center}
+%       \includegraphics[width=\linewidth,page=1]{slides-figures/PPAM-data/bandwidth.pdf}
+
+%       Read
+%     \end{center}
+
+%     \column{.49\linewidth}
+
+%     \begin{center}
+%       \includegraphics[width=\linewidth,page=2]{slides-figures/PPAM-data/bandwidth.pdf}
+
+%       Write
+%     \end{center}
+
+%   \end{columns}
+
+%   \begin{itemize}
+%   \item Using the appropriate vectorial instructions gives significant
+%     improvements.
+%   \item Read Peak: 183GB/s.
+%   \item Write Peak: 160GB/s.
+%   \end{itemize}
+
+%   \tiny Source: PPAM 2013
+% \end{frame}
+
 \begin{frame}
-  \frametitle{Have to use vector operations to get high bandwidth}
+  \frametitle{Compute all the simple derivative at once}
 
-  \begin{columns}
-    \column{.49\linewidth}
+%  \begin{columns}
+%    \column{.5\linewidth}
+    \begin{block}{Idea}
+      \scriptsize
+      \begin{equation}
+        \left( \begin{array}{cccc}
+            \underline{u}_x     & \underline{v}_x     & \underline{w}_x     & \underline{h}_x \\
+            \underline{u}_y     & \underline{v}_y     & \underline{w}_y     & \underline{h}_y \\
+            \underline{u}_z     & \underline{v}_z     & \underline{w}_z     & \underline{h}_z \\
+            \underline{u}_{hyp} & \underline{v}_{hyp} & \underline{w}_{hyp} & \underline{h}_{hyp}
+          \end{array} \right)
+        = \left(
+          \begin{array}{c}
+            D_{x}\\ D_{y}\\ D_{z}\\ D_{hyp}
+          \end{array}\right)
+        \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) 
+        \vspace{-1ex}
+      \end{equation}
+    \end{block}  
+%    \column{.5\linewidth}
 
-    \begin{center}
-      \includegraphics[width=\linewidth,page=1]{slides-figures/PPAM-data/bandwidth.pdf}
+    \begin{block}{Why would it work}
+      All the derivative matrices have the same sparsity pattern. In
+      other words, the non zeros are at the same positions in the
+      matrix, only the values of the non zero change.
+    \end{block}
 
-      Read
-    \end{center}
-
-    \column{.49\linewidth}
-
-    \begin{center}
-      \includegraphics[width=\linewidth,page=2]{slides-figures/PPAM-data/bandwidth.pdf}
-
-      Write
-    \end{center}
-
-  \end{columns}
-
-  \begin{itemize}
-  \item Using the appropriate vectorial instructions gives significant
-    improvements.
-  \item Read Peak: 183GB/s.
-  \item Write Peak: 160GB/s.
-  \end{itemize}
-
-  \tiny Source: PPAM 2013
+%  \end{columns}
 \end{frame}

### 2014-06-12 slides/2014_ics/slides-ics.tex
slides

@@
 
+\section{Radial Basis Function - Finite Difference}
+
 \AtBeginSection[]
@@
 
-\section{Introduction}
-
-\subsection{RBF-FD}
-
 \begin{frame}[fragile]
-  \frametitle{Derivative as linear algebra operators}
+  \frametitle{Derivative as (sparse) linear algebra operators}
 
@@
 
-\begin{frame}
-  \frametitle{RBF-FD methods}
-
-  \todo{Add definition of RBF}
-
+\begin{frame}[fragile]
+  \frametitle{Radial Basis Function}
   \begin{columns}
     \column{.45\linewidth}
-    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/RBFStencil_n32_a}
 
-    An example of an RBF-FD stencil of size $n=32$ on a sphere of
-    $N=1024$ nodes to approximate any derivative operator at the point
-    in the square.
+    \begin{block}{RBF}
+      \includegraphics[width=\linewidth]{slides-figures/rbf.pdf}
 
-    \column{.55\linewidth}
+      \begin{itemize}
+      \item Radial symmetry.
+      \item RBFs approximate a function $f$ sampled at N distinct location $$f(\vec{r}) = \sum_i w_i \phi(||\vec{r}-\vec{r_i}||)$$ 
+      \end{itemize}
+    \end{block}
 
-    \begin{block}{Shallow waters equations of a rotating fluid}
-      \tiny
+    \column{.45\linewidth}
+    \begin{block}{The RBF-FD method}
+      \includegraphics[width=.8\linewidth]{slides-figures/ICS-figures/RBFStencil_n32_a}
+
+      Derivative operators of an RBF are approximated using the $n=32$
+      closest sampling points on a sphere of $N=1024$ points.
+
+      The DM has exactly $n$ non zero per row.
+    \end{block}
+
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Computing derivatives. One at a time?}
+
+  \begin{block}{Shallow waters equations}
+
     \begin{eqnarray}
@@
       \nonumber \\
-        \dfrac{\partial v}{\partial t} &=& - \left( u\dfrac{\partial v}{\partial x} + v\dfrac{\partial v}{\partial y} + w\dfrac{\partial v}{\partial z} + f(zu - xw) + g\dfrac{\partial h}{\partial x} \right)
+      \dfrac{\partial v}{\partial t} &=& - \left( u\dfrac{\partial v}{\partial x} + v\dfrac{\partial v}{\partial y} + w\dfrac{\partial v}{\partial z} + f(zu - xw) + g\dfrac{\partial h}{\partial y} \right)
       \nonumber\\
-        \dfrac{\partial w}{\partial t} &=& - \left( u\dfrac{\partial w}{\partial x} + v\dfrac{\partial w}{\partial y} + w\dfrac{\partial w}{\partial z} + f(xv - yu) + g\dfrac{\partial h}{\partial x} \right)
+      \dfrac{\partial w}{\partial t} &=& - \left( u\dfrac{\partial w}{\partial x} + v\dfrac{\partial w}{\partial y} + w\dfrac{\partial w}{\partial z} + f(xv - yu) + g\dfrac{\partial h}{\partial z} \right)
       \nonumber\\
-        \frac{\partial h}{\partial t}& =&-\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right) , \label{height} \nonumber
+      \frac{\partial h}{\partial t} &=& -\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right)  \nonumber
     \end{eqnarray}
+    $u,v,w$ velocity. $h$ geopotential height.
   \end{block}
 
-    How to compute one simple derivative: sparse matrix
-    multiplication. Bound by (roughly) 25GFlop/s. Can we do better?
-  \end{columns}
-\end{frame}
+  One at a time? $\dfrac{\partial u}{\partial x}$, then $\dfrac{\partial v}{\partial y}$
 
-\subsection{MIC}
-
-% \begin{frame}
-%   \frametitle{What is Intel MIC}
-%   Intel Many integrated Core (MIC) is Intel's response to GPUs
-%   becoming popular in High Performance Computing.
-
-%   \begin{block}{What GPUs do well?}
-%     \begin{itemize}
-%       \item Get a lot of GFlops by using hundreds of cores
-%       \item Each core has large SIMD-like abilities
-%       \item Hide memory latency by using fast cycle context switch
-%     \end{itemize}
-%   \end{block}
-
-%   \begin{block}{What GPUs do not do well?}
-%     \begin{itemize}
-%       \item Alien to program
-%       \item Poor support for legacy applications
-%       \item Inter thread communications
-%       \item Branching
-%     \end{itemize}
-%   \end{block}
-
-%   Goal of \MIC : do all of it well!
-% \end{frame}
-
-\begin{frame}
-  \frametitle{Intel MIC: Overall Architecture}
-  \begin{columns}
-    \column{18em}    
-    \includegraphics[width=18em]{slides-figures/MIC-overview.pdf}
-    \column{17em}
-
-    \begin{itemize}
-      \item 8 memory controllers
-        \begin{itemize}
-          \item GDDR5
-          \item 2 channels (32-bit)
-          \item 5.5GT/s
-          \item 352GB/s aggregated peak
-            \begin{itemize}
-              \item about same as a K40 
-              \item but you typically get half
-            \end{itemize}
-        \end{itemize}
-      \item 50+ cores
-        \begin{itemize}
-          \item 32KB of L1 cache
-          \item 512KB of L2 cache
-          \item LRU
-          \item 8-way associative
-        \end{itemize}
-      \item 1 PCI-e controller
-        \begin{itemize}
-          \item to the host (2GB/s guaranteed to memory)
-        \end{itemize}
-    \end{itemize}
-  \end{columns}
+  \only<2>{\vspace{-19em}\hspace{-1em}\begin{minipage}{.95\linewidth}
+      \colorbox{white}{
+        \includegraphics[width=\linewidth]{slides-figures/PPAM-figures/all_compare_spmv.pdf}
+        \tiny PPAM 2013
+      }
+    \end{minipage}}
 
@@
 \begin{frame}
-  \frametitle{Core architecture}
-  \begin{columns}
-    \column{20em}
-    \includegraphics[width=20em,height=20em]{slides-figures/MIC-core.png}%from intelsdg  
-    \column{12em}
-    \begin{itemize}
-      \small
-    \item Clocked at 1Ghz
-    \item 64-bit
-    \item 4 hardware threads
-      \begin{itemize}
-        \footnotesize
-        \item no context switching
-        \item no 2 instructions for the same thread
-      \end{itemize}
-    \item A vectorial unit
-      \begin{itemize}
-        \footnotesize
-        \item 512-bit registers
-        \item 16 SP floats
-        \item 8 DP floats
-        \item support FMA
-      \end{itemize}
-    \item Two instruction pipes:
-      \begin{itemize}
-        \footnotesize
-        \item 2 ALU ops
-        \item ALU + MEM ops
-        \item ALU + VFP ops
-        \item VFP + MEM ops
-      \end{itemize}
-    \item In-order execution
-    \end{itemize}
-  \end{columns}
-\end{frame}
+  \frametitle{Compute all the simple derivatives at once}
 
-% \begin{frame}
-%   \frametitle{Vectorial Unit (SIMD) : The classics}
+\only<1-2>{  \begin{block}{Shallow waters equations}
+    \begin{eqnarray}
+      \dfrac{\partial u}{\partial t} &=& - \left( u\dfrac{ \textcolor{blue}{\partial}  u}{\textcolor{blue}{\partial x}} + v\dfrac{\textcolor{red}{\partial} u}{\textcolor{red}{\partial y}} + w\dfrac{\textcolor{green}{\partial} u}{\textcolor{green}{\partial z}} + f(yw - zv) + g\dfrac{\textcolor{blue}{\partial} h}{\textcolor{blue}{\partial x}} \right)
+      \nonumber \\
+      \dfrac{\partial v}{\partial t} &=& - \left( u\dfrac{ \textcolor{blue}{\partial} v}{\textcolor{blue}{\partial x}} + v\dfrac{\textcolor{red}{\partial} v}{\textcolor{red}{\partial y}} + w\dfrac{\textcolor{green}{\partial} v}{\textcolor{green}{\partial z}} + f(zu - xw) + g\dfrac{\textcolor{red}{\partial} h}{\textcolor{red}{\partial} y} \right)
+      \nonumber \\
+      \dfrac{\partial w}{\partial t} &=& - \left( u\dfrac{\textcolor{blue}{\partial} w}{\textcolor{blue}{\partial x}} + v\dfrac{\textcolor{red}{\partial} w}{\textcolor{red}{\partial y}} + w\dfrac{\textcolor{green}{\partial} w}{\textcolor{green}{\partial z}} + f(xv - yu) + g\dfrac{\textcolor{green}{\partial} h}{\textcolor{green}{\partial z}} \right)
+      \nonumber\\
+      \frac{\partial h}{\partial t} &=& -\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right)  \nonumber
+    \end{eqnarray}
+  \end{block}}
 
-%   \begin{columns}
-%     \column{15em}
+\pause
 
-%     \includegraphics[width=15em]{slides-figures/vect-basic.png}%from intelsdg
-
-%     \column{15em}
-
-%     {\small 32 512-bits registers (16x32, 8x64)}
-
-%     \begin{block}{Binary operators (1 cycle throughput)}
-%       \small mul, add, sub, and, xor, or, fma
-%     \end{block}
-
-%     \begin{block}{Short Vector Math Library}
-%       sqrt, rsqrt, log, exp, div
-%     \end{block}
-
-%   \end{columns}
-% \end{frame}
-
-\begin{frame}
-  \frametitle{When Should I Use \MIC ?}
-
-  \begin{block}{Key points of SE10P}
-    \begin{itemize}
-      \item Large memory bandwidth (peak: 220GB/s)
-      \item 61 cores with mandatory use of hardware threading
-      \item 512-bit wide SIMD registers: 
-        \begin{itemize} 
-          \item FMA: up to 2x16 SP Flop/c (2x8 DP Flop/c)
-          \item otherwise: up to 16 SP Flop/c (8 DP Flop/c)
-        \end{itemize}
-      \item On a 61-core configuration at 1.05Ghz:
-        \begin{itemize} 
-        \item FMA: 2x16x61x1.05Ghz = 2.048 TFlop/s SP (1.024TFlop/s DP)
-        \item otherwise: 16x61x1.05Ghz = 1.024 GFlop/s SP (512GFlop/s DP)
-        \end{itemize}
-    \end{itemize}
-  \end{block}
-
-  Lots of bandwidth? Fused Multiply Add? Large vector registers?
-
-  Sounds like the perfect system for Sparse Matrix operations!
-\end{frame}
-
-% \begin{frame}
-%   \frametitle{Have to use vector operations to get high bandwidth}
-
-%   \begin{columns}
-%     \column{.49\linewidth}
-
-%     \begin{center}
-%       \includegraphics[width=\linewidth,page=1]{slides-figures/PPAM-data/bandwidth.pdf}
-
-%       Read
-%     \end{center}
-
-%     \column{.49\linewidth}
-
-%     \begin{center}
-%       \includegraphics[width=\linewidth,page=2]{slides-figures/PPAM-data/bandwidth.pdf}
-
-%       Write
-%     \end{center}
-
-%   \end{columns}
-
-%   \begin{itemize}
-%   \item Using the appropriate vectorial instructions gives significant
-%     improvements.
-%   \item Read Peak: 183GB/s.
-%   \item Write Peak: 160GB/s.
-%   \end{itemize}
-
-%   \tiny Source: PPAM 2013
-% \end{frame}
-
-\begin{frame}
-  \frametitle{Compute all the simple derivative at once}
-
-%  \begin{columns}
-%    \column{.5\linewidth}
   \begin{block}{Idea}
@@
         \begin{array}{c}
-            D_{x}\\ D_{y}\\ D_{z}\\ D_{hyp}
+          \textcolor{blue}{D_{x}}\\ \textcolor{red}{D_{y}}\\ \textcolor{green}{D_{z}}\\ D_{hyp}
         \end{array}\right)
-        \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) 
+      \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) \nonumber
       \vspace{-1ex}
@@
   \end{block}
-%    \column{.5\linewidth}
+
+\pause
 
   \begin{block}{Why would it work}
-      All the derivative matrices have the same sparsity pattern. In
-      other words, the non zeros are at the same positions in the
+    All the derivative matrices have the same sparsity pattern.
+
+    In other words, the non zeros are at the same positions in the
     matrix, only the values of the non zero change.
+
+    Benefits:
+    \begin{itemize}
+    \item Matrix compression: 1 column ID, 4 nonzero value. (gain: 38\%)
+    \item Vector reusage.
+    \end{itemize}
   \end{block}
-
-%  \end{columns}
 \end{frame}
@@
   \begin{columns}
-    \column{.49\linewidth}
+\only<1>{    \column{.49\linewidth}
     \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_best-crop.pdf}
@@
     \end{center}
-
+}
     \column{.49\linewidth}
@@
 
-    % \column{.33\linewidth}
-    % \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_no_cache-crop.pdf}
-    % \begin{center}
-    %   Worst case \\no cacheline
-    % \end{center}
+\only<2>{
+    \column{.49\linewidth}
+    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_no_cache-crop.pdf}
+    \begin{center}
+      Worst case \\no cacheline
+    \end{center}
+}
   \end{columns}
 
-  \tiny Also, taking cacheline into account makes a large difference (see paper for details).
+%  \tiny Also, taking cacheline into account makes a large difference (see paper for details).
 \end{frame}
@@
 
-\section{Implementation (4 vectors, 4 matrices)}
+\section{Implementation on Xeon Phi (4 vectors, 4 matrices, Single Precision)}
+
+% \begin{frame}
+%   \frametitle{Intel MIC: Overall Architecture}
+%   \begin{columns}
+%     \column{18em}    
+%     \includegraphics[width=18em]{slides-figures/MIC-overview.pdf}
+%     \column{17em}
+
+%     \begin{itemize}
+%       \item 8 memory controllers
+%         \begin{itemize}
+%           \item GDDR5
+%           \item 2 channels (32-bit)
+%           \item 5.5GT/s
+%           \item 352GB/s aggregated peak
+%             \begin{itemize}
+%               \item about same as a K40 
+%               \item but you typically get half
+%             \end{itemize}
+%         \end{itemize}
+%       \item 50+ cores
+%         \begin{itemize}
+%           \item 32KB of L1 cache
+%           \item 512KB of L2 cache
+%           \item LRU
+%           \item 8-way associative
+%         \end{itemize}
+%       \item 1 PCI-e controller
+%         \begin{itemize}
+%           \item to the host (2GB/s guaranteed to memory)
+%         \end{itemize}
+%     \end{itemize}
+%   \end{columns}
+
+% \end{frame}
+
+% \begin{frame}
+%   \frametitle{Core architecture}
+%   \begin{columns}
+%     \column{20em}
+%     \includegraphics[width=20em,height=20em]{slides-figures/MIC-core.png}%from intelsdg  
+%     \column{12em}
+%     \begin{itemize}
+%       \small
+%     \item Clocked at 1Ghz
+%     \item 64-bit
+%     \item 4 hardware threads
+%       \begin{itemize}
+%         \footnotesize
+%         \item no context switching
+%         \item no 2 instructions for the same thread
+%       \end{itemize}
+%     \item A vectorial unit
+%       \begin{itemize}
+%         \footnotesize
+%         \item 512-bit registers
+%         \item 16 SP floats
+%         \item 8 DP floats
+%         \item support FMA
+%       \end{itemize}
+%     \item Two instruction pipes:
+%       \begin{itemize}
+%         \footnotesize
+%         \item 2 ALU ops
+%         \item ALU + MEM ops
+%         \item ALU + VFP ops
+%         \item VFP + MEM ops
+%       \end{itemize}
+%     \item In-order execution
+%     \end{itemize}
+%   \end{columns}
+% \end{frame}
 
 \begin{frame}
-  \frametitle{Observation on the 4 vectors, 4 matrices case}
+  \frametitle{Summary of \MIC}
+
+  \begin{block}{Key points of Xeon Phi SE10P}
+    \begin{itemize}
+    \item Accelerator card (soon main processor with Knights Landing) 
+    \item Large memory bandwidth (peak: 220GB/s)
+    \item 61 cores with mandatory use of hardware threading
+    \item Mandatory use of 512-bit wide SIMD registers: 
+      \begin{itemize} 
+      \item FMA: up to 2x16 SP Flop/c (2x8 DP Flop/c)
+      \item otherwise: up to 16 SP Flop/c (8 DP Flop/c)
+      \end{itemize}
+    \item On a 61-core configuration at 1.05Ghz:
+      \begin{itemize} 
+      \item FMA: 2x16x61x1.05Ghz = 2.048 TFlop/s SP (1.024TFlop/s DP)
+      \item otherwise: 16x61x1.05Ghz = 1.024 GFlop/s SP (512GFlop/s DP)
+      \end{itemize}
+    \end{itemize}
+  \end{block}
+
+  Lots of bandwidth? Fused Multiply Add? Large vector registers?
+
+  Sounds like the perfect system for scientific computing!
+\end{frame}
+
+\begin{frame}
+  \frametitle{Observations on the 4 vectors, 4 matrices case}
 
   \begin{itemize}
-  \item We aim at over 200Gflop/s in single precision. (\~{}  10\% of peak.)
+  \item We aim at over 200Gflop/s in single precision.
+  \item Peak of Xeon Phi is 2TFlop/s.
+  \item We want 10\% of peak performance.
   \item One instruction in ten must be a Fused Multiply Add.
-  \item The kernel must be fairly instruction efficient
+  \item The kernel must be fairly instruction-efficient.
   \end{itemize}
@@
       \centering
-      \includegraphics[width=.6\linewidth]{slides-figures/ICS-figures/kd-tree-3d-norcm-crop.png}
+      \includegraphics[width=.6\linewidth]{slides-figures/ICS-figures/kd-tree-3d-rcm-crop.png}
     \end{block}
@@
 
+\begin{frame}
+  \frametitle{Thank you}
+
+  \begin{block}{Other related works}
+    \begin{itemize}
+    \item Scalability of graph algorithms on Intel MIC. (MTAAP 2012)
+    \item Sparse Matrix Vector Multiplication on Xeon Phi. (PPAM 2013)
+    \item Vectorization of Graph Centrality on Xeon Phi. (MTAAP 2014)
+    \end{itemize}
+  \end{block}
+
+  \begin{block}{More information}
+    Contact : esaule@uncc.edu
+
+    Visit: \url{http://webpages.uncc.edu/~esaule}
+  \end{block}
+\end{frame}
+
 \bibliographystyle{alpha}

### 2014-06-12 slides/2014_ics/slides-ics.tex
Modified a few plural errors, and replaced N by $N$ in a figure caption.

@@
       \item Radial symmetry.
-      \item RBFs approximate a function $f$ sampled at N distinct location $$f(\vec{r}) = \sum_i w_i \phi(||\vec{r}-\vec{r_i}||)$$ 
+      \item RBFs approximate a function $f$ sampled at $N$ distinct locations $$f(\vec{r}) = \sum_i w_i \phi(||\vec{r}-\vec{r_i}||)$$ 
       \end{itemize}
@@
 
-      The DM has exactly $n$ non zero per row.
+      The DM has exactly $n$ non-zeros per row.
     \end{block}
@@
     In other words, the non zeros are at the same positions in the
-    matrix, only the values of the non zero change.
+    matrix, only the non-zero values change.
 
@@
     \begin{itemize}
-    \item Matrix compression: 1 column ID, 4 nonzero value. (gain: 38\%)
+    \item Matrix compression: 1 column ID, 4 non-zero values. (gain: 38\%)
     \item Vector reusage.
@@
 
-    Number of floating point operation:\\$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$
+    Number of floating point operations:\\$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$
   \end{columns}

### 2014-06-12 figure_files/figure5/gflops_speedup_figure5.py

@@ -184,4 +184,3 @@ print ratios_32
 
 labels = [(v+m) for v in ["v1","v4"] for m in ["m1","m4"]]
 
-

### 2014-06-12 slides/2014_ics/slides-ics.tex
slieds

@@
 \only<1-2>{  \begin{block}{Shallow waters equations}
+    \vspace{-1em}
     \begin{eqnarray}
@@
     \end{equation}
+    \only<2>{$D_{hyp}$ is essentially a Laplacian operator used to adjust time stepping.}
   \end{block}
@@
 
-  \begin{block}{Why would it work}
+  \begin{block}{Why would it work ?}
     All the derivative matrices have the same sparsity pattern.
@@
     \column{.4\linewidth}
-    \scalebox{.55}{
+    \scalebox{.7}{
       \begin{tabular}{|c|l|}
@@
         $n_m$ & number of matrices \\
-        $s_M$ & size of the $n_m$ matrices in bytes\\
-        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
-        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
-        \hline
+%        $s_M$ & size of the $n_m$ matrices in bytes\\
+%        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
+%        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
+%        \hline
         $cl$    & size of a cache line in bytes\\
-        $b_{wT}$ & number of bytes written to memory  \\
-        $b_{rT}$ & minimum number of bytes read from memory  \\
-        $b_T$   & minimum number of bytes transferred \\
-        $B_{rT}$ & maximum number of bytes read from memory \\
-        $B_T$   & maximum number of bytes transferred \\
-        $O$     & number of floating point operations \\
-        $I_b$   & maximum computational intensity\\
-        $I_w$   & minimum computational intensity\\        
+%        $b_{wT}$ & number of bytes written to memory  \\
+%        $b_{rT}$ & minimum number of bytes read from memory  \\
+%        $b_T$   & minimum number of bytes transferred \\
+%        $B_{rT}$ & maximum number of bytes read from memory \\
+%        $B_T$   & maximum number of bytes transferred \\
+%        $O$     & number of floating point operations \\
+%        $I_b$   & maximum computational intensity\\
+%        $I_w$   & minimum computational intensity\\        
         \hline
@@
     }
+
+    \includegraphics[width=\linewidth]{slides-figures/PPAM-figures/spmv.pdf}
     \column{.6\linewidth}
@@
     Number of floating point operations:\\$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$
-  \end{columns}
 
-  \begin{center}
     Intensity (Best case):\\
-    $I_b = \frac{O}{b_T+ b_{wT}} = \frac{2 n_v n_m}{ (b_i + b_x n_m) + n_v n_m b_x n_z^{-1} + n_v b_x n_z^{-1} }$
+%    $I_b = \frac{O}{b_T+ b_{wT}} = \frac{2 n_v n_m}{ (b_i + b_x n_m) + n_v n_m b_x n_z^{-1} + n_v b_x n_z^{-1} }$
+    $I_b = \frac{O}{b_T+ b_{wT}}$
 
     Intensity (Worst case):\\
-    $I_w = \frac{O}{B_T+ b_{wT}} = \frac{2 n_v n_m}{(b_i+b_x n_m) + n_v n_m b_x n_z^{-1} + cl \ceil{\frac{n_vb_x}{cl}} }$
-  \end{center}
+%    $I_w = \frac{O}{B_T+ b_{wT}} = \frac{2 n_v n_m}{(b_i+b_x n_m) + n_v n_m b_x n_z^{-1} + cl \ceil{\frac{n_vb_x}{cl}} }$
+    $I_w = \frac{O}{B_T+ b_{wT}}$
+  \end{columns}
 \end{frame}
